### 2013-11-20 document/ms.tex
adding graphical model

\documentclass[12pt,preprint]{aastex}

% has to be before amssymb it seems
\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
            filecolor=linkcolor,urlcolor=linkcolor}

\usepackage{url}
\usepackage{algorithmic,algorithm}

\usepackage{listings}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{language=Python,
        basicstyle=\footnotesize\ttfamily,
        showspaces=false,
        showstringspaces=false,
        tabsize=2,
        breaklines=false,
        breakatwhitespace=true,
        identifierstyle=\ttfamily,
        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
    }

\input{vars}

\begin{document}

\title{%
    Probabilistic hierarchical modeling of exoplanet populations using
    incomplete catalogs
}

\newcommand{\nyu}{2}
\newcommand{\mpia}{3}
\author{%
    Daniel~Foreman-Mackey\altaffilmark{1,\nyu},
    David~W.~Hogg\altaffilmark{\nyu,\mpia},
    \etal
}
\altaffiltext{1}{To whom correspondence should be addressed:
                        \url{danfm@nyu.edu}}
\altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
                        Department of Physics, New York University,
                        4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
                        K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\begin{abstract}

\end{abstract}

\keywords{%
}

\section{Introduction}

blah.

\acknowledgments
SAMSI. %
It is a pleasure to thank
    \ldots
for helpful contributions to the ideas and code presented here.
This project was partially supported by the NSF (grant AST-0908357), and NASA
(grant NNX08AJ48G).

\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}
\begin{thebibliography}{}\raggedright

\bibitem[Fang \& Margot(2012)]{fang}
Fang, J., \& Margot, J.-L.\ 2012, \apj, 761, 92
\arxiv{1207.5250}

\bibitem[Tremaine \& Dong(2012)]{tremaine}
Tremaine, S., \& Dong, S.\ 2012, \aj, 143, 94
\arxiv{1106.5403}

\end{thebibliography}

\end{document}

### 2013-11-20 document/gm.py

@@ -1,0 +1,43 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import os
+from matplotlib import rc
+rc("font", family="serif", size=12)
+rc("text", usetex=True)
+rc("text.latex",
+   preamble=r"\input{{{0}}}"
+   .format(os.path.join(os.path.dirname(os.path.abspath(__file__)), "vars")))
+
+import daft
+
+pgm = daft.PGM([4.5, 3], origin=[-0.5, 1])
+
+pgm.add_node(daft.Node("population", r"\population", 0.0, 3.0, fixed=True))
+
+pgm.add_node(daft.Node("planet", r"\planet", 1.0, 3.0))
+pgm.add_node(daft.Node("planetobs", r"\planetobs", 2.0, 3.0, observed=True))
+
+pgm.add_node(daft.Node("stellar", r"\stellar", 3.0, 3.0))
+pgm.add_node(daft.Node("stellarobs", r"$\stellarobs$", 3.0, 2.0,
+                       observed=True))
+
+pgm.add_node(daft.Node("isobs", r"\isobs", 1.5, 2.0, observed=True))
+
+pgm.add_node(daft.Node("selection", r"\selection", 0.0, 2.0, fixed=True))
+
+pgm.add_edge("population", "planet")
+pgm.add_edge("planet", "planetobs")
+pgm.add_edge("planet", "isobs")
+pgm.add_edge("stellar", "planetobs")
+pgm.add_edge("stellar", "isobs")
+pgm.add_edge("stellar", "stellarobs")
+pgm.add_edge("selection", "isobs")
+
+pgm.add_plate(daft.Plate([0.4, 1.4, 3.1, 2.2], label=r"stars",
+                         position="bottom right"))
+pgm.add_plate(daft.Plate([0.5, 1.5, 2, 2], label=r"planets"))
+
+pgm.render()
+pgm.figure.savefig("gm.png", dpi=150)
+pgm.figure.savefig("gm.pdf")

### 2013-11-20 document/ms.tex
adding list of parameters and updating PGM

@@
 \usepackage{algorithmic,algorithm}
+\usepackage{multirow}
 
@@
 
-blah.
+\section{A toy model}
+
+We start by assuming that ever target star has some $K$ orbiting planets.
+Most of these planets are unobserved because they are two small to be
+detectable or because of geometry.
+The graphical model in \fig{gm} is a sketch of the hypothetical generative
+procedure for a catalog of exoplanet observations.
+The way it is shown, this model is extremely general but we will consider a
+very specific form here.
+In \fig{gm} the global ``hyper'' parameters are:
+\begin{itemize}
+\item{the parameters describing the distributions of the exoplanet population
+\population, and}
+\item{the parameters of a model of the catalog's exoplanet selection function
+\selection.}
+\end{itemize}
+Similarly, the ``hidden variables'' are:
+\begin{itemize}
+\item{the physical parameters of the star and the stellar system
+\stellar---including the stellar mass and radius, a measure of the stellar
+variability, and the mean inclination of the orbiting exoplanets---and}
+\item{the physical parameters of an exoplanet \planet---period, radius, and
+relative inclination.}
+\end{itemize}
+Finally, the observables are
+
+\begin{figure}[htbp]
+\begin{center}
+    \includegraphics{gm.pdf}
+\end{center}
+\caption{%
+\figlabel{gm}}
+\end{figure}
+
+\begin{deluxetable}{ccl}
+\tablecaption{%
+A description of the model parameters.
+\tablabel{parameters}}
+\tablewidth{0pt}
+\tablehead{& & \colhead{Description}}
+\startdata
+\multirow{3}{*}{\population}
+& $\population_\period$ & The period distribution \\
+& $\population_\radius$ & The exoplanet radius distribution \\
+& $\population_\relincl$ & The relative inclination distribution \\
+\hline
+\selection & --- & The parameters of the catalog selection function \\
+\hline
+$\isobs_{kn}$ & --- & A flag indicating the presence of exoplanet $k$ orbiting
+star $n$ in the catalog \\
+\hline
+\multirow{3}{*}{$\planet_{kn}$}
+& $\period_{kn}$ & The period of exoplanet $k$ orbiting star $n$ (or null) \\
+& $\radius_{kn}$ & The exoplanet's radius \\
+& $\relincl_{kn}$ & The inclination of this orbit away from the systemic mean \\
+\hline
+\multirow{3}{*}{$\planetobs_{kn}$}
+& $\periodobs_{kn}$ & The observed exoplanet period and uncertainty \\
+& $(\rorobs)_{kn}$ & The observed radius ratio and uncertainty \\
+& $\impactobs_{kn}$ & A constraint on the observed impact parameter \\
+\hline
+\multirow{4}{*}{$\stellar_{n}$}
+& $\smass_{n}$ & The mass of star $n$ \\
+& $\sradius_{n}$ & The radius of star $n$ \\
+& $\snoise_{n}$ & The ``variability'' of star $n$ \\
+& $\incl_{n}$ & The mean inclination of the exoplanets in the system $n$ \\
+\hline
+\multirow{2}{*}{$\stellarobs_{n}$}
+& $\sloggobs_{n}$ & The observed surface gravity (and uncertainty) of star $n$ \\
+& $\snoiseobs_{n}$ & The estimated variability of star $n$ \\
+\enddata
+\end{deluxetable}
 

### 2013-11-20 document/gm.py

@@ -11,18 +11,23 @@ rc("text.latex",
 
 import daft
 
-pgm = daft.PGM([4.5, 3], origin=[-0.5, 1])
+pgm = daft.PGM([4.5, 2.6], origin=[-0.5, 1])
+
+pgm.add_plate(daft.Plate([0.4, 1.1, 3.1, 2.4], label=r"stars $n=1,\ldots,N$",
+                         position="bottom right"))
+pgm.add_plate(daft.Plate([0.5, 1.4, 2, 2], label=r"planets $k=1,\ldots,K$"))
 
 pgm.add_node(daft.Node("population", r"\population", 0.0, 3.0, fixed=True))
 
-pgm.add_node(daft.Node("planet", r"\planet", 1.0, 3.0))
-pgm.add_node(daft.Node("planetobs", r"\planetobs", 2.0, 3.0, observed=True))
-
-pgm.add_node(daft.Node("stellar", r"\stellar", 3.0, 3.0))
-pgm.add_node(daft.Node("stellarobs", r"$\stellarobs$", 3.0, 2.0,
+pgm.add_node(daft.Node("planet", r"$\planet_{kn}$", 1.0, 3.0))
+pgm.add_node(daft.Node("planetobs", r"$\planetobs_{kn}$", 2.0, 3.0,
                        observed=True))
 
-pgm.add_node(daft.Node("isobs", r"\isobs", 1.5, 2.0, observed=True))
+pgm.add_node(daft.Node("stellar", r"$\stellar_n$", 3.0, 3.0))
+pgm.add_node(daft.Node("stellarobs", r"$\stellarobs_n$", 3.0, 2.0,
+                       observed=True))
+
+pgm.add_node(daft.Node("isobs", r"$\isobs_{kn}$", 1.5, 2.0, observed=True))
 
 pgm.add_node(daft.Node("selection", r"\selection", 0.0, 2.0, fixed=True))
 
@@ -33,10 +38,6 @@ pgm.add_edge("stellar", "planetobs")
 pgm.add_edge("stellar", "isobs")
 pgm.add_edge("stellar", "stellarobs")
 pgm.add_edge("selection", "isobs")
-
-pgm.add_plate(daft.Plate([0.4, 1.4, 3.1, 2.2], label=r"stars",
-                         position="bottom right"))
-pgm.add_plate(daft.Plate([0.5, 1.5, 2, 2], label=r"planets"))
 
 pgm.render()
 pgm.figure.savefig("gm.png", dpi=150)

### 2013-11-21 document/ms.tex
starting description of importance sampling scheme

@@
 very specific form here.
-In \fig{gm} the global ``hyper'' parameters are:
-\begin{itemize}
-\item{the parameters describing the distributions of the exoplanet population
-\population, and}
-\item{the parameters of a model of the catalog's exoplanet selection function
-\selection.}
-\end{itemize}
-Similarly, the ``hidden variables'' are:
-\begin{itemize}
-\item{the physical parameters of the star and the stellar system
-\stellar---including the stellar mass and radius, a measure of the stellar
-variability, and the mean inclination of the orbiting exoplanets---and}
-\item{the physical parameters of an exoplanet \planet---period, radius, and
-relative inclination.}
-\end{itemize}
-Finally, the observables are
+The parameters of this model are listed and described in \tab{parameters}.
+
+Using this model, we want to find---conditioned on the catalog of exoplanet
+candidates---joint constraints on the parameters describing the population of
+exoplanets and the selection function of the survey.
+To this end, we need to be able to efficiently marginalize over the parameters
+of all of the exoplanet systems (observed and unobserved) to compute the
+marginalized likelihood function
+\begin{eqnarray}
+p(\planetobs,\stellarobs,\isobs\,|\,\population,\selection) &=&
+\prod_{n=1}^N \int \left [
+p(\stellar_n)\,p(\stellarobs_n\,|\,\stellar_n) \,
+ \prod_{k=1}^K
+p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection)\right ]
+\dd \stellar_n
+\end{eqnarray}
+where
+\begin{eqnarray}\eqlabel{planet-marg}
+p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection) &=&
+\int
+p(\planet_{kn}\,|\,\population) \,
+p(\planetobs_{kn}\,|\,\planet_{kn},\stellar_{n}) \,
+p(\isobs_{kn}\,|\,\planet_{kn},\stellar_{n},\selection)
+\dd \planet_{kn} \quad.
+\end{eqnarray}
+
+We can evaluate the integral in \eq{planet-marg} using importance sampling.
+If the exoplanet is not observed ($\isobs_{kn}=0$), the most sensible proposal
+density is simply $\tilde{p}(\planet_{kn})=p(\planet_{kn}\,|\,\population)$
+and the integral becomes
+\begin{eqnarray}
+p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection) &=&
+\frac{1}{\sum_l \tilde{p}(x_l)}\,\sum_l
+p(\isobs_{kn}\,|\,\planet_{kn}=x_l,\stellar_{n},\selection)
+\end{eqnarray}
+for $x_l \sim p(\planet_{kn}\,|\,\population)$.
 

### 2013-11-21 simple/simple.py
starting simple example

@@ -1,0 +1,64 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import (division, print_function, absolute_import,
+                        unicode_literals)
+
+__all__ = []
+
+import numpy as np
+
+
+class LogNormal(object):
+
+    def __init__(self, mean, std):
+        self.mean = mean
+        self.std = std
+
+    def sample(self, size=1):
+        return np.exp(self.mean + self.std * np.random.randn(size))
+
+    def __call__(self, x):
+        return np.exp(-0.5 * (np.log(x)-self.mean)**2/self.std**2) \
+            / np.sqrt(2*np.pi) / self.std
+
+
+def generate_catalog(N, K, delta, rdist):
+    # Generate stellar radii.
+    R = 5 * np.random.rand(N)
+    R_err = R * (0.1 + 0.1 * np.random.rand(N))
+    R_obs = R + R_err * np.random.randn(N)
+
+    # Generate planetary radii.
+    r = rdist.sample(N*K).reshape((N, K))
+    ror = r / R[:, None]
+    ror_err = 1e-4 * np.random.rand(*ror.shape)
+    ror_obs = ror + ror_err * np.random.randn(*ror.shape)
+    star_id, tmp = np.meshgrid(np.arange(K), np.arange(N))
+
+    # Apply selection.
+    q = ror**2 > delta
+
+    return (
+        np.array(zip(R_obs, R_err)),
+        np.array(zip(star_id[q], ror_obs[q], ror_err[q]))
+    )
+
+
+if __name__ == "__main__":
+    import matplotlib.pyplot as pl
+
+    rdist = LogNormal(-5, 1)
+    stars, planets = generate_catalog(5000, 20, 1e-3, rdist)
+
+    # Plot observed histogram.
+    rstar = stars[:, 0][np.array(planets[:, 0], dtype=int)]
+    ror = planets[:, 1]
+    r = ror * rstar
+    pl.hist(np.log(r), 50, normed=True, histtype="step", color="k")
+
+    # Plot true distribution.
+    logr = np.linspace(-8, 1, 5000)
+    pl.plot(logr, rdist(np.exp(logr)), "k", lw=2)
+
+    pl.savefig("dist.png")

### 2013-11-21 document/ms.tex
title, abstract, and two (2)

@@
 \title{%
-    Probabilistic hierarchical modeling of exoplanet populations using
-    incomplete catalogs
+    Exoplanet population inferences
+    from a hierarchical model
+    of the \Kepler\ ``Object of Interest'' Catalog
 }
@@
 \begin{abstract}
-
+The catalog of \Kepler\ Objects of Interest (KOI) is a list of putative periodic exoplanet transit signals
+  discovered in the photometric data from the \Kepler Satellite.
+In principle this catalog contains a wealth of statistical information about the full population of exoplanets.
+Here we construct a model of the population that explains the content of the catalog.
+Our approach is novel in that
+  it takes the form of a rigorous (if approximate) hierarchical Bayesian inference,
+  it includes a parameterization of a non-trivial selection function (incompleteness),
+  and we fully marginalize out all latent and nuisance parameters using sampling.
+Because the model is properly hierarchical,
+  it can be constrained by multiple data sources;
+  we demonstrate this by adding the Solar System as a constraint,
+  but extensions to radial velocity, direct detection, and astrometric surveys would also be straightforward.
+Approximations abound:
+  Simple parameterized forms are assumed for the multivariate distribution of exoplanet parameters,
+  incompleteness is assumed to depend only on a scalar related to signal-to-noise,
+  and the KOI catalog is assumed to be dominated by real (not false-positive) exoplanet signals.
+We find XXX and YYY.
 \end{abstract}
@@
 \keywords{%
+methods:statistical
 }
@@
 We start by assuming that ever target star has some $K$ orbiting planets.
-Most of these planets are unobserved because they are two small to be
+Most of these planets are unobserved because they are two (2) small to be
 detectable or because of geometry.

### 2013-11-21 document/ms.tex
typo

@@
     from a hierarchical model
-    of the \Kepler\ ``Object of Interest'' Catalog
+    of the \kepler\ ``Object of Interest'' Catalog
 }
@@
 \begin{abstract}
-The catalog of \Kepler\ Objects of Interest (KOI) is a list of putative periodic exoplanet transit signals
-  discovered in the photometric data from the \Kepler Satellite.
+The catalog of \kepler\ Objects of Interest (KOI) is a list of putative periodic exoplanet transit signals
+  discovered in the photometric data from the \kepler Satellite.
 In principle this catalog contains a wealth of statistical information about the full population of exoplanets.

### 2013-11-22 simple/simple.py
hacking away at simple model

@@ -9,23 +9,48 @@ __all__ = []
 import numpy as np
 
 
-class LogNormal(object):
+class Normal(object):
 
     def __init__(self, mean, std):
-        self.mean = mean
         self.std = std
 
     def sample(self, size=1):
-        return np.exp(self.mean + self.std * np.random.randn(size))
+        return np.abs(self.std * np.random.randn(size))
 
     def __call__(self, x):
-        return np.exp(-0.5 * (np.log(x)-self.mean)**2/self.std**2) \
-            / np.sqrt(2*np.pi) / self.std
+        return 2*np.exp(-0.5*x*x/self.std**2) / np.sqrt(2*np.pi) / self.std
+
+
+class Histogram(object):
+
+    def __init__(self, bins, values=None):
+        nbins = len(bins) - 1
+        self.bins = np.array(bins)
+        if values is not None:
+            assert len(values) == nbins
+            self.values = np.array(values)
+        else:
+            self.values = np.ones(nbins)
+        self.normalize()
+
+    def normalize(self):
+        norm = np.sum(self.values * (self.bins[1:] - self.bins[:-1]))
+        self.values /= norm
+
+    def sample(self, size=1):
+        rands = np.random.rand(size)
+        cs = np.cumsum(self.values * (self.bins[1:] - self.bins[:-1]))
+        return np.interp(rands, np.append(0, cs), self.bins)
+
+    def __call__(self, x):
+        n = len(self.bins)
+        loc = np.interp(x, self.bins, np.arange(n), left=n-1, right=n-1)
+        return np.append(self.values, 0.0)[np.array(np.floor(loc), dtype=int)]
 
 
 def generate_catalog(N, K, delta, rdist):
     # Generate stellar radii.
-    R = 5 * np.random.rand(N)
+    R = 1.0 + 2 * np.random.rand(N)
     R_err = R * (0.1 + 0.1 * np.random.rand(N))
     R_obs = R + R_err * np.random.randn(N)
 
@@ -33,7 +58,7 @@ def generate_catalog(N, K, delta, rdist):
     r = rdist.sample(N*K).reshape((N, K))
     ror = r / R[:, None]
     ror_err = 1e-4 * np.random.rand(*ror.shape)
-    ror_obs = ror + ror_err * np.random.randn(*ror.shape)
+    ror_obs = np.abs(ror + ror_err * np.random.randn(*ror.shape))
[…]

### 2013-11-22 document/ms.tex

@@
 The catalog of \kepler\ Objects of Interest (KOI) is a list of putative periodic exoplanet transit signals
-  discovered in the photometric data from the \kepler Satellite.
+  discovered in the photometric data from the \kepler\ Satellite.
 In principle this catalog contains a wealth of statistical information about the full population of exoplanets.
@@
 
-\keywords{%
-methods:statistical
+\keywords{% CUT THESE DOWN TO SIX
+  astronomical~databases:~miscellaneous
+  ---
+  catalogs
+  ---
+  eclipses
+  ---
+  methods:~statistical
+  ---
+  planetary~systems
+  ---
+  planets~and~satellites:~detection
+  ---
+  planets~and~satellites:~fundamental~parameters
+  ---
+  stars:~statistics
+  ---
+  surveys
 }

### 2013-11-26 document/gm.py
simpler radius distribution PGM

@@ -11,32 +11,36 @@ rc("text.latex",
 
 import daft
 
-pgm = daft.PGM([4.5, 2.6], origin=[-0.5, 1])
+pgm = daft.PGM([6, 2.6], origin=[-0.5, 1])
 
-pgm.add_plate(daft.Plate([0.4, 1.1, 3.1, 2.4], label=r"stars $n=1,\ldots,N$",
+pgm.add_plate(daft.Plate([0.4, 1.1, 4.1, 2.4], label=r"stars $n=1,\ldots,N$",
                          position="bottom right"))
-pgm.add_plate(daft.Plate([0.5, 1.4, 2, 2], label=r"planets $k=1,\ldots,K$"))
+pgm.add_plate(daft.Plate([0.5, 1.4, 2.5, 2], label=r"planets $k=1,\ldots,K$"))
 
-pgm.add_node(daft.Node("population", r"\population", 0.0, 3.0, fixed=True))
-
-pgm.add_node(daft.Node("planet", r"$\planet_{kn}$", 1.0, 3.0))
-pgm.add_node(daft.Node("planetobs", r"$\planetobs_{kn}$", 2.0, 3.0,
-                       observed=True))
-
-pgm.add_node(daft.Node("stellar", r"$\stellar_n$", 3.0, 3.0))
-pgm.add_node(daft.Node("stellarobs", r"$\stellarobs_n$", 3.0, 2.0,
-                       observed=True))
-
-pgm.add_node(daft.Node("isobs", r"$\isobs_{kn}$", 1.5, 2.0, observed=True))
-
+pgm.add_node(daft.Node("rpop", r"\rpop", 0.0, 3.0, fixed=True))
 pgm.add_node(daft.Node("selection", r"\selection", 0.0, 2.0, fixed=True))
+pgm.add_node(daft.Node("rp", r"$\rp_{kn}$", 1.0, 3.0))
+pgm.add_node(daft.Node("ror", r"$\ror_{kn}$", 2.0, 3.0))
+pgm.add_node(daft.Node("rorobs", r"$\rorobs_{kn}$", 2.5, 2.5, observed=True))
+pgm.add_node(daft.Node("isobs", r"$\isobs_{kn}$", 2, 2, observed=True))
 
-pgm.add_edge("population", "planet")
-pgm.add_edge("planet", "planetobs")
-pgm.add_edge("planet", "isobs")
[…]

### 2013-11-26 document/gm.py
index order

@@ -19,10 +19,10 @@ pgm.add_plate(daft.Plate([0.5, 1.4, 2.5, 2], label=r"planets $k=1,\ldots,K$"))
 
 pgm.add_node(daft.Node("rpop", r"\rpop", 0.0, 3.0, fixed=True))
 pgm.add_node(daft.Node("selection", r"\selection", 0.0, 2.0, fixed=True))
-pgm.add_node(daft.Node("rp", r"$\rp_{kn}$", 1.0, 3.0))
-pgm.add_node(daft.Node("ror", r"$\ror_{kn}$", 2.0, 3.0))
-pgm.add_node(daft.Node("rorobs", r"$\rorobs_{kn}$", 2.5, 2.5, observed=True))
-pgm.add_node(daft.Node("isobs", r"$\isobs_{kn}$", 2, 2, observed=True))
+pgm.add_node(daft.Node("rp", r"$\rp_{nk}$", 1.0, 3.0))
+pgm.add_node(daft.Node("ror", r"$\ror_{nk}$", 2.0, 3.0))
+pgm.add_node(daft.Node("rorobs", r"$\rorobs_{nk}$", 2.5, 2.5, observed=True))
+pgm.add_node(daft.Node("isobs", r"$\isobs_{nk}$", 2, 2, observed=True))
 
 pgm.add_node(daft.Node("Rpop", r"\Rpop", 5, 3.0, fixed=True))
 pgm.add_node(daft.Node("Rs", r"$\Rs_{n}$", 3.5, 3.0))

### 2013-11-26 document/ms.tex
new title

@@
 \title{%
-    Exoplanet population inferences
-    from a hierarchical model
-    of the \kepler\ ``Object of Interest'' Catalog
+Inferring the exoplanet radius distribution from a noisy, incomplete catalog
 }
@@
 
-\section{A toy model}
-
-We start by assuming that ever target star has some $K$ orbiting planets.
-Most of these planets are unobserved because they are two (2) small to be
-detectable or because of geometry.
-The graphical model in \fig{gm} is a sketch of the hypothetical generative
-procedure for a catalog of exoplanet observations.
-The way it is shown, this model is extremely general but we will consider a
-very specific form here.
-The parameters of this model are listed and described in \tab{parameters}.
-
-Using this model, we want to find---conditioned on the catalog of exoplanet
-candidates---joint constraints on the parameters describing the population of
-exoplanets and the selection function of the survey.
-To this end, we need to be able to efficiently marginalize over the parameters
-of all of the exoplanet systems (observed and unobserved) to compute the
-marginalized likelihood function
-\begin{eqnarray}
-p(\planetobs,\stellarobs,\isobs\,|\,\population,\selection) &=&
-\prod_{n=1}^N \int \left [
-p(\stellar_n)\,p(\stellarobs_n\,|\,\stellar_n) \,
- \prod_{k=1}^K
-p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection)\right ]
-\dd \stellar_n
-\end{eqnarray}
-where
-\begin{eqnarray}\eqlabel{planet-marg}
-p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection) &=&
-\int
-p(\planet_{kn}\,|\,\population) \,
-p(\planetobs_{kn}\,|\,\planet_{kn},\stellar_{n}) \,
-p(\isobs_{kn}\,|\,\planet_{kn},\stellar_{n},\selection)
-\dd \planet_{kn} \quad.
-\end{eqnarray}
-
-We can evaluate the integral in \eq{planet-marg} using importance sampling.
-If the exoplanet is not observed ($\isobs_{kn}=0$), the most sensible proposal
-density is simply $\tilde{p}(\planet_{kn})=p(\planet_{kn}\,|\,\population)$
-and the integral becomes
-\begin{eqnarray}
-p(\planetobs_{kn},\isobs_{kn}\,|\,\stellar_n,\population,\selection) &=&
-\frac{1}{\sum_l \tilde{p}(x_l)}\,\sum_l
-p(\isobs_{kn}\,|\,\planet_{kn}=x_l,\stellar_{n},\selection)
-\end{eqnarray}
-for $x_l \sim p(\planet_{kn}\,|\,\population)$.
-
 \begin{figure}[htbp]
@@
 
-\begin{deluxetable}{ccl}
-\tablecaption{%
-A description of the model parameters.
-\tablabel{parameters}}
-\tablewidth{0pt}
-\tablehead{& & \colhead{Description}}
-\startdata
-\multirow{3}{*}{\population}
-& $\population_\period$ & The period distribution \\
-& $\population_\radius$ & The exoplanet radius distribution \\
-& $\population_\relincl$ & The relative inclination distribution \\
-\hline
-\selection & --- & The parameters of the catalog selection function \\
-\hline
-$\isobs_{kn}$ & --- & A flag indicating the presence of exoplanet $k$ orbiting
-star $n$ in the catalog \\
-\hline
-\multirow{3}{*}{$\planet_{kn}$}
-& $\period_{kn}$ & The period of exoplanet $k$ orbiting star $n$ (or null) \\
-& $\radius_{kn}$ & The exoplanet's radius \\
-& $\relincl_{kn}$ & The inclination of this orbit away from the systemic mean \\
-\hline
-\multirow{3}{*}{$\planetobs_{kn}$}
-& $\periodobs_{kn}$ & The observed exoplanet period and uncertainty \\
-& $(\rorobs)_{kn}$ & The observed radius ratio and uncertainty \\
-& $\impactobs_{kn}$ & A constraint on the observed impact parameter \\
-\hline
-\multirow{4}{*}{$\stellar_{n}$}
-& $\smass_{n}$ & The mass of star $n$ \\
-& $\sradius_{n}$ & The radius of star $n$ \\
-& $\snoise_{n}$ & The ``variability'' of star $n$ \\
-& $\incl_{n}$ & The mean inclination of the exoplanets in the system $n$ \\
-\hline
-\multirow{2}{*}{$\stellarobs_{n}$}
-& $\sloggobs_{n}$ & The observed surface gravity (and uncertainty) of star $n$ \\
-& $\snoiseobs_{n}$ & The estimated variability of star $n$ \\
-\enddata
-\end{deluxetable}
+% \begin{deluxetable}{ccl}
+% \tablecaption{%
+% A description of the model parameters.
+% \tablabel{parameters}}
+% \tablewidth{0pt}
+% \tablehead{& & \colhead{Description}}
+% \startdata
+% \multirow{3}{*}{\population}
+% & $\population_\period$ & The period distribution \\
+% & $\population_\radius$ & The exoplanet radius distribution \\
+% & $\population_\relincl$ & The relative inclination distribution \\
+% \hline
+% \selection & --- & The parameters of the catalog selection function \\
+% \hline
+% $\isobs_{kn}$ & --- & A flag indicating the presence of exoplanet $k$ orbiting
+% star $n$ in the catalog \\
+% \hline
+% \multirow{3}{*}{$\planet_{kn}$}
+% & $\period_{kn}$ & The period of exoplanet $k$ orbiting star $n$ (or null) \\
+% & $\radius_{kn}$ & The exoplanet's radius \\
+% & $\relincl_{kn}$ & The inclination of this orbit away from the systemic mean \\
+% \hline
+% \multirow{3}{*}{$\planetobs_{kn}$}
+% & $\periodobs_{kn}$ & The observed exoplanet period and uncertainty \\
+% & $(\rorobs)_{kn}$ & The observed radius ratio and uncertainty \\
+% & $\impactobs_{kn}$ & A constraint on the observed impact parameter \\
+% \hline
+% \multirow{4}{*}{$\stellar_{n}$}
+% & $\smass_{n}$ & The mass of star $n$ \\
+% & $\sradius_{n}$ & The radius of star $n$ \\
+% & $\snoise_{n}$ & The ``variability'' of star $n$ \\
+% & $\incl_{n}$ & The mean inclination of the exoplanets in the system $n$ \\
+% \hline
+% \multirow{2}{*}{$\stellarobs_{n}$}
+% & $\sloggobs_{n}$ & The observed surface gravity (and uncertainty) of star $n$ \\
+% & $\snoiseobs_{n}$ & The estimated variability of star $n$ \\
+% \enddata
+% \end{deluxetable}
 

### 2013-11-26 document/ms.tex
adding general discussion

@@
 
+\section{Hierarchical inference}
+
+If we assume that the exoplanetary systems that we have observed are drawn (in
+noisy and biased way) from some global distribution then it would be
+interesting to infer the shape of this global distribution conditioned on all
+the available data and our understanding of survey selection effects.
+This problem is called the \emph{hierarchical inference problem}.
+In theory, this problem is no different than a standard inference problem and
+the same algorithms might be applicable but in general, this class of problems
+refer to the cases where forward modelling directly to the raw data is
+computationally intractable.
+
+To be concrete, it is useful to consider a simple abstract example.
+\Fig{abstract} shows the probabilistic graphical model for a set of
+observations $\data_n$ taken some $N$ systems that are each parameterized by a
+set of physical parameters $\local_n$ drawn from some global distribution
+parameterized by the ``hyperparameters'' \hyper.
+Note that each $\data_n$ and $\local_n$ are, in general, large vectors (or
+blobs) with many entries.
+Given this structure, we would like to compute the marginalized likelihood of
+the full set of observations $\data = \{ \data_n \}$ given a specific
+parameterization \hyper
+\begin{eqnarray}
+p(\data\,|\,\hyper) &=& \int \dd \local\,p(\data,\,\local\,|\,\hyper)
+\nonumber\\
+&=& \prod_{n=1}^N\int \dd \local_n\,p(\data_n,\,\local_n\,|\,\hyper) \quad.
+\eqlabel{marg-like}
+\end{eqnarray}
+This integral isn't generally analytic but it could be solved using MCMC or
+another approximate numerical technique.
+A problem becomes apparent when we think a little harder about what might want
+to \emph{do} with this marginalized likelihood.
+In general, we will want to optimize it to find the maximum likelihood
+population distribution of apply a prior function---parameterized by
+\hyperhyper\ in \fig{abstract}---and draw posterior samples
+$\hyper \sim p(\hyper\,|\,\data)$.
+In both of these cases, we would have to perform the (probably extremely
+expensive) integration in \eq{marg-like} some impractically huge number of
+times.
+
+It is possible, however, to greatly simplify this problem and make it
+computationally tractable if we have a \emph{catalog}.
+In this context, a catalog is a set of posterior probabilistic constraints on
+the parameters $\local_n$ under some (known) \emph{interim priors}
+$p(w_n\,|\,\hyper_0)$.
+In particular, the best representation of these constraints is a set of
+posterior samples $\local_n^{(j)} \sim p(\local_n\,|\,\data_n,\,\hyper_0)$.
+Given a set of $J$ such samples, the integral in \eq{marg-like} can be
+re-written as
+\begin{eqnarray}
+\int \dd \local_n\,p(\data_n,\,\local_n\,|\,\hyper) &=&
+\int \dd \local_n\,p(\local_n\,|\,\hyper)\,p(\data_n\,|\,\local_n,\,\hyper)\,
+    \frac{p(\local_n\,|\,\data_n,\,\hyper_0)}
+         {p(\local_n\,|\,\data_n,\,\hyper_0)} \nonumber\\
+&=& Z_n\,\int \dd\local_n\,
+    \frac{p(\local_n\,|\,\hyper)}{p(\local_n\,|\,\hyper_0)} \,
+    \frac{p(\data_n\,|\,\local_n,\,\hyper)}{p(\data_n\,|\,\local_n,\,\hyper_0)}
+    \,p(\local_n\,|\,\data_n,\,\hyper_0) \nonumber\\
+&\approx& \frac{Z_n}{J} \, \sum_{j=1}^J
+    \frac{p(\local_n^{(j)}\,|\,\hyper)}{p(\local_n^{(j)}\,|\,\hyper_0)} \,
+    \frac{p(\data_n\,|\,\local_n^{(j)},\,\hyper)}
+         {p(\data_n\,|\,\local_n^{(j)},\,\hyper_0)}
+\eqlabel{full-import}
+\end{eqnarray}
+where
+\begin{eqnarray}
+Z_n &=& \int \dd\local_n\,p(\local_n,\,\data_n\,|\,\hyper_0)
+\end{eqnarray}
+is the ``evidence'' for the data under the interim priors.
+For our purposes, $Z_n$ is a constant that we (thankfully) never need to
+compute.
+
+There is a further simplification that we can make in the case where the
+likelihood function does not depend on the hyperparameters.
+In this case, \eq{full-import} becomes
+\begin{eqnarray}
+\int \dd \local_n\,p(\data_n,\,\local_n\,|\,\hyper) &\propto&
+\int \dd\local_n\,
+    \frac{p(\local_n^{(j)}\,|\,\hyper)}{p(\local_n^{(j)}\,|\,\hyper_0)} \,
+    p(\local_n\,|\,\data_n,\,\hyper_0) \nonumber\\
+&\approx& \frac{1}{J} \, \sum_{j=1}^J
+    \frac{p(\local_n^{(j)}\,|\,\hyper)}{p(\local_n^{(j)}\,|\,\hyper_0)} \quad.
+\end{eqnarray}
+This simplification won't always be completely applicable and we'll come back
+to it when we discuss the effects of survey selection functions below but most
+of the expensive computations can normally be cancelled.
+
+\begin{figure}[htbp]
+\begin{center}
+    \includegraphics{abstract.pdf}
+\end{center}
+\caption{%
+\figlabel{abstract}}
+\end{figure}
+
+In this \paper, we will focus on the problem of inferring the population
+parameters from a noisy incomplete \emph{catalog}.
+This catalog is defined as some set of posterior probability
+functions---preferably represented as samples---for some ``observable'' set of
+systems.
+
+\section{The model}
+
+\paragraph{Selection function}
+We assume that the selection function is specified by a probability
+distribution function that gives the probability of observing a transit with a
+given signal-to-noise ratio.
+Although we don't perform the integral explicitly, this assumption requires a
+marginalization over the period and inclination distribution.
+Instead, we choose to model the selection function as a scaled logistic
+function of signal-to-noise \sn:
+\begin{eqnarray}
+p(\isobs=1\,|\,\sn,\,\selection) &=&
+\frac{\Pmax}{1+\exp \left [ -(\sn-\snmin) / \selectwidth^2 \right ]} \quad.
+\end{eqnarray}
+Particular period and inclination distributions imply different values for
+\Pmax.
+
 \begin{figure}[htbp]

### 2013-11-26 document/abstract.py

@@ -1,0 +1,31 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import os
+from matplotlib import rc
+rc("font", family="serif", size=12)
+rc("text", usetex=True)
+rc("text.latex",
+   preamble=r"\input{{{0}}}"
+   .format(os.path.join(os.path.dirname(os.path.abspath(__file__)), "vars")))
+
+import daft
+
+pgm = daft.PGM([3, 2.5], origin=[-1, 1])
+
+pgm.add_plate(daft.Plate([-0.5, 1.4, 2.0, 1.1], label=r"$n=1,\cdots,N$",
+                         position="bottom right"))
+
+pgm.add_node(daft.Node("hyperhyper", r"\hyperhyper", -0.5, 3.0, fixed=True))
+pgm.add_node(daft.Node("hyper", r"$\hyper$", 0.5, 3.0))
+pgm.add_node(daft.Node("local", r"$\local_n$", 0, 2))
+pgm.add_node(daft.Node("data", r"$\data_n$", 1, 2, observed=True))
+
+pgm.add_edge("hyperhyper", "hyper")
+pgm.add_edge("hyper", "local")
+pgm.add_edge("hyper", "data")
+pgm.add_edge("local", "data")
+
+pgm.render()
+pgm.figure.savefig("abstract.png", dpi=150)
+pgm.figure.savefig("abstract.pdf")

### 2013-12-20 generate_catalog.py
generating a synthetic catalog

@@ -1,0 +1,30 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import (division, print_function, absolute_import,
+                        unicode_literals)
+
+__all__ = []
+
+import numpy as np
+np.random.seed(1000004)
+
+# Newton's constant in $R_\odot^3 M_\odot^{-1} {days}^{-2}$.
+G = 2945.4625385377644
+
+
+def generate_catalog(N, rate=5):
+    Rs = 0.5 + 0.7 * np.random.rand(N)
+    Ms = 0.5 + 0.7 * np.random.rand(N)
+    incl = np.arccos(1-2*np.random.rand(N))
+
+    K = np.random.poisson(rate, N)
+    r = [np.random.rand(k) for k in K]
+    a = [np.exp(8*np.random.rand(k)) for k in K]
+    P = [2*np.pi*np.sqrt(a0**3/(G*M)) for a0, M in zip(a, Ms)]
+    di = [np.radians(30*np.random.randn(k)) for k in K]
+    b = [a0*np.sin(i+di0) for i, di0, a0 in zip(incl, di, a)]
+
+
+if __name__ == "__main__":
+    generate_catalog(1000)

### 2013-12-20 generate_catalog.py
more catalogs

@@ -16,14 +16,19 @@ G = 2945.4625385377644
 def generate_catalog(N, rate=5):
     Rs = 0.5 + 0.7 * np.random.rand(N)
     Ms = 0.5 + 0.7 * np.random.rand(N)
+    sigs = 5e-4 + 1e-4 * np.random.rand(N)
     incl = np.arccos(1-2*np.random.rand(N))
 
     K = np.random.poisson(rate, N)
     r = [np.random.rand(k) for k in K]
+    z = [r0/R for r0, R in zip(r, Rs)]
     a = [np.exp(8*np.random.rand(k)) for k in K]
     P = [2*np.pi*np.sqrt(a0**3/(G*M)) for a0, M in zip(a, Ms)]
     di = [np.radians(30*np.random.randn(k)) for k in K]
-    b = [a0*np.sin(i+di0) for i, di0, a0 in zip(incl, di, a)]
+    b = [a0*np.sin(i+di0)/R for i, di0, a0, R in zip(incl, di, a, Rs)]
+    signal = [z0**2*np.sqrt(4.2/P0)*(b0+z0 < 1.0)
+              for z0, P0, b0 in zip(z, P, b)]
+    print([sum(s/sig>7) for s, sig in zip(signal, sigs)])
 
 
 if __name__ == "__main__":

### 2014-03-22 document/abstract.py
re-starting

@@ -1,31 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-import os
-from matplotlib import rc
-rc("font", family="serif", size=12)
-rc("text", usetex=True)
-rc("text.latex",
-   preamble=r"\input{{{0}}}"
-   .format(os.path.join(os.path.dirname(os.path.abspath(__file__)), "vars")))
-
-import daft
-
-pgm = daft.PGM([3, 2.5], origin=[-1, 1])
-
-pgm.add_plate(daft.Plate([-0.5, 1.4, 2.0, 1.1], label=r"$n=1,\cdots,N$",
-                         position="bottom right"))
-
-pgm.add_node(daft.Node("hyperhyper", r"\hyperhyper", -0.5, 3.0, fixed=True))
-pgm.add_node(daft.Node("hyper", r"$\hyper$", 0.5, 3.0))
-pgm.add_node(daft.Node("local", r"$\local_n$", 0, 2))
-pgm.add_node(daft.Node("data", r"$\data_n$", 1, 2, observed=True))
-
-pgm.add_edge("hyperhyper", "hyper")
-pgm.add_edge("hyper", "local")
-pgm.add_edge("hyper", "data")
-pgm.add_edge("local", "data")
-
-pgm.render()
-pgm.figure.savefig("abstract.png", dpi=150)
-pgm.figure.savefig("abstract.pdf")

### 2014-03-22 document/gm.py

@@ -1,48 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-import os
-from matplotlib import rc
-rc("font", family="serif", size=12)
-rc("text", usetex=True)
-rc("text.latex",
-   preamble=r"\input{{{0}}}"
-   .format(os.path.join(os.path.dirname(os.path.abspath(__file__)), "vars")))
-
-import daft
-
-pgm = daft.PGM([6, 2.6], origin=[-0.5, 1])
-
-pgm.add_plate(daft.Plate([0.4, 1.1, 4.1, 2.4], label=r"stars $n=1,\ldots,N$",
-                         position="bottom right"))
-pgm.add_plate(daft.Plate([0.5, 1.4, 2.5, 2], label=r"planets $k=1,\ldots,K$"))
-
-pgm.add_node(daft.Node("rpop", r"\rpop", 0.0, 3.0, fixed=True))
-pgm.add_node(daft.Node("selection", r"\selection", 0.0, 2.0, fixed=True))
-pgm.add_node(daft.Node("rp", r"$\rp_{nk}$", 1.0, 3.0))
-pgm.add_node(daft.Node("ror", r"$\ror_{nk}$", 2.0, 3.0))
-pgm.add_node(daft.Node("rorobs", r"$\rorobs_{nk}$", 2.5, 2.5, observed=True))
-pgm.add_node(daft.Node("isobs", r"$\isobs_{nk}$", 2, 2, observed=True))
-
-pgm.add_node(daft.Node("Rpop", r"\Rpop", 5, 3.0, fixed=True))
-pgm.add_node(daft.Node("Rs", r"$\Rs_{n}$", 3.5, 3.0))
-pgm.add_node(daft.Node("Rsobs", r"$\Rsobs_{n}$", 4, 2.5, observed=True))
-pgm.add_node(daft.Node("sigmasobs", r"$\sigmasobs_{n}$", 3.5, 2.0,
-                       observed=True))
-
-pgm.add_edge("rpop", "rp")
-pgm.add_edge("Rpop", "Rs")
-
-pgm.add_edge("Rs", "ror")
-pgm.add_edge("rp", "ror")
-pgm.add_edge("ror", "rorobs")
-
-pgm.add_edge("Rs", "Rsobs")
-
-pgm.add_edge("ror", "isobs")
-pgm.add_edge("sigmasobs", "isobs")
[…]

### 2014-03-22 generate_catalog.py

@@ -1,35 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import (division, print_function, absolute_import,
-                        unicode_literals)
-
-__all__ = []
-
-import numpy as np
-np.random.seed(1000004)
-
-# Newton's constant in $R_\odot^3 M_\odot^{-1} {days}^{-2}$.
-G = 2945.4625385377644
-
-
-def generate_catalog(N, rate=5):
-    Rs = 0.5 + 0.7 * np.random.rand(N)
-    Ms = 0.5 + 0.7 * np.random.rand(N)
-    sigs = 5e-4 + 1e-4 * np.random.rand(N)
-    incl = np.arccos(1-2*np.random.rand(N))
-
-    K = np.random.poisson(rate, N)
-    r = [np.random.rand(k) for k in K]
-    z = [r0/R for r0, R in zip(r, Rs)]
-    a = [np.exp(8*np.random.rand(k)) for k in K]
-    P = [2*np.pi*np.sqrt(a0**3/(G*M)) for a0, M in zip(a, Ms)]
-    di = [np.radians(30*np.random.randn(k)) for k in K]
-    b = [a0*np.sin(i+di0)/R for i, di0, a0, R in zip(incl, di, a, Rs)]
-    signal = [z0**2*np.sqrt(4.2/P0)*(b0+z0 < 1.0)
-              for z0, P0, b0 in zip(z, P, b)]
-    print([sum(s/sig>7) for s, sig in zip(signal, sigs)])
-
-
-if __name__ == "__main__":
-    generate_catalog(1000)

### 2014-03-22 hierarchical.py

@@ -1,202 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import (division, print_function, absolute_import,
-                        unicode_literals)
-
-__all__ = []
-
-import numpy as np
-
-_G = 2945.4625385377644
-
-
-class Uniform(object):
-
-    def __init__(self, mn, mx):
-        assert mx > mn
-        self.mn = mn
-        self.mx = mx
-
-    def __call__(self, value):
-        if np.isscalar(value):
-            if self.mn <= value <= self.mx:
-                return -np.log(self.mx-self.mn)
-            return -np.inf
-        m = (self.mn > value) + (value > self.mx)
-        r = -np.log(self.mx-self.mn) + np.zeros_like(value)
-        r[m] = -np.inf
-        return r
-
-    def sample(self, size=1):
-        return self.mn + (self.mx-self.mn)*np.random.rand(size)
-
-
-class Normal(object):
-
-    def __init__(self, mean, std):
-        self.mean = mean
-        self.std = std
-
-    @property
-    def std(self):
-        return self._std
-
-    @std.setter
-    def std(self, std):
-        var = std*std
-        self._std = std
-        self._mh_inv_var = -0.5/var
-        self._mh_ln_2pi_var = -0.5 * np.log(2*np.pi*var)
-
-    def __call__(self, value):
-        return (value-self.mean)**2*self._mh_inv_var+self._mh_ln_2pi_var
-
-    def sample(self, size=1):
-        return self.mean + self._std * np.random.randn(size)
-
-
-class Power(object):
-
-    def __init__(self, mn, mx, slope):
-        self.mn = mn
-        self.mx = mx
-        self.slope = slope
-
-    def __call__(self, value):
-        np1 = self.slope+1
-        x0, x1 = self.mn**np1, self.mx**np1
-        v = np.log(np1) - np.log(x1-x0) + self.slope * np.log(value)
-        if np.isscalar(value):
-            if self.mn <= value <= self.mx:
-                return v
-            return -np.inf
-        m = (self.mn > value) + (value > self.mx)
-        v[m] = -np.inf
-        return v
-
-    def sample(self, size=1):
-        y = np.random.rand(size)
-        np1 = self.slope+1
-        x0, x1 = self.mn**np1, self.mx**np1
-        return ((x1 - x0) * y + x0) ** (1.0/np1)
-
-
-class Selection(object):
-
[…]

### 2014-03-22 scripts/download_cdpp.py

@@ -1,35 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-import gevent
-from gevent import monkey
-monkey.patch_all()
-
-import os
-import urllib2
-
-url = "http://archive.stsci.edu/pub/kepler/catalogs/cdpp_quarter{0:d}.txt.gz"
-outdir = "data"
-try:
-    os.makedirs(outdir)
-except os.error:
-    print("Output directory already exists.")
-
-
-def download_quarter(n):
-    print("Downloading quarter {0}".format(n))
-    u = url.format(n)
-    fn = os.path.split(u)[1]
-    try:
-        data = urllib2.urlopen(u).read()
-    except Exception as e:
-        print("Quarter {0} failed with error:".format(n))
-        print(e)
-    else:
-        with open(os.path.join(outdir, fn), "wb") as f:
-            f.write(data)
-
-
-if __name__ == "__main__":
-    jobs = [gevent.spawn(download_quarter, n) for n in range(0, 17)]
-    gevent.joinall(jobs)

### 2014-03-22 population.py
Adding demo notebook

@@ -1,0 +1,375 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+__all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",
+           "NormalizedPopulation", "Dataset", "CensoringFunction"]
+
+import numpy as np
+from itertools import izip
+import matplotlib.pyplot as pl
+from scipy.misc import logsumexp
+
+
+class BrokenPowerLaw(object):
+
+    def __init__(self, bins, alpha_range=(-10, 10), beta_range=(-10, 10),
+                 logc_range=None):
+        self.bins = np.atleast_1d(bins)
+        self.logx = 0.5*(bins[1:] + bins[:-1])
+        self.ln_bin_widths = np.log(self.bins[1:] - self.bins[:-1])
+        self.alpha_range = alpha_range
+        self.beta_range = beta_range
+        if logc_range is None:
+            logc_range = (np.min(bins)-2, np.max(bins)+2)
+        self.logc_range = logc_range
+
+    def __len__(self):
+        return 3
+
+    def initial(self):
+        return np.array([0.0, 0.0, np.mean(self.bins)])
+
+    def lnprior(self, theta):
+        alpha, beta, logc = theta
+        if not self.alpha_range[0] < alpha < self.alpha_range[1]:
+            return -np.inf
+        if not self.beta_range[0] < beta < self.beta_range[1]:
+            return -np.inf
+        if not self.logc_range[0] < logc < self.logc_range[1]:
+            return -np.inf
+        return 0.0
+
+    def __call__(self, theta):
+        alpha, beta, logc = theta
+        x, xc = np.exp(self.logx), np.exp(logc)
+        v = alpha*self.logx + (beta-alpha)*np.log(0.5*(xc+x)) - beta*logc
+        v -= logsumexp(v + self.ln_bin_widths)
+        return v
+
+
+class Histogram(object):
+
+    def __init__(self, bins, resample=1):
+        self.bins = np.atleast_1d(bins)
+        self.resampled = self.bins[::resample]
+        assert self.resampled[-1] == self.bins[-1]
+
+        self.ln_bin_widths = np.log(self.resampled[1:] - self.resampled[:-1])
+        self.nbins = len(self.ln_bin_widths)
+        self.inds = np.arange(len(self.bins)-1) // resample
+
+    def __len__(self):
+        return self.nbins - 1
+
+    def initial(self):
[…]

### 2014-03-23 document/ms.tex
starting document

--- 2013-11-26 document/ms.tex
@@
 
-% has to be before amssymb it seems
 \usepackage{color,hyperref}
@@
             filecolor=linkcolor,urlcolor=linkcolor}
+\usepackage{url}
+\usepackage{amssymb,amsmath}
 
-\usepackage{url}
-\usepackage{algorithmic,algorithm}
-\usepackage{multirow}
+\newcommand{\project}[1]{{\sffamily #1}}
+\newcommand{\emcee}{\project{emcee}}
+\newcommand{\kepler}{\project{Kepler}}
+\newcommand{\license}{MIT License}
 
-\usepackage{listings}
-\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
-\lstset{language=Python,
-        basicstyle=\footnotesize\ttfamily,
-        showspaces=false,
-        showstringspaces=false,
-        tabsize=2,
-        breaklines=false,
-        breakatwhitespace=true,
-        identifierstyle=\ttfamily,
-        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
-        commentstyle=\color[rgb]{0.133,0.545,0.133},
-        stringstyle=\color[rgb]{0.627,0.126,0.941},
-    }
+\newcommand{\paper}{\emph{Article}}
 
-\input{vars}
+\newcommand{\foreign}[1]{\emph{#1}}
+\newcommand{\etal}{\foreign{et\,al.}}
+\newcommand{\etc}{\foreign{etc.}}
+
+\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
+\newcommand{\fig}[1]{\Fig{#1}}
+\newcommand{\figlabel}[1]{\label{fig:#1}}
+\newcommand{\Tab}[1]{Table~\ref{tab:#1}}
+\newcommand{\tab}[1]{\Tab{#1}}
+\newcommand{\tablabel}[1]{\label{tab:#1}}
+\newcommand{\Eq}[1]{Equation~(\ref{eq:#1})}
+\newcommand{\eq}[1]{\Eq{#1}}
+\newcommand{\eqlabel}[1]{\label{eq:#1}}
+\newcommand{\Sect}[1]{Section~\ref{sect:#1}}
+\newcommand{\sect}[1]{\Sect{#1}}
+\newcommand{\App}[1]{Appendix~\ref{sect:#1}}
+\newcommand{\app}[1]{\App{#1}}
+\newcommand{\sectlabel}[1]{\label{sect:#1}}
+
+\newcommand{\dd}{\ensuremath{\,\mathrm{d}}}
+\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
+
+% TO DOS
+\newcommand{\todo}[3]{{\color{#2} \emph{#1} TODO: #3}}
+\newcommand{\dfmtodo}[1]{\todo{DFM}{red}{#1}}
+\newcommand{\hoggtodo}[1]{\todo{HOGG}{blue}{#1}}
+\newcommand{\mortontodo}[1]{\todo{MORTON}{green}{#1}}
+
+% Document specific variables.
+\newcommand{\rate}{\ensuremath{\eta}}
+\newcommand{\ratepars}{\ensuremath{\theta}}
+\newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
+\newcommand{\radius}{\ensuremath{r}}
+\newcommand{\period}{\ensuremath{T}}
+\newcommand{\completeness}{\ensuremath{P_\mathrm{c}}}
+\newcommand{\transitprob}{\ensuremath{P_\mathrm{t}}}
+
+\newcommand{\interim}{\ensuremath{\alpha}}
 
@@
 \title{%
-Inferring the exoplanet radius distribution from a noisy, incomplete catalog
+    Inferring the rate of Earth-like exoplanets from noisy data
 }
@@
 \newcommand{\mpia}{3}
+\newcommand{\princeton}{4}
 \author{%
@@
     David~W.~Hogg\altaffilmark{\nyu,\mpia},
+    Timothy~D.~Morton\altaffilmark{\princeton},
     \etal
 }
-\altaffiltext{1}{To whom correspondence should be addressed:
+\altaffiltext{1}            {To whom correspondence should be addressed:
                              \url{danfm@nyu.edu}}
-\altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
+\altaffiltext{\nyu}         {Center for Cosmology and Particle Physics,
                              Department of Physics, New York University,
                              4 Washington Place, New York, NY, 10003, USA}
-\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
+\altaffiltext{\mpia}        {Max-Planck-Institut f\"ur Astronomie,
                              K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{\princeton}   {Department of Astrophysics, Princeton University,
+                             Princeton, NJ}
 
 \begin{abstract}
-The catalog of \kepler\ Objects of Interest (KOI) is a list of putative periodic exoplanet transit signals
-  discovered in the photometric data from the \kepler\ Satellite.
-In principle this catalog contains a wealth of statistical information about the full population of exoplanets.
-Here we construct a model of the population that explains the content of the catalog.
-Our approach is novel in that
-  it takes the form of a rigorous (if approximate) hierarchical Bayesian inference,
-  it includes a parameterization of a non-trivial selection function (incompleteness),
-  and we fully marginalize out all latent and nuisance parameters using sampling.
-Because the model is properly hierarchical,
-  it can be constrained by multiple data sources;
-  we demonstrate this by adding the Solar System as a constraint,
-  but extensions to radial velocity, direct detection, and astrometric surveys would also be straightforward.
-Approximations abound:
-  Simple parameterized forms are assumed for the multivariate distribution of exoplanet parameters,
-  incompleteness is assumed to depend only on a scalar related to signal-to-noise,
-  and the KOI catalog is assumed to be dominated by real (not false-positive) exoplanet signals.
-We find XXX and YYY.
+
+We infer the distribution and rate of Earth-like exoplanets around Sun-like
+stars.
+
 \end{abstract}
 
-\keywords{% CUT THESE DOWN TO SIX
-  astronomical~databases:~miscellaneous
+\keywords{%
+exoplanets: sickness
 ---
-  catalogs
+code: open-source
 ---
-  eclipses
-  ---
-  methods:~statistical
-  ---
-  planetary~systems
-  ---
-  planets~and~satellites:~detection
-  ---
-  planets~and~satellites:~fundamental~parameters
-  ---
-  stars:~statistics
-  ---
-  surveys
+keywords: made-up-by-Hogg
 }
@@
 
-\begin{figure}[htbp]
-\begin{center}
-    \includegraphics{gm.pdf}
-\end{center}
-\caption{%
-\figlabel{gm}}
-\end{figure}
+\section{The model}
 
-% \begin{deluxetable}{ccl}
-% \tablecaption{%
-% A description of the model parameters.
-% \tablabel{parameters}}
-% \tablewidth{0pt}
-% \tablehead{& & \colhead{Description}}
-% \startdata
-% \multirow{3}{*}{\population}
-% & $\population_\period$ & The period distribution \\
-% & $\population_\radius$ & The exoplanet radius distribution \\
-% & $\population_\relincl$ & The relative inclination distribution \\
-% \hline
-% \selection & --- & The parameters of the catalog selection function \\
-% \hline
-% $\isobs_{kn}$ & --- & A flag indicating the presence of exoplanet $k$ orbiting
-% star $n$ in the catalog \\
-% \hline
-% \multirow{3}{*}{$\planet_{kn}$}
-% & $\period_{kn}$ & The period of exoplanet $k$ orbiting star $n$ (or null) \\
-% & $\radius_{kn}$ & The exoplanet's radius \\
-% & $\relincl_{kn}$ & The inclination of this orbit away from the systemic mean \\
-% \hline
-% \multirow{3}{*}{$\planetobs_{kn}$}
-% & $\periodobs_{kn}$ & The observed exoplanet period and uncertainty \\
-% & $(\rorobs)_{kn}$ & The observed radius ratio and uncertainty \\
-% & $\impactobs_{kn}$ & A constraint on the observed impact parameter \\
-% \hline
-% \multirow{4}{*}{$\stellar_{n}$}
-% & $\smass_{n}$ & The mass of star $n$ \\
-% & $\sradius_{n}$ & The radius of star $n$ \\
-% & $\snoise_{n}$ & The ``variability'' of star $n$ \\
-% & $\incl_{n}$ & The mean inclination of the exoplanets in the system $n$ \\
-% \hline
-% \multirow{2}{*}{$\stellarobs_{n}$}
-% & $\sloggobs_{n}$ & The observed surface gravity (and uncertainty) of star $n$ \\
-% & $\snoiseobs_{n}$ & The estimated variability of star $n$ \\
-% \enddata
-% \end{deluxetable}
+\emph{Note: I think that we should use $T$ instead of $P$ for period because
+$P$ is too heavily overloaded by the probabilities\ldots}
+
+We have some parametric function \rate\ describing the true rate of exoplanets
+with periods \period\ and radii \radius
+\begin{eqnarray}\eqlabel{true-rate}
+\rate (\period,\,\radius;\,\ratepars)
+&=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}\quad.
+\end{eqnarray}
+Here we have not made any assumptions about the functional form of \rate, we
+are only stating that it can be modeled by some---possibly very large---set of
+parameters \ratepars.
+
+In this \paper, we will assume that the completeness function of the survey is
+known and it can be represented as a probability at every period and radius
+$\completeness (\period,\,\radius)$.
+Similarly, we can consider the marginalized transit probability $\transitprob
+(\period)$ as only a function of period.
+These selection effects can be incorporated into the rate from \eq{true-rate}
+to get the \emph{observed rate}
+\begin{eqnarray}\eqlabel{obs-rate}
+\obs{\rate} (\period,\,\radius;\,\ratepars)
+&=& \rate (\period,\,\radius;\,\ratepars)\,\completeness(\period,\,\radius)\,
+\transitprob(\period)\quad.
+\end{eqnarray}
+
+Now we will make the further assumption that every candidate in our sample of
+detected exoplanets is an independent Poisson sample from the population
+described by our rate function.
+Under this assumption, the likelihood of the catalog conditioned on a set of
+parameters \ratepars\ is
+\begin{eqnarray}
+\ln p(\period,\,\radius\,|\,\ratepars) &\propto&
+\sum_{k=1}^K \ln\obs{\rate} (\period_k,\,\radius_k;\,\ratepars)
+- \int \obs{\rate} (\period,\,\radius;\,\ratepars) \dd\period\dd\radius
+\quad.
+\end{eqnarray}
+
+\hoggtodo{How is this related to the version of the likelihood that doesn't
+include the normalization. What is the best way to introduce both forms?}
+
+\section{Including observational uncertainties}
+
+In the previous section, we implicitly assumed that the measurements of
+\period\ and \radius\ given by the catalog were perfect, without any
+observational uncertainties.
+This isn't a terrible approximation for the period measurements but not for
+the radii.
+Luckily, it is not hard to include these effects in our probabilistic
+framework.
+
+In general, catalogs produce posterior probability distributions for
+quantities conditioned on data.
+In particular, the catalog that we're considering here provides estimates of
+the posterior distribution of \period\ and \radius\ conditioned on the
+observables $\obs{\period}$ and $\obs{\radius}$ under some \emph{interim
+priors} parameterized by \interim
+\begin{eqnarray}
+p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim) \quad.
+\end{eqnarray}
+This estimate can be propagated as \emph{posterior samples}
+\begin{eqnarray}
+\{\period^{(n)},\,\radius^{(n)}\} &\sim&
+p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim) \quad.
+\end{eqnarray}
+
+Now the quantity that we need to compute is again the (marginalized)
+likelihood of the data $\obs{\period}$ and $\obs{\radius}$ conditioned on a
+set of rate function parameters \ratepars
+\begin{eqnarray}
+p(\obs{\period},\,\obs{\radius}\,|\,\ratepars) &=&
+\int p(\period,\,\radius,\,\obs{\period},\,\obs{\radius}\,|\,\ratepars)
+\dd\period\dd\radius \\
+&=&
+\int p(\period,\,\radius\,|\,\ratepars) \,
+p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)
+\dd\period\dd\radius \\
+&=&
+\int p(\period,\,\radius\,|\,\ratepars) \,
+p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)\,
+\frac{p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)}
+     {p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)}
+\dd\period\dd\radius \\
+&=&
+p(\obs{\period},\,\obs{\radius}\,|\,\interim)\,
+\int
+\frac{p(\period,\,\radius\,|\,\ratepars) \,
+      p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)}
+     {p(\period,\,\radius\,|\,\interim) \,
+      p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)}\,
+p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)
+\dd\period\dd\radius \\
+\frac{p(\obs{\period},\,\obs{\radius}\,|\,\ratepars)}
+     {p(\obs{\period},\,\obs{\radius}\,|\,\interim)}
+&=&
+\int
+\frac{p(\period,\,\radius\,|\,\ratepars)}
+     {p(\period,\,\radius\,|\,\interim)} \,
+p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)
+\dd\period\dd\radius \\
+&\approx&
+\frac{1}{N}
+\sum_{n=1}^N
+\frac{p(\period^{(n)},\,\radius^{(n)}\,|\,\ratepars)}
+     {p(\period^{(n)},\,\radius^{(n)}\,|\,\interim)}
+\end{eqnarray}
 
 \acknowledgments
-SAMSI. %
 It is a pleasure to thank
+    Erik Petigura (Berkeley),
     \ldots
@@
 
-\bibitem[Fang \& Margot(2012)]{fang}
-Fang, J., \& Margot, J.-L.\ 2012, \apj, 761, 92
-\arxiv{1207.5250}
-
-\bibitem[Tremaine \& Dong(2012)]{tremaine}
-Tremaine, S., \& Dong, S.\ 2012, \aj, 143, 94
-\arxiv{1106.5403}
+\bibitem[Petigura \etal(2013)]{petigura}
+Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013,
+Proceedings of the National Academy of Science, 110, 19273
 

### 2014-03-24 document/ms.tex
adding Gaussian process smoothness prior on bin heights

@@
 \newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
-\newcommand{\radius}{\ensuremath{r}}
+\newcommand{\radius}{\ensuremath{R}}
 \newcommand{\period}{\ensuremath{T}}
@@
 Under this assumption, the likelihood of the catalog conditioned on a set of
-parameters \ratepars\ is
+parameters \ratepars\ is (see Appendix XXXXX for a derivation)
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-\hoggtodo{How is this related to the version of the likelihood that doesn't
-include the normalization. What is the best way to introduce both forms?}
 

### 2014-03-24 population.py

@@ -10,6 +10,7 @@ import numpy as np
 from itertools import izip
 import matplotlib.pyplot as pl
 from scipy.misc import logsumexp
+from scipy.linalg import cho_factor, cho_solve
 
 
 class BrokenPowerLaw(object):
@@ -105,33 +106,64 @@ class Population(object):
         # Compute the cell areas.
         p = self.log_per_bins[::log_per_resample]
         rp = self.log_rp_bins[::log_rp_resample]
+        ir, ip = np.meshgrid(0.5*(rp[1:]+rp[:-1]), 0.5*(p[1:]+p[:-1]))
+        self.cell_coords = np.vstack((ip.flatten(), ir.flatten())).T
         self.ln_cell_area = (np.log(p[1:]-p[:-1])[:, None] +
                              np.log(rp[1:]-rp[:-1])[None, :])
         self.grid_shape = self.ln_cell_area.shape
         self.ln_cell_area = self.ln_cell_area.flatten()
         self.ncells = np.prod(len(self.ln_cell_area))
 
+        # Allocate the cache as empty.
+        self._cache_key = None
+        self._cache_val = None
+
     def __len__(self):
-        return self.ncells - 1
+        return self.ncells + 2
 
     def initial(self):
-        v = np.zeros(self.ncells)
-        v -= logsumexp(v + self.ln_cell_area)
+        v = np.zeros(self.ncells+3)
+        v[3:] -= logsumexp(v[3:] + self.ln_cell_area)
+        v[:3] = [0.5, -3.0, -3.0]
         return v[:-1]
 
-    def evaluate(self, theta):
+    def _get_grid(self, theta):
+        k = tuple(theta[3:])
+        if k == self._cache_key:
+            return self._cache_val
+
         # Compute the integral over the first N-1 cells.
-        norm = logsumexp(theta + self.ln_cell_area[:-1])
+        norm = logsumexp(theta[3:] + self.ln_cell_area[:-1])
         if norm >= 0.0:
             return None
 
-        # Compute the height of the last cell.
+        # Compute the height of the last cell and cache the heights.
         v = np.log(1.0 - np.exp(norm)) - self.ln_cell_area[-1]
-        ln_heights = np.append(theta, v).reshape(self.grid_shape)
-        return ln_heights[self.ix, self.iy]
+        self._cache_key = k
+        self._cache_val = np.append(theta[3:], v).reshape(self.grid_shape)
+
+        return self._cache_val
+
[…]

### 2014-03-25 document/ms.tex
an abstract attempt

@@
 \begin{abstract}
-
-We infer the distribution and rate of Earth-like exoplanets around Sun-like
-stars.
-
+Although no true extra-solar Earth analog is known,
+  hundreds of planets have been found around Sun-like stars
+  that are either Earth-sized but on shorter periods,
+  or else on year-long orbits but somewhat bigger than Earth.
+These populations ought to permit a
+  (possibly highly uncertain)
+  extrapolated measurement of the probability or rate at which Sun-like stars host true Earth analogs.
+Here we build a set of probabilistic models of the population of known small planets around G dwarfs.
+The models all take into account the transit probability and the completeness (or detection efficiency)
+  as a function of planet radius and orbital period.
+Because we forward model the data,
+  we do not ever have to weight the data by the inverse detection efficiency,
+  which leads to higher-variance population estimators.
+Our most sophisticated models do not assume that the joint radius--period distribution is separable,
+  and correctly take into account the significant observational uncertainties on the planet radii
+  using the machinery of hierarchical inference.
+One output of these models is fully marginalized estimates---%
+  marginalizing out observational uncertainties and all distribution parameters---%
+  of the rate density of true Earth analogs,
+  expressed as a number or Earths per star per natural logarithmic interval of radius and period.
+These rate densities depend on the permitted flexibility of the distribution model,
+  but tend to come out around ten percent,
+  with large uncertainties.
 \end{abstract}

### 2014-03-25 document/ms.tex
abstract tweaks

--- 2014-03-24 document/ms.tex
@@
 \begin{abstract}
-
-We infer the distribution and rate of Earth-like exoplanets around Sun-like
-stars.
-
+Although no true extra-solar Earth analog is known,
+  hundreds of planets have been found around Sun-like stars
+  that are either Earth-sized but on shorter periods,
+  or else on year-long orbits but somewhat bigger than Earth.
+These populations ought to permit a
+  (possibly highly uncertain)
+  extrapolated measurement of the probability or rate at which Sun-like stars host true Earth analogs.
+Here we build a set of probabilistic models of the population of known small planets around G dwarfs.
+The models all take into account the transit probability and the completeness (or detection efficiency)
+  as a function of planet radius and orbital period.
+Because we forward model the data,
+  we do not ever have to perform the traditional weighting of the data by the inverse detection efficiency;
+  our inferences are more precise than those produced by traditional methods.
+Our most sophisticated models do not assume that the joint radius--period distribution is separable,
+  and correctly take into account the significant observational uncertainties on the planet radii,
+  using the machinery of hierarchical inference.
+One output of these models is fully marginalized estimates---%
+  marginalizing out observational uncertainties and all distribution parameters---%
+  of the rate density of true Earth analogs,
+  expressed as a number of planets per star per natural logarithmic interval of radius and period,
+  evaluated at the properties of Earth.
+The rate density we infer depends strongly on the permitted flexibility of the distribution model,
+  but tends to come out around ten percent, with large uncertainty.
 \end{abstract}
@@
 exoplanets: sickness
+---
+exoplanets: eta-Earth
 ---

### 2014-03-25 document/ms.tex
introduction text

@@
 \section{Introduction}
+
+STUFF HERE ABOUT EARTH ANALOGS.
+
+This study is novel in a number of ways.
+We create probabilistic information about the full exoplanet population
+  \emph{not} by weighting the detected objects by the inverse detection efficiency (CITE PETIGURA),
+  but rather by forward modeling the observed distribution subject to the detection efficienty.
+Re-weighting the data by inverse selection probability---%
+  a method called V-max in the quasar and galaxy luminosity function contexts---%
+  is not wrong,
+  but it produces a higher-variance estimate of the population than a justified likelihood approach.
+Here our likelihood function is a Poisson rate likelihood;
+  it makes the fundamental assumption that the data points are independently drawn from the model,
+  but no other assumptions.
+Indeed, the assumptions behind this study are weaker than those of any previous study
+  of the Earth-analog population (CITE PETIGURA, DONG).
+
+Another new aspect of this study is that it makes use of very flexible distribution function models.
+Indeed most of the models we consider are ``non-parametric'' in the weak sense
+  that they have enormous numbers of free parameters
+  (not the strong sense of having an \emph{infinite} number of free parameters).
+The models are protected from degeneracies and over-fitting by priors.
+In some models we assume that the two-dimensional distribution of planets
+  in radius and period is separable;
+  in others we don't.
+We treat all of these parameters as ``nuisance parameters'' when we ask questions
+  about the occurrence of Earth analogs.
+That is, our rate conclusions are fully marginalized,
+  and those marginalizations are over non-trivial numbers of parameters.
+
+When we make conclusions about the radius disribution of exoplanets,
+  the distribution parameters are not nuisances but in fact our object.
+Even then, however, our results will have the period parameters marginalized out.
+
+Perhaps the most important novelty of this study is in its treatment of errors.
+Although planet periods are usually precisely measured,
+  planet radius measurements near Earth radius have large uncertainties.
+The justified probabilistic approach to including these uncertainties in a distribution analysis
+  is to introduce the ``true'' radii as latent variables and infer and marginalize them out
+  along with all other nuisance parameters.
+Although this sounds expensive---and it can be---%
+  there is a simple approach we have been advocating (CITE HOGG) that makes use of importance sampling.
+We use this approach here;
+  our results are fully marginalized over posterior uncertainties in all of the planet radii.
+More will be said about this below;
+  the importance sampling looks like a product over data points
+  of sums (averages) over individual-datum radius samplings.
 

### 2014-03-25 document/ms.tex
more abstract

@@
   using the machinery of hierarchical inference.
-One output of these models is fully marginalized estimates---%
+We reconsider and confirm with weaker assumptions
+  prior results on the distribution of exoplanet radii
+  that show a flattening of the distribution below a few Earth radii.
+One output of our models is fully marginalized estimates---%
   marginalizing out observational uncertainties and all distribution parameters---%

### 2014-03-25 document/ms.tex
title
intro text tweaks

@@
 \title{%
-    Inferring the rate of Earth-like exoplanets from noisy data
+  Inferring the distribution and rate of Earth-like exoplanets \\
+  from noisy individual-planet inferences
 }
@@
 
-When we make conclusions about the radius disribution of exoplanets,
-  the distribution parameters are not nuisances but in fact our object.
-Even then, however, our results will have the period parameters marginalized out.
+When, in what follows, we ask questions about the radius disribution of exoplanets,
+  the distribution parameters are no longer nuisances
+The radius-distribution parameters are what we are trying to measure.
+Even in this case, however, our results will have the period-distribution parameters marginalized out.
 
-Perhaps the most important novelty of this study is in its treatment of errors.
-Although planet periods are usually precisely measured,
-  planet radius measurements near Earth radius have large uncertainties.
+Perhaps the most important novelty of this study is in its treatment of observational uncertainties.
+Although planet periods in our data set are very precisely measured,
+  the planet radius measurements have large uncertainties.
 The justified probabilistic approach to including these uncertainties in a distribution analysis

### 2014-03-25 document/ms.tex
discussion start

@@
 \altaffiltext{\princeton}   {Department of Astrophysics, Princeton University,
-                             Princeton, NJ}
+                             Princeton, NJ, 08540, USA} % CHECK THIS ZIP CODE
 
@@
   but it produces a higher-variance estimate of the population than a justified likelihood approach.
-Here our likelihood function is a Poisson rate likelihood;
+One simple intuition along these lines is that the inverse probability weights can be large in dynamic range,
+  so the effective number of samples, after weighting, can be much smaller than the actual number of samples.
+Another is that it tends to upweight most strongly the most marginal detections in the data sample,
+  since these tend to be the samples that lie in the region of lowest completeness.
+In this work, we build a justified likelihood of the observed data,
+  taking into account the completeness function or detection efficiency.
+Our likelihood function is a variable-rate Poisson likelihood;
   it makes the fundamental assumption that the data points are independently drawn from the model,
-  but no other assumptions.
+  but is agnostic about all other properties of the data or distribution functions.
 Indeed, the assumptions behind this study are weaker than those of any previous study
@@
 
+\section{Results}
+
+\section{Discussion}
+
+We have made the first measurement of the Earth-like exoplanet population
+  that is based on forward modeling,
+  takes proper account of the observational uncertainties on planet radii,
+  and makes use of extremely flexible non-separable distributions for radius and period.
+We find BLAH BLAH BLAH.
+
+Although our method makes weaker assumptions than any previous work in this area,
+  of course it \emph{does} make strong assumptions.
+It assumes that the data set (provided by CITE PETIGURA)
+  contains planets and only planets (no false positives).
+It assumes that the individual-planet radius uncertainties
+  are also correctly determined.
+It assumes that the completeness calculation (CITE PETIGURA)
+  is accurate and smoothly varying with radius and period;
+  it makes no attempt to adjust the completeness star-by-star for brightness or variability.
+It assumes that the exoplanets are all independently drawn from the same distribution function,
+  with no regard to age, metallicity, or multiplicity.
+
+Although we have taken account of the radius uncertainties in a principled way,
+  we did make one important approximation:
+We ignored planets with central values far outside the radius and period box in which we built the model.
+That is, we assumed that there are no very large outliers in radius determination.
+This assumption is a complex one;
+  it involves the effective, end-to-end error distribution in the data set,
+  but also smoothness of the distribution function at or near the edges of the analysis box.
+
+If our inferences are right about true Earth analogs,
+  then we expect that there are about XXX transiting Earth analogs
+  hidden in the \kepler\ data set.
+We would be able to find some of these Earth analogs if the effective noise level in the data could be reduced
+  by a factor of about XXX.
+We have identified a number of directions along which this noise reduction might happen,
+  including but not limited to
+  retrospective data-driven or physics-driven recalibrations of the \kepler\ focal plane (CITE HOGG WHITEPAPER),
+  new methods for performing aperture photometry (NOTHING TO CITE),
+  and better models for stochastic stellar variability that take account of frequency structure (CITE CARTER, BREWER, AIGRAIN).
+If any (or all) of these deliver substantial improvements for \kepler\ archival data reanalyses,
+  it's Nobel Prize time!  (Or at least \textsl{The Colbert Report}?)
+
 \newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}

### 2014-03-25 document/ms.tex
petigura

@@
     Timothy~D.~Morton\altaffilmark{\princeton},
+    Erik~Petigura\altaffilmark{\berkeley},
     \etal
@@
                           K\"onigstuhl 17, D-69117 Heidelberg, Germany}
-\altaffiltext{\princeton}   {Department of Astrophysics, Princeton University,
+\altaffiltext{\princeton}{Department of Astrophysics, Princeton University,
                           Princeton, NJ, 08540, USA} % CHECK THIS ZIP CODE
+\altaffiltext{\berkeley} {Department of Astronomy, University of California at Berkeley,
+                          Berkeley, CA, XXXXX, USA} % CHECK THIS ZIP CODE
 

### 2014-03-26 document/ms.tex
bugfix
formatting shit

@@
 \newcommand{\princeton}{4}
+\newcommand{\berkeley}{5}
 \author{%
@@
     Timothy~D.~Morton\altaffilmark{\princeton},
-    Erik~Petigura\altaffilmark{\berkeley},
+    Erik~A.~Petigura\altaffilmark{\berkeley},
     \etal
@@
                           Princeton, NJ, 08540, USA} % CHECK THIS ZIP CODE
-\altaffiltext{\berkeley} {Department of Astronomy, University of California at Berkeley,
+\altaffiltext{\berkeley} {Department of Astronomy,
+                          University of California at Berkeley,
                           Berkeley, CA, XXXXX, USA} % CHECK THIS ZIP CODE
@@
 \begin{abstract}
-Although no true extra-solar Earth analog is known,
-  hundreds of planets have been found around Sun-like stars
-  that are either Earth-sized but on shorter periods,
-  or else on year-long orbits but somewhat bigger than Earth.
-These populations ought to permit a
-  (possibly highly uncertain)
-  extrapolated measurement of the probability or rate at which Sun-like stars host true Earth analogs.
-Here we build a set of probabilistic models of the population of known small planets around G dwarfs.
-The models all take into account the transit probability and the completeness (or detection efficiency)
-  as a function of planet radius and orbital period.
-Because we forward model the data,
-  we do not ever have to perform the traditional weighting of the data by the inverse detection efficiency;
-  our inferences are more precise than those produced by traditional methods.
-Our most sophisticated models do not assume that the joint radius--period distribution is separable,
-  and correctly take into account the significant observational uncertainties on the planet radii,
-  using the machinery of hierarchical inference.
-We reconsider and confirm with weaker assumptions
-  prior results on the distribution of exoplanet radii
-  that show a flattening of the distribution below a few Earth radii.
-One output of our models is fully marginalized estimates---%
-  marginalizing out observational uncertainties and all distribution parameters---%
-  of the rate density of true Earth analogs,
-  expressed as a number of planets per star per natural logarithmic interval of radius and period,
-  evaluated at the properties of Earth.
-The rate density we infer depends strongly on the permitted flexibility of the distribution model,
-  but tends to come out around ten percent, with large uncertainty.
+Although no true extra-solar Earth analog is known, hundreds of planets have
+been found around Sun-like stars that are either Earth-sized but on shorter
+periods, or else on year-long orbits but somewhat bigger than Earth.
+These populations ought to permit a (possibly highly uncertain) extrapolated
+measurement of the probability or rate at which Sun-like stars host true Earth
+analogs.
+Here we build a set of probabilistic models of the population of known small
+planets around G dwarfs.
+The models all take into account the transit probability and the completeness
+(or detection efficiency) as a function of planet radius and orbital period.
+Because we forward model the data, we do not ever have to perform the
+traditional weighting of the data by the inverse detection efficiency.
+Our most sophisticated models do not assume that the joint radius--period
+distribution is separable, and correctly take into account the significant
+observational uncertainties on the planet radii, using the machinery of
+hierarchical inference.
+We reconsider and confirm with weaker assumptions prior results on the
+distribution of exoplanet radii that show a flattening of the distribution
+below a few Earth radii.
+One output of our models is fully marginalized estimates---marginalizing out
+observational uncertainties and all distribution parameters---of the rate
+density of true Earth analogs, expressed as a number of planets per star per
+natural logarithmic interval of radius and period, evaluated at the properties
+of Earth.
+The rate density we infer depends strongly on the permitted flexibility of the
+distribution model, but tends to come out around ten percent, with large
+uncertainty.
 \end{abstract}

### 2014-03-26 document/ms.tex
the rules
extrapolation

@@
 of Earth.
-The rate density we infer depends strongly on the permitted flexibility of the
-distribution model, but tends to come out around ten percent, with large
-uncertainty.
+The rate density we infer is an extrapolation; it depends strongly on the
+permitted flexibility of the distribution model, but it tends to come out
+at around ten percent, with large uncertainty.
 \end{abstract}

### 2014-03-26 document/ms.tex
more formatting

--- 2014-03-26 document/ms.tex
@@
 \newcommand{\interim}{\ensuremath{\alpha}}
+
+%
+%  RULES OF THE GAME, MOTHAFUCKERS.
+%
+%  * 80 characters
+%  * line breaks at the ends of sentences
+%  * that is all.
+%
 
@@
 We create probabilistic information about the full exoplanet population
-  \emph{not} by weighting the detected objects by the inverse detection efficiency (CITE PETIGURA),
-  but rather by forward modeling the observed distribution subject to the detection efficienty.
-Re-weighting the data by inverse selection probability---%
-  a method called V-max in the quasar and galaxy luminosity function contexts---%
-  is not wrong,
-  but it produces a higher-variance estimate of the population than a justified likelihood approach.
-One simple intuition along these lines is that the inverse probability weights can be large in dynamic range,
-  so the effective number of samples, after weighting, can be much smaller than the actual number of samples.
-Another is that it tends to upweight most strongly the most marginal detections in the data sample,
-  since these tend to be the samples that lie in the region of lowest completeness.
-In this work, we build a justified likelihood of the observed data,
-  taking into account the completeness function or detection efficiency.
-Our likelihood function is a variable-rate Poisson likelihood;
-  it makes the fundamental assumption that the data points are independently drawn from the model,
-  but is agnostic about all other properties of the data or distribution functions.
-Indeed, the assumptions behind this study are weaker than those of any previous study
-  of the Earth-analog population (CITE PETIGURA, DONG).
+\emph{not} by weighting the detected objects by the inverse detection
+efficiency (CITE PETIGURA), but rather by forward modeling the observed
+distribution subject to the detection efficienty.
+Re-weighting the data by inverse selection probability---a method called
+V-max in the quasar and galaxy luminosity function contexts---is not wrong,
+but it produces a higher-variance estimate of the population than a justified
+likelihood approach.
+One simple intuition along these lines is that the inverse probability weights
+can be large in dynamic range, so the effective number of samples, after
+weighting, can be much smaller than the actual number of samples.
+Another is that it tends to upweight most strongly the most marginal
+detections in the data sample, since these tend to be the samples that lie in
+the region of lowest completeness.  In this work, we build a justified
+likelihood of the observed data, taking into account the completeness function
+or detection efficiency.
+Our likelihood function is a variable-rate Poisson likelihood; it makes the
+fundamental assumption that the data points are independently drawn from the
+model, but is agnostic about all other properties of the data or distribution
+functions.
+Indeed, the assumptions behind this study are weaker than those of any
+previous study of the Earth-analog population (CITE PETIGURA, DONG).
 
-Another new aspect of this study is that it makes use of very flexible distribution function models.
+Another new aspect of this study is that it makes use of very flexible
+distribution function models.
 Indeed most of the models we consider are ``non-parametric'' in the weak sense
-  that they have enormous numbers of free parameters
-  (not the strong sense of having an \emph{infinite} number of free parameters).
+that they have enormous numbers of free parameters (not the strong sense of
+having an \emph{infinite} number of free parameters).
 The models are protected from degeneracies and over-fitting by priors.
-In some models we assume that the two-dimensional distribution of planets
-  in radius and period is separable;
-  in others we don't.
-We treat all of these parameters as ``nuisance parameters'' when we ask questions
-  about the occurrence of Earth analogs.
-That is, our rate conclusions are fully marginalized,
-  and those marginalizations are over non-trivial numbers of parameters.
+In some models we assume that the two-dimensional distribution of planets in
+radius and period is separable; in others we don't.
+We treat all of these parameters as ``nuisance parameters'' when we ask
+questions about the occurrence of Earth analogs.
+That is, our rate conclusions are fully marginalized, and those
+marginalizations are over non-trivial numbers of parameters.
 
-When, in what follows, we ask questions about the radius disribution of exoplanets,
-  the distribution parameters are no longer nuisances
-The radius-distribution parameters are what we are trying to measure.
-Even in this case, however, our results will have the period-distribution parameters marginalized out.
+When, in what follows, we ask questions about the radius disribution of
+exoplanets, the distribution parameters are no longer nuisances The
+radius-distribution parameters are what we are trying to measure.
+Even in this case, however, our results will have the period-distribution
+parameters marginalized out.
 
-Perhaps the most important novelty of this study is in its treatment of observational uncertainties.
-Although planet periods in our data set are very precisely measured,
-  the planet radius measurements have large uncertainties.
-The justified probabilistic approach to including these uncertainties in a distribution analysis
-  is to introduce the ``true'' radii as latent variables and infer and marginalize them out
-  along with all other nuisance parameters.
-Although this sounds expensive---and it can be---%
-  there is a simple approach we have been advocating (CITE HOGG) that makes use of importance sampling.
-We use this approach here;
-  our results are fully marginalized over posterior uncertainties in all of the planet radii.
-More will be said about this below;
-  the importance sampling looks like a product over data points
-  of sums (averages) over individual-datum radius samplings.
+Perhaps the most important novelty of this study is in its treatment of
+observational uncertainties.
+Although planet periods in our data set are very precisely measured, the
+planet radius measurements have large uncertainties.
+The justified probabilistic approach to including these uncertainties in a
+distribution analysis is to introduce the ``true'' radii as latent variables
+and infer and marginalize them out along with all other nuisance parameters.
+Although this sounds expensive---and it can be---there is a simple approach
+we have been advocating (CITE HOGG) that makes use of importance sampling.
+We use this approach here; our results are fully marginalized over posterior
+uncertainties in all of the planet radii.
+More will be said about this below; the importance sampling looks like a
+product over data points of sums (averages) over individual-datum radius
+samplings.
 
@@
   by a factor of about XXX.
-We have identified a number of directions along which this noise reduction might happen,
-  including but not limited to
-  retrospective data-driven or physics-driven recalibrations of the \kepler\ focal plane (CITE HOGG WHITEPAPER),
-  new methods for performing aperture photometry (NOTHING TO CITE),
-  and better models for stochastic stellar variability that take account of frequency structure (CITE CARTER, BREWER, AIGRAIN).
-If any (or all) of these deliver substantial improvements for \kepler\ archival data reanalyses,
-  it's Nobel Prize time!  (Or at least \textsl{The Colbert Report}?)
+We have identified a number of directions along which this noise reduction
+might happen, including but not limited to retrospective data-driven or
+physics-driven recalibrations of the \kepler\ focal plane (CITE HOGG
+WHITEPAPER), new methods for performing aperture photometry (NOTHING TO CITE),
+and better models for stochastic stellar variability that take account of
+frequency structure (CITE CARTER, BREWER, AIGRAIN).
+If any (or all) of these deliver substantial improvements for \kepler\
+archival data reanalyses, it's Nobel Prize time!  (Or at least \textsl{The
+Colbert Report}?)
 

### 2014-03-26 document/ms.tex
starting figures and various paper things

--- 2014-03-26 document/ms.tex
@@
 %  * line breaks at the ends of sentences
+%  * eqnarrys ONLY
 %  * that is all.
@@
 
-\acknowledgments
-It is a pleasure to thank
-    Erik Petigura (Berkeley),
-    \ldots
-for helpful contributions to the ideas and code presented here.
-This project was partially supported by the NSF (grant AST-0908357), and NASA
-(grant NNX08AJ48G).
+\section{Separable Distribution}
 
-\section{Results}
+The first model that we consider for $\rate(\period,\,\radius)$ is a simple
+separable distribution.
+This assumption means that the rate can be written as
+\begin{eqnarray}
+\rate(\period,\,\radius) &=& \rate(\period)\,\rate(\radius)
+\end{eqnarray}
+and, more specifically, we model the period distribution as a broken power law
+\dfmtodo{CHECK ME!!}
+\begin{eqnarray}
+\rate(\period) &\propto& \frac{\period^\alpha \,
+(\period+\period_0)^{\beta-\alpha}}{\period_0^\beta}
+\end{eqnarray}
+and the radius distribution as a piecewise constant function (or histogram)
+where the parameters are the bin heights.
 
@@
 
-We have made the first measurement of the Earth-like exoplanet population
-  that is based on forward modeling,
-  takes proper account of the observational uncertainties on planet radii,
-  and makes use of extremely flexible non-separable distributions for radius and period.
+We have made the first measurement of the Earth-like exoplanet population that
+is based on forward modeling, takes proper account of the observational
+uncertainties on planet radii, and makes use of extremely flexible
+non-separable distributions for radius and period.
 We find BLAH BLAH BLAH.
 
-Although our method makes weaker assumptions than any previous work in this area,
-  of course it \emph{does} make strong assumptions.
-It assumes that the data set (provided by CITE PETIGURA)
-  contains planets and only planets (no false positives).
-It assumes that the individual-planet radius uncertainties
-  are also correctly determined.
-It assumes that the completeness calculation (CITE PETIGURA)
-  is accurate and smoothly varying with radius and period;
-  it makes no attempt to adjust the completeness star-by-star for brightness or variability.
-It assumes that the exoplanets are all independently drawn from the same distribution function,
-  with no regard to age, metallicity, or multiplicity.
+Although our method makes weaker assumptions than any previous work in this
+area, of course it \emph{does} make strong assumptions.
+It assumes that the data set (provided by CITE PETIGURA) contains planets and
+only planets (no false positives).
+It assumes that the individual-planet radius uncertainties are also correctly
+determined.
+It assumes that the completeness calculation (CITE PETIGURA) is accurate and
+smoothly varying with radius and period; it makes no attempt to adjust the
+completeness star-by-star for brightness or variability.
+It assumes that the exoplanets are all independently drawn from the same
+distribution function, with no regard to age, metallicity, or multiplicity.
 
-Although we have taken account of the radius uncertainties in a principled way,
-  we did make one important approximation:
-We ignored planets with central values far outside the radius and period box in which we built the model.
-That is, we assumed that there are no very large outliers in radius determination.
-This assumption is a complex one;
-  it involves the effective, end-to-end error distribution in the data set,
-  but also smoothness of the distribution function at or near the edges of the analysis box.
+Although we have taken account of the radius uncertainties in a principled
+way, we did make one important approximation: We ignored planets with central
+values far outside the radius and period box in which we built the model.
+That is, we assumed that there are no very large outliers in radius
+determination.
+This assumption is a complex one; it involves the effective, end-to-end error
+distribution in the data set, but also smoothness of the distribution function
+at or near the edges of the analysis box.
 
-If our inferences are right about true Earth analogs,
-  then we expect that there are about XXX transiting Earth analogs
-  hidden in the \kepler\ data set.
-We would be able to find some of these Earth analogs if the effective noise level in the data could be reduced
-  by a factor of about XXX.
+If our inferences are right about true Earth analogs, then we expect that
+there are about XXX transiting Earth analogs hidden in the \kepler\ data set.
+We would be able to find some of these Earth analogs if the effective noise
+level in the data could be reduced by a factor of about XXX.
 We have identified a number of directions along which this noise reduction
@@
 Colbert Report}?)
+
+\acknowledgments
+It is a pleasure to thank
+    \ldots
+for helpful contributions to the ideas and code presented here.
+This project was partially supported by the NSF (grant AST-0908357), and NASA
+(grant NNX08AJ48G).
 

### 2014-03-26 document/ms.tex
generalities

@@
 %  * eqnarrys ONLY
+%  * ALWAYS cite inside parens () and use \citealt{}; no in-line cites
 %  * that is all.
@@
 
-\section{The model}
+\section{Model generalities}
+
+From our perspective, a ``model'' is a likelihood function---a probability
+density (or distribution) function (PDF) for the data given
+parameteters---and a set of prior PDFs for the parameters.
+In what follows, we will treat as ``data'' a certain set of exoplanet radius
+and period measurements (\citealt{petigura}).
+We will treat the uncertainties on those measurements as ``prior
+information''; that is, we will treat those as true as given and not as being
+generated by the parameterized likelihood function.
+
+In the models that follow, the most important parameters will be parameters
+of the PDF for the exoplanet radii and periods.
+However, there will also be ``latent parameters'' that represent the ``true''
+values of the exoplanet radii and periods.
+That is, for each exoplanet, there is a ``measured'' or ``observed'' radius
+and period, which constitute our data.
+These relate to the unobserved ``true'' radius and period by a noise model.
+For the conclusions of this study, the true values of radius and period are
+nuisance parameters; we will marginalize them out.
 

### 2014-03-26 document/ms.tex
bibitem
cite Dong

@@
 Indeed, the assumptions behind this study are weaker than those of any
-previous study of the Earth-analog population (CITE PETIGURA, DONG).
+previous study of the radius and period distribution of the Earth-analog
+population (\citealt{dong, petigura}).
 
@@
 \begin{thebibliography}{}\raggedright
-
+\bibitem[Dong \& Zhu(2013)]{dong}
+Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 
 \bibitem[Petigura \etal(2013)]{petigura}
@@
 Proceedings of the National Academy of Science, 110, 19273
-
 \end{thebibliography}

### 2014-03-26 document/code/load_data.py
figures and results for first figure

@@ -19,9 +19,10 @@ bp = os.path.join(base, "data")
 stars = pd.read_hdf(os.path.join(bp, "stlr.h5"), "stlr")
 G = 2945.4625385377644
 P = 10.0
-transit_lnprobs = np.log(stars.Rstar) + (np.log(4*np.pi*np.pi)
-                                         - np.log(G*P*P*stars.Mstar)) / 3
-transit_lnprob0 = np.median(transit_lnprobs)
+# transit_lnprobs = np.log(stars.Rstar) + (np.log(4*np.pi*np.pi)
+#                                          - np.log(G*P*P*stars.Mstar)) / 3
+# transit_lnprob0 = np.median(transit_lnprobs)
+transit_lnprob0 = -2.98353340397
 ln_period0 = np.log(P)
 
 
@@ -46,3 +47,19 @@ def load_candidates(censor, samples=0, rp_func=np.log):
     log_per_obs = np.array(log_per_obs, dtype=float)
     log_rp_obs = np.array(log_rp_obs, dtype=float)
     return population.Dataset(log_per_obs, log_rp_obs, censor)
+
+
+def load_petigura_bins(mylog=np.log):
+    ep_Rp_logbins = 0.5 * np.log10(2) * np.arange(9)
+    ep_Rp_lnbins = mylog(10 ** ep_Rp_logbins)
+    ep_Rp_values = np.array([12., 14.2, 18.6, 5.9, 1.9, 1.0, 0.9, 0.7])
+    norm = np.sum(ep_Rp_values * (ep_Rp_lnbins[1:] - ep_Rp_lnbins[:-1]))
+    ep_Rp_pdf = ep_Rp_values / norm
+
+    ep_p_logbins = np.log10([6.25, 12.5, 25, 50, 100])
+    ep_p_lnbins = mylog(10 ** ep_p_logbins)
+    ep_p_values = np.array([8.9, 13.7, 15.8, 15.2])
+    norm = np.sum(ep_p_values * (ep_p_lnbins[1:] - ep_p_lnbins[:-1]))
+    ep_p_pdf = ep_p_values / norm
+
+    return ep_p_lnbins, ep_Rp_lnbins, ep_p_pdf, ep_Rp_pdf

### 2014-03-26 document/code/separable1.py

@@ -3,21 +3,26 @@
 
 from __future__ import division, print_function
 
+import emcee
 import numpy as np
-import matplotlib.pyplot as pl
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates)
+import cPickle as pickle
+from multiprocessing import Pool
 
+from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
+                       load_candidates, load_petigura_bins)
 from population import (CensoringFunction, BrokenPowerLaw, Histogram,
                         SeparablePopulation, NormalizedPopulation,
                         ProbabilisticModel)
 
 # Load the data.
+ep = load_petigura_bins()
 ln_P_inj, ln_R_inj, recovered = load_completenes_sim()
 censor = CensoringFunction(ln_P_inj, ln_R_inj, recovered,
-                           nbin_log_per=48, nbin_log_rp=48,
-                           log_per_range=np.log([5, 400]),
-                           log_rp_range=np.log([0.5, 16]),
+                           nbin_log_per=48, nbin_log_rp=32,
+                           log_per_range=(ep[0].min(), ep[0].max()),
+                           log_rp_range=(ep[1].min(), ep[1].max()),
+                           # log_per_range=np.log([5, 400]),
+                           # log_rp_range=np.log([0.5, 16]),
                            transit_lnprob0=transit_lnprob0,
                            ln_period0=ln_period0)
 dataset = load_candidates(censor)
@@ -37,10 +42,34 @@ dataset = load_candidates(censor)
 
 # Build the probabilistic model.
 log_per_dist = BrokenPowerLaw(censor.log_per_bins)
-log_rp_dist = Histogram(censor.log_rp_bins, resample=8)
+log_rp_dist = Histogram(censor.log_rp_bins, resample=4)
 pop = NormalizedPopulation(11., SeparablePopulation(log_per_dist, log_rp_dist))
 model = ProbabilisticModel(dataset, pop)
 
 # Plot the initial distribution.
-fig = pop.plot(pop.initial())
+fig = pop.plot(pop.initial() + 0.1 * np.random.randn(len(pop.initial())),
+               ep=ep)
 fig.savefig("separable1-initial.pdf")
+
+# Set up the sampler.
+p0 = pop.initial()
+ndim, nwalkers = len(p0), 32
[…]

### 2014-03-26 document/ms.tex

@@
 \emph{not} by weighting the detected objects by the inverse detection
-efficiency (CITE PETIGURA), but rather by forward modeling the observed
-distribution subject to the detection efficienty.
+efficiency (\citealt{dong, petigura}), but rather by forward modeling the
+observed distribution subject to the detection efficiency.
 Re-weighting the data by inverse selection probability---a method called
@@
 Although this sounds expensive---and it can be---there is a simple approach
-we have been advocating (CITE HOGG) that makes use of importance sampling.
+we have been advocating (\cite{hogge}) that makes use of importance sampling.
 We use this approach here; our results are fully marginalized over posterior
@@
 area, of course it \emph{does} make strong assumptions.
-It assumes that the data set (provided by CITE PETIGURA) contains planets and
+It assumes that the data set (from \citealt{petigura}) contains planets and
 only planets (no false positives).
@@
 determined.
-It assumes that the completeness calculation (CITE PETIGURA) is accurate and
-smoothly varying with radius and period; it makes no attempt to adjust the
-completeness star-by-star for brightness or variability.
+It assumes that the completeness calculation (again from \citealt{petigura})
+is accurate and smoothly varying with radius and period; it makes no attempt
+to adjust the completeness star-by-star for brightness or variability.
 It assumes that the exoplanets are all independently drawn from the same
@@
 might happen, including but not limited to retrospective data-driven or
-physics-driven recalibrations of the \kepler\ focal plane (CITE HOGG
-WHITEPAPER), new methods for performing aperture photometry (NOTHING TO CITE),
-and better models for stochastic stellar variability that take account of
-frequency structure (CITE CARTER, BREWER, AIGRAIN).
+physics-driven recalibrations of the \kepler\ focal plane
+(\citealt{hoggwhitepaper}), new methods for performing aperture photometry,
+and better models for stochastic instrument variability or stochastic
+intrinsic stellar variability that take account of frequency structure
+(\citealt{brewer, carter, roberts}).
 If any (or all) of these deliver substantial improvements for \kepler\
@@
 (grant NNX08AJ48G).
+This research made use of the NASA \project{Astrophysics Data System}.
 
@@
 \begin{thebibliography}{}\raggedright
+\bibitem[Brewer \& Stello(2009)]{brewer}
+Brewer, B.~J., \& Stello, D.\ 2009, \mnras, 395, 2226 (\arxiv{0902.3907})
+\bibitem[Carter \& Winn(2009)]{carter}
+Carter, J.~A., \& Winn, J.~N.\ 2009, \apj, 704, 51 (\arxiv{0909.0747})
 \bibitem[Dong \& Zhu(2013)]{dong}
-Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 
+Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 (\arxiv{1212.4853})
+\bibitem[Hogg et \etal(2013)]{hoggwhitepaper}
+Hogg, D.~W., Angus, R., Barclay, T., et al.\ 2013, \arxiv{1309.0653}
+\bibitem[Hogg \etal(2010)]{hogge}
+Hogg, D.~W., Myers, A.~D., \& Bovy, J.\ 2010, \apj, 725, 2166 (\arxiv{1008.4146})
 \bibitem[Petigura \etal(2013)]{petigura}
 Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013,
-Proceedings of the National Academy of Science, 110, 19273
+Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})
+\bibitem[Roberts \etal(2013)]{roberts}
+Roberts, S., McQuillan, A., Reece, S., \& Aigrain, S.\ 2013, \mnras, 435, 3639 (\arxiv{1308.3644})
 \end{thebibliography}

### 2014-03-27 document/ms.tex
notes to self

@@
 
-STUFF HERE ABOUT EARTH ANALOGS.
+\hoggtodo{
+\kepler\ has been busting out exoplanets (CITE).
+Many of them even appear possibly rocky and habitable (CITE).
+Nonetheless, none yet known are true Earth analogs.
+}
+
+\hoggtodo{
+This hasn't stopped anyone from estimating the rate at which Sun-like stars
+host Earth-like planets on year-ish period orbits (CITE).
+All of these rate estimates are extrapolations of one kind or another.
+Here we perform an extremely conservative extrapolation of this kind.
+}
+
+\hoggtodo{
+What would make an extrapolation like this conservative?
+Using \emph{huge bins} does not a conservative estimate make.
+You don't want to assume separabilities you can't justify.
+You want to explore an enormous function space for the exoplanet rates or pdf.
+We are going to do be conservative in these senses here, far more conservative
+than anyone who has come before.
+}
 
@@
 observed distribution subject to the detection efficiency.
+\hoggtodo{Check that Dong \& Zhu do indeed weight by inverse probability.}
 Re-weighting the data by inverse selection probability---a method called
@@
 non-separable distributions for radius and period.
-We find BLAH BLAH BLAH.
+We find \hoggtodo{BLAH BLAH BLAH.}
 

### 2014-03-27 document/code/load_data.py
more efficient resampling of uncertainties

@@ -32,15 +32,22 @@ def load_completenes_sim(rp_func=np.log):
     return np.log(sim.inj_P), rp_func(sim.inj_Rp), m
 
 
-def load_candidates(censor, samples=0, rp_func=np.log):
+def load_candidates(censor, samples=0, N=10000, rp_func=np.log):
     log_per_obs = []
     log_rp_obs = []
     for line in open(os.path.join(bp, "table_ekoi836.tex")):
         cols = line.split("&")
-        log_per_obs.append(np.log(float(cols[2])))
+        log_per = np.log(float(cols[2]))
+        log_per_obs.append(log_per)
         if samples > 0:
-            log_rp_obs.append(rp_func(float(cols[12]) +
-                              float(cols[13])*np.random.randn(samples)))
+            m, s = map(float, cols[12:14])
+            r = m + s*np.sort(np.random.randn(N))
+            lnw = censor.evaluate([log_per], rp_func(r)) - 0.5 * ((r-m)/s)**2
+            cummulative = np.cumsum(np.exp(lnw))
+            cummulative /= cummulative[-1]
+            rnd = np.random.rand(samples)
+            r = np.interp(rnd, cummulative, r)
+            log_rp_obs.append(rp_func(r))
         else:
             log_rp_obs.append(rp_func(float(cols[12])))
 

### 2014-03-27 document/code/plot_results.py

@@ -1,0 +1,28 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import os
+import sys
+import numpy as np
+import cPickle as pickle
+
+from load_data import load_petigura_bins
+
+mylog = np.log
+rp_label = "\ln R"
+if "--linear" in sys.argv:
+    mylog = lambda x: x
+    rp_label = "R"
+
+ep = load_petigura_bins(mylog=mylog)
+censor, dataset, pop, sampler = pickle.load(open(sys.argv[1]))
+
+samples = sampler.chain
+samples = samples.reshape((-1, samples.shape[-1]))
+
+fig_per, fig_rp = pop.plot(samples[np.random.randint(len(samples), size=24)],
+                           rp_label=rp_label, ep=ep)
+fig_per.savefig(os.path.splitext(sys.argv[1])[0]+"-period.png")
+fig_rp.savefig(os.path.splitext(sys.argv[1])[0]+"-radius.png")

### 2014-03-27 document/code/separable-linear.py

@@ -1,0 +1,48 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import emcee
+import numpy as np
+import cPickle as pickle
+
+from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
+                       load_candidates, load_petigura_bins)
+from population import (CensoringFunction, BrokenPowerLaw, Histogram,
+                        SeparablePopulation, NormalizedPopulation,
+                        ProbabilisticModel)
+
+# Load the data.
+mylog = lambda x: x
+ep = load_petigura_bins(mylog=mylog)
+ln_P_inj, R_inj, recovered = load_completenes_sim(rp_func=mylog)
+censor = CensoringFunction(ln_P_inj, R_inj, recovered,
+                           nbin_log_per=48, nbin_log_rp=48,
+                           log_per_range=np.log([5, 400]),
+                           log_rp_range=mylog([0.5, 16]),
+                           transit_lnprob0=transit_lnprob0,
+                           ln_period0=ln_period0)
+dataset = load_candidates(censor, rp_func=mylog)
+
+# Build the probabilistic model.
+log_per_dist = BrokenPowerLaw(censor.log_per_bins)
+log_rp_dist = Histogram(censor.log_rp_bins, resample=4)
+pop = NormalizedPopulation(11., SeparablePopulation(log_per_dist, log_rp_dist))
+model = ProbabilisticModel(dataset, pop)
+
+# Set up the sampler.
+p0 = pop.initial()
+ndim, nwalkers = len(p0), 32
+pos = [p0 + 1e-8 * np.random.randn(ndim) for i in range(nwalkers)]
+
+# Make sure that all the initial positions have finite probability.
+assert np.all(np.isfinite(map(model.lnprob, pos)))
+
+# Set up the sampler.
+sampler = emcee.EnsembleSampler(nwalkers, ndim, model)
+pos, lp, state = sampler.run_mcmc(pos, 3000)
+sampler.reset()
+sampler.run_mcmc(pos, 2000)
+
+pickle.dump((censor, dataset, pop, sampler), open("separable-linear.pkl", "wb"), -1)

### 2014-03-27 document/code/separable2.py

@@ -3,11 +3,9 @@
 
 from __future__ import division, print_function
 
-import sys
 import emcee
 import numpy as np
 import cPickle as pickle
-import matplotlib.pyplot as pl
 
 from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates, load_petigura_bins)
@@ -30,32 +28,13 @@ dataset = load_candidates(censor)
 
 # Build the probabilistic model.
 log_per_dist = BrokenPowerLaw(censor.log_per_bins)
-log_rp_dist = Histogram(censor.log_rp_bins, resample=4)
+log_rp_dist = Histogram(censor.log_rp_bins, resample=2)
 pop = NormalizedPopulation(11., SeparablePopulation(log_per_dist, log_rp_dist))
 model = ProbabilisticModel(dataset, pop)
 
-if "--plot" in sys.argv:
-    # Plot the completeness function.
-    pl.figure(figsize=(10, 6))
-    x = censor.log_per_bins
-    y = censor.log_rp_bins
-    pl.pcolor(x, y, np.exp(censor.lncompleteness).T, cmap="gray")
-    pl.plot(dataset.log_per_obs, dataset.log_rp_obs, ".r", ms=3)
-    pl.xlim(censor.log_per_bins.min(), censor.log_per_bins.max())
-    pl.ylim(censor.log_rp_bins.min(), censor.log_rp_bins.max())
-    pl.xlabel(r"$\ln P$")
-    pl.ylabel(r"$\ln R_P$")
-    pl.colorbar()
-    pl.savefig("separable1-completeness.pdf")
-
-    # Plot the initial distribution.
-    fig = pop.plot(pop.initial() + 0.1 * np.random.randn(len(pop.initial())),
-                   ep=ep)
-    fig.savefig("separable1-initial.pdf")
-
 # Set up the sampler.
 p0 = pop.initial()
-ndim, nwalkers = len(p0), 32
+ndim, nwalkers = len(p0), 64
 pos = [p0 + 1e-8 * np.random.randn(ndim) for i in range(nwalkers)]
 
 # Make sure that all the initial positions have finite probability.
@@ -67,4 +46,4 @@ pos, lp, state = sampler.run_mcmc(pos, 3000)
 sampler.reset()
 sampler.run_mcmc(pos, 2000)
 
-pickle.dump((censor, dataset, pop, sampler), open("separable1.pkl", "wb"), -1)
[…]

### 2014-03-27 document/code/separable3.py

@@ -3,11 +3,9 @@
 
 from __future__ import division, print_function
 
-import sys
 import emcee
 import numpy as np
 import cPickle as pickle
-import matplotlib.pyplot as pl
 
 from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates, load_petigura_bins)
@@ -18,23 +16,29 @@ from population import (CensoringFunction, BrokenPowerLaw, Histogram,
 # Load the data.
 ep = load_petigura_bins()
 ln_P_inj, ln_R_inj, recovered = load_completenes_sim()
+
+per_rng = np.log([5.0, 400.0])
+rp_rng = np.log([16.0, 64.0])
+
+K = 50000
+ln_P_inj = np.append(ln_P_inj, per_rng[0] +
+                     (per_rng[1] - per_rng[0]) * np.random.rand(K))
+ln_R_inj = np.append(ln_R_inj, rp_rng[0] +
+                     (rp_rng[1]-rp_rng[0]) * np.random.rand(K))
+recovered = np.append(recovered, np.ones(K, dtype=bool))
+
 censor = CensoringFunction(ln_P_inj, ln_R_inj, recovered,
-                           nbin_log_per=48, nbin_log_rp=32,
-                           log_per_range=(ep[0].min(), ep[0].max()),
-                           log_rp_range=(ep[1].min(), ep[1].max()),
-                           # log_per_range=np.log([5, 400]),
-                           # log_rp_range=np.log([0.5, 16]),
+                           nbin_log_per=48, nbin_log_rp=70,
+                           log_per_range=np.log([5, 400]),
+                           log_rp_range=np.log([0.5, 64]),
+                           # log_per_range=(ep[0].min(), ep[0].max()),
+                           # log_rp_range=(ep[1].min(), 2 * ep[1].max()),
                            transit_lnprob0=transit_lnprob0,
                            ln_period0=ln_period0)
-dataset = load_candidates(censor)
+dataset = load_candidates(censor, samples=64)
 
-# Build the probabilistic model.
-log_per_dist = BrokenPowerLaw(censor.log_per_bins)
-log_rp_dist = Histogram(censor.log_rp_bins, resample=4)
[…]

### 2014-03-27 population.py

@@ -18,6 +18,7 @@ class BrokenPowerLaw(object):
     def __init__(self, bins, alpha_range=(-10, 10), beta_range=(-10, 10),
                  logc_range=None):
         self.bins = np.atleast_1d(bins)
+        self.base = np.array(self.bins)
         self.logx = 0.5*(bins[1:] + bins[:-1])
         self.ln_bin_widths = np.log(self.bins[1:] - self.bins[:-1])
         self.alpha_range = alpha_range
@@ -52,14 +53,20 @@ class BrokenPowerLaw(object):
 
 class Histogram(object):
 
-    def __init__(self, bins, resample=1):
+    def __init__(self, bins, base=None):
         self.bins = np.atleast_1d(bins)
-        self.resampled = self.bins[::resample]
-        assert self.resampled[-1] == self.bins[-1]
+        if base is None:
+            self.base = np.array(bins)
+        else:
+            self.base = np.atleast_1d(base)
+        assert self.bins[0] == self.base[0]
+        assert self.bins[-1] == self.base[-1]
 
-        self.ln_bin_widths = np.log(self.resampled[1:] - self.resampled[:-1])
+        self.ln_bin_widths = np.log(self.bins[1:] - self.bins[:-1])
         self.nbins = len(self.ln_bin_widths)
-        self.inds = np.arange(len(self.bins)-1) // resample
+        self.inds = np.digitize(0.5*(self.base[1:]+self.base[:-1]),
+                                self.bins) - 1
+        assert np.all((self.inds >= 0) * (self.inds < len(self.bins)))
 
     def __len__(self):
         return self.nbins - 1
@@ -88,24 +95,37 @@ class Population(object):
     poisson = False
 
     def __init__(self, log_per_bins, log_rp_bins,
-                 log_per_resample=1, log_rp_resample=1):
+                 log_per_base=None, log_rp_base=None):
         self.log_per_bins = np.atleast_1d(log_per_bins)
         self.log_rp_bins = np.atleast_1d(log_rp_bins)
+        if log_per_base is None:
+            self.log_per_base = np.array(self.log_per_bins)
+        else:
+            self.log_per_base = np.atleast_1d(log_per_base)
+        if log_rp_base is None:
+            self.log_rp_base = np.array(self.log_rp_bins)
[…]

### 2014-03-27 document/ms.tex
typo
order of period and radius

@@
 density of true Earth analogs, expressed as a number of planets per star per
-natural logarithmic interval of radius and period, evaluated at the properties
+natural logarithmic interval of period and radius, evaluated at the properties
 of Earth.
@@
 Indeed, the assumptions behind this study are weaker than those of any
-previous study of the radius and period distribution of the Earth-analog
+previous study of the period and radius distribution of the Earth-analog
 population (\citealt{dong, petigura}).
@@
 In some models we assume that the two-dimensional distribution of planets in
-radius and period is separable; in others we don't.
+period and radius is separable; in others we don't.
 We treat all of these parameters as ``nuisance parameters'' when we ask
@@
 parameteters---and a set of prior PDFs for the parameters.
-In what follows, we will treat as ``data'' a certain set of exoplanet radius
-and period measurements (\citealt{petigura}).
+In what follows, we will treat as ``data'' a certain set of exoplanet period
+and radius measurements (\citealt{petigura}).
 We will treat the uncertainties on those measurements as ``prior
@@
 In the models that follow, the most important parameters will be parameters
-of the PDF for the exoplanet radii and periods.
+of the PDF for the exoplanet periods and radii.
 However, there will also be ``latent parameters'' that represent the ``true''
-values of the exoplanet radii and periods.
-That is, for each exoplanet, there is a ``measured'' or ``observed'' radius
-and period, which constitute our data.
-These relate to the unobserved ``true'' radius and period by a noise model.
-For the conclusions of this study, the true values of radius and period are
+values of the exoplanet periods and radii.
+That is, for each exoplanet, there is a ``measured'' or ``observed'' period
+and radius, which constitute our data.
+These relate to the unobserved ``true'' period and radius by a noise model.
+For the conclusions of this study, the true values of period and radius are
 nuisance parameters; we will marginalize them out.
@@
 uncertainties on planet radii, and makes use of extremely flexible
-non-separable distributions for radius and period.
+non-separable distributions for period and radius.
 We find \hoggtodo{BLAH BLAH BLAH.}
@@
 It assumes that the completeness calculation (again from \citealt{petigura})
-is accurate and smoothly varying with radius and period; it makes no attempt
+is accurate and smoothly varying with period and radius; it makes no attempt
 to adjust the completeness star-by-star for brightness or variability.
@@
 way, we did make one important approximation: We ignored planets with central
-values far outside the radius and period box in which we built the model.
+values far outside the period and radius box in which we built the model.
 That is, we assumed that there are no very large outliers in radius
@@
 Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 (\arxiv{1212.4853})
-\bibitem[Hogg et \etal(2013)]{hoggwhitepaper}
+\bibitem[Hogg \etal(2013)]{hoggwhitepaper}
 Hogg, D.~W., Angus, R., Barclay, T., et al.\ 2013, \arxiv{1309.0653}

### 2014-03-27 document/ms.tex
some notes in the generalities section

@@
 
+\emph{Note: I think that we should use $T$ instead of $P$ for period because
+$P$ is too heavily overloaded by the probabilities\ldots}
+
 From our perspective, a ``model'' is a likelihood function---a probability
@@
 parameteters---and a set of prior PDFs for the parameters.
-In what follows, we will treat as ``data'' a certain set of exoplanet period
+In what follows, we will treat as ``data'' a catalog of exoplanet period
 and radius measurements (\citealt{petigura}).
+\emph{(I don't think need this next sentence).}
 We will treat the uncertainties on those measurements as ``prior
@@
 However, there will also be ``latent parameters'' that represent the ``true''
-values of the exoplanet periods and radii.
+values of the exoplanet periods \period\ and radii \radius.
 That is, for each exoplanet, there is a ``measured'' or ``observed'' period
-and radius, which constitute our data.
+\obs{\period}\ and radius \obs{\radius}, which constitute our data.
 These relate to the unobserved ``true'' period and radius by a noise model.
@@
 
-\emph{Note: I think that we should use $T$ instead of $P$ for period because
-$P$ is too heavily overloaded by the probabilities\ldots}
-
-We have some parametric function \rate\ describing the true rate of exoplanets
-with periods \period\ and radii \radius
+The purpose of this \paper\ is to determine constraints on the \emph{rate
+function} of exoplanets
 \begin{eqnarray}\eqlabel{true-rate}
-\rate (\period,\,\radius;\,\ratepars)
-&=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}\quad.
+\rate_\ratepars (\period,\,\radius)
+&=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}
 \end{eqnarray}
+given our dataset.
 Here we have not made any assumptions about the functional form of \rate, we
@@
 In this \paper, we will assume that the completeness function of the survey is
-known and it can be represented as a probability at every period and radius
-$\completeness (\period,\,\radius)$.
+known and it can be represented as a probability at every (true) period and
+radius $\completeness (\period,\,\radius)$.
 Similarly, we can consider the marginalized transit probability $\transitprob

### 2014-03-27 document/ms.tex
catalog notation

@@
 \newcommand{\rate}{\ensuremath{\eta}}
-\newcommand{\ratepars}{\ensuremath{\theta}}
+\newcommand{\ratepars}{{\ensuremath{\bvec{\theta}}}}
 \newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
@@
 \newcommand{\period}{\ensuremath{T}}
-\newcommand{\completeness}{\ensuremath{P_\mathrm{c}}}
-\newcommand{\transitprob}{\ensuremath{P_\mathrm{t}}}
+\newcommand{\completeness}{{\ensuremath{P_\mathrm{c}}}}
+\newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
+\newcommand{\entry}{\ensuremath{S}}
+\newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}
 
-\newcommand{\interim}{\ensuremath{\alpha}}
+\newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}
 
@@
 
+Let's also introduce more succinct notation for a \emph{catalog}.
+We'll refer to the set of period and radius measurements for a set of $K$
+candidate exoplanets as
+\begin{eqnarray}
+\catalog &=& \{\entry_k\}_{k=1}^K
+\end{eqnarray}
+where each entry $\entry_k = (\period_k,\,\radius_k)$ is a tuple containing a
+period and radius.
+
 The purpose of this \paper\ is to determine constraints on the \emph{rate
@@
 \begin{eqnarray}\eqlabel{true-rate}
-\rate_\ratepars (\period,\,\radius)
+\rate_\ratepars (\entry)
 &=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}
@@
 known and it can be represented as a probability at every (true) period and
-radius $\completeness (\period,\,\radius)$.
+radius $\completeness (\entry)$.
 Similarly, we can consider the marginalized transit probability $\transitprob
@@
 These selection effects can be incorporated into the rate from \eq{true-rate}
-to get the \emph{observed rate}
+to get the \emph{observable rate}
 \begin{eqnarray}\eqlabel{obs-rate}
-\obs{\rate} (\period,\,\radius;\,\ratepars)
-&=& \rate (\period,\,\radius;\,\ratepars)\,\completeness(\period,\,\radius)\,
+\obs{\rate}_\ratepars (\entry)
+&=& \rate_\ratepars(\entry)\,\completeness(\entry)\,
 \transitprob(\period)\quad.
@@
 described by our rate function.
-Under this assumption, the likelihood of the catalog conditioned on a set of
-parameters \ratepars\ is (see Appendix XXXXX for a derivation)
-\begin{eqnarray}
-\ln p(\period,\,\radius\,|\,\ratepars) &\propto&
-\sum_{k=1}^K \ln\obs{\rate} (\period_k,\,\radius_k;\,\ratepars)
-- \int \obs{\rate} (\period,\,\radius;\,\ratepars) \dd\period\dd\radius
+Under this assumption, the likelihood of a catalog \catalog\ conditioned on a
+set of parameters \ratepars\ is (see Appendix XXXXX for a derivation)
+\begin{eqnarray}\eqlabel{true-like}
+\ln p(\catalog\,|\,\ratepars) &\propto&
+\sum_{k=1}^K \ln\obs{\rate}_\ratepars (\entry_k)
+- \int \obs{\rate}_\ratepars (\entry) \dd\entry
 \quad.
@@
 \section{Including observational uncertainties}
+
+The ``likelihood function'' given in \eq{true-like}\ldots
 

### 2014-03-27 document/ms.tex
corrected description of uncertainties

@@
 \newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
+\newcommand{\appropto}{\mathrel{\vcenter{
+  \offinterlineskip\halign{\hfil$##$\cr
+    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
 
@@
 radius $\completeness (\entry)$.
-Similarly, we can consider the marginalized transit probability $\transitprob
-(\period)$ as only a function of period.
+This probability will also include the geometric transit probability as a
+function of period $\transitprob (\period)$.
 These selection effects can be incorporated into the rate from \eq{true-rate}
@@
 \obs{\rate}_\ratepars (\entry)
-&=& \rate_\ratepars(\entry)\,\completeness(\entry)\,
-\transitprob(\period)\quad.
+&=& \rate_\ratepars(\entry)\,\completeness(\entry)\quad.
 \end{eqnarray}
 
-Now we will make the further assumption that every candidate in our sample of
-detected exoplanets is an independent Poisson sample from the population
-described by our rate function.
-Under this assumption, the likelihood of a catalog \catalog\ conditioned on a
-set of parameters \ratepars\ is (see Appendix XXXXX for a derivation)
+Given this parameterized rate $\obs{\rate}_\ratepars$, we can compute the
+Poisson likelihood of a \catalog\ (see Appendix XXXXX for the derivation)
 \begin{eqnarray}\eqlabel{true-like}
@@
 - \int \obs{\rate}_\ratepars (\entry) \dd\entry
-\quad.
 \end{eqnarray}
+where we have assumed that every object in \catalog\ is an independent
+Poisson sample from the rate.
+
+If we had perfect noiseless measurements of the periods and radii for every
+object in our catalog, \eq{true-like} provides a metric for determining
+constraints on the parameters of the rate function.
+In practice, the measurements of \catalog\ are noisy so we need to include
+these effects.
 
@@
 
-The ``likelihood function'' given in \eq{true-like}\ldots
+The ``likelihood function'' given in \eq{true-like} is a model of the
+probability of observing a catalog with \emph{true} periods and radii.
+If we want to include the effects of measurement uncertainties, we need to
+take a moment to consider the meaning of measurement uncertainties.
 
-In the previous section, we implicitly assumed that the measurements of
-\period\ and \radius\ given by the catalog were perfect, without any
-observational uncertainties.
-This isn't a terrible approximation for the period measurements but not for
-the radii.
-Luckily, it is not hard to include these effects in our probabilistic
-framework.
+\dfmtodo{Introduce the idea of catalogs with uncertainties being a description
+of a posterior sampling over catalogs: $\catalog^{(n)} \sim
+p(\catalog\,|\,\obs{\catalog},\,\interim)$.}
 
-In general, catalogs produce posterior probability distributions for
-quantities conditioned on data.
-In particular, the catalog that we're considering here provides estimates of
-the posterior distribution of \period\ and \radius\ conditioned on the
-observables $\obs{\period}$ and $\obs{\radius}$ under some \emph{interim
-priors} parameterized by \interim
+The quantity that we need to compute is the \emph{marginalized likelihood} of
+the data \obs{\catalog}\ given a particular setting of the rate function
+parameters \ratepars.
+This is achieved by marginalizing over the true periods and radii
 \begin{eqnarray}
-p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim) \quad.
+p(\obs{\catalog}\,|\,\ratepars) &=&
+\int p(\catalog,\,\obs{\catalog}\,|\,\ratepars) \dd\catalog \\
+&=&
+\int p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog) \dd\catalog
 \end{eqnarray}
-This estimate can be propagated as \emph{posterior samples}
+where the second line is true because the rate function is defined on the true
+parameters.
+This integral can then be written in terms of the sampling under interim
+priors
 \begin{eqnarray}
-\{\period^{(n)},\,\radius^{(n)}\} &\sim&
-p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim) \quad.
+p(\obs{\catalog}\,|\,\ratepars)
+&=&
+\int p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog)\,
+\frac{p(\catalog\,|\,\obs{\catalog},\,\interim)}
+     {p(\catalog\,|\,\obs{\catalog},\,\interim)}
+\dd\catalog
+\\
+&=&
+p(\obs{\catalog}\,|\,\interim)\,
+\int \frac{p(\catalog\,|\,\ratepars)}{p(\catalog\,|\,\interim)}\,
+p(\catalog\,|\,\obs{\catalog},\,\interim)
+\dd\catalog \quad.
 \end{eqnarray}
-
-Now the quantity that we need to compute is again the (marginalized)
-likelihood of the data $\obs{\period}$ and $\obs{\radius}$ conditioned on a
-set of rate function parameters \ratepars
-\begin{eqnarray}
-p(\obs{\period},\,\obs{\radius}\,|\,\ratepars) &=&
-\int p(\period,\,\radius,\,\obs{\period},\,\obs{\radius}\,|\,\ratepars)
-\dd\period\dd\radius \\
-&=&
-\int p(\period,\,\radius\,|\,\ratepars) \,
-p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)
-\dd\period\dd\radius \\
-&=&
-\int p(\period,\,\radius\,|\,\ratepars) \,
-p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)\,
-\frac{p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)}
-     {p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)}
-\dd\period\dd\radius \\
-&=&
-p(\obs{\period},\,\obs{\radius}\,|\,\interim)\,
-\int
-\frac{p(\period,\,\radius\,|\,\ratepars) \,
-      p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)}
-     {p(\period,\,\radius\,|\,\interim) \,
-      p(\obs{\period},\,\obs{\radius}\,|\,\period,\,\radius)}\,
-p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)
-\dd\period\dd\radius \\
-\frac{p(\obs{\period},\,\obs{\radius}\,|\,\ratepars)}
-     {p(\obs{\period},\,\obs{\radius}\,|\,\interim)}
-&=&
-\int
-\frac{p(\period,\,\radius\,|\,\ratepars)}
-     {p(\period,\,\radius\,|\,\interim)} \,
-p(\period,\,\radius\,|\,\obs{\period},\,\obs{\radius},\,\interim)
-\dd\period\dd\radius \\
-&\approx&
-\frac{1}{N}
+We can approximate this integral using the samples $\catalog^{(n)}$ defined
+above and find the marginalized likelihood up to an unimportant normalization
+constant
+\begin{eqnarray}\eqlabel{marg-like}
+p(\obs{\catalog}\,|\,\ratepars) &\appropto&
 \sum_{n=1}^N
-\frac{p(\period^{(n)},\,\radius^{(n)}\,|\,\ratepars)}
-     {p(\period^{(n)},\,\radius^{(n)}\,|\,\interim)}
+\frac{p(\catalog^{(n)}\,|\,\ratepars)}{p(\catalog^{(n)}\,|\,\interim)}
 \end{eqnarray}
+where the numerator inside the sum is given by \eq{true-like} and the
+denominator is the interim prior evaluated at the samples.
 

### 2014-03-27 document/ms.tex
a few more citations
citing the people that need to be cited

@@
 Given this parameterized rate $\obs{\rate}_\ratepars$, we can compute the
-Poisson likelihood of a \catalog\ (see Appendix XXXXX for the derivation)
+Poisson likelihood of a \catalog\ (\citealt{tabachnik,youdin,dong})
 \begin{eqnarray}\eqlabel{true-like}
@@
 \begin{thebibliography}{}\raggedright
+
 \bibitem[Brewer \& Stello(2009)]{brewer}
 Brewer, B.~J., \& Stello, D.\ 2009, \mnras, 395, 2226 (\arxiv{0902.3907})
+
 \bibitem[Carter \& Winn(2009)]{carter}
 Carter, J.~A., \& Winn, J.~N.\ 2009, \apj, 704, 51 (\arxiv{0909.0747})
+
 \bibitem[Dong \& Zhu(2013)]{dong}
 Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 (\arxiv{1212.4853})
+
 \bibitem[Hogg \etal(2013)]{hoggwhitepaper}
 Hogg, D.~W., Angus, R., Barclay, T., et al.\ 2013, \arxiv{1309.0653}
+
 \bibitem[Hogg \etal(2010)]{hogge}
 Hogg, D.~W., Myers, A.~D., \& Bovy, J.\ 2010, \apj, 725, 2166 (\arxiv{1008.4146})
+
 \bibitem[Petigura \etal(2013)]{petigura}
@@
 Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})
+
 \bibitem[Roberts \etal(2013)]{roberts}
-Roberts, S., McQuillan, A., Reece, S., \& Aigrain, S.\ 2013, \mnras, 435, 3639 (\arxiv{1308.3644})
+Roberts, S., McQuillan, A., Reece, S., \& Aigrain, S.\ 2013, \mnras, 435, 3639
+(\arxiv{1308.3644})
+
+\bibitem[Tabachnik \& Tremaine(2002)]{tabachnik}
+Tabachnik, S., \& Tremaine, S.\ 2002, \mnras, 335, 151
+(\arxiv{astro-ph/0107482})
+
+\bibitem[Youdin(2011)]{youdin}
+Youdin, A.~N.\ 2011, \apj, 742, 38 (\arxiv{1105.1782})
+
 \end{thebibliography}

### 2014-03-27 document/code/separable3.py
moving to more sensible 'catalog-level' formulation

@@ -11,11 +11,42 @@ from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates, load_petigura_bins)
 from population import (CensoringFunction, BrokenPowerLaw, Histogram,
                         SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel)
+                        ProbabilisticModel, Population)
 
 # Load the data.
 ep = load_petigura_bins()
 ln_P_inj, ln_R_inj, recovered = load_completenes_sim()
+tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
+
+# Set up the censoring function.
+censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
+                           bins=(32, 64),
+                           range=[np.log([5, 400]), np.log([0.5, 16])],
+                           transit_lnprob_function=tlp)
+
+# Define the population.
+lpb, lrb = censor.bins
+pop = Population([lpb[::4], lrb[::4]], base=censor.bins)
+print(pop.evaluate(pop.initial()))
+pop.evaluate(pop.initial())
+pop.evaluate(pop.initial())
+
+if True:
+    import matplotlib.pyplot as pl
+    # Plot the completeness function.
+    pl.figure(figsize=(10, 6))
+    x = censor.bins[0]
+    y = censor.bins[1]
+    pl.pcolor(x, y, np.exp(censor.lncompleteness[1:-1, 1:-1]).T, cmap="gray")
+    # pl.plot(dataset.log_per_obs, dataset.log_rp_obs, ".r", ms=3)
+    pl.xlim(x.min(), x.max())
+    pl.ylim(y.min(), y.max())
+    pl.xlabel(r"$\ln P$")
+    pl.ylabel(r"$\ln R_P$")
+    pl.colorbar()
+    pl.savefig("separable3-completeness.png")
+
+assert 0
 
 per_rng = np.log([5.0, 400.0])
 rp_rng = np.log([16.0, 64.0])

### 2014-03-27 population.py

@@ -6,8 +6,9 @@ from __future__ import division, print_function
 __all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",
            "NormalizedPopulation", "Dataset", "CensoringFunction"]
 
+from itertools import izip, product
+
 import numpy as np
-from itertools import izip
 import matplotlib.pyplot as pl
 from scipy.misc import logsumexp
 from scipy.linalg import cho_factor, cho_solve
@@ -94,73 +95,60 @@ class Histogram(object):
 class Population(object):
     poisson = False
 
-    def __init__(self, log_per_bins, log_rp_bins,
-                 log_per_base=None, log_rp_base=None):
-        self.log_per_bins = np.atleast_1d(log_per_bins)
-        self.log_rp_bins = np.atleast_1d(log_rp_bins)
-        if log_per_base is None:
-            self.log_per_base = np.array(self.log_per_bins)
-        else:
-            self.log_per_base = np.atleast_1d(log_per_base)
-        if log_rp_base is None:
-            self.log_rp_base = np.array(self.log_rp_bins)
-        else:
-            self.log_rp_base = np.atleast_1d(log_rp_base)
+    def __init__(self, bins, base=None):
+        self.bins = bins
+        if base is None:
+            base = bins
+        self.base = bins
+        self.shape = np.array(map(len, bins)) - 1
 
         # Make sure that the resampling is sane.
-        assert (self.log_per_bins[0] == self.log_per_base[0])
-        assert (self.log_per_bins[-1] == self.log_per_base[-1])
-        assert (self.log_rp_bins[0] == self.log_rp_base[0])
-        assert (self.log_rp_bins[-1] == self.log_rp_base[-1])
+        assert all([b1[0] == b2[0] for b1, b2 in izip(bins, base)])
+        assert all([b1[-1] == b2[-1] for b1, b2 in izip(bins, base)])
 
         # Figure out which bins the base grid sits in. This will only be exact
         # when the grids line up perfectly.
-        p = self.log_per_base
-        rp = self.log_rp_base
-        ix = np.digitize(0.5*(p[1:]+p[:-1]), self.log_per_bins) - 1
-        assert np.all((ix >= 0) * (ix < len(self.log_per_bins)))
-        iy = np.digitize(0.5*(rp[1:]+rp[:-1]), self.log_rp_bins) - 1
-        assert np.all((iy >= 0) * (iy < len(self.log_rp_bins)))
-        self.iy, self.ix = np.meshgrid(iy, ix)
[…]

### 2014-03-28 document/ms.tex
some notes to self

@@
 
-\section{Separable Distribution}
+In practice, the variance of this approximation depends on the choice of
+interim prior and, in particular, it is useful to include the completeness
+function.
+\dfmtodo{BLAH BLAH BLAH.}
+
+\section{The completeness function \& dataset}
+
+\dfmtodo{%
+Describe Erik's data products and our model of the completeness in boxes
+extended artificially beyond the original range.
+Also describe the re-sampling procedure for the candidate parameters.
+}
+
+\section{Separable distribution}
 

### 2014-03-28 document/code/load_data.py
lotsa things

@@ -26,34 +26,53 @@ transit_lnprob0 = -2.98353340397
 ln_period0 = np.log(P)
 
 
-def load_completenes_sim(rp_func=np.log):
+def load_completenes_sim(rp_func=np.log, per_rng=np.log([5, 400]),
+                         rp_rng=None, K=50000):
+    if rp_rng is None:
+        rp_rng = rp_func([0.5, 16])
     sim = pd.read_hdf(os.path.join(bp, "mcDV.h5"), "mcDV")
     m = sim.found * sim.bDV
-    return np.log(sim.inj_P), rp_func(sim.inj_Rp), m
+    x, y, z = np.log(sim.inj_P), rp_func(sim.inj_Rp), m
+    if rp_rng[1] > rp_func(16):
+        x = np.append(x, np.random.uniform(per_rng[0], per_rng[1], K))
+        y = np.append(y, np.random.uniform(rp_func(16.), rp_rng[1], K))
+        z = np.append(z, np.ones(K, dtype=bool))
+    return x, y, z
 
 
-def load_candidates(censor, samples=0, N=10000, rp_func=np.log):
-    log_per_obs = []
-    log_rp_obs = []
-    for line in open(os.path.join(bp, "table_ekoi836.tex")):
+def load_candidates(censor=None, samples=0, N=10000, rp_func=np.log):
+    lines = open(os.path.join(bp, "table_ekoi836.tex")).readlines()
+    K = len(lines)
+    good = np.ones(K, dtype=bool)
+    catalogs = np.empty((max(samples, 1), K, 2))
+    lnweights = np.zeros(max(samples, 1))
+    for k, line in enumerate(lines):
         cols = line.split("&")
         log_per = np.log(float(cols[2]))
-        log_per_obs.append(log_per)
+
         if samples > 0:
             m, s = map(float, cols[12:14])
-            r = m + s*np.sort(np.random.randn(N))
-            lnw = censor.evaluate([log_per], rp_func(r)) - 0.5 * ((r-m)/s)**2
-            cummulative = np.cumsum(np.exp(lnw))
-            cummulative /= cummulative[-1]
-            rnd = np.random.rand(samples)
-            r = np.interp(rnd, cummulative, r)
-            log_rp_obs.append(rp_func(r))
+            catalogs[:, k, 0] = log_per
+            if censor is None:
+                catalogs[:, k, 1] = rp_func(m+s*np.random.randn(samples))
+                continue
+            r = m + s*np.random.randn(N)
[…]

### 2014-03-28 document/code/separable3.py

@@ -13,23 +13,36 @@ from population import (CensoringFunction, BrokenPowerLaw, Histogram,
                         SeparablePopulation, NormalizedPopulation,
                         ProbabilisticModel, Population)
 
+# Define the box that we're going to work in.
+per_rng = np.log([5.0, 400.0])
+rp_rng = np.log([0.5, 64.0])
+
 # Load the data.
 ep = load_petigura_bins()
-ln_P_inj, ln_R_inj, recovered = load_completenes_sim()
+ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
+                                                     rp_rng=rp_rng)
 tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
 
 # Set up the censoring function.
 censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                           bins=(32, 64),
-                           range=[np.log([5, 400]), np.log([0.5, 16])],
+                           bins=(32, 70),
+                           range=[per_rng, rp_rng],
                            transit_lnprob_function=tlp)
+
+# Load the dataset and resample under the censoring function.
+dataset = load_candidates(censor, 24)
 
 # Define the population.
 lpb, lrb = censor.bins
-pop = Population([lpb[::4], lrb[::4]], base=censor.bins)
-print(pop.evaluate(pop.initial()))
-pop.evaluate(pop.initial())
-pop.evaluate(pop.initial())
+b = np.concatenate(([lrb[0]], ep[1], [lrb[-1]]))
+pdist = BrokenPowerLaw(lpb)
+rdist = Histogram(b, lrb)
+print(rdist.inds)
+pop = NormalizedPopulation(11., SeparablePopulation([pdist, rdist]))
+
+# Define the probabilistic model.
+model = ProbabilisticModel(dataset, pop, censor)
+print("Initial ln-prob = {0}".format(model.lnprob(pop.initial())))
 
 if True:
     import matplotlib.pyplot as pl
@@ -38,58 +51,14 @@ if True:
     x = censor.bins[0]
     y = censor.bins[1]
     pl.pcolor(x, y, np.exp(censor.lncompleteness[1:-1, 1:-1]).T, cmap="gray")
[…]

### 2014-03-28 population.py

@@ -6,7 +6,7 @@ from __future__ import division, print_function
 __all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",
            "NormalizedPopulation", "Dataset", "CensoringFunction"]
 
-from itertools import izip, product
+from itertools import izip
 
 import numpy as np
 import matplotlib.pyplot as pl
@@ -63,7 +63,8 @@ class Histogram(object):
         assert self.bins[0] == self.base[0]
         assert self.bins[-1] == self.base[-1]
 
-        self.ln_bin_widths = np.log(self.bins[1:] - self.bins[:-1])
+        self.ln_bin_widths = np.log(np.diff(self.bins))
+        assert np.all(np.isfinite(self.ln_bin_widths))
         self.nbins = len(self.ln_bin_widths)
         self.inds = np.digitize(0.5*(self.base[1:]+self.base[:-1]),
                                 self.bins) - 1
@@ -73,8 +74,7 @@ class Histogram(object):
         return self.nbins - 1
 
     def initial(self):
-        v = np.zeros(self.nbins)
-        v -= logsumexp(v + self.ln_bin_widths)
+        v = np.zeros(self.nbins) - logsumexp(self.ln_bin_widths)
         return v[:-1]
 
     def lnprior(self, theta):
@@ -135,11 +135,7 @@ class Population(object):
         # v[:3] = [0.5, -3.0, -3.0]
         return v[:-1]
 
-    def _get_grid(self, theta):
-        k = tuple(theta)
-        if k == self._cache_key:
-            return self._cache_val
-
+    def compute_grid(self, theta):
         # Compute the integral over the first N-1 cells.
         norm = logsumexp(theta + self.ln_cell_area[:-1])
         if norm >= 0.0:
@@ -147,9 +143,14 @@ class Population(object):
 
         # Compute the height of the last cell and cache the heights.
         v = np.log(1.0 - np.exp(norm)) - self.ln_cell_area[-1]
-        self._cache_key = k
-        self._cache_val = np.append(theta, v).reshape(self.shape)
+        return np.append(theta, v).reshape(self.shape)
 
+    def _get_grid(self, theta):
+        k = tuple(theta)
+        if k == self._cache_key:
+            return self._cache_val
+        self._cache_key = k
+        self._cache_val = self.compute_grid(theta)
         return self._cache_val
 
     def evaluate(self, theta):
@@ -238,42 +239,37 @@ class Population(object):
 class SeparablePopulation(Population):
     poisson = False
 
[…]

### 2014-03-28 document/code/plot_results.py
need to COPY the lnprob array

@@ -7,22 +7,28 @@ import os
 import sys
 import numpy as np
 import cPickle as pickle
+import matplotlib.pyplot as pl
+from scipy.misc import logsumexp
 
 from load_data import load_petigura_bins
 
-mylog = np.log
-rp_label = "\ln R"
-if "--linear" in sys.argv:
-    mylog = lambda x: x
-    rp_label = "R"
-
-ep = load_petigura_bins(mylog=mylog)
+ep = load_petigura_bins()
 censor, dataset, pop, sampler = pickle.load(open(sys.argv[1]))
 
 samples = sampler.chain
-samples = samples.reshape((-1, samples.shape[-1]))
+# samples = samples.reshape((-1, samples.shape[-1]))
 
-fig_per, fig_rp = pop.plot(samples[np.random.randint(len(samples), size=24)],
-                           rp_label=rp_label, ep=ep)
-fig_per.savefig(os.path.splitext(sys.argv[1])[0]+"-period.png")
-fig_rp.savefig(os.path.splitext(sys.argv[1])[0]+"-radius.png")
+# # Subsample the chain to get some posterior samples to display.
+# subsamples = samples[np.random.randint(len(samples), size=24)]
+# lnrates = map(pop.evaluate, subsamples)
+# y = logsumexp(lnrates[10] + censor.ln_bin_widths[0][:, None], axis=0)
+# y -= logsumexp(y + censor.ln_bin_widths[1])
+# x = censor.bins[1]
+# pl.plot(x[:-1], np.exp(y))
+pl.plot(samples[:, :, 1].T)
+pl.savefig("test.png")
+
+# fig_per, fig_rp = pop.plot(,
+#                            rp_label=rp_label, ep=ep)
+# fig_per.savefig(os.path.splitext(sys.argv[1])[0]+"-period.png")
+# fig_rp.savefig(os.path.splitext(sys.argv[1])[0]+"-radius.png")

### 2014-03-28 document/code/separable3.py

@@ -43,6 +43,7 @@ pop = NormalizedPopulation(11., SeparablePopulation([pdist, rdist]))
 # Define the probabilistic model.
 model = ProbabilisticModel(dataset, pop, censor)
 print("Initial ln-prob = {0}".format(model.lnprob(pop.initial())))
+print("Initial ln-prob = {0}".format(model.lnprob(pop.initial())))
 
 if True:
     import matplotlib.pyplot as pl

### 2014-03-28 population.py

@@ -176,65 +176,6 @@ class Population(object):
         logdet = np.sum(2*np.log(np.diag(factor)))
         return -0.5 * (np.dot(y, cho_solve((factor, flag), y)) + logdet)
 
-    def plot(self, thetas, ep=None, alpha=0.5, rp_label="\ln R"):
-        thetas = np.atleast_2d(thetas)
-
-        # Set up figures and axes.
-        fig_per = pl.figure(figsize=(6, 5))
-        ax_per = fig_per.add_subplot(111)
-        fig_rp = pl.figure(figsize=(6, 5))
-        ax_rp = fig_rp.add_subplot(111)
-        for fig in [fig_per, fig_rp]:
-            fig.subplots_adjust(left=0.16, bottom=0.15, right=0.98, top=0.97)
-
-        if ep is None:
-            rinds = np.ones(len(self.log_rp_bins), dtype=bool)
-        else:
-            rinds = ((self.log_rp_bins <= ep[1].max())
-                     * (self.log_rp_bins >= ep[1].min()))
-
-        # Loop over samples and plot the projections.
-        for theta in thetas:
-            grid = self.evaluate(theta)
-
-            z_per = self.log_rp_bins
-            y_per = logsumexp(grid + np.log(z_per[1:]-z_per[:-1])[None, :],
-                              axis=1)
-            z_rp = self.log_per_bins
-            y_rp = logsumexp(grid + np.log(z_rp[1:]-z_rp[:-1])[:, None],
-                             axis=0)
-            y_rp = y_rp[rinds[1:] * rinds[:-1]]
-
-            for ax, x, y in izip([ax_per, ax_rp],
-                                 [self.log_per_bins, self.log_rp_bins[rinds]],
-                                 [y_per, y_rp]):
-                y -= logsumexp(y + np.log(x[1:] - x[:-1]))
-                x = np.array(zip(x[:-1], x[1:])).flatten()
-                y = np.array(zip(y, y)).flatten()
-                ax.plot(x, np.exp(y), "k", alpha=alpha)
-
-        # Plot Erik's values if given.
-        if ep is not None:
-            ax_rp.plot(0.5*(ep[1][1:]+ep[1][:-1]), ep[3], ".", color="r",
-                       ms=8)
-            ax_per.plot(0.5*(ep[0][1:]+ep[0][:-1]), ep[2], ".", color="r",
-                        ms=8)
-
[…]

### 2014-03-28 document/ms.tex
text stuff

--- 2014-03-27 document/ms.tex
@@
 planets around G dwarfs.
-The models all take into account the transit probability and the completeness
+The models take into account the transit probability and the completeness
 (or detection efficiency) as a function of planet radius and orbital period.
-Because we forward model the data, we do not ever have to perform the
-traditional weighting of the data by the inverse detection efficiency.
+Because we forward model the data, we do not ever have to perform any kind of
+inverse-detection-efficiency weighting of the data.
 Our most sophisticated models do not assume that the joint radius--period
-distribution is separable, and correctly take into account the significant
-observational uncertainties on the planet radii, using the machinery of
-hierarchical inference.
+distribution is separable.
+They also correctly take into account the significant observational
+uncertainties on the planet radii, using the machinery of hierarchical
+inference.
 We reconsider and confirm with weaker assumptions prior results on the
@@
 \emph{not} by weighting the detected objects by the inverse detection
-efficiency (\citealt{dong, petigura}), but rather by forward modeling the
-observed distribution subject to the detection efficiency.
-\hoggtodo{Check that Dong \& Zhu do indeed weight by inverse probability.}
+efficiency (\citealt{petigura}), but rather by forward modeling the
+observed distribution subject to the detection efficiency
+(\dfmtodo{CITE TREMAINE?} \citealt{dong}).
 Re-weighting the data by inverse selection probability---a method called
@@
 but it produces a higher-variance estimate of the population than a justified
-likelihood approach.
-One simple intuition along these lines is that the inverse probability weights
-can be large in dynamic range, so the effective number of samples, after
+likelihood approach; it is higher variance because
+the effective number of samples, after
 weighting, can be much smaller than the actual number of samples.
-Another is that it tends to upweight most strongly the most marginal
-detections in the data sample, since these tend to be the samples that lie in
-the region of lowest completeness.  In this work, we build a justified
+In this work, we build a justified
 likelihood of the observed data, taking into account the completeness function
@@
 functions.
-Indeed, the assumptions behind this study are weaker than those of any
-previous study of the period and radius distribution of the Earth-analog
-population (\citealt{dong, petigura}).
+Indeed, the assumptions behind this study here are weaker than those of any
+previous study of the period and radius distribution of the Earth-sized
+exoplanet population (\dfmtodo{CITE TREMAINE?} \citealt{dong, petigura}).
 

### 2014-03-28 document/code/load_data.py
generating nice figures

@@ -88,4 +88,4 @@ def load_petigura_bins(mylog=np.log):
     norm = np.sum(ep_p_values * (ep_p_lnbins[1:] - ep_p_lnbins[:-1]))
     ep_p_pdf = ep_p_values / norm
 
-    return ep_p_lnbins, ep_Rp_lnbins, ep_p_pdf, ep_Rp_pdf
+    return (ep_p_lnbins, ep_p_pdf), (ep_Rp_lnbins, ep_Rp_pdf)

### 2014-03-28 document/code/plot_results.py

@@ -3,32 +3,27 @@
 
 from __future__ import division, print_function
 
-import os
 import sys
 import numpy as np
 import cPickle as pickle
-import matplotlib.pyplot as pl
-from scipy.misc import logsumexp
 
 from load_data import load_petigura_bins
 
 ep = load_petigura_bins()
 censor, dataset, pop, sampler = pickle.load(open(sys.argv[1]))
 
-samples = sampler.chain
-# samples = samples.reshape((-1, samples.shape[-1]))
+burnin = 4000
+samples = sampler.chain[:, burnin:, :]
+samples = samples.reshape((-1, samples.shape[-1]))
 
-# # Subsample the chain to get some posterior samples to display.
-# subsamples = samples[np.random.randint(len(samples), size=24)]
-# lnrates = map(pop.evaluate, subsamples)
-# y = logsumexp(lnrates[10] + censor.ln_bin_widths[0][:, None], axis=0)
-# y -= logsumexp(y + censor.ln_bin_widths[1])
-# x = censor.bins[1]
-# pl.plot(x[:-1], np.exp(y))
-pl.plot(samples[:, :, 1].T)
-pl.savefig("test.png")
-
-# fig_per, fig_rp = pop.plot(,
-#                            rp_label=rp_label, ep=ep)
-# fig_per.savefig(os.path.splitext(sys.argv[1])[0]+"-period.png")
-# fig_rp.savefig(os.path.splitext(sys.argv[1])[0]+"-radius.png")
+# Subsample the chain to get some posterior samples to display.
+subsamples = samples[np.random.randint(len(samples), size=100)]
+figs = pop.plot(subsamples,
+                ranges=[np.log([6.25, 100]), np.log([1.0, 16.0])],
+                labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
+                top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
+                literature=ep)
+figs[0].savefig("period.png")
+figs[0].savefig("period.pdf")
+figs[1].savefig("radius.png")
+figs[1].savefig("radius.pdf")

### 2014-03-28 document/code/separable3.py

@@ -30,7 +30,7 @@ censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
                            transit_lnprob_function=tlp)
 
 # Load the dataset and resample under the censoring function.
-dataset = load_candidates(censor, 24)
+dataset = load_candidates(censor, 120)
 
 # Define the population.
 lpb, lrb = censor.bins
@@ -72,8 +72,6 @@ assert np.all(finite), "{0}".format(np.sum(finite))
 
 # Set up the sampler.
 sampler = emcee.EnsembleSampler(nwalkers, ndim, model)
-pos, lp, state = sampler.run_mcmc(pos, 3000)
-sampler.reset()
-sampler.run_mcmc(pos, 2000)
+sampler.run_mcmc(pos, 5000)
 
 pickle.dump((censor, dataset, pop, sampler), open("separable3.pkl", "wb"), -1)

### 2014-03-28 population.py

@@ -9,9 +9,12 @@ __all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",
 from itertools import izip
 
 import numpy as np
-import matplotlib.pyplot as pl
+
 from scipy.misc import logsumexp
 from scipy.linalg import cho_factor, cho_solve
+
+import matplotlib.pyplot as pl
+from matplotlib.ticker import ScalarFormatter
 
 
 class BrokenPowerLaw(object):
@@ -99,7 +102,7 @@ class Population(object):
         self.bins = bins
         if base is None:
             base = bins
-        self.base = bins
+        self.base = base
         self.shape = np.array(map(len, bins)) - 1
 
         # Make sure that the resampling is sane.
@@ -176,6 +179,83 @@ class Population(object):
         logdet = np.sum(2*np.log(np.diag(factor)))
         return -0.5 * (np.dot(y, cho_solve((factor, flag), y)) + logdet)
 
+    def plot(self, thetas, labels=None, ranges=None, alpha=0.3,
+             literature=None, top_axes=None):
+        # Pre-compute the ranges and allowed ranges.
+        if ranges is None:
+            ranges = [(b.min(), b.max()) for b in self.base]
+        bm = []
+        vm = []
+        for rng, b in izip(ranges, self.base):
+            if rng is None:
+                rng = (b.min(), b.max())
+            m = (b >= rng[0]) * (b <= rng[1])
+            bm.append(m)
+            vm.append(np.arange(len(m) - 1)[m[:-1] * m[1:]])
+        vm = np.meshgrid(*vm, indexing="ij")
+
+        thetas = np.atleast_2d(thetas)
+        figs = [pl.figure(figsize=(6, 5)) for b in self.base]
+        axes = [f.add_subplot(111) for f in figs]
+
+        lbw0 = map(np.log, map(np.diff,
+                               [b[m] for b, m in izip(self.base, bm)]))
+        N = len(self.base)
+        for theta in thetas:
+            grid = self._get_grid(theta)[vm]
+            for i, (x, m) in enumerate(izip(self.base, bm)):
+                # Compute the volumes along the marginalized axes.
+                lbw = list(lbw0)
+                lbw.pop(i)
+                lnvol = reduce(np.add, np.ix_(*lbw))
+
+                # Build a broadcasting object and axis list to perform the
+                # marginalization.
+                bc = [slice(None)] * N
+                bc[i] = None
+                a = range(N)
[…]

### 2014-03-28 document/code/plot_results.py
output format

@@ -3,6 +3,7 @@
 
 from __future__ import division, print_function
 
+import os
 import sys
 import numpy as np
 import cPickle as pickle
@@ -10,9 +11,17 @@ import cPickle as pickle
 from load_data import load_petigura_bins
 
 ep = load_petigura_bins()
-censor, dataset, pop, sampler = pickle.load(open(sys.argv[1]))
 
-burnin = 4000
+# Load the MCMC results.
+bp = sys.argv[1]
+fn = os.path.join(bp, "results.pkl")
+censor, dataset, pop, sampler = pickle.load(open(fn))
+
+# Remove a burn-in and flatten the chain.
+if len(sys.argv) > 2:
+    burnin = int(sys.argv[2])
+else:
+    burnin = 4000
 samples = sampler.chain[:, burnin:, :]
 samples = samples.reshape((-1, samples.shape[-1]))
 
@@ -23,7 +32,7 @@ figs = pop.plot(subsamples,
                 labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
                 top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
                 literature=ep)
-figs[0].savefig("period.png")
-figs[0].savefig("period.pdf")
-figs[1].savefig("radius.png")
-figs[1].savefig("radius.pdf")
+figs[0].savefig(os.path.join(bp, "period.png"))
+figs[0].savefig(os.path.join(bp, "period.pdf"))
+figs[1].savefig(os.path.join(bp, "radius.png"))
+figs[1].savefig(os.path.join(bp, "radius.pdf"))

### 2014-03-28 document/code/separable.py

@@ -3,6 +3,7 @@
 
 from __future__ import division, print_function
 
+import os
 import emcee
 import numpy as np
 import cPickle as pickle
@@ -11,7 +12,12 @@ from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates, load_petigura_bins)
 from population import (CensoringFunction, BrokenPowerLaw, Histogram,
                         SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel, Population)
+                        ProbabilisticModel)
+
+try:
+    os.makedirs("separable")
+except os.error:
+    pass
 
 # Define the box that we're going to work in.
 per_rng = np.log([5.0, 400.0])
@@ -43,7 +49,6 @@ pop = NormalizedPopulation(11., SeparablePopulation([pdist, rdist]))
 # Define the probabilistic model.
 model = ProbabilisticModel(dataset, pop, censor)
 print("Initial ln-prob = {0}".format(model.lnprob(pop.initial())))
-print("Initial ln-prob = {0}".format(model.lnprob(pop.initial())))
 
 if True:
     import matplotlib.pyplot as pl
@@ -59,7 +64,7 @@ if True:
     pl.xlabel(r"$\ln P$")
     pl.ylabel(r"$\ln R_P$")
     pl.colorbar()
-    pl.savefig("separable3-completeness.png")
+    pl.savefig("separable/completeness.png")
 
 # Set up the sampler.
 p0 = pop.initial()
@@ -74,4 +79,5 @@ assert np.all(finite), "{0}".format(np.sum(finite))
 sampler = emcee.EnsembleSampler(nwalkers, ndim, model)
 sampler.run_mcmc(pos, 5000)
 
-pickle.dump((censor, dataset, pop, sampler), open("separable3.pkl", "wb"), -1)
+pickle.dump((censor, dataset, pop, sampler),
+            open("separable/results.pkl", "wb"), -1)

### 2014-03-28 document/code/separable1.py

@@ -1,70 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-import sys
-import emcee
-import numpy as np
-import cPickle as pickle
-import matplotlib.pyplot as pl
-
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates, load_petigura_bins)
-from population import (CensoringFunction, BrokenPowerLaw, Histogram,
-                        SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel)
-
-# Load the data.
-ep = load_petigura_bins()
-ln_P_inj, ln_R_inj, recovered = load_completenes_sim()
-censor = CensoringFunction(ln_P_inj, ln_R_inj, recovered,
-                           nbin_log_per=48, nbin_log_rp=32,
-                           log_per_range=(ep[0].min(), ep[0].max()),
-                           log_rp_range=(ep[1].min(), ep[1].max()),
-                           # log_per_range=np.log([5, 400]),
-                           # log_rp_range=np.log([0.5, 16]),
-                           transit_lnprob0=transit_lnprob0,
-                           ln_period0=ln_period0)
-dataset = load_candidates(censor)
-
-# Build the probabilistic model.
-log_per_dist = BrokenPowerLaw(censor.log_per_bins)
-log_rp_dist = Histogram(censor.log_rp_bins, resample=4)
-pop = NormalizedPopulation(11., SeparablePopulation(log_per_dist, log_rp_dist))
-model = ProbabilisticModel(dataset, pop)
-
-if "--plot" in sys.argv:
-    # Plot the completeness function.
-    pl.figure(figsize=(10, 6))
-    x = censor.log_per_bins
-    y = censor.log_rp_bins
-    pl.pcolor(x, y, np.exp(censor.lncompleteness).T, cmap="gray")
-    pl.plot(dataset.log_per_obs, dataset.log_rp_obs, ".r", ms=3)
-    pl.xlim(censor.log_per_bins.min(), censor.log_per_bins.max())
-    pl.ylim(censor.log_rp_bins.min(), censor.log_rp_bins.max())
-    pl.xlabel(r"$\ln P$")
-    pl.ylabel(r"$\ln R_P$")
-    pl.colorbar()
-    pl.savefig("separable1-completeness.pdf")
-
[…]

### 2014-03-28 document/code/separable2.py

@@ -1,49 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-import emcee
-import numpy as np
-import cPickle as pickle
-
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates, load_petigura_bins)
-from population import (CensoringFunction, BrokenPowerLaw, Histogram,
-                        SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel)
-
-# Load the data.
-ep = load_petigura_bins()
-ln_P_inj, ln_R_inj, recovered = load_completenes_sim()
-censor = CensoringFunction(ln_P_inj, ln_R_inj, recovered,
-                           nbin_log_per=48, nbin_log_rp=32,
-                           log_per_range=(ep[0].min(), ep[0].max()),
-                           log_rp_range=(ep[1].min(), ep[1].max()),
-                           # log_per_range=np.log([5, 400]),
-                           # log_rp_range=np.log([0.5, 16]),
-                           transit_lnprob0=transit_lnprob0,
-                           ln_period0=ln_period0)
-dataset = load_candidates(censor)
-
-# Build the probabilistic model.
-log_per_dist = BrokenPowerLaw(censor.log_per_bins)
-log_rp_dist = Histogram(censor.log_rp_bins, resample=2)
-pop = NormalizedPopulation(11., SeparablePopulation(log_per_dist, log_rp_dist))
-model = ProbabilisticModel(dataset, pop)
-
-# Set up the sampler.
-p0 = pop.initial()
-ndim, nwalkers = len(p0), 64
-pos = [p0 + 1e-8 * np.random.randn(ndim) for i in range(nwalkers)]
-
-# Make sure that all the initial positions have finite probability.
-assert np.all(np.isfinite(map(model.lnprob, pos)))
-
-# Set up the sampler.
-sampler = emcee.EnsembleSampler(nwalkers, ndim, model)
-pos, lp, state = sampler.run_mcmc(pos, 3000)
-sampler.reset()
-sampler.run_mcmc(pos, 2000)
-
-pickle.dump((censor, dataset, pop, sampler), open("separable2.pkl", "wb"), -1)

### 2014-03-29 document/code/fake.py
generating synthetic catalogs

@@ -1,0 +1,73 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import os
+import emcee
+import numpy as np
+import cPickle as pickle
+from itertools import product
+import matplotlib.pyplot as pl
+
+from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
+                       load_candidates, load_petigura_bins)
+from population import (CensoringFunction, BrokenPowerLaw, Histogram,
+                        SeparablePopulation, NormalizedPopulation,
+                        ProbabilisticModel)
+
+try:
+    os.makedirs("fake")
+except os.error:
+    pass
+
+# Define the box that we're going to work in.
+per_rng = np.log([5.0, 400.0])
+rp_rng = np.log([0.5, 64.0])
+
+# Load the data.
+ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
+                                                     rp_rng=rp_rng)
+tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
+
+# Set up the censoring function.
+censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
+                           bins=(32, 36),
+                           range=[per_rng, rp_rng],
+                           transit_lnprob_function=tlp)
+
+# Build a synthetic population.
+lpb, lrb = censor.bins
+pdist = BrokenPowerLaw(lpb)
+rdist = BrokenPowerLaw(lrb)
+pop0 = SeparablePopulation([pdist, rdist])
+pop0 = NormalizedPopulation(10.8, pop0)
+
+# Plot the true distributions.
+truth = [10.8, 0.5, -0.2, 4.0, 3.0, -3.0, 1.0]
+figs = pop0.plot(truth, alpha=1,
+                 labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
+                 top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"])
+figs[0].savefig(os.path.join("fake", "true-period.png"))
+figs[1].savefig(os.path.join("fake", "true-radius.png"))
+
+# Sample from this censored population.
+lnrate = np.array(censor.lnprob[1:-1, 1:-1])
[…]

### 2014-03-29 document/code/general.py

@@ -3,19 +3,35 @@
 
 from __future__ import division, print_function
 
-import os
-import emcee
-import numpy as np
-import cPickle as pickle
+from IPython.parallel import Client
+c = Client()
+view = c[:]
+
+with view.sync_imports():
+    import os
+    import sys
+    import time
+    import emcee
+    import numpy as np
+    import cPickle as pickle
+
+# Monkey patch the path.
+f = __file__
+d = os.path.dirname
+p = d(d(d(os.path.abspath(f))))
+sys.path.insert(0, p)
+view.execute("if \"{0}\" not in sys.path: sys.path.insert(0, \"{0}\")"
+             .format(p), block=True)
+
+with view.sync_imports():
+    from population import (CensoringFunction, Population, SmoothPopulation,
+                            NormalizedPopulation, ProbabilisticModel)
 
 from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates, load_petigura_bins)
-from population import (CensoringFunction, BrokenPowerLaw, Histogram,
-                        SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel)
 
 try:
-    os.makedirs("separable")
+    os.makedirs("general")
 except os.error:
     pass
 
@@ -31,23 +47,22 @@ tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
 
 # Set up the censoring function.
 censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                           bins=(32, 70),
+                           bins=(32, 32),
                            range=[per_rng, rp_rng],
                            transit_lnprob_function=tlp)
 
 # Load the dataset and resample under the censoring function.
-dataset = load_candidates(censor, 120)
+dataset = load_candidates(censor)  # , 120)
 
 # Define the population.
 lpb, lrb = censor.bins
-b = np.concatenate(([lrb[0]], ep[1], [lrb[-1]]))
-pdist = BrokenPowerLaw(lpb)
-rdist = Histogram(b, lrb)
-print(rdist.inds)
-pop = NormalizedPopulation(11., SeparablePopulation([pdist, rdist]))
+pop = Population([lpb[::8], lrb[::4]], censor.bins)
[…]

### 2014-03-29 document/code/load_data.py

@@ -69,8 +69,14 @@ def load_candidates(censor=None, samples=0, N=10000, rp_func=np.log):
             catalogs[:, k, 1] = r[inds[:samples]]
             lnweights += lnp[inds[:samples]]
         else:
+            r = rp_func(float(cols[12]))
+            lnp = censor.get_lncompleteness([log_per, r])
+            if not np.isfinite(lnp):
+                print("Dropping candidate at R={0}".format(float(cols[12])))
+                good[k] = 0
+                continue
             catalogs[0, k, 0] = log_per
-            catalogs[0, k, 1] = rp_func(float(cols[12]))
+            catalogs[0, k, 1] = r
 
     return population.Dataset(catalogs[:, good], lnweights)
 

### 2014-03-29 document/code/plot_results.py

@@ -15,20 +15,25 @@ ep = load_petigura_bins()
 # Load the MCMC results.
 bp = sys.argv[1]
 fn = os.path.join(bp, "results.pkl")
-censor, dataset, pop, sampler = pickle.load(open(fn))
+censor, dataset, pop, samples, lnprob = pickle.load(open(fn))
 
 # Remove a burn-in and flatten the chain.
 if len(sys.argv) > 2:
     burnin = int(sys.argv[2])
 else:
     burnin = 4000
-samples = sampler.chain[:, burnin:, :]
+
+# import matplotlib.pyplot as pl
+# pl.plot(samples[:, :, 0].T)
+# pl.savefig("test.png")
+# assert 0
+samples = samples[:, burnin:, :]
 samples = samples.reshape((-1, samples.shape[-1]))
 
 # Subsample the chain to get some posterior samples to display.
 subsamples = samples[np.random.randint(len(samples), size=100)]
 figs = pop.plot(subsamples,
-                ranges=[np.log([6.25, 100]), np.log([1.0, 16.0])],
+                # ranges=[np.log([6.25, 100]), np.log([1.0, 16.0])],
                 labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
                 top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
                 literature=ep)

### 2014-03-29 population.py

@@ -165,20 +165,6 @@ class Population(object):
     def lnprior(self, theta):
         return 0.0
 
-        grid = self._get_grid(theta)
-        if grid is None:
-            return -np.inf
-
-        # Compute the Gaussian process prior.
-        y = grid.flatten()
-        y -= np.mean(y)
-        d = (self.cell_coords[:, None, :] - self.cell_coords[None, :, :])**2
-        K = np.exp(theta[0] - 0.5 * np.sum(d / np.exp(theta[1:3]), axis=2))
-        K += np.diag(1e-10 * np.ones_like(y))
-        factor, flag = cho_factor(K)
-        logdet = np.sum(2*np.log(np.diag(factor)))
-        return -0.5 * (np.dot(y, cho_solve((factor, flag), y)) + logdet)
-
     def plot(self, thetas, labels=None, ranges=None, alpha=0.3,
              literature=None, top_axes=None):
         # Pre-compute the ranges and allowed ranges.
@@ -202,7 +188,7 @@ class Population(object):
                                [b[m] for b, m in izip(self.base, bm)]))
         N = len(self.base)
         for theta in thetas:
-            grid = self._get_grid(theta)[vm]
+            grid = self.evaluate(theta)[vm]
             for i, (x, m) in enumerate(izip(self.base, bm)):
                 # Compute the volumes along the marginalized axes.
                 lbw = list(lbw0)
@@ -291,6 +277,58 @@ class SeparablePopulation(Population):
                 return -np.inf
             n += len(d)
         return lp
+
+
+class SmoothPopulation(object):
+    poisson = False
+
+    def __init__(self, pars, base_population, eps=1e-10):
+        self.pars = np.atleast_1d(pars)
+        self.base_population = base_population
+        self.base = base_population.base
+        self.eps = eps
+
+        # Pre-compute the distance vectors.
+        bins = self.base_population.bins
+        coords = np.meshgrid(*[0.5*(b[1:] + b[:-1]) for b in bins],
+                             indexing="ij")
+        coords = np.vstack([c.flatten() for c in coords]).T
+        self.dvec = (coords[:, None, :] - coords[None, :, :])**2
+        self.ndim = self.dvec.shape[2] + 1
+
+        assert len(self.pars) == self.ndim
+
+    def __len__(self):
+        return len(self.base_population) + self.ndim
+
+    def initial(self):
[…]

### 2014-03-29 document/code/fake.py
running inference on synthetic catalogs

@@ -14,7 +14,7 @@ from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates, load_petigura_bins)
 from population import (CensoringFunction, BrokenPowerLaw, Histogram,
                         SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel)
+                        ProbabilisticModel, Dataset)
 
 try:
     os.makedirs("fake")
@@ -44,7 +44,7 @@ pop0 = SeparablePopulation([pdist, rdist])
 pop0 = NormalizedPopulation(10.8, pop0)
 
 # Plot the true distributions.
-truth = [10.8, 0.5, -0.2, 4.0, 3.0, -3.0, 1.0]
+truth = [10.8, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
 figs = pop0.plot(truth, alpha=1,
                  labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
                  top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"])
@@ -64,10 +64,35 @@ for i, j in product(xrange(len(lpb)-1), xrange(len(lrb)-1)):
     entry = np.vstack((np.random.uniform(lpb[i], lpb[i+1], k),
                        np.random.uniform(lrb[j], lrb[j+1], k))).T
     catalog = np.concatenate((catalog, entry), axis=0)
+dataset = Dataset([catalog])
+print("{0} entries in catalog".format(len(catalog)))
 
 # Plot the actual rate function.
 pl.figure()
-pl.pcolor(lpb, lrb, np.exp(lnrate.T), cmap="gray")
+pl.pcolor(lpb, lrb, np.exp(censor.lncompleteness[1:-1, 1:-1].T), cmap="gray")
+# pl.pcolor(lpb, lrb, np.exp(lnrate.T), cmap="gray")
 pl.plot(catalog[:, 0], catalog[:, 1], ".r", ms=3)
 pl.colorbar()
+pl.xlim(min(lpb), max(lpb))
+pl.ylim(min(lrb), max(lrb))
 pl.savefig("fake/true-rate.png")
+
+# Run inference with the true population.
+model = ProbabilisticModel(dataset, pop0, censor)
+print("Initial ln-prob = {0}".format(model.lnprob(pop0.initial())))
+
+# Set up the sampler.
+p0 = pop0.initial()
+ndim, nwalkers = len(p0), 32
[…]

### 2014-03-29 document/code/fake.py
bin-to-bin smoothness prior

@@ -9,12 +9,13 @@ import numpy as np
 import cPickle as pickle
 from itertools import product
 import matplotlib.pyplot as pl
+import scipy.optimize as op
 
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates, load_petigura_bins)
-from population import (CensoringFunction, BrokenPowerLaw, Histogram,
+from load_data import transit_lnprob0, ln_period0, load_completenes_sim
+from population import (CensoringFunction, BrokenPowerLaw,
                         SeparablePopulation, NormalizedPopulation,
-                        ProbabilisticModel, Dataset)
+                        ProbabilisticModel, Dataset, Population,
+                        SmoothPopulation, BinToBinPopulation)
 
 try:
     os.makedirs("fake")
@@ -24,6 +25,7 @@ except os.error:
 # Define the box that we're going to work in.
 per_rng = np.log([5.0, 400.0])
 rp_rng = np.log([0.5, 64.0])
+truth = [11.0, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
 
 # Load the data.
 ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
@@ -41,13 +43,15 @@ lpb, lrb = censor.bins
 pdist = BrokenPowerLaw(lpb)
 rdist = BrokenPowerLaw(lrb)
 pop0 = SeparablePopulation([pdist, rdist])
-pop0 = NormalizedPopulation(10.8, pop0)
+pop0 = NormalizedPopulation(truth[0], pop0)
 
 # Plot the true distributions.
-truth = [10.8, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
-figs = pop0.plot(truth, alpha=1,
+literature = [(pdist.bins, np.exp(pdist(truth[1:4]))),
+              (rdist.bins, np.exp(rdist(truth[4:])))]
+figs = pop0.plot(truth,
                  labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
-                 top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"])
+                 top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
+                 literature=literature)
 figs[0].savefig(os.path.join("fake", "true-period.png"))
[…]

### 2014-03-29 population.py

@@ -163,10 +163,10 @@ class Population(object):
         return grid[self.inds]
 
     def lnprior(self, theta):
-        return 0.0
+        return np.sum(theta)
 
     def plot(self, thetas, labels=None, ranges=None, alpha=0.3,
-             literature=None, top_axes=None):
+             literature=None, lit_style={}, top_axes=None):
         # Pre-compute the ranges and allowed ranges.
         if ranges is None:
             ranges = [(b.min(), b.max()) for b in self.base]
@@ -212,9 +212,12 @@ class Population(object):
                 axes[i].plot(x, np.exp(y), "k", alpha=0.3)
 
         # Plot literature values.
+        lit_style["marker"] = lit_style.get("marker", ".")
+        lit_style["color"] = lit_style.get("color", "r")
+        lit_style["ls"] = lit_style.get("ls", "None")
         if literature is not None:
             for ax, (x, y) in izip(axes, literature):
-                ax.plot(0.5*(x[1:]+x[:-1]), y, ".r")
+                ax.plot(0.5*(x[1:]+x[:-1]), y, **lit_style)
 
         # Set the axis limits.
         for i, (ax, rng) in enumerate(izip(axes, ranges)):
@@ -317,14 +320,54 @@ class SmoothPopulation(object):
             return -np.inf
 
         # Compute the Gaussian process prior.
-        y = grid.flatten()
-        y -= np.mean(y)
+        y = np.exp(grid.flatten())
+        # y -= np.mean(y)
         chi2 = np.sum(self.dvec/np.exp(theta[1:self.ndim]), axis=2)
         K = np.exp(theta[0] - 0.5 * chi2)
         K += np.diag(self.eps * np.ones_like(y))
         factor, flag = cho_factor(K)
         logdet = np.sum(2*np.log(np.diag(factor)))
         return -0.5 * (np.dot(y, cho_solve((factor, flag), y)) + logdet)
+
+    def plot(self, thetas, **kwargs):
+        thetas = np.atleast_2d(thetas)
+        return self.base_population.plot(thetas[:, self.ndim:], **kwargs)
+
+
+class BinToBinPopulation(object):
+    poisson = False
+
+    def __init__(self, pars, base_population):
+        self.pars = np.atleast_1d(pars)
+        self.base_population = base_population
+        self.base = base_population.base
+        self.ndim = len(self.base)
+        assert len(self.pars) == self.ndim
+
+    def __len__(self):
[…]

### 2014-03-30 document/code/fake_noise.py
synthetic catalogs with noise

@@ -18,7 +18,7 @@ from population import (CensoringFunction, BrokenPowerLaw,
                         SmoothPopulation, BinToBinPopulation)
 
 try:
-    os.makedirs("fake")
+    os.makedirs("fake_noise")
 except os.error:
     pass
 
@@ -52,8 +52,8 @@ figs = pop0.plot(truth,
                  labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
                  top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
                  literature=literature)
-figs[0].savefig(os.path.join("fake", "true-period.png"))
-figs[1].savefig(os.path.join("fake", "true-radius.png"))
+figs[0].savefig(os.path.join("fake_noise", "true-period.png"))
+figs[1].savefig(os.path.join("fake_noise", "true-radius.png"))
 
 # Sample from this censored population.
 lnrate = np.array(censor.lnprob[1:-1, 1:-1])
@@ -68,23 +68,31 @@ for i, j in product(xrange(len(lpb)-1), xrange(len(lrb)-1)):
     entry = np.vstack((np.random.uniform(lpb[i], lpb[i+1], k),
                        np.random.uniform(lrb[j], lrb[j+1], k))).T
     catalog = np.concatenate((catalog, entry), axis=0)
-dataset = Dataset([catalog])
+
+# Add in some observational uncertainties.
+catalog = np.exp(catalog)
+err = np.vstack([np.zeros(len(catalog)), 0.33 * catalog[:, 1]]).T
+catalog += err * np.random.randn(*(err.shape))
+
+dataset = Dataset.sample(catalog, err, samples=100, censor=censor,
+                         functions=[np.log, np.log])
 print("{0} entries in catalog".format(len(catalog)))
 
 # Plot the actual rate function.
 pl.figure()
 pl.pcolor(lpb, lrb, np.exp(censor.lncompleteness[1:-1, 1:-1].T), cmap="gray")
-pl.plot(catalog[:, 0], catalog[:, 1], ".r", ms=3)
+pl.plot(dataset.catalogs[:, :, 0], dataset.catalogs[:, :, 1], ".r", ms=3,
+        alpha=0.3)
 pl.colorbar()
 pl.xlim(min(lpb), max(lpb))
 pl.ylim(min(lrb), max(lrb))
-pl.savefig("fake/true-rate.png")
[…]

### 2014-03-30 document/code/fake_noise.py
computing vmax style histogtrams

@@ -23,6 +23,8 @@ except os.error:
     pass
 
 # Define the box that we're going to work in.
+# per_rng = np.log([6.25, 100.0])
+# rp_rng = np.log([1.0, 16.0])
 per_rng = np.log([5.0, 400.0])
 rp_rng = np.log([0.5, 64.0])
 truth = [11.0, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
@@ -88,15 +90,53 @@ pl.xlim(min(lpb), max(lpb))
 pl.ylim(min(lrb), max(lrb))
 pl.savefig("fake_noise/true-rate.png")
 
-# Run inference on a binned model.
+# Build the binned model.
 bins = [lpb[::8], lrb[::4]]
 print("Run inference on a grid with shape: {0}".format(map(len, bins)))
 pop = Population(bins, censor.bins)
 pop = BinToBinPopulation([-1, -1], pop)
 pop = NormalizedPopulation(truth[0], pop)
-
-# Run inference with the true population.
 model = ProbabilisticModel(dataset, pop, censor)
+
+# Compute the vmax histogram.
+ix0 = np.digitize(np.log(catalog[:, 0]), bins[0])
+iy0 = np.digitize(np.log(catalog[:, 1]), bins[1])
+ix = np.digitize(np.log(catalog[:, 0]), lpb)
+iy = np.digitize(np.log(catalog[:, 1]), lrb)
+lp = censor.lnprob[ix, iy]
+grid = np.zeros((len(bins[0])-1, len(bins[1])-1))
+for i, j in product(range(len(bins[0])-1), range(len(bins[1])-1)):
+    v = lp[(ix0 == i) * (iy0 == j)]
+    grid[i, j] = np.sum(np.exp(-v[np.isfinite(v)]))
+grid[np.isinf(grid)] = 0.0
+
+# Plot the vmax results.
+pl.figure()
+x = bins[0]
+y = np.sum(grid, axis=1)
+y /= np.sum(y * np.diff(x))
+x = np.array(zip(x[:-1], x[1:])).flatten()
+y = np.array(zip(y, y)).flatten()
+pl.plot(x, y, "k")
+x, y = literature[0]
+pl.plot(0.5*(x[1:]+x[:-1]), y, ".r")
+pl.xlim(x[0], x[-1])
+pl.xlabel("$\ln T$")
[…]

### 2014-03-30 document/ms.tex

@@
+%
+%  RULES OF THE GAME
+%
+%  * 80 characters
+%  * line breaks at the ends of sentences
+%  * eqnarrys ONLY
+%  * ALWAYS cite inside parens () and use \citealt{}; no in-line cites
+%  * that is all.
+%
+
 \documentclass[12pt,preprint]{aastex}
+
+\include{vc}
 
@@
 \newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}
-
-%
-%  RULES OF THE GAME, MOTHAFUCKERS.
-%
-%  * 80 characters
-%  * line breaks at the ends of sentences
-%  * eqnarrys ONLY
-%  * ALWAYS cite inside parens () and use \citealt{}; no in-line cites
-%  * that is all.
-%
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics{figures/separable-radius.pdf}
+\end{center}
+\caption{%
+The inferred distribution of exoplanet radii assuming a separable rate
+function.
+The black lines are draws from the posterior probability distribution of bin
+heights and the red points are previous measurements based on the same
+dataset (\citealt{petigura}).
+\figlabel{separable-radius}}
+\end{figure}
+
+\section{General distribution}
+
+It has been shown that the joint period--radius distribution for exoplanets is
+not separable (\dfmtodo{CITE}).
+In this section, we discuss non-separable model in our analysis framework.
+
+We model the rate function $\rate(\period,\,\radius)$ as a piecewise constant
+function in rectangular bins in $\ln\period$ and $\ln\radius$.
+The parameters \ratepars\ that we're fitting are the values of each of these
+bins (subject to a normalization constraint).
+
 \section{Discussion}
@@
 Colbert Report}?)
+
+All of the code used in this project is available from
+\url{http://github.com/dfm/exopop} under the MIT open-source software license.
+This code (plus some dependencies) can be run to re-generate all of the
+figures and results in this \paper; this version of the paper was generated
+with git commit \texttt{\githash} (\gitdate).
 

### 2014-03-30 document/code/fake_noise.py
correct indexing

@@ -85,6 +85,7 @@ pl.figure()
 pl.pcolor(lpb, lrb, np.exp(censor.lncompleteness[1:-1, 1:-1].T), cmap="gray")
 pl.plot(dataset.catalogs[:, :, 0], dataset.catalogs[:, :, 1], ".r", ms=3,
         alpha=0.3)
+pl.plot(np.log(catalog[:, 0]), np.log(catalog[:, 1]), ".b", ms=5)
 pl.colorbar()
 pl.xlim(min(lpb), max(lpb))
 pl.ylim(min(lrb), max(lrb))
@@ -106,7 +107,7 @@ iy = np.digitize(np.log(catalog[:, 1]), lrb)
 lp = censor.lnprob[ix, iy]
 grid = np.zeros((len(bins[0])-1, len(bins[1])-1))
 for i, j in product(range(len(bins[0])-1), range(len(bins[1])-1)):
-    v = lp[(ix0 == i) * (iy0 == j)]
+    v = lp[(ix0 == i+1) * (iy0 == j+1)]
     grid[i, j] = np.sum(np.exp(-v[np.isfinite(v)]))
 grid[np.isinf(grid)] = 0.0
 
@@ -136,6 +137,13 @@ pl.plot(0.5*(x[1:]+x[:-1]), y, ".r")
 pl.xlim(x[0], x[-1])
 pl.xlabel("$\ln R$")
 pl.savefig("fake_noise/vmax-radius.png")
+
+pl.clf()
+pl.pcolor(bins[0], bins[1], grid.T, cmap="gray")
+pl.colorbar()
+pl.xlim(min(lpb), max(lpb))
+pl.ylim(min(lrb), max(lrb))
+pl.savefig("fake_noise/vmax-grid.png")
 assert 0
 
 

### 2014-03-31 document/ms.tex
adding discussion of synthetic catalogs

@@
 \usepackage{amssymb,amsmath}
+\usepackage{subfigure}
 
@@
 
+\dfmtodo{Discuss binned model and choice of priors.}
+
+\dfmtodo{Show where the normalization comes in and describe how it acts
+qualitatively.}
+
 \section{The completeness function \& dataset}
@@
 }
+
+\section{Synthetic catalogs}
+
+The likelihood function in \eq{true-like} provides a \emph{probability
+distribution over catalogs} conditioned on the parameters of the rate
+function.
+Therefore, it is also the description of a \emph{generative process} for
+synthetic catalogs.
+To demonstrate the effectiveness of our model, we generate synthetic catalogs
+from known distributions in period and radius, and test the performance of our
+method on the resulting data.
+For comparison, we also implement the one-over-efficiency weighting procedure
+(see, for example, \citealt{howard, dressing, petigura}) and apply it to the
+same datasets.
+
+\paragraph{The generative process}
+The variable rate Poisson likelihood function in \eq{true-like} can be sampled
+generally for arbitrary forms of the rate function using a procedure called
+``thinning'' (\citealt{poisson}).
+In this \paper, however, we will only consider binned representations of
+\rate, greatly simplifying the generative process.
+Conditioned on a set of rate function parameters and a given completeness
+function (here we'll use the injections from \citealt{petigura}), we compute
+the observable rate grid from \eq{obs-rate}.
+Then in each bin, we sample $J$ candidates uniformly in period and
+radius (as defined by the bin edges) where $J$ is a Poisson random variable
+sampled with the rate given by the observable rate in this bin.
+\dfmtodo{Make this description not be the most confusing thing ever.}
+Then, we add 30\% measurement uncertainties to the radius measurements where
+30\% was chosen to match to properties of the real catalog
+(\citealt{petigura}).
+
+\paragraph{Results}
+See \fig{synthetic-radius}.
+As expected, the forward model is a much lower variance estimator.
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[one-over-efficiency]{%
+\includegraphics[width=0.45\textwidth]{figures/synthetic-vmax-radius.pdf}
+}
+\subfigure[fully-marginalized forward model]{%
+\includegraphics[width=0.45\textwidth]{figures/synthetic-radius.pdf}
+}
+\end{center}
+\caption{%
+The inferred distribution of exoplanet radii based on a noisy observation of a
+synthetic catalog.
+The red dotted line shows the true distribution function and the black
+histograms are the constraints from the different inference techniques.
+\figlabel{synthetic-radius}}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 \begin{center}
@@
 
+\bibitem[Dressing  \& Charbonneau(2013)]{dressing}
+Dressing, C.~D., \& Charbonneau, D.\ 2013, \apj, 767, 95 (\arxiv{1302.1647})
+
 \bibitem[Hogg \etal(2013)]{hoggwhitepaper}
@@
 \bibitem[Hogg \etal(2010)]{hogge}
-Hogg, D.~W., Myers, A.~D., \& Bovy, J.\ 2010, \apj, 725, 2166 (\arxiv{1008.4146})
+Hogg, D.~W., Myers, A.~D., \& Bovy, J.\ 2010, \apj, 725, 2166
+(\arxiv{1008.4146})
 
-\bibitem[Petigura \etal(2013)]{petigura}
-Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013,
+\bibitem[Howard \etal(2012)]{howard}
+Howard, A.~W., Marcy, G.~W., Bryson, S.~T., et al.\ 2012, \apjs, 201, 15
+(\arxiv{1103.2541})
+
+\bibitem[Lewis \& Shedler(1979)]{poisson}
+Lewis, P.~A.~W., \& Shedler, G.~S.\ 1979, Naval Research Logistics Quarterly,
+26, 403
+
+\bibitem[Petigura \etal(2013a)]{petigura-a}
+Petigura, E.~A., Marcy, G.~W., \& Howard, A.~W.\ 2013a, \apj, 770, 69
+(\arxiv{1304.0460})
+
+\bibitem[Petigura \etal(2013b)]{petigura}
+Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013b,
 Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})

### 2014-03-31 document/code/fake_noise.py
running many test cases

@@ -3,6 +3,13 @@
 
 from __future__ import division, print_function
 
+try:
+    from savefig import monkey_patch
+except ImportError:
+    pass
+else:
+    monkey_patch()
+
 import os
 import emcee
 import numpy as np
@@ -10,188 +17,221 @@ import cPickle as pickle
 from itertools import product
 import matplotlib.pyplot as pl
 import scipy.optimize as op
+from scipy.misc import logsumexp
 
 from load_data import transit_lnprob0, ln_period0, load_completenes_sim
 from population import (CensoringFunction, BrokenPowerLaw,
                         SeparablePopulation, NormalizedPopulation,
                         ProbabilisticModel, Dataset, Population,
-                        SmoothPopulation, BinToBinPopulation)
-
-try:
-    os.makedirs("fake_noise")
-except os.error:
-    pass
-
-# Define the box that we're going to work in.
-# per_rng = np.log([6.25, 100.0])
-# rp_rng = np.log([1.0, 16.0])
-per_rng = np.log([5.0, 400.0])
-rp_rng = np.log([0.5, 64.0])
-truth = [11.0, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
-
-# Load the data.
-ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
-                                                     rp_rng=rp_rng)
-tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
-
-# Set up the censoring function.
-censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                           bins=(32, 36),
-                           range=[per_rng, rp_rng],
-                           transit_lnprob_function=tlp)
-
-# Build a synthetic population.
-lpb, lrb = censor.bins
-pdist = BrokenPowerLaw(lpb)
-rdist = BrokenPowerLaw(lrb)
-pop0 = SeparablePopulation([pdist, rdist])
-pop0 = NormalizedPopulation(truth[0], pop0)
-
-# Plot the true distributions.
-literature = [(pdist.bins, np.exp(pdist(truth[1:4]))),
-              (rdist.bins, np.exp(rdist(truth[4:])))]
[…]

### 2014-03-31 document/code/mc_fake.py

@@ -1,0 +1,13 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import numpy as np
+from fake_noise import main
+from multiprocessing import Pool
+
+fns = map("fake_noise/{0:03d}".format, range(16))
+seeds = np.random.randint(50000, size=len(fns))
+pool = Pool()
+pool.map(main, zip(fns, seeds))

### 2014-03-31 population.py

@@ -141,7 +141,11 @@ class Population(object):
     def compute_grid(self, theta):
         # Compute the integral over the first N-1 cells.
         norm = logsumexp(theta + self.ln_cell_area[:-1])
-        if norm >= 0.0:
+        if not np.isfinite(norm):
+            print(norm)
+            print(theta)
+            return None
+        if norm > 0.0:
             return None
 
         # Compute the height of the last cell and cache the heights.
@@ -384,6 +388,9 @@ class NormalizedPopulation(object):
 
     def __len__(self):
         return len(self.base_population) + 1
+
+    def _get_grid(self, theta):
+        return self.base_population._get_grid(theta[1:])
 
     def initial(self):
         return np.append(self.ln_norm, self.base_population.initial())

### 2014-04-02 document/ms.tex
discussion of posterior samplings over catalogs

@@
 
-The ``likelihood function'' given in \eq{true-like} is a model of the
+The likelihood function given in \eq{true-like} is a model of the
 probability of observing a catalog with \emph{true} periods and radii.
@@
 
-\dfmtodo{Introduce the idea of catalogs with uncertainties being a description
-of a posterior sampling over catalogs: $\catalog^{(n)} \sim
-p(\catalog\,|\,\obs{\catalog},\,\interim)$.}
+The catalog of planet candidates considered in this \paper\
+(\citealt{petigura}) was generated by drawing samples from a posterior
+probability distribution for physical parameters (period, radius, impact
+parameter, \etc) conditioned on \kepler\ light curves.
+The $k$th entry in this catalog is, therefore, a representation of the
+posterior PDF
+\begin{eqnarray}\eqlabel{measurement}
+p(\entry_k\,|\,\obs{\entry}_k,\,\interim)
+\end{eqnarray}
+using the notation from the previous section.
+In \eq{measurement}, the notations \interim\ indicates that the measurement
+was made with a specific choice of prior over parameters
+$p(\entry_k\,|\,\interim)$.
+We will refer to these priors as \emph{interim priors} because our goal is to
+infer the distribution that is consistent with the data.
+
+More generally, a published catalog represents a \emph{posterior PDF over
+plausible catalogs}
+\begin{eqnarray}\eqlabel{catalog-pdf}
+p(\catalog\,|\,\obs{\catalog},\,\interim) \quad.
+\end{eqnarray}
+It is normally safe to assume that each entry in the catalog is conditionally
+independent given the choices of interim priors.
+\dfmtodo{Why and when is this a good/bad/necessary assumption?}
+Under this assumption, \eq{catalog-pdf} factorizes as
+\begin{eqnarray}\eqlabel{catalog-factor}
+p(\catalog\,|\,\obs{\catalog},\,\interim) &=&
+\prod_{k=1}^K p(\entry_k\,|\,\obs{\entry}_k,\,\interim) \quad.
+\end{eqnarray}
+
+It will be useful in the following discussion to represent the catalog PDF as
+a set of $N$ samples
+\begin{eqnarray}\eqlabel{catalog-samples}
+\catalog^{(n)} &\sim& p(\catalog\,|\,\obs{\catalog},\,\interim) \quad.
+\end{eqnarray}
+For the general catalog, these could be generated using MCMC or a similar
+sampling algorithm.
+Under the conditional independence assumption of \eq{catalog-factor}, however,
+the problem is much simpler.
+Each entry $\entry_k$ can be sampled independently and then the catalog
+samples $\catalog^{(n)}$ are given by the outer product of the independent
+sample lists.
 
@@
 
-\section{Separable distribution}
-
-The first model that we consider for $\rate(\period,\,\radius)$ is a simple
-separable distribution.
-This assumption means that the rate can be written as
-\begin{eqnarray}
-\rate(\period,\,\radius) &=& \rate(\period)\,\rate(\radius)
-\end{eqnarray}
-and, more specifically, we model the period distribution as a broken power law
-\dfmtodo{CHECK ME!!}
-\begin{eqnarray}
-\rate(\period) &\propto& \frac{\period^\alpha \,
-(\period+\period_0)^{\beta-\alpha}}{\period_0^\beta}
-\end{eqnarray}
-and the radius distribution as a piecewise constant function (or histogram)
-where the parameters are the bin heights.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics{figures/separable-radius.pdf}
-\end{center}
-\caption{%
-The inferred distribution of exoplanet radii assuming a separable rate
-function.
-The black lines are draws from the posterior probability distribution of bin
-heights and the red points are previous measurements based on the same
-dataset (\citealt{petigura}).
-\figlabel{separable-radius}}
-\end{figure}
-
-\section{General distribution}
+\section{Real data}
 

### 2014-04-02 document/ms.tex
clearer description of importance sampling approximation

@@
 
-The quantity that we need to compute is the \emph{marginalized likelihood} of
-the data \obs{\catalog}\ given a particular setting of the rate function
-parameters \ratepars.
-This is achieved by marginalizing over the true periods and radii
+Now that we have a representation of the catalog, we want to infer constraints
+on the occurrence rate of exoplanets conditioned on all the data.
+To perform this inference, we need to compute the  \emph{marginalized
+likelihood}
+\begin{eqnarray}\eqlabel{marg-likelihood}
+p(\obs{\catalog}\,|\,\ratepars) &=&
+\int p(\catalog,\,\obs{\catalog}\,|\,\ratepars) \dd\catalog \quad.
+\end{eqnarray}
+The marginalized likelihood measures the probability of all the \kepler\ light
+curves given a particular setting of the occurrence rate parameters.
+If we can efficiently compute this then we can optimize to find the
+maximum likelihood occurrence rate or we can apply priors $p(\ratepars)$ and
+sample to find posterior constraints on the occurrence rate.
+Note that any results obtained through \eq{marg-likelihood} will
+\emph{correctly account for all the measurement uncertainties}.
+
+The integral in \eq{marg-likelihood} cannot be computed analytically and, in
+fact, it will generally be extremely computationally expensive to evaluate
+numerically (\dfmtodo{CITE}).
+In our case, however, we only need to evaluate this integral up to a constant
+and there is a simple approximation---using the sampling representation of the
+catalog---that can be sufficiently accurate for our purposes.
+To derive this approximation, first recall that the joint distribution inside
+the integral in \eq{marg-likelihood} factorizes as
 \begin{eqnarray}
-p(\obs{\catalog}\,|\,\ratepars) &=&
-\int p(\catalog,\,\obs{\catalog}\,|\,\ratepars) \dd\catalog \\
-&=&
-\int p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog) \dd\catalog
+p(\catalog,\,\obs{\catalog}\,|\,\ratepars) &=&
+p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog)
 \end{eqnarray}
-where the second line is true because the rate function is defined on the true
-parameters.
-This integral can then be written in terms of the sampling under interim
-priors
+because the occurrence rate generates the \emph{true} distribution of periods
+and radii.
+Then, we can multiply the integrand by
 \begin{eqnarray}
-p(\obs{\catalog}\,|\,\ratepars)
-&=&
-\int p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog)\,
 \frac{p(\catalog\,|\,\obs{\catalog},\,\interim)}
      {p(\catalog\,|\,\obs{\catalog},\,\interim)}
-\dd\catalog
-\\
+\end{eqnarray}
+and simplify to find
+\begin{eqnarray}
+\frac{p(\obs{\catalog}\,|\,\ratepars)}
+     {p(\obs{\catalog}\,|\,\interim)}
 &=&
-p(\obs{\catalog}\,|\,\interim)\,
 \int \frac{p(\catalog\,|\,\ratepars)}{p(\catalog\,|\,\interim)}\,
@@
 \end{eqnarray}
-We can approximate this integral using the samples $\catalog^{(n)}$ defined
-above and find the marginalized likelihood up to an unimportant normalization
-constant
-\begin{eqnarray}\eqlabel{marg-like}
+Using the samples over catalogs $\catalog^{(n)}$ as defined above, we can
+approximate this integral---up to an irrelevant normalization constant---as
+\begin{eqnarray}\eqlabel{approx-marg-likelihood}
 p(\obs{\catalog}\,|\,\ratepars) &\appropto&
@@
 where the numerator inside the sum is given by \eq{true-like} and the
-denominator is the interim prior evaluated at the samples.
+denominator is the interim prior evaluated at the coordinates of the samples.
 
-In practice, the variance of this approximation depends on the choice of
-interim prior and, in particular, it is useful to include the completeness
-function.
-\dfmtodo{BLAH BLAH BLAH.}
+This is a generalization of the approximation presented previously in the
+exoplanet literature (\citealt{hogge}) but in this case, the likelihood
+function \eq{true-like} doesn't factorize over entries in the catalog.
+As a result, the variance of \eq{approx-marg-likelihood} is more sensitive to
+the choice of interim prior.
+In practice, we have found that including the completeness function
+$\completeness(\entry)$ in $p(\entry\,|\,\interim)$ greatly improves
+performance.
+This term can be incorporated into existing samplings---generated using less
+informative priors---using rejection sampling.
+
+\section{Parameterization of the occurrence rate}
 

### 2014-04-02 document/ms.tex
re-focus abstract on the method, including uncertainties

@@
 analogs.
-Here we build a set of probabilistic models of the population of known small
-planets around G dwarfs.
-The models take into account the transit probability and the completeness
-(or detection efficiency) as a function of planet radius and orbital period.
-Because we forward model the data, we do not ever have to perform any kind of
-inverse-detection-efficiency weighting of the data.
-Our most sophisticated models do not assume that the joint radius--period
-distribution is separable.
-They also correctly take into account the significant observational
-uncertainties on the planet radii, using the machinery of hierarchical
-inference.
-We reconsider and confirm with weaker assumptions prior results on the
-distribution of exoplanet radii that show a flattening of the distribution
-below a few Earth radii.
+Here we present a general probabilistic framework for making justified
+inferences about the occurrence rate of exoplanets taking into account survey
+completeness \emph{and observational uncertainties}.
+We demonstrate that our method produces lower variance estimates than the
+standard inverse-detection-efficiency weighting procedure.
+When applied to an existing catalog of small planet candidates around G dwarfs
+(Petigura \etal 2013), our method confirms previous measurements of a
+flattening radius distribution near Earth's radius.
 One output of our models is fully marginalized estimates---marginalizing out

### 2014-04-02 document/ms.tex
typos

@@
 
-\hoggtodo{
+\hoggtodo{%
 \kepler\ has been busting out exoplanets (CITE).
@@
 
-\hoggtodo{
+\hoggtodo{%
 This hasn't stopped anyone from estimating the rate at which Sun-like stars
@@
 
-\hoggtodo{
+\hoggtodo{%
 What would make an extrapolation like this conservative?
@@
 density (or distribution) function (PDF) for the data given
-parameteters---and a set of prior PDFs for the parameters.
+parameters---and a set of prior PDFs for the parameters.
 In what follows, we will treat as ``data'' a catalog of exoplanet period

### 2014-04-02 document/ms.tex
smoothness prior

@@
 \newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}
+\newcommand{\binheight}{{\ensuremath{w}}}
+\newcommand{\binheights}{{\ensuremath{\bvec{\binheight}}}}
 
+\newcommand{\smoothreg}{{\ensuremath{\lambda}}}
 \newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}
@@
 
-\dfmtodo{Discuss binned model and choice of priors.}
+In the following sections, we will model the occurrence rate
+$\rate(\period,\,\radius)$ as a two dimensional grid with $N_\period$ bins in
+$\ln\period$ and $N_\radius$ bins in $\ln\radius$.
+We choose to fix the bin locations leaving only the bin heights
+$\binheights = \{\binheight_{ij}\}_{i=1,j=1}^{N_\period,N_\radius}$ as
+free parameters.
+In practice, the results seem to be more numerically stable---but the model is
+exactly equivalent---when we model the rate as
+\begin{eqnarray}
+\rate(\period,\,\radius) &=& \rate_0\,p(\period,\,\radius)
+\end{eqnarray}
+where $\rate_0$ is the total rate of planets in the sample and the relative
+probability factor is subject to the constraint
+\begin{eqnarray}
+1 &=& \int p(\period,\,\radius) \dd\period\dd\radius\quad.
+\end{eqnarray}
+In this model, the parameter set \ratepars\ includes $\rate_0$ and the heights
+of all but one of the bins in $p(\period,\,\radius)$.
 
-\dfmtodo{Show where the normalization comes in and describe how it acts
-qualitatively.}
+Since the choice of bin locations and number is arbitrary, the most
+conservative strategy is to choose very small bins---yielding an effectively
+non-parametric model---and choose a prior (or regularization) that encodes any
+prior expectations about the distribution.
+In particular, it seems reasonable to expect the distribution to be
+``smooth''.
+The most extreme version of this model would be a completely non-parametric
+rate function---generated by, for example, a Gaussian process
+(\citealt{poiss-gp})---but this discussion is beyond the scope of this
+\paper.
+Instead, we'll simply choose small bins and add a term to the likelihood
+function that penalizes large changes between neighboring bins.
+Specifically, we choose a term of the form
+\begin{eqnarray}
+\ln p(\binheights\,|\,\smoothreg_\period,\,\smoothreg_\radius) &=&
+-\frac{1}{2}\sum_{i=2}^{N_\period}
+\frac{|\binheights_{i,\cdot} - \binheights_{i-1,\cdot}|}{\smoothreg_\period}
+-\frac{1}{2}\sum_{j=2}^{N_\radius}
+\frac{|\binheights_{\cdot,j} - \binheights_{\cdot,j-1}|}{\smoothreg_\radius}
+-\frac{1}{2}\ln (\smoothreg_\period\,\smoothreg_\radius)
+\quad.
+\end{eqnarray}
+In our final analysis, we'll marginalize over the regularization strengths
+$\smoothreg_\period$ and $\smoothreg_\radius$ but we'll also explore the
+effects of this choice below.
 
@@
 
+\bibitem[Adams, Murray \& MacKay(2009)]{poiss-gp}
+Adams, R.~P., Murray, I., \& MacKay, D.~J.~C.\ 2009, ICML, 2009, 9
+(\href{http://homepages.inf.ed.ac.uk/imurray2/pub/09poisson/}{online})
+
 \bibitem[Brewer \& Stello(2009)]{brewer}

### 2014-04-03 document/ms.tex
smoothness term should be squared
normaliztion of smoothness prior?
GP prior instead

@@
 free parameters.
-In practice, the results seem to be more numerically stable---but the model is
-exactly equivalent---when we model the rate as
-\begin{eqnarray}
-\rate(\period,\,\radius) &=& \rate_0\,p(\period,\,\radius)
-\end{eqnarray}
-where $\rate_0$ is the total rate of planets in the sample and the relative
-probability factor is subject to the constraint
-\begin{eqnarray}
-1 &=& \int p(\period,\,\radius) \dd\period\dd\radius\quad.
-\end{eqnarray}
-In this model, the parameter set \ratepars\ includes $\rate_0$ and the heights
-of all but one of the bins in $p(\period,\,\radius)$.
+
+\dfmtodo{What about the normalization?}
 
@@
 \paper.
-Instead, we'll simply choose small bins and add a term to the likelihood
-function that penalizes large changes between neighboring bins.
-Specifically, we choose a term of the form
+Instead, we'll simply choose small bins and add a Gaussian process prior
+(\dfmtodo{CITE}) on the logarithms of the bin heights
 \begin{eqnarray}
-\ln p(\binheights\,|\,\smoothreg_\period,\,\smoothreg_\radius) &=&
--\frac{1}{2}\sum_{i=2}^{N_\period}
-\frac{|\binheights_{i,\cdot} - \binheights_{i-1,\cdot}|}{\smoothreg_\period}
--\frac{1}{2}\sum_{j=2}^{N_\radius}
-\frac{|\binheights_{\cdot,j} - \binheights_{\cdot,j-1}|}{\smoothreg_\radius}
--\frac{1}{2}\ln (\smoothreg_\period\,\smoothreg_\radius)
-\quad.
+\ln p(\binheights\,|\,) &=&
+-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}\,\bvec{r}
+-\frac{1}{2}\ln \det \bvec{K} - \frac{N_\period\,N_\radius}{2}\ln 2\pi
 \end{eqnarray}
-In our final analysis, we'll marginalize over the regularization strengths
-$\smoothreg_\period$ and $\smoothreg_\radius$ but we'll also explore the
-effects of this choice below.
+where $\bvec{r}$ is the vector of residuals
+\begin{eqnarray}
+\bvec{r} &=& \left(\begin{array}{cccc}
+\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
+\end{array}\right)^\mathrm{T}
+\end{eqnarray}
+and $\bvec{K}$ is the kernel matrix generated by
+\begin{eqnarray}
+K_{ij} &=&
+\end{eqnarray}
 

### 2014-04-03 document/ms.tex
correct formulation of GP prior

@@
 \newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}
+
+\newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}
+
+\newcommand{\bincenter}{{\ensuremath{\bvec{x}}}}
 \newcommand{\binheight}{{\ensuremath{w}}}
@@
 
-\newcommand{\smoothreg}{{\ensuremath{\lambda}}}
-\newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}
+\newcommand{\smooth}{{\ensuremath{\lambda}}}
 
@@
 In the following sections, we will model the occurrence rate
-$\rate(\period,\,\radius)$ as a two dimensional grid with $N_\period$ bins in
-$\ln\period$ and $N_\radius$ bins in $\ln\radius$.
+$\rate(\period,\,\radius)$ as a two dimensional grid with $N$ bins in
+$\ln\period$ and $\ln\radius$.
 We choose to fix the bin locations leaving only the bin heights
-$\binheights = \{\binheight_{ij}\}_{i=1,j=1}^{N_\period,N_\radius}$ as
-free parameters.
+$\binheights = \{\binheight_i\}_{i=1}^{N}$ as free parameters.
 
@@
 \begin{eqnarray}
-\ln p(\binheights\,|\,) &=&
--\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}\,\bvec{r}
--\frac{1}{2}\ln \det \bvec{K} - \frac{N_\period\,N_\radius}{2}\ln 2\pi
+\ln p(\binheights\,|\,\mu,\,\bvec{\smooth}) &=&
+-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}^{-1}\,\bvec{r}
+-\frac{1}{2}\ln \det \bvec{K} - \frac{N}{2}\ln 2\pi
 \end{eqnarray}
@@
 \end{eqnarray}
-and $\bvec{K}$ is the kernel matrix generated by
+and $\bvec{K}$ is the kernel matrix evaluated \emph{at the bin centers}
+$\bincenter_i$
 \begin{eqnarray}
-K_{ij} &=&
+K_{ij}(\bvec{\smooth}) &\equiv& \smooth_0^2\,\exp \left(
+-\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}
+\left[\begin{array}{cc}
+1/\smooth_1 & 0 \\ 0 & 1/\smooth_2
+\end{array}\right]
+[\bincenter_i-\bincenter_j] \right) \quad.
 \end{eqnarray}
+
+Below, we will explore the effects of this choice of prior but when we
+generate our final results, we will actually apply hyperpriors to the
+parameters $\mu$ and $\bvec{\smooth}$ and marginalize over their values.
 

### 2014-04-04 document/code/fake_noise.py
working with smaller bins

@@ -21,9 +21,9 @@ from scipy.misc import logsumexp
 
 from load_data import transit_lnprob0, ln_period0, load_completenes_sim
 from population import (CensoringFunction, BrokenPowerLaw,
-                        SeparablePopulation, NormalizedPopulation,
+                        SeparablePopulation,
                         ProbabilisticModel, Dataset, Population,
-                        BinToBinPopulation)
+                        SmoothPopulation)
 
 
 def main(args, state=None):
@@ -55,7 +55,7 @@ def main(args, state=None):
 
     # Set up the censoring function.
     censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                               bins=(32, 36),
+                               bins=(32, 40),
                                range=[per_rng, rp_rng],
                                transit_lnprob_function=tlp)
 
@@ -63,8 +63,7 @@ def main(args, state=None):
     lpb, lrb = censor.bins
     pdist = BrokenPowerLaw(lpb)
     rdist = BrokenPowerLaw(lrb)
-    pop0 = SeparablePopulation([pdist, rdist])
-    pop0 = NormalizedPopulation(truth[0], pop0)
+    pop0 = SeparablePopulation([pdist, rdist], lnnorm=truth[0])
 
     # Plot the true distributions.
     literature = [(pdist.bins, np.exp(pdist(truth[1:4]))),
@@ -119,10 +118,11 @@ def main(args, state=None):
     pl.savefig(os.path.join(bp, "true-rate.pdf"))
 
     # Build the binned model.
-    bins = [lpb[::8], lrb[::4]]
-    print("Run inference on a grid with shape: {0}".format(map(len, bins)))
-    pop = Population(bins, censor.bins)
-    pop = NormalizedPopulation(truth[0], pop)
+    bins = [lpb[::4], lrb[::4]]
+    print("Run inference on a grid with shape: {0}"
+          .format([len(b)-1 for b in bins]))
+    pop = Population(bins, censor.bins, lnnorm=truth[0])
+    pop = SmoothPopulation([7.0, 0.2, 0.0, 0.2], pop)
     model = ProbabilisticModel(dataset, pop, censor)
 
     # Compute the vmax histogram.
@@ -143,15 +143,11 @@ def main(args, state=None):
[…]

### 2014-04-04 population.py

@@ -96,14 +96,14 @@ class Histogram(object):
 
 
 class Population(object):
-    poisson = False
 
-    def __init__(self, bins, base=None):
+    def __init__(self, bins, base=None, lnnorm=0.0):
         self.bins = bins
         if base is None:
             base = bins
         self.base = base
         self.shape = np.array(map(len, bins)) - 1
+        self.lnnorm = lnnorm
 
         # Make sure that the resampling is sane.
         assert all([b1[0] == b2[0] for b1, b2 in izip(bins, base)])
@@ -130,27 +130,15 @@ class Population(object):
         self._cache_val = None
 
     def __len__(self):
-        return self.ncells - 1
+        return self.ncells
 
     def initial(self):
         v = np.zeros(self.ncells)
-        v -= logsumexp(v + self.ln_cell_area)
-        # v[:3] = [0.5, -3.0, -3.0]
-        return v[:-1]
+        v += self.lnnorm - logsumexp(v + self.ln_cell_area)
+        return v
 
     def compute_grid(self, theta):
-        # Compute the integral over the first N-1 cells.
-        norm = logsumexp(theta + self.ln_cell_area[:-1])
-        if not np.isfinite(norm):
-            print(norm)
-            print(theta)
-            return None
-        if norm > 0.0:
-            return None
-
-        # Compute the height of the last cell and cache the heights.
-        v = np.log(1.0 - np.exp(norm)) - self.ln_cell_area[-1]
-        return np.append(theta, v).reshape(self.shape)
+        return theta.reshape(self.shape)
 
     def _get_grid(self, theta):
         k = tuple(theta)
@@ -251,32 +239,32 @@ class Population(object):
 
 
 class SeparablePopulation(Population):
-    poisson = False
 
-    def __init__(self, distributions):
+    def __init__(self, distributions, lnnorm=0.0):
         self.distributions = distributions
-        self.npars = sum(map(len, distributions))
+        self.npars = sum(map(len, distributions)) + 1
         super(SeparablePopulation, self).__init__(
-            [d.base for d in distributions])
+            [d.base for d in distributions], lnnorm=lnnorm)
 
     def __len__(self):
         return self.npars
 
     def initial(self):
-        return np.concatenate([d.initial() for d in self.distributions])
+        return np.concatenate([self.lnnorm]
[…]

### 2014-04-04 document/code/fake_noise.py
synthetic 'petigura-style' catalogs

@@ -20,7 +20,7 @@ import scipy.optimize as op
 from scipy.misc import logsumexp
 
 from load_data import transit_lnprob0, ln_period0, load_completenes_sim
-from population import (CensoringFunction, BrokenPowerLaw,
+from population import (CensoringFunction, Histogram,
                         SeparablePopulation,
                         ProbabilisticModel, Dataset, Population,
                         SmoothPopulation)
@@ -28,7 +28,8 @@ from population import (CensoringFunction, BrokenPowerLaw,
 
 def main(args, state=None):
     bp, seed = args
-    np.random.seed(seed)
+    if seed is not None:
+        np.random.seed(seed)
     try:
         os.makedirs(bp)
     except os.error:
@@ -44,9 +45,12 @@ def main(args, state=None):
     # Define the box that we're going to work in.
     # per_rng = np.log([6.25, 100.0])
     # rp_rng = np.log([1.0, 16.0])
-    per_rng = np.log([5.0, 400.0])
-    rp_rng = np.log([0.5, 64.0])
-    truth = [11.0, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
+    # per_rng = np.log([5.0, 400.0])
+    # rp_rng = np.log([0.5, 64.0])
+
+    # Bins chosen to include EP's bins.
+    per_rng, np_bins = np.log([6.25, 400.0]), 4*6
+    rp_rng, nr_bins = np.log([0.5, 32]), 4*12
 
     # Load the data.
     ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
@@ -55,19 +59,33 @@ def main(args, state=None):
 
     # Set up the censoring function.
     censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                               bins=(32, 40),
+                               bins=(np_bins, nr_bins),
                                range=[per_rng, rp_rng],
                                transit_lnprob_function=tlp)
 
-    # Build a synthetic population.
+    # print(np.exp(censor.bins[0][::4]))
+    # print(np.exp(censor.bins[1][::4]))
+
+    # The values from EP's paper (+some made up numbers).
     lpb, lrb = censor.bins
-    pdist = BrokenPowerLaw(lpb)
-    rdist = BrokenPowerLaw(lrb)
[…]

### 2014-04-04 population.py

@@ -4,7 +4,7 @@
 from __future__ import division, print_function
 
 __all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",
-           "NormalizedPopulation", "Dataset", "CensoringFunction"]
+           "Dataset", "CensoringFunction"]
 
 from itertools import izip
 
@@ -74,25 +74,17 @@ class Histogram(object):
         assert np.all((self.inds >= 0) * (self.inds < len(self.bins)))
 
     def __len__(self):
-        return self.nbins - 1
+        return self.nbins
 
     def initial(self):
         v = np.zeros(self.nbins) - logsumexp(self.ln_bin_widths)
-        return v[:-1]
+        return v
 
     def lnprior(self, theta):
         return 0.0
 
     def __call__(self, theta):
-        norm = logsumexp(theta + self.ln_bin_widths[:-1])
-        if norm >= 0.0:
-            return None
-
-        # Compute the height of the last cell.
-        v = np.log(1.0 - np.exp(norm)) - self.ln_bin_widths[-1]
-        ln_heights = np.append(theta, v)
-
-        return ln_heights[self.inds]
+        return theta[self.inds]
 
 
 class Population(object):

### 2014-04-04 document/ms.tex
description of completeness model

@@
 
-\dfmtodo{%
-Describe Erik's data products and our model of the completeness in boxes
-extended artificially beyond the original range.
-Also describe the re-sampling procedure for the candidate parameters.
-}
+Erik Petigura and collaborators (\citealt{petigura-a,petigura}) heroically
+determined the empirical end-to-end completeness of their transit search
+pipeline by injecting synthetic signals of known period and radius into the
+raw aperture photometry and testing recovery rates.
+They also published a catalog of XXX small planet candidates around G type
+stars; YYY of them previously unknown.
+Given the unprecedented quality of the completeness characterization, we will
+focus on this dataset in this \paper.
+
+We use all the injected samples from \citet{petigura} to compute the mean
+(marginalized) detection efficiency in bins of $\ln\period$ and $\ln\radius$.
+In each bin, this efficiency is simply the fraction of recovered injections.
+For the purposes of this \paper, we will neglect the uncertainties introduced
+by the fact that this estimate is being made with a finite number of samples.
+
+The largest injected radius is $16\,R_\oplus$ but, because of the measurement
+uncertainties on the radii, it will be necessary for us to model the
+distribution to larger radii.
+To do this, we approximate the survey completeness for $\radius > 16\,R\oplus$
+as 1.
+
+\dfmtodo{Hopefully Erik will send us MCMC samples. If not, describe the
+re-sampling procedure that we used. Either way, we'll need to re-sample under
+the completeness probability for the purposes of the importance sampling.}
 

### 2014-04-05 document/ms.tex
discussing inverse-detection-frequency

@@
 
-\section{Synthetic catalogs}
+\section{Validation using synthetic catalogs}
+
+To confirm that the probabilistic framework developed in the previous sections
+provides correct inferences about the population of exoplanets given noisy
+data, we'll start by generating synthetic catalogs and running our full
+inference pipeline on those data.
+In this section, we also demonstrate the key benefits of our method over the
+standard intuitive inverse-detection-efficiency procedure.
+
+\paragraph{Inverse-detection-efficiency}
+Authors in the exoplanet literature (for example, \citealt{howard, dressing,
+petigura}) have built population level models using this procedure.
+The basic idea is to make a histogram of planet candidates in the model where
+each sample has a weight proportional to the inverse detection efficiency.
+This method makes intuitive sense and, in the limit of infinitely precise
+measurements, it is not wrong.
 

### 2014-04-10 document/ms.tex
note to remember to discuss multiplicity
working on real data

@@
 than anyone who has come before.
+}
+
+\dfmtodo{%
+Cite \citet{tremaine} for likelihood function including multiplicity.
 }
@@
 K_{ij}(\bvec{\smooth}) &\equiv& \smooth_0^2\,\exp \left(
--\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}
-\left[\begin{array}{cc}
-1/\smooth_1 & 0 \\ 0 & 1/\smooth_2
-\end{array}\right]
+-\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}\,
+\mathrm{diag}(1/\smooth_k^2)\,
 [\bincenter_i-\bincenter_j] \right) \quad.
@@
 measurements, it is not wrong.
+Even in this limit, this estimator of the occurrence distributions will have
+large variance; especially as the detection efficiency becomes low.
+In the presence of observational uncertainties, the inferred distributions
+will be biased.
 
@@
 
+\bibitem[Tremaine \& Dong(2012)]{tremaine}
+Tremaine, S., \& Dong, S.\ 2012, \aj, 143, 94 (\arxiv{1106.5403})
+
 \bibitem[Youdin(2011)]{youdin}

### 2014-04-10 document/code/fake_noise.py

@@ -86,14 +86,14 @@ def main(args, state=None):
     # Plot the true distributions.
     literature = [(pdist.base, np.exp(pdist(truth[1:1+len(p_vals)]))),
                   (rdist.base, np.exp(rdist(truth[1+len(p_vals):])))]
-    figs = pop0.plot(truth,
-                     labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
-                     top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
-                     literature=literature)
-    figs[0].savefig(os.path.join(bp, "true-period.png"))
-    figs[0].savefig(os.path.join(bp, "true-period.pdf"))
-    figs[1].savefig(os.path.join(bp, "true-radius.png"))
-    figs[1].savefig(os.path.join(bp, "true-radius.pdf"))
+    # figs = pop0.plot(truth,
+    #                  labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
+    #                  top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
+    #                  literature=literature)
+    # figs[0].savefig(os.path.join(bp, "true-period.png"))
+    # figs[0].savefig(os.path.join(bp, "true-period.pdf"))
+    # figs[1].savefig(os.path.join(bp, "true-radius.png"))
+    # figs[1].savefig(os.path.join(bp, "true-radius.pdf"))
 
     # Sample from this censored population.
     lnrate = np.array(censor.lnprob[1:-1, 1:-1])
@@ -122,6 +122,13 @@ def main(args, state=None):
     dataset = Dataset.sample(catalog, err, samples=100, censor=censor,
                              functions=[np.log, np.log])
     print("{0} entries in catalog".format(len(catalog)))
+
+    # Plot the initial population.
+    fig = pop0.plot_2d(truth, censor=censor, catalog=np.log(catalog),
+                       labels=["$\ln T/\mathrm{days}$", "$\ln R/R_\oplus$"],
+                       top_axes=["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"],
+                       true=literature, alpha=1)
+    fig.savefig(os.path.join(bp, "initial.png"))
 
     # Plot the actual rate function.
     pl.figure()
[…]

### 2014-04-10 document/code/load_data.py

@@ -40,45 +40,13 @@ def load_completenes_sim(rp_func=np.log, per_rng=np.log([5, 400]),
     return x, y, z
 
 
-def load_candidates(censor=None, samples=0, N=10000, rp_func=np.log):
+def load_candidates():
     lines = open(os.path.join(bp, "table_ekoi836.tex")).readlines()
-    K = len(lines)
-    good = np.ones(K, dtype=bool)
-    catalogs = np.empty((max(samples, 1), K, 2))
-    lnweights = np.zeros(max(samples, 1))
-    for k, line in enumerate(lines):
-        cols = line.split("&")
-        log_per = np.log(float(cols[2]))
-
-        if samples > 0:
-            m, s = map(float, cols[12:14])
-            catalogs[:, k, 0] = log_per
-            if censor is None:
-                catalogs[:, k, 1] = rp_func(m+s*np.random.randn(samples))
-                continue
-            r = m + s*np.random.randn(N)
-            r = rp_func(r[r > 0.0])
-            v = np.vstack([log_per+np.zeros_like(r), r]).T
-            lnp = censor.get_lncompleteness(v)
-            mask = np.random.rand(len(lnp)) < np.exp(lnp)
-            if np.sum(mask) < samples:
-                print("Dropping candidate at R={0}".format(m))
-                good[k] = 0
-                continue
-            inds = np.arange(len(lnp))[mask]
-            catalogs[:, k, 1] = r[inds[:samples]]
-            lnweights += lnp[inds[:samples]]
-        else:
-            r = rp_func(float(cols[12]))
-            lnp = censor.get_lncompleteness([log_per, r])
-            if not np.isfinite(lnp):
-                print("Dropping candidate at R={0}".format(float(cols[12])))
-                good[k] = 0
-                continue
-            catalogs[0, k, 0] = log_per
-            catalogs[0, k, 1] = r
-
-    return population.Dataset(catalogs[:, good], lnweights)
+    data = np.array([[l.split("&")[i] for i in (0, 2, 12, 13)] for l in lines
+                     if l.split("&")[4].strip() == "P"],
+                    dtype=float)
+    return (np.array(data[:, 0], dtype=int), data[:, 1:3],
+            np.vstack([np.zeros(len(data)), data[:, 3]]).T)
 
 
 def load_petigura_bins(mylog=np.log):

### 2014-04-10 document/code/main.py

@@ -1,0 +1,113 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+try:
+    from savefig import monkey_patch
+except ImportError:
+    pass
+else:
+    monkey_patch()
+
+import os
+import emcee
+import numpy as np
+import cPickle as pickle
+from itertools import product
+import matplotlib.pyplot as pl
+import scipy.optimize as op
+from scipy.misc import logsumexp
+
+from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
+                       load_candidates)
+from population import (CensoringFunction, Histogram,
+                        SeparablePopulation,
+                        ProbabilisticModel, Dataset, Population,
+                        SmoothPopulation)
+
+
+def main():
+    bp = "main"
+    try:
+        os.makedirs(bp)
+    except os.error:
+        pass
+
+    # Bins chosen to include EP's bins.
+    per_rng, np_bins = np.log([6.25, 400.0]), 4*6
+    rp_rng, nr_bins = np.log([0.5, 32]), 4*12
+
+    # Load the data.
+    ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
+                                                         rp_rng=rp_rng)
+    tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
+
+    # Set up the censoring function.
+    censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
+                               bins=(np_bins, nr_bins),
+                               range=[per_rng, rp_rng],
+                               transit_lnprob_function=tlp)
+
+    # The values from EP's paper (+some made up numbers).
+    lpb, lrb = censor.bins
+    x, y = lpb[::4], lrb[::4]
+    p_vals = np.array([8.9, 13.7, 15.8, 15.2, 15., 14.8])
+    r_vals = np.array([11, 11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
+                       0.5, 0.5])
+    literature = [(x, p_vals / np.sum(p_vals*np.diff(x))),
+                  (y, r_vals / np.sum(r_vals*np.diff(y)))]
+
+    # Load the candidates.
+    ids, catalog, err = load_candidates()
+    print(catalog)
[…]

### 2014-04-10 population.py

@@ -14,6 +14,7 @@ from scipy.misc import logsumexp
 from scipy.linalg import cho_factor, cho_solve
 
 import matplotlib.pyplot as pl
+from matplotlib.ticker import MaxNLocator
 from matplotlib.ticker import ScalarFormatter
 
 
@@ -148,6 +149,138 @@ class Population(object):
 
     def lnprior(self, theta):
         return np.sum(theta)
+
+    def plot_2d(self, thetas, ranges=None, censor=None, catalog=None,
+                labels=None, top_axes=None, literature=None, lit_style={},
+                true=None, true_style={}, alpha=0.3):
+        assert len(self.base) == 2
+
+        # Pre-compute the ranges and allowed ranges.
+        if ranges is None:
+            ranges = [(b.min(), b.max()) for b in self.base]
+        bm = []
+        vm = []
+        for rng, b in izip(ranges, self.base):
+            if rng is None:
+                rng = (b.min(), b.max())
+            m = (b >= rng[0]) * (b <= rng[1])
+            bm.append(m)
+            vm.append(np.arange(len(m) - 1)[m[:-1] * m[1:]])
+        vm = np.meshgrid(*vm, indexing="ij")
+
+        # Compute the grid images.
+        thetas = np.atleast_2d(thetas)
+        grids = np.array([self.evaluate(t)[vm] for t in thetas])
+
+        fig = pl.figure(figsize=(10, 10))
+        ax = pl.axes([0.1, 0.1, 0.6, 0.6])
+
+        ax_top = pl.axes([0.1, 0.7, 0.6, 0.2])
+        ax_top.set_xticklabels([])
+        ax_right = pl.axes([0.7, 0.1, 0.2, 0.6])
+        ax_right.set_yticklabels([])
+
+        # Plot the occurence image.
+        img = np.median(grids, axis=0).T
+        ax.pcolor(self.base[0], self.base[1], np.exp(img),
+                  cmap="gray", alpha=0.8)
+
+        # Plot the occurence histograms.
+        ys = [logsumexp(grids
+                        + np.log(np.diff(self.base[1][bm[1]]))[None, None, :],
+                        axis=2),
+              logsumexp(grids
+                        + np.log(np.diff(self.base[0][bm[0]]))[None, :, None],
+                        axis=1)]
+        for i, a in enumerate([ax_top, ax_right]):
+            x0 = self.base[i][bm[i]]
[…]

### 2014-04-10 document/code/main.py
working on real data

@@ -36,7 +36,8 @@ def main():
 
     # Bins chosen to include EP's bins.
     per_rng, np_bins = np.log([6.25, 400.0]), 4*6
-    rp_rng, nr_bins = np.log([0.5, 32]), 4*12
+    rp_rng, nr_bins = np.log([1.0, 32]), 4*11
+    # rp_rng, nr_bins = np.log([0.5, 32]), 4*12
 
     # Load the data.
     ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
@@ -53,8 +54,10 @@ def main():
     lpb, lrb = censor.bins
     x, y = lpb[::4], lrb[::4]
     p_vals = np.array([8.9, 13.7, 15.8, 15.2, 15., 14.8])
-    r_vals = np.array([11, 11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
+    r_vals = np.array([11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
                        0.5, 0.5])
+    # r_vals = np.array([11, 11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
+    #                    0.5, 0.5])
     literature = [(x, p_vals / np.sum(p_vals*np.diff(x))),
                   (y, r_vals / np.sum(r_vals*np.diff(y)))]
 

### 2014-04-10 document/code/main.py
sampling for real data

@@ -11,23 +11,20 @@ else:
     monkey_patch()
 
 import os
+import h5py
 import emcee
 import numpy as np
 import cPickle as pickle
 from itertools import product
-import matplotlib.pyplot as pl
-import scipy.optimize as op
-from scipy.misc import logsumexp
 
 from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
                        load_candidates)
-from population import (CensoringFunction, Histogram,
-                        SeparablePopulation,
+from population import (CensoringFunction,
                         ProbabilisticModel, Dataset, Population,
                         SmoothPopulation)
 
 
-def main():
+def main(ep_bins=True):
     bp = "main"
     try:
         os.makedirs(bp)
@@ -35,9 +32,17 @@ def main():
         pass
 
     # Bins chosen to include EP's bins.
-    per_rng, np_bins = np.log([6.25, 400.0]), 4*6
-    rp_rng, nr_bins = np.log([1.0, 32]), 4*11
-    # rp_rng, nr_bins = np.log([0.5, 32]), 4*12
+    if ep_bins:
+        per_rng, np_bins = np.log([6.25, 100.0]), 4*4
+        rp_rng, nr_bins = np.log([1.0, 16]), 4*8
+        p_vals = np.array([8.9, 13.7, 15.8, 15.2])
+        r_vals = np.array([12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7])
+    else:
+        per_rng, np_bins = np.log([6.25, 400.0]), 4*6
+        rp_rng, nr_bins = np.log([0.5, 32]), 4*12
+        p_vals = np.array([8.9, 13.7, 15.8, 15.2, 15., 14.8])
+        r_vals = np.array([11, 11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
+                           0.5, 0.5])
 
     # Load the data.
     ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
@@ -53,20 +58,12 @@ def main():
     # The values from EP's paper (+some made up numbers).
     lpb, lrb = censor.bins
     x, y = lpb[::4], lrb[::4]
-    p_vals = np.array([8.9, 13.7, 15.8, 15.2, 15., 14.8])
-    r_vals = np.array([11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
-                       0.5, 0.5])
[…]

### 2014-04-10 population.py

@@ -202,14 +202,8 @@ class Population(object):
 
                 if i:
                     a.plot(y, x, "k", alpha=alpha)
-                    a.set_xlim(np.array((-0.1, 1.05))*(a.get_xlim()[1]))
-                    a.axvline(0.0, color="k", alpha=0.3)
-                    a.xaxis.set_major_locator(MaxNLocator(4))
                 else:
                     a.plot(x, y, "k", alpha=alpha)
-                    a.set_ylim(np.array((-0.1, 1.05))*(a.get_ylim()[1]))
-                    a.axhline(0.0, color="k", alpha=0.3)
-                    a.yaxis.set_major_locator(MaxNLocator(4))
 
         # Plot the completeness contours.
         if censor is not None:
@@ -253,6 +247,13 @@ class Population(object):
         ax.set_ylim(ranges[1])
         ax_top.set_xlim(ranges[0])
         ax_right.set_ylim(ranges[1])
+
+        ax_top.set_ylim(np.array((-0.1, 1.05))*(ax_top.get_ylim()[1]))
+        ax_top.axhline(0.0, color="k", alpha=0.3)
+        ax_top.yaxis.set_major_locator(MaxNLocator(4))
+        ax_right.set_xlim(np.array((-0.1, 1.05))*(ax_right.get_xlim()[1]))
+        ax_right.axvline(0.0, color="k", alpha=0.3)
+        ax_right.xaxis.set_major_locator(MaxNLocator(4))
 
         if labels is not None:
             ax.set_xlabel(labels[0])

### 2014-04-11 population.py
not linear prior

@@ -148,7 +148,8 @@ class Population(object):
         return grid[self.inds]
 
     def lnprior(self, theta):
-        return np.sum(theta)
+        return 0.0
+        # return np.sum(theta)
 
     def plot_2d(self, thetas, ranges=None, censor=None, catalog=None,
                 labels=None, top_axes=None, literature=None, lit_style={},

### 2014-04-11 document/code/main.py
moar samples

@@ -132,7 +132,7 @@ def main(ep_bins=True):
     assert np.all(finite), "{0}".format(np.sum(finite))
 
     # Run the sampler.
-    N = 1000
+    N = 50000
     fn = os.path.join(bp, "results.h5")
     with h5py.File(fn, "w") as f:
         f.create_dataset("chain", shape=(nwalkers, N, ndim), dtype=np.float64)

### 2014-04-19 document/ms.tex
adding derivation of maximum likelihood ide procedure

@@
 % Document specific variables.
-\newcommand{\rate}{\ensuremath{\eta}}
-\newcommand{\ratepars}{{\ensuremath{\bvec{\theta}}}}
+\newcommand{\rate}{\ensuremath{\Gamma}}
+\newcommand{\ratepar}{{\ensuremath{\theta}}}
+\newcommand{\ratepars}{{\ensuremath{\bvec{\ratepar}}}}
 \newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
@@
 \newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
-\newcommand{\entry}{\ensuremath{S}}
+\newcommand{\entry}{\ensuremath{w}}
 \newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}
@@
 
+\newcommand{\binarea}{{\ensuremath{\Delta}}}
 \newcommand{\bincenter}{{\ensuremath{\bvec{x}}}}
@@
 
+\appendix
+
+\section{Maximum likelihood estimators for noiseless catalogs}
+
+Equation XXXX gives the likelihood of a catalog $\catalog =
+\{\entry_k\}_{k=1}^K$, conditioned on a particular choice of occurence rate
+parameters \ratepars:
+\begin{eqnarray}
+\ln p(\catalog\,|\,\ratepars) &=& \sum_{k=1}^K \ln \rate(\entry_k)
+    - \int \rate(\entry) \dd\entry \quad.
+\end{eqnarray}
+If we model the occurence rate as a piecewise constant function
+\begin{eqnarray}
+\rate (\entry) &=& \left \{\begin{array}{ll}
+\ratepar_1 & \mathrm{if\,} \entry \in \binarea_1, \\
+\cdots & \cdots \\
+\ratepar_N & \mathrm{if\,} \entry \in \binarea_N, \\
+0 & \mathrm{otherwise}
+\end{array}\right.
+\end{eqnarray}
+where $\binarea_n$ is the volume of the bin $n$, the log likelihood becomes
+\begin{eqnarray}
+\ln p(\catalog\,|\,\ratepars) &=&
+    \sum_{k=1}^K \sum_{n=1}^N \mathbf{1}[\entry_k \in \binarea_n]
+        \ln\ratepar_n
+    - \sum_{n=1}^N \ratepar_n\,\binarea_n \quad.
+\end{eqnarray}
+The maximum likelihood solution to this can be computed analytically
+\begin{eqnarray}
+\frac{\dd \ln p(\catalog\,|\,\ratepars)}{\dd \ratepar_m} &=&
+    \frac{N_m}{\ratepar_m} - \binarea_m \\
+\to \quad \ratepar_m &=& \frac{N_m}{\binarea_m}
+\end{eqnarray}
+where $N_m$ is the number of objects in the bin $m$.
+This result is simply a \emph{normalized histogram} of the points.
+
+Including the completeness function $\completeness(\entry)$ is as simple as
+modeling the observable occurence rate as
+\begin{eqnarray}
+\rate (\entry) &=& \completeness(\entry) \, \left \{\begin{array}{ll}
+\ratepar_1 & \mathrm{if\,} \entry \in \binarea_1, \\
+\cdots & \cdots \\
+\ratepar_N & \mathrm{if\,} \entry \in \binarea_N, \\
+0 & \mathrm{otherwise}
+\end{array}\right. \quad.
+\end{eqnarray}
+This leads to a similar analytic maximum likelihood solution
+\begin{eqnarray}
+\frac{\dd \ln p(\catalog\,|\,\ratepars)}{\dd \ratepar_m} &=&
+    \frac{N_m}{\ratepar_m}
+    - \int_{\binarea_m} \completeness(\entry)\dd\entry \\
+\to \quad \ratepar_m &=& N_m \,
+    \left [\int_{\binarea_m} \completeness(\entry)\dd\entry \right ]^{-1}
+\end{eqnarray}
+where the remaining integral is over the bin $m$.
+This result is \emph{almost} the same as the inverse-detection-efficiency
+procedure but each point is weighted by the bin-averaged efficiency instead of
+the per-object efficiency.
+
 \newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}

### 2014-04-19 document/ms.tex
some incomplete general discussion of hierarchical inference

@@
 \newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
-\newcommand{\entry}{\ensuremath{w}}
+\newcommand{\data}{{\ensuremath{\bvec{x}}}}
+\newcommand{\entry}{{\ensuremath{\bvec{w}}}}
 \newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}
@@
 When applied to an existing catalog of small planet candidates around G dwarfs
-(Petigura \etal 2013), our method confirms previous measurements of a
+(Petigura \etal\ 2013), our method confirms previous measurements of a
 flattening radius distribution near Earth's radius.
@@
 samplings.
+
+\section{A brief introduction to hierarchical inference}
+
+In this \paper, we are asking the question: \emph{what constraints can we put
+on the occurence rate of exoplanets given all the light curves measured by
+\project{Kepler}?}
+Mathematically, this involves evaluating the so-called \emph{marginalized
+likelihood function}
+\begin{eqnarray}\eqlabel{crazylike}
+p(\{\data_k\}\,|\,\ratepars) &=&
+    \int\left[\prod_{k=1}^K p(\data_k\,|\,\entry_k)\right ]
+    \,p(\{\entry_k\}\,|\,\ratepars)
+    \dd\{\entry_k\}
+\end{eqnarray}
+where $\{\data_k\}$ is the set of all light curves, \ratepars\ is a vector of
+parameters describing the occurence rate function $\rate_\ratepars(\entry)$
+and \entry\ is a vector of physical parameters describing a planetary system
+(period, radius ratio, \etc).
+
+It is generally computationally intractable to compute the likelihood in
+\eq{crazylike} because it involves the numerical evaluation of an extremely
+ high dimensional---the number of physical parameters times the number of
+targets---and computationally costly integral.
+Instead, we can use existing catalogs of planet candidates to approximate the
+marginalization integral.
+For our purposes, this is the definition of \emph{hierarchical inference}.
+
+A very simple example of such an inference is a procedure that should be very
+familiar: the histogram.
+In the limit of negligible uncertainties, the maximum marginalized likelihood
+for the occurence rate is a histogram of the objects in the catalog (see
+\app{histogram} for the derivation).
 
@@
 \section{Maximum likelihood estimators for noiseless catalogs}
+\sectlabel{histogram}
 

### 2014-04-20 document/ms.tex
better discussion of into materials

@@
 on the occurence rate of exoplanets given all the light curves measured by
-\project{Kepler}?}
+\kepler?}
 Mathematically, this involves evaluating the so-called \emph{marginalized
@@
 It is generally computationally intractable to compute the likelihood in
-\eq{crazylike} because it involves the numerical evaluation of an extremely
- high dimensional---the number of physical parameters times the number of
-targets---and computationally costly integral.
+\eq{crazylike} because it involves a high dimensional numerical integration of
+a function where each evaluation of the integrand touches \emph{every data
+point}.
 Instead, we can use existing catalogs of planet candidates to approximate the
@@
 
-A very simple example of such an inference is a procedure that should be very
-familiar: the histogram.
-In the limit of negligible uncertainties, the maximum marginalized likelihood
-for the occurence rate is a histogram of the objects in the catalog (see
-\app{histogram} for the derivation).
+A catalog is a dimensionality reduction of the data that it is based on where
+all the (relevant) information has been retained (we hope!).
+In the context of \kepler, the catalog reduces the set of downloaded pixel
+time series (approximately XXXX for the average \kepler\ target) to
+probabilistic constraints on a handful of physical parameters---we'll call
+them \entry---like the orbital period and radius.
+If we take this set of parameters as \emph{sufficient statistics} of the data
+then we can, in theory, compute \eq{crazylike}---up to an unimportant
+constant---without ever looking at the raw data again.
+This is true because the catalog provides constraints on the posterior
+probability of the parameters $\{\entry_k\}$ under some ``interim prior''
+\interim
+\begin{eqnarray}\eqlabel{crazypost}
+p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
+\frac{p(\{\entry_k\}\,|\,\interim)}{p(\{\data_k\}\,|\,\interim)}\,
+\prod_{k=1}^K p(\data_k\,|\,\entry_k) \quad.
+\end{eqnarray}
+We can then multiply the integrand in \eq{crazylike} by
+\begin{eqnarray}
+\frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
+     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
+\end{eqnarray}
+and use \eq{crazypost} to find
+\begin{eqnarray}
+\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
+    \int
+    \frac{p(\{\entry_k\}\,|\,\ratepars)}{p(\{\entry_k\}\,|\,\interim)}\,
+    p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)
+    \dd\{\entry_k\} \quad.
+\end{eqnarray}
+On the right hand side of this equation, the data come in only through the
+quantity $p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)$ that is provided by the
+catalog!
+
+A very simple example is a procedure that will be very familiar: the histogram.
+If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
+constant rate function---where the bin heights are the parameters---and if the
+uncertainties on the catalog are negligible compared to the bin widths then
+the maximum marginalized solution for \ratepars\ is a histogram of the
+catalog entries (see \app{histogram} for the derivation).
+The case of non-negligible uncertainties is described by \citet{hogge} using a
+method similar to the one discussed below.
+The limiting noiseless case for an incomplete catalog also has an analytic
+maximum likelihood result that we derive in \app{histogram}.
 

### 2014-04-30 document/ms.tex
reworking model generalities and removing redundant sections

@@
     Timothy~D.~Morton\altaffilmark{\princeton},
-    Erik~A.~Petigura\altaffilmark{\berkeley},
     \etal
@@
                           Princeton, NJ, 08540, USA} % CHECK THIS ZIP CODE
-\altaffiltext{\berkeley} {Department of Astronomy,
-                          University of California at Berkeley,
-                          Berkeley, CA, XXXXX, USA} % CHECK THIS ZIP CODE
 
@@
 and \entry\ is a vector of physical parameters describing a planetary system
-(period, radius ratio, \etc).
+(orbital periods, radius ratios, stellar radius, \etc).
+In this equation, we have only made one assumption ...
 
@@
 
-A catalog is a dimensionality reduction of the data that it is based on where
-all the (relevant) information has been retained (we hope!).
+A catalog is a dimensionality reduction of the raw data that with all the
+relevant information retained (we hope!).
 In the context of \kepler, the catalog reduces the set of downloaded pixel
-time series (approximately XXXX for the average \kepler\ target) to
-probabilistic constraints on a handful of physical parameters---we'll call
-them \entry---like the orbital period and radius.
+time series (approximately XXXX data points for the average \kepler\ target)
+to probabilistic constraints on a handful of physical parameters---\entry\
+from above---like the orbital period and radius.
 If we take this set of parameters as \emph{sufficient statistics} of the data
@@
 This is true because the catalog provides constraints on the posterior
-probability of the parameters $\{\entry_k\}$ under some ``interim prior''
-\interim
+probability of the parameters $\{\entry_k\}$ under some choice of ``interim
+prior'' \interim
 \begin{eqnarray}\eqlabel{crazypost}
@@
 \end{eqnarray}
-We can then multiply the integrand in \eq{crazylike} by
+It turns out that we can use this posterior to simplify \eq{crazylike} to a
+form that can, in many common cases, be evaluated efficiently.
+To find this result, multiply the integrand in \eq{crazylike} by
 \begin{eqnarray}
@@
 \end{eqnarray}
-On the right hand side of this equation, the data come in only through the
-quantity $p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)$ that is provided by the
-catalog!
+The data only enter this equation through the posterior constraints provided
+by the catalog!
 
@@
 
-\emph{Note: I think that we should use $T$ instead of $P$ for period because
-$P$ is too heavily overloaded by the probabilities\ldots}
+In the inference procedure described above, we only discussed an abstract
+occurence rate function.
+To make it more concrete, we'll model the catalog as a Poisson draw from the
+\emph{observable} rate function $\obs{\rate}_\ratepars$.
+This leads to the well known result (see \citealt{tabachnik,youdin} for some
+of the examples from the exoplanet literature)
+\begin{eqnarray}\eqlabel{poisson-like}
+p(\{\entry_k\}\,|\,\ratepars) &=&
+    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
+    \prod_{k=1}^K \obs{\rate}_\ratepars (\entry_k)\quad.
+\end{eqnarray}
+The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate of
+exoplanets that you would expect to observe taking into account the geometric
+transit probability and any other detection efficiencies.
+In practice, it probably makes sense to model the observable rate as
+\begin{eqnarray}
+\obs{\rate}_\ratepars(\entry) &=&
+    \completeness(\entry)\,\rate_\ratepars(\entry)
+\end{eqnarray}
+where $\completeness(\entry)$ is the detection efficiency at \entry\ and
+$\rate_\ratepars(\entry)$ is the object that we want to infer: the true
+occurence rate.
+Up to this point, we haven't discussed any specific functional form for
+$\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
+whether we model the occurence rate as (for example) a broken power law or
+a histogram.
 
-From our perspective, a ``model'' is a likelihood function---a probability
-density (or distribution) function (PDF) for the data given
-parameters---and a set of prior PDFs for the parameters.
-In what follows, we will treat as ``data'' a catalog of exoplanet period
-and radius measurements (\citealt{petigura}).
-\emph{(I don't think need this next sentence).}
-We will treat the uncertainties on those measurements as ``prior
-information''; that is, we will treat those as true as given and not as being
-generated by the parameterized likelihood function.
+For the results in this \paper, we will assume that the completeness function
+$\completeness(\entry)$ is known empirically but that is not a requirement for
+the validity of this method.
+Instead, we could use a functional form for it and infer the parameters
+(\dfmtodo{cite all the papers that do the ramp, step function, whatever}).
 
-In the models that follow, the most important parameters will be parameters
-of the PDF for the exoplanet periods and radii.
-However, there will also be ``latent parameters'' that represent the ``true''
-values of the exoplanet periods \period\ and radii \radius.
-That is, for each exoplanet, there is a ``measured'' or ``observed'' period
-\obs{\period}\ and radius \obs{\radius}, which constitute our data.
-These relate to the unobserved ``true'' period and radius by a noise model.
-For the conclusions of this study, the true values of period and radius are
-nuisance parameters; we will marginalize them out.
+\paragraph{Inverse-detection-efficiency}
+It's now interesting to take a brief aside and discuss the connection between
+our model and the commonly used inverse-detection-efficiency procedure
+(\citealt{howard,dressing,petigura}).
+This procedure involves making a weighted histogram of the catalog entries
+where the weight for object $\entry_k$ is $1/\completeness(\entry_k)$.
+This makes intuitive sense but (to our knowledge) does not have a clear
+probabilistic justification or interpretation.
 
-Let's also introduce more succinct notation for a \emph{catalog}.
-We'll refer to the set of period and radius measurements for a set of $K$
-candidate exoplanets as
+If we model the occurence rate as a histogram with $J$ fixed bin volumes
+$\binarea_j$
 \begin{eqnarray}
-\catalog &=& \{\entry_k\}_{k=1}^K
+\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
+\ratepar_1 & \entry \in \binarea_1,\\
+\ratepar_2 & \entry \in \binarea_2,\\
+\cdots \\
+\ratepar_J & \entry \in \binarea_J,\\
+0 & \mathrm{otherwise}
+\end{array}\right.
 \end{eqnarray}
-where each entry $\entry_k = (\period_k,\,\radius_k)$ is a tuple containing a
-period and radius.
-
-The purpose of this \paper\ is to determine constraints on the \emph{rate
-function} of exoplanets
-\begin{eqnarray}\eqlabel{true-rate}
-\rate_\ratepars (\entry)
-&=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}
+then \eq{poisson-like} becomes
+\begin{eqnarray}
+\ln p(\{\entry_k\}\,|\,\ratepars) &=&
+    \sum_{k=1}^K \sum_{j=1}^J \mathbf{1}[\entry_k \in
+        \binarea_j]\,\ln\completeness(\entry_k)\,\ratepar_j
+    -\sum_{j=1}^J\ratepar_j\,\int_{\binarea_j} \completeness(\entry)\dd\entry
 \end{eqnarray}
-given our dataset.
-Here we have not made any assumptions about the functional form of \rate, we
-are only stating that it can be modeled by some---possibly very large---set of
-parameters \ratepars.
-
-In this \paper, we will assume that the completeness function of the survey is
-known and it can be represented as a probability at every (true) period and
-radius $\completeness (\entry)$.
-This probability will also include the geometric transit probability as a
-function of period $\transitprob (\period)$.
-These selection effects can be incorporated into the rate from \eq{true-rate}
-to get the \emph{observable rate}
-\begin{eqnarray}\eqlabel{obs-rate}
-\obs{\rate}_\ratepars (\entry)
-&=& \rate_\ratepars(\entry)\,\completeness(\entry)\quad.
+where the indicator function $\mathbf{1}[\cdot]$ is one if $\cdot$ is true and
+zero otherwise.
+If we take the gradient of this function with respect to \ratepars\ and set it
+equal to zero, we find the maximum likelihood result
+\begin{eqnarray}
+\ratepar_j &=& \frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
 \end{eqnarray}
-
-Given this parameterized rate $\obs{\rate}_\ratepars$, we can compute the
-Poisson likelihood of a \catalog\ (\citealt{tabachnik,youdin,dong})
-\begin{eqnarray}\eqlabel{true-like}
-\ln p(\catalog\,|\,\ratepars) &\propto&
-\sum_{k=1}^K \ln\obs{\rate}_\ratepars (\entry_k)
-- \int \obs{\rate}_\ratepars (\entry) \dd\entry
-\end{eqnarray}
-where we have assumed that every object in \catalog\ is an independent
-Poisson sample from the rate.
-
-If we had perfect noiseless measurements of the periods and radii for every
-object in our catalog, \eq{true-like} provides a metric for determining
-constraints on the parameters of the rate function.
-In practice, the measurements of \catalog\ are noisy so we need to include
-these effects.
-
-\section{Including observational uncertainties}
-
-The likelihood function given in \eq{true-like} is a model of the
-probability of observing a catalog with \emph{true} periods and radii.
-If we want to include the effects of measurement uncertainties, we need to
-take a moment to consider the meaning of measurement uncertainties.
-
-The catalog of planet candidates considered in this \paper\
-(\citealt{petigura}) was generated by drawing samples from a posterior
-probability distribution for physical parameters (period, radius, impact
-parameter, \etc) conditioned on \kepler\ light curves.
-The $k$th entry in this catalog is, therefore, a representation of the
-posterior PDF
-\begin{eqnarray}\eqlabel{measurement}
-p(\entry_k\,|\,\obs{\entry}_k,\,\interim)
-\end{eqnarray}
-using the notation from the previous section.
-In \eq{measurement}, the notations \interim\ indicates that the measurement
-was made with a specific choice of prior over parameters
-$p(\entry_k\,|\,\interim)$.
-We will refer to these priors as \emph{interim priors} because our goal is to
-infer the distribution that is consistent with the data.
-
-More generally, a published catalog represents a \emph{posterior PDF over
-plausible catalogs}
-\begin{eqnarray}\eqlabel{catalog-pdf}
-p(\catalog\,|\,\obs{\catalog},\,\interim) \quad.
-\end{eqnarray}
-It is normally safe to assume that each entry in the catalog is conditionally
-independent given the choices of interim priors.
-\dfmtodo{Why and when is this a good/bad/necessary assumption?}
-Under this assumption, \eq{catalog-pdf} factorizes as
-\begin{eqnarray}\eqlabel{catalog-factor}
-p(\catalog\,|\,\obs{\catalog},\,\interim) &=&
-\prod_{k=1}^K p(\entry_k\,|\,\obs{\entry}_k,\,\interim) \quad.
-\end{eqnarray}
-
-It will be useful in the following discussion to represent the catalog PDF as
-a set of $N$ samples
-\begin{eqnarray}\eqlabel{catalog-samples}
-\catalog^{(n)} &\sim& p(\catalog\,|\,\obs{\catalog},\,\interim) \quad.
-\end{eqnarray}
-For the general catalog, these could be generated using MCMC or a similar
-sampling algorithm.
-Under the conditional independence assumption of \eq{catalog-factor}, however,
-the problem is much simpler.
-Each entry $\entry_k$ can be sampled independently and then the catalog
-samples $\catalog^{(n)}$ are given by the outer product of the independent
-sample lists.
-
-Now that we have a representation of the catalog, we want to infer constraints
-on the occurrence rate of exoplanets conditioned on all the data.
-To perform this inference, we need to compute the  \emph{marginalized
-likelihood}
-\begin{eqnarray}\eqlabel{marg-likelihood}
-p(\obs{\catalog}\,|\,\ratepars) &=&
-\int p(\catalog,\,\obs{\catalog}\,|\,\ratepars) \dd\catalog \quad.
-\end{eqnarray}
-The marginalized likelihood measures the probability of all the \kepler\ light
-curves given a particular setting of the occurrence rate parameters.
-If we can efficiently compute this then we can optimize to find the
-maximum likelihood occurrence rate or we can apply priors $p(\ratepars)$ and
-sample to find posterior constraints on the occurrence rate.
-Note that any results obtained through \eq{marg-likelihood} will
-\emph{correctly account for all the measurement uncertainties}.
-
-The integral in \eq{marg-likelihood} cannot be computed analytically and, in
-fact, it will generally be extremely computationally expensive to evaluate
-numerically (\dfmtodo{CITE}).
-In our case, however, we only need to evaluate this integral up to a constant
-and there is a simple approximation---using the sampling representation of the
-catalog---that can be sufficiently accurate for our purposes.
-To derive this approximation, first recall that the joint distribution inside
-the integral in \eq{marg-likelihood} factorizes as
-\begin{eqnarray}
-p(\catalog,\,\obs{\catalog}\,|\,\ratepars) &=&
-p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog)
-\end{eqnarray}
-because the occurrence rate generates the \emph{true} distribution of periods
-and radii.
-Then, we can multiply the integrand by
-\begin{eqnarray}
-\frac{p(\catalog\,|\,\obs{\catalog},\,\interim)}
-     {p(\catalog\,|\,\obs{\catalog},\,\interim)}
-\end{eqnarray}
-and simplify to find
-\begin{eqnarray}
-\frac{p(\obs{\catalog}\,|\,\ratepars)}
-     {p(\obs{\catalog}\,|\,\interim)}
-&=&
-\int \frac{p(\catalog\,|\,\ratepars)}{p(\catalog\,|\,\interim)}\,
-p(\catalog\,|\,\obs{\catalog},\,\interim)
-\dd\catalog \quad.
-\end{eqnarray}
-Using the samples over catalogs $\catalog^{(n)}$ as defined above, we can
-approximate this integral---up to an irrelevant normalization constant---as
-\begin{eqnarray}\eqlabel{approx-marg-likelihood}
-p(\obs{\catalog}\,|\,\ratepars) &\appropto&
-\sum_{n=1}^N
-\frac{p(\catalog^{(n)}\,|\,\ratepars)}{p(\catalog^{(n)}\,|\,\interim)}
-\end{eqnarray}
-where the numerator inside the sum is given by \eq{true-like} and the
-denominator is the interim prior evaluated at the coordinates of the samples.
-
-This is a generalization of the approximation presented previously in the
-exoplanet literature (\citealt{hogge}) but in this case, the likelihood
-function \eq{true-like} doesn't factorize over entries in the catalog.
-As a result, the variance of \eq{approx-marg-likelihood} is more sensitive to
-the choice of interim prior.
-In practice, we have found that including the completeness function
-$\completeness(\entry)$ in $p(\entry\,|\,\interim)$ greatly improves
-performance.
-This term can be incorporated into existing samplings---generated using less
-informative priors---using rejection sampling.
-
-\section{Parameterization of the occurrence rate}
-
-In the following sections, we will model the occurrence rate
-$\rate(\period,\,\radius)$ as a two dimensional grid with $N$ bins in
-$\ln\period$ and $\ln\radius$.
-We choose to fix the bin locations leaving only the bin heights
-$\binheights = \{\binheight_i\}_{i=1}^{N}$ as free parameters.
-
-\dfmtodo{What about the normalization?}
-
-Since the choice of bin locations and number is arbitrary, the most
-conservative strategy is to choose very small bins---yielding an effectively
-non-parametric model---and choose a prior (or regularization) that encodes any
-prior expectations about the distribution.
-In particular, it seems reasonable to expect the distribution to be
-``smooth''.
-The most extreme version of this model would be a completely non-parametric
-rate function---generated by, for example, a Gaussian process
-(\citealt{poiss-gp})---but this discussion is beyond the scope of this
-\paper.
-Instead, we'll simply choose small bins and add a Gaussian process prior
-(\dfmtodo{CITE}) on the logarithms of the bin heights
-\begin{eqnarray}
-\ln p(\binheights\,|\,\mu,\,\bvec{\smooth}) &=&
--\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}^{-1}\,\bvec{r}
--\frac{1}{2}\ln \det \bvec{K} - \frac{N}{2}\ln 2\pi
-\end{eqnarray}
-where $\bvec{r}$ is the vector of residuals
-\begin{eqnarray}
-\bvec{r} &=& \left(\begin{array}{cccc}
-\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
-\end{array}\right)^\mathrm{T}
-\end{eqnarray}
-and $\bvec{K}$ is the kernel matrix evaluated \emph{at the bin centers}
-$\bincenter_i$
-\begin{eqnarray}
-K_{ij}(\bvec{\smooth}) &\equiv& \smooth_0^2\,\exp \left(
--\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}\,
-\mathrm{diag}(1/\smooth_k^2)\,
-[\bincenter_i-\bincenter_j] \right) \quad.
-\end{eqnarray}
-
-Below, we will explore the effects of this choice of prior but when we
-generate our final results, we will actually apply hyperpriors to the
-parameters $\mu$ and $\bvec{\smooth}$ and marginalize over their values.
+where $N_j$ is the number of objects that fall within the bin $j$.
+Since it is the maximum likelihood solution, this result is the minimum
+variance estimator for the occurence rate in the limit of negligible catalog
+uncertainties.
+This is \emph{almost} the same as the inverse-detection-efficiency result and
+in various limiting cases, they would provide the same result.
+\dfmtodo{Finish explaining why this result is better under the specified
+assumptions.}
 
@@
 distribution to larger radii.
-To do this, we approximate the survey completeness for $\radius > 16\,R\oplus$
+To do this, we approximate the survey completeness for $\radius>16\,R_\oplus$
 as 1.
-
-\dfmtodo{Hopefully Erik will send us MCMC samples. If not, describe the
-re-sampling procedure that we used. Either way, we'll need to re-sample under
-the completeness probability for the purposes of the importance sampling.}
 
@@
 It is a pleasure to thank
-    \ldots
+    Erik Petigura (Berkeley),
+    \etc
 for helpful contributions to the ideas and code presented here.

### 2014-05-02 document/ms.tex
some discussion of simulations

@@
 
-To confirm that the probabilistic framework developed in the previous sections
-provides correct inferences about the population of exoplanets given noisy
-data, we'll start by generating synthetic catalogs and running our full
-inference pipeline on those data.
-In this section, we also demonstrate the key benefits of our method over the
-standard intuitive inverse-detection-efficiency procedure.
-
-\paragraph{Inverse-detection-efficiency}
-Authors in the exoplanet literature (for example, \citealt{howard, dressing,
-petigura}) have built population level models using this procedure.
-The basic idea is to make a histogram of planet candidates in the model where
-each sample has a weight proportional to the inverse detection efficiency.
-This method makes intuitive sense and, in the limit of infinitely precise
-measurements, it is not wrong.
-Even in this limit, this estimator of the occurrence distributions will have
-large variance; especially as the detection efficiency becomes low.
-In the presence of observational uncertainties, the inferred distributions
-will be biased.
-
-The likelihood function in \eq{true-like} provides a \emph{probability
-distribution over catalogs} conditioned on the parameters of the rate
-function.
-Therefore, it is also the description of a \emph{generative process} for
-synthetic catalogs.
-To demonstrate the effectiveness of our model, we generate synthetic catalogs
-from known distributions in period and radius, and test the performance of our
-method on the resulting data.
-For comparison, we also implement the one-over-efficiency weighting procedure
-(see, for example, \citealt{howard, dressing, petigura}) and apply it to the
-same datasets.
-
-\paragraph{The generative process}
-The variable rate Poisson likelihood function in \eq{true-like} can be sampled
-generally for arbitrary forms of the rate function using a procedure called
-``thinning'' (\citealt{poisson}).
-In this \paper, however, we will only consider binned representations of
-\rate, greatly simplifying the generative process.
-Conditioned on a set of rate function parameters and a given completeness
-function (here we'll use the injections from \citealt{petigura}), we compute
-the observable rate grid from \eq{obs-rate}.
-Then in each bin, we sample $J$ candidates uniformly in period and
-radius (as defined by the bin edges) where $J$ is a Poisson random variable
-sampled with the rate given by the observable rate in this bin.
-\dfmtodo{Make this description not be the most confusing thing ever.}
-Then, we add 30\% measurement uncertainties to the radius measurements where
-30\% was chosen to match to properties of the real catalog
-(\citealt{petigura}).
+In order to get a feeling for the constraints provided by our method and to
+explore any biases introduced by ignoring the observational uncertainties, we
+start by ``observing'' a few synthetic catalogs from qualitatively different
+known occurence rate functions.
+For each of these simulations, we'll take the completeness function computed
+by \citet{petigura} as given.
+In general, \eq{poisson-like} can be sampled using a procedure called thinning
+(\citealt{poisson}) but for our purposes, we'll simply consider a piecewise
+constant rate function evaluated on a fine grid in (log-) period and radius.
+For this discrete function, the generative procedure is simply:
+\begin{enumerate}
+{\item loop over each grid cell $i$,}
+{\item draw Poisson random integer $K_i\sim\mathrm{Poissson}(\obs{\rate}_i)$
+with the observable rate evaluated in the cell, and}
+{\item distribute $K_i$ catalog entries in the cell randomly.}
+\end{enumerate}
+We then choose fractional observational uncertainties on the radii from the
+\citet{petigura} catalog and apply them to the catalog assuming that they are
+Gaussian.
 

### 2014-05-04 document/ms.tex
don't forget about the priors
a few words about assumptions

@@
 p(\{\data_k\}\,|\,\ratepars) &=&
-    \int\left[\prod_{k=1}^K p(\data_k\,|\,\entry_k)\right ]
+    \int p(\{\data_k\}\,|\,\{\entry_k\})
     \,p(\{\entry_k\}\,|\,\ratepars)
@@
 (orbital periods, radius ratios, stellar radius, \etc).
-In this equation, we have only made one assumption ...
+In this equation, our only assumption is that the datasets depend on the
+occurence rate of exoplanets only through the catalog.
+In our case, this assumption qualitatively means that the signals found in the
+light curves depend only on the actual planet properties and not on their
+distributions.
 
-It is generally computationally intractable to compute the likelihood in
-\eq{crazylike} because it involves a high dimensional numerical integration of
-a function where each evaluation of the integrand touches \emph{every data
-point}.
-Instead, we can use existing catalogs of planet candidates to approximate the
-marginalization integral.
-For our purposes, this is the definition of \emph{hierarchical inference}.
-
-A catalog is a dimensionality reduction of the raw data that with all the
-relevant information retained (we hope!).
+In other words, a catalog is a dimensionality reduction of the raw data with
+all the relevant information retained.
 In the context of \kepler, the catalog reduces the set of downloaded pixel
@@
 p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
-\frac{p(\{\entry_k\}\,|\,\interim)}{p(\{\data_k\}\,|\,\interim)}\,
-\prod_{k=1}^K p(\data_k\,|\,\entry_k) \quad.
+\frac{p(\{\data_k\}\,|\,\{\entry_k\})\,p(\{\entry_k\}\,|\,\interim)}
+     {p(\{\data_k\}\,|\,\interim)} \quad.
 \end{eqnarray}
@@
 uncertainties on the catalog are negligible compared to the bin widths then
-the maximum marginalized solution for \ratepars\ is a histogram of the
-catalog entries (see \app{histogram} for the derivation).
+the maximum marginalized likelihood solution for \ratepars\ is a histogram of
+the catalog entries (see \app{histogram} for the derivation).
 The case of non-negligible uncertainties is described by \citet{hogge} using a
 method similar to the one discussed below.
-The limiting noiseless case for an incomplete catalog also has an analytic
-maximum likelihood result that we derive in \app{histogram}.
 
@@
 assumptions.}
+
+\paragraph{Occurence rate parameterization \& priors}
+\dfmtodo{describe the Gaussian process prior on the bin heights.}
 

### 2014-05-06 document/ms.tex
a few notes

@@
 
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[one-over-efficiency]{%
-\includegraphics[width=0.45\textwidth]{figures/synthetic-vmax-radius.pdf}
-}
-\subfigure[fully-marginalized forward model]{%
-\includegraphics[width=0.45\textwidth]{figures/synthetic-radius.pdf}
-}
-\end{center}
-\caption{%
-The inferred distribution of exoplanet radii based on a noisy observation of a
-synthetic catalog.
-The red dotted line shows the true distribution function and the black
-histograms are the constraints from the different inference techniques.
-\figlabel{synthetic-radius}}
-\end{figure}
+% \begin{figure}[htbp]
+% \begin{center}
+% \subfigure[one-over-efficiency]{%
+% \includegraphics[width=0.45\textwidth]{figures/synthetic-vmax-radius.pdf}
+% }
+% \subfigure[fully-marginalized forward model]{%
+% \includegraphics[width=0.45\textwidth]{figures/synthetic-radius.pdf}
+% }
+% \end{center}
+% \caption{%
+% The inferred distribution of exoplanet radii based on a noisy observation of a
+% synthetic catalog.
+% The red dotted line shows the true distribution function and the black
+% histograms are the constraints from the different inference techniques.
+% \figlabel{synthetic-radius}}
+% \end{figure}
+
+\section{Extrapolation to Earth}
 
 \section{Real data}
-
-It has been shown that the joint period--radius distribution for exoplanets is
-not separable (\dfmtodo{CITE}).
-In this section, we discuss non-separable model in our analysis framework.
-
-We model the rate function $\rate(\period,\,\radius)$ as a piecewise constant
-function in rectangular bins in $\ln\period$ and $\ln\radius$.
-The parameters \ratepars\ that we're fitting are the values of each of these
-bins (subject to a normalization constraint).
 

### 2014-05-06 document/ms.tex
actually mention the importance sampling approximation

@@
 and use \eq{crazypost} to find
-\begin{eqnarray}
+\begin{eqnarray}\eqlabel{simplemarglike}
 \frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
@@
 
+The constraints in \eq{crazypost} can always be---and often are---propagated
+as a list of $N$ samples $\{\entry_k^{(n)}\}$ from the posterior
+\begin{eqnarray}
+\entry_k^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
+\end{eqnarray}
+We can use these samples to approximately compute the integral in
+\eq{simplemarglike} as
+\begin{eqnarray}
+\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
+    \frac{1}{N} \sum_{n=1}^N
+    \frac{p(\{\entry_k^{(n)}\}\,|\,\ratepars)}
+         {p(\{\entry_k^{(n)}\}\,|\,\interim)} \quad.
+\end{eqnarray}
+This is very efficient to compute as long as an evaluation of
+$p(\{\entry_k^{(n)}\}\,|\,\ratepars)$ is not very expensive.
+That being said, this could be a high variance estimator depending on the
+initial choice of $p(\{\entry_k^{(n)}\}\,|\,\interim)$ but in our case, we'll
+show that not many samples are needed to get an accurate estimate.
+
 A very simple example is a procedure that will be very familiar: the histogram.
@@
 The case of non-negligible uncertainties is described by \citet{hogge} using a
-method similar to the one discussed below.
+method similar to the one discussed here.
 
@@
 \paragraph{Occurence rate parameterization \& priors}
-\dfmtodo{describe the Gaussian process prior on the bin heights.}
+In this paper, we will also model
 

### 2014-05-07 document/ms.tex
comments from @davidwhogg

@@
 
-\newcommand{\paper}{\emph{Article}}
+\newcommand{\paper}{\textsl{Article}}
 
@@
 \end{eqnarray}
-where $\{\data_k\}$ is the set of all light curves, \ratepars\ is a vector of
-parameters describing the occurence rate function $\rate_\ratepars(\entry)$
-and \entry\ is a vector of physical parameters describing a planetary system
-(orbital periods, radius ratios, stellar radius, \etc).
+where $\{\data_k\}$ is the set of all light curves, one light curve $\data_k$
+per target $k$, \ratepars\ is a vector of parameters describing the occurence
+rate function $\rate_\ratepars(\entry)$ and $\entry_k$ is the vector of
+physical parameters describing the planetary system (orbital periods,
+radius ratios, stellar radius, \etc) around target $k$.
 In this equation, our only assumption is that the datasets depend on the
-occurence rate of exoplanets only through the catalog.
+occurence rate of exoplanets only through the catalog $\{\entry_k\}$.
 In our case, this assumption qualitatively means that the signals found in the
-light curves depend only on the actual planet properties and not on their
-distributions.
+light curves depend only on the actual planet properties and not on the
+distributions from which they are drawn.
+It is worth noting that---as we will discuss further below---the catalog only
+provides \emph{probabilistic} constraints on $\{\entry_k\}$.
 
@@
 In the context of \kepler, the catalog reduces the set of downloaded pixel
-time series (approximately XXXX data points for the average \kepler\ target)
+time series (approximately 70,000 data points for the typical \kepler\ target)
 to probabilistic constraints on a handful of physical parameters---\entry\
-from above---like the orbital period and radius.
-If we take this set of parameters as \emph{sufficient statistics} of the data
-then we can, in theory, compute \eq{crazylike}---up to an unimportant
-constant---without ever looking at the raw data again.
+from above---like the orbital period and planetary radius.
+If we take this set of parameters $\{\entry_k\}$ as \emph{sufficient
+statistics} of the data then we can, in theory, compute \eq{crazylike}---up to
+an unimportant constant---without ever looking at the raw data again.
 This is true because the catalog provides constraints on the posterior
@@
 The data only enter this equation through the posterior constraints provided
-by the catalog!
+by the catalog $\{\entry_k\}$!
+For our purposes, this is the \emph{definition} of hierarchical inference.
 
@@
 
-A very simple example is a procedure that will be very familiar: the histogram.
+A very simple example is the familiar procedure of making a histogram.
 If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
@@
 \end{eqnarray}
+In this equation, the integral in the normalization term is the expected
+number of observable exoplanets in the sample.
+A further assumption that we have made in \eq{poisson-like} is that each
+object is an independent sample from the rate function.
+This assumption---though it is often made---cannot be correct when we consider
+systems of multiple planets.
+It is possible to relax this assumption (see \citealt{tremaine} for a related
+technique) but this generalization is beyond the scope of the current \paper.
+
 The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate of
@@
 \paragraph{Occurence rate parameterization \& priors}
-In this paper, we will also model
+For the remainder of this \paper, we will also model the occurence rate as a
+histogram in logarithmic bins of period $\ln\period$ and radius $\ln\radius$.
 
-\section{The completeness function \& dataset}
+\section{Data and completeness function}
 

### 2014-05-08 document/ms.tex
notes about implicit priors

@@
 
-Erik Petigura and collaborators (\citealt{petigura-a,petigura}) heroically
-determined the empirical end-to-end completeness of their transit search
-pipeline by injecting synthetic signals of known period and radius into the
-raw aperture photometry and testing recovery rates.
-They also published a catalog of XXX small planet candidates around G type
-stars; YYY of them previously unknown.
-Given the unprecedented quality of the completeness characterization, we will
-focus on this dataset in this \paper.
+Using an independent exoplanet search and characterization pipeline,
+\citet{petigura} published a catalog of 603 planet candidates orbiting stars
+in their ``Sun-like'' sample of \kepler\ targets.
+For each candidate, \citet{petigura} used Markov chain Monte Carlo to sample
+the posterior probability density---presumably assuming uniform priors---for
+the radius ratio, transit duration, and impact parameter.
+They then incorporated the uncertainties in the stellar radius and published
+constraints on the physical radii of their candidates.
+The exact shape of the implied prior on planet radius depends on the assumed
+priors on radius ratio and stellar radius but in the limit of a broad uniform
+prior on each quantity, the implicit prior at small physical radius is
+approximately uniform; this is the assumption that we make.
+
+A huge benefit of this dataset is that Erik Petigura and collaborators
+published a rigorous analysis the empirical end-to-end completeness of their
+transit search pipeline.
+Instead of choosing a functional form for the detection efficiency of the
+pipeline as a function of the parameters of interest, \citet{petigura}
+injected synthetic signals of known period and radius into the raw aperture
+photometry and determining the empirical recovery rate after the full analysis.
 
@@
 In each bin, this efficiency is simply the fraction of recovered injections.
-For the purposes of this \paper, we will neglect the uncertainties introduced
-by the fact that this estimate is being made with a finite number of samples.
+For the purposes of this \paper, we neglect the counting uncertainties
+introduced by the finite number of samples.
 
-The largest injected radius is $16\,R_\oplus$ but, because of the measurement
-uncertainties on the radii, it will be necessary for us to model the
-distribution to larger radii.
+The largest injected signal had a radius of $16\,R_\oplus$ but, because of the
+measurement uncertainties on the radii, we need to model the distribution at
+larger radii.
 To do this, we approximate the survey completeness for $\radius>16\,R_\oplus$
 as 1.
+
+Given our domain knowledge of how detection efficiency depends on the physical
+parameters, it might seem silly to measure the survey completeness in period
+and radius instead of radius ratio or signal-to-noise.
+It is also likely that a change of coordinates would yield a higher precision
+result.
+That being said, it is still correct to measure the completeness in period and
+radius, and there are a few practical reasons for our choice.
+First, since the radius uncertainties are dominated by uncertainties in the
+stellar parameters, it is not possible to use the published catalog
+(\citealt{petigura}) to compute constraints on radius ratios.
+\dfmtodo{finish this}
 

### 2014-05-08 document/ms.tex
what can future catalogs produce?

@@
 radius, and there are a few practical reasons for our choice.
-First, since the radius uncertainties are dominated by uncertainties in the
-stellar parameters, it is not possible to use the published catalog
-(\citealt{petigura}) to compute constraints on radius ratios.
-\dfmtodo{finish this}
+The main argument is that since the radius uncertainties are dominated by
+uncertainties in the stellar parameters, it is not possible to use the
+published catalog (\citealt{petigura}) to compute constraints on radius
+ratios.
+In future catalogs, this problem would be solved by publishing representations
+of the posterior probabilities that are being measured.
+In this case, these parameters are radius ratio and stellar radius instead of
+just the physical radius.
 

### 2014-05-10 document/ms.tex
sweeping things under the rug

@@
 \eq{simplemarglike} as
-\begin{eqnarray}
+\begin{eqnarray}\eqlabel{importance}
 \frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
@@
 constraints on the physical radii of their candidates.
-The exact shape of the implied prior on planet radius depends on the assumed
-priors on radius ratio and stellar radius but in the limit of a broad uniform
-prior on each quantity, the implicit prior at small physical radius is
-approximately uniform; this is the assumption that we make.
+Given this data reduction and since we don't have access to the individual
+posterior constraints on radius ratio and stellar radius, we can't directly
+compute the importance weights $p(\{\entry_k^\}\,|\,\interim)$ needed for
+\eq{importance}.
+For the rest of this \paper, we'll make the simplifying assumption that these
+weights are constant in log-period and log-radius but the results don't seem
+to be sensitive to this specific choice.
 
@@
 In future catalogs, this problem would be solved by publishing representations
-of the posterior probabilities that are being measured.
-In this case, these parameters are radius ratio and stellar radius instead of
-just the physical radius.
+of all the posterior probabilities that are being measured.
+In this case, the most useful data product would be \emph{posterior samples
+for each target's radius ratio and stellar radius}.
 

### 2014-05-12 document/ms.tex
descriptions of synthetic catalogs

@@
 
+\dfmtodo{Talk about how we apply the geometric transit probability.}
+
 \section{Validation using synthetic catalogs}
@@
 (\citealt{poisson}) but for our purposes, we'll simply consider a piecewise
-constant rate function evaluated on a fine grid in (log-) period and radius.
-For this discrete function, the generative procedure is simply:
+constant rate function evaluated on a fine grid in log-period and log-radius.
+For this discrete function, the generative procedure is simple;
 \begin{enumerate}
@@
 We then choose fractional observational uncertainties on the radii from the
-\citet{petigura} catalog and apply them to the catalog assuming that they are
-Gaussian.
+\citet{petigura} catalog and apply them to the true catalog as Gaussian noise.
+
+We generate synthetic catalogs from two qualitatively different rate
+functions.
+Both distributions are generated by a separable model
+\begin{eqnarray}
+\rate_\ratepars (\ln\period,\,\ln\radius) &=&
+    \rate_\ratepars^{(\period)}(\ln\period)\,
+    \rate_\ratepars^{(\radius)}(\ln\radius)
+\end{eqnarray}
+but fit using the full general model.
+The first model has a smooth occurence surface where both distributions are
+smooth broken power laws.
+The second is designed to be exactly the distribution inferred by
+\citet{petigura} in the range that they considered and then smoothly
+extrapolated outside that range.
+
+For each catalog, we directly applied the inverse-detection-efficiency
+procedure as implemented by \citealt{petigura}\footnote{our implementation
+exactly reproduces their results when applied to the published catalog} and
+our probabilistic method.
 

### 2014-05-12 document/ms.tex
figures and captions

@@
 \newcommand{\radius}{\ensuremath{R}}
-\newcommand{\period}{\ensuremath{T}}
+\newcommand{\period}{\ensuremath{P}}
 \newcommand{\completeness}{{\ensuremath{P_\mathrm{c}}}}
@@
 \newcommand{\smooth}{{\ensuremath{\lambda}}}
+
+\newcommand{\modela}{\emph{Model A}}
+\newcommand{\modelb}{\emph{Model B}}
+
+\newcommand{\gammaearth}{{\ensuremath{\rate_\oplus}}}
 
@@
 but fit using the full general model.
-The first model has a smooth occurence surface where both distributions are
-smooth broken power laws.
-The second is designed to be exactly the distribution inferred by
+The first model---\modela---has a smooth occurence surface where both
+distributions are smooth broken power laws.
+The second---\modelb---is designed to be exactly the distribution inferred by
 \citet{petigura} in the range that they considered and then smoothly
@@
 
-For each catalog, we directly applied the inverse-detection-efficiency
-procedure as implemented by \citealt{petigura}\footnote{our implementation
-exactly reproduces their results when applied to the published catalog} and
+For each catalog, we directly apply the inverse-detection-efficiency
+procedure as implemented by \citealt{petigura}\footnote{Our implementation
+exactly reproduces their results when applied to the published catalog.} and
 our probabilistic method.
@@
 
-% \begin{figure}[htbp]
-% \begin{center}
-% \subfigure[one-over-efficiency]{%
-% \includegraphics[width=0.45\textwidth]{figures/synthetic-vmax-radius.pdf}
-% }
-% \subfigure[fully-marginalized forward model]{%
-% \includegraphics[width=0.45\textwidth]{figures/synthetic-radius.pdf}
-% }
-% \end{center}
-% \caption{%
-% The inferred distribution of exoplanet radii based on a noisy observation of a
-% synthetic catalog.
-% The red dotted line shows the true distribution function and the black
-% histograms are the constraints from the different inference techniques.
-% \figlabel{synthetic-radius}}
-% \end{figure}
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
+\end{center}
+\caption{%
+Inferences about the occurence rate function based on the simulated catalog
+\modela.
+\emph{Center:} the points with error bars show the exoplanet candidates in the
+simulated incomplete catalog, the contours show the survey completeness
+function (\citealt{petigura}), and the grayscale shows the median posterior
+occurence surface.
+\emph{Top and left:} the red dashed line shows the \emph{true} distribution
+that was used to generate the catalog, the points with error bars show the
+results of the inverse-detection-efficiency procedure, and the histograms are
+posterior samples from the marginalized occurence rate as inferred by our
+method.
+\figlabel{smooth-results}}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
+\end{center}
+\caption{%
+The same as \fig{smooth-results} for \modelb.
+\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
+\figlabel{simulation-results}}
+\end{figure}
 
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figures/smooth/rate.pdf}
+\end{center}
+\caption{%
+The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
+different techniques.
+Applying the method used by \citet{petigura} gives a constraint indicated by
+the vertical black line with error bars shown as dashed lines.
+The histogram is the MCMC estimate of our posterior constraint on this rate
+and the true value is indicated as the thick gray vertical line.
+\figlabel{smooth-rate}}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figures/simulation/rate.pdf}
+\end{center}
+\caption{%
+The same as \fig{smooth-rate} for \modelb.
+\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
+\figlabel{simulation-rate}}
+\end{figure}
+
 \section{Real data}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/results/results.pdf}
+\end{center}
+\caption{%
+The same as \fig{smooth-results} when applied to the observed data from
+\citet{petigura}.
+\figlabel{real-results}}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figures/results/rate.pdf}
+\end{center}
+\caption{%
+The same as \fig{smooth-rate} but applied to the catalog from \citet{petigura}.
+\figlabel{real-rate}}
+\end{figure}
 

### 2014-05-12 document/ms.tex
summary of simulated results

@@
 extrapolated outside that range.
+The catalogs generated from these two models are shown in \fig{smooth-results}
+and \fig{simulation-results}, respectively and the data are available online
+at \dfmtodo{ADD SOME URL}.
 
@@
 exactly reproduces their results when applied to the published catalog.} and
-our probabilistic method.
+our probabilistic method, marginalizing over the hyperparameters of the
+Gaussian process regularization.
+\Fig{smooth-results} and \fig{simulation-results} show the results of this
+analysis in both cases.
+In particular, the side panels compare the marginalized occurence rate
+functions in period and radius to the true functions that were used to
+generate the catalogs.
+\Fig{smooth-results} shows that even if the true occurence rate is a smooth
+function, the rate inferred by the inverse-detection-efficiency method can
+appear to have sharp features.
+In this first example---where the true distribution is well described by our
+Gaussian process model---the probabilistic inference of the occurence rate is
+both more precise and accurate.
 
-\paragraph{Results}
-See \fig{synthetic-radius}.
-As expected, the forward model is a much lower variance estimator.
+In the second example, the true rate function includes a sharp feature chosen
+to reproduce the result published by \citet{petigura}.
+In this case, \fig{simulation-results} shows that the probabilistic
+constraints on the occurence rate are less precise \emph{but more accurate}
+than results using the inverse-detection-efficiency method.
+At this point, the distributions inferred by these two different techniques
+don't seem drastically inconsistent but the difference are most obvious as the
+survey completeness goes to zero: exactly the region of interest.
 

### 2014-05-12 document/ms.tex
starting to describe the extrapolation technique

@@
 
+As well as inferring the ocurrence distribution of exoplanets, this dataset
+can also be used to constrain the rate of Earth analogs.
+Even though no true analog has been observed, we can make some assumptions
+about the distribution (as described below) and extrapolate to the location of
+the Earth.
+Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
+location of Earth}
+\begin{eqnarray}
+\gammaearth &=& \rate (\ln\period_\oplus,\,\ln\radius_\oplus) \\
+&=&
+\left.\frac{\dd N}{\dd\ln\period\,\dd\ln\radius}\right|
+_{\radius=\radius_\oplus,\,\period=\period_\oplus}\quad.
+\end{eqnarray}
+
 \begin{figure}[htbp]

### 2014-05-12 document/code/general.py
plotting style changes

@@ -1,112 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-from IPython.parallel import Client
-c = Client()
-view = c[:]
-
-with view.sync_imports():
-    import os
-    import sys
-    import time
-    import emcee
-    import numpy as np
-    import cPickle as pickle
-
-# Monkey patch the path.
-f = __file__
-d = os.path.dirname
-p = d(d(d(os.path.abspath(f))))
-sys.path.insert(0, p)
-view.execute("if \"{0}\" not in sys.path: sys.path.insert(0, \"{0}\")"
-             .format(p), block=True)
-
-with view.sync_imports():
-    from population import (CensoringFunction, Population, SmoothPopulation,
-                            NormalizedPopulation, ProbabilisticModel)
-
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates, load_petigura_bins)
-
-try:
-    os.makedirs("general")
-except os.error:
-    pass
-
-# Define the box that we're going to work in.
-per_rng = np.log([5.0, 400.0])
-rp_rng = np.log([0.5, 64.0])
-
-# Load the data.
-ep = load_petigura_bins()
-ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
-                                                     rp_rng=rp_rng)
-tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
-
-# Set up the censoring function.
-censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                           bins=(32, 32),
-                           range=[per_rng, rp_rng],
-                           transit_lnprob_function=tlp)
-
-# Load the dataset and resample under the censoring function.
-dataset = load_candidates(censor)  # , 120)
-
-# Define the population.
-lpb, lrb = censor.bins
-pop = Population([lpb[::8], lrb[::4]], censor.bins)
-# pop = SmoothPopulation([0.5, -3.0, -3.0], pop)
-pop = NormalizedPopulation(11., pop)
-
-# Define the probabilistic model.
-model = ProbabilisticModel(dataset, pop, censor)
-print("{0} parameters".format(len(pop.initial())))
[…]

### 2014-05-13 document/ms.tex
updated figures without NaN errorbars
notes about priors

@@
 \begin{eqnarray}
-\entry_k^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
+\{\entry_k\}^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
 \end{eqnarray}
@@
 
+Now, we can substitute \eq{poisson-like} into \eq{simplemarglike} and apply
+the importance sampling approximation to derive the following expression for
+the marginalized likelihood
+\begin{eqnarray}\eqlabel{money}
+\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
+    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
+    \prod_{k=1}^K
+    \frac{1}{N_k} \sum_{n=1}^{N_k}
+    \frac{\obs{\rate}_\ratepars (\entry_k^{(n)})}
+         {p(\entry_k^{(n)}\,|\,\interim)} \quad.
+\end{eqnarray}
+In this equation, we're making the further assumption that the catalog treated
+the objects independently.
+In other words, we have per-object posterior samples
+\begin{eqnarray}
+\entry_k^{(n)} &\sim& p(\entry_k\,|\,\data_k,\,\interim) \quad.
+\end{eqnarray}
+\Eq{money} is the \emph{money equation} for our method.
+It lets us efficiently compute the \emph{marginalized likelihood of the entire
+set of light curves for a particular occurence rate function}.
+
 \paragraph{Inverse-detection-efficiency}
@@
 \paragraph{Occurence rate parameterization \& priors}
-For the remainder of this \paper, we will also model the occurence rate as a
-histogram in logarithmic bins of period $\ln\period$ and radius $\ln\radius$.
+For the remainder of this \paper, we also model the occurence rate as a
+two-dimensional histogram with fixed logarithmic bins in period and radius.
+When we include observational uncertainties---using \eq{money}---the maximum
+likelihood result is no longer analytic.
+Therefore, if we want to compute the ``best-fit'' occurence rate, we can use a
+standard non-linear optimization algorithm.
 
@@
 \end{eqnarray}
+This is also the quantity that \citet{petigura} estimated in different units.
 
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{figures/smooth/rate.pdf}
+\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{figures/simulation/rate.pdf}
+\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{figures/results/rate.pdf}
+\includegraphics[width=0.6\textwidth]{figures/results/rate.pdf}
 \end{center}

### 2014-05-13 document/ms.tex
words about ess

@@
 
+\newcommand{\mean}{{\ensuremath{\mu}}}
 \newcommand{\smooth}{{\ensuremath{\lambda}}}
+\newcommand{\smoothpars}{{\ensuremath{\bvec{\smooth}}}}
+\newcommand{\cov}{{\ensuremath{\mathrm{K}}}}
 
@@
 standard non-linear optimization algorithm.
+In the regions of parameter space that we tend to care about, the completeness
+is low and there are only a few observations with large uncertainties.
+In this case, we're especially interested in probabilistic constraints on the
+occurence rate; not just the best-fit model.
+To do this, we apply a prior\footnote{This is a slight misnomer because
+we also we also marginalize over the parameters of this model.} $p(\ratepars)$
+on the occurence rate parameters and generate samples from the posterior
+probability
+\begin{eqnarray}\eqlabel{posterior}
+p(\ratepars\,|\,\{\data_k\}) &\propto&
+    p(\ratepars)\,p(\{\data_k\}\,|\,\ratepars)
+\end{eqnarray}
+using Markov chain Monte Carlo (MCMC).
+
+Given our model of the rate function as a collection of bins, we choose to
+model the prior as a Gaussian process (\citealt{gp,gibson-gp}) on the
+logarithmic bin heights.
+This model encodes our prior belief that, on the grid scale that we consider,
+the occurence rate should be smooth but it is otherwise very flexible about
+the form of the function.
+Mathematically, the prior is written as
+\begin{eqnarray}
+p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \\
+&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{I},\,
+\cov(\{\binarea_j\},\,\smoothpars)\right]
+\end{eqnarray}
+where $\mathcal{N}(\cdot;\,\mean\,\bvec{I},\,\cov)$ is a $J$-dimensional
+Gaussian with a constant mean \mean\ and covariance matrix \cov\ that depends
+on the bin locations $\{\binarea_j\}$ and a set of hyperparameters
+$\smoothpars = (\smooth_0,\,\smooth_\period,\,\smooth_\radius)$.
+The covariance function that we use is an anisotropic, axis-aligned
+exponential-squared kernel so elements of the matrix are
+\begin{eqnarray}
+\cov_{ij} &=& \smooth_0\,\exp\left(-\frac{1}{2}\,
+    [\binarea_i-\binarea_j]^\mathrm{T}\,\Sigma^{-1}\,[\binarea_i-\binarea_j]
+\right)
+\end{eqnarray}
+where $\Sigma^{-1}$ is the diagonal matrix
+\begin{eqnarray}
+\Sigma^{-1} &=& \left(\begin{array}{cc}
+1/\smooth_\period & 0 \\
+0 & 1/\smooth_\radius
+\end{array}\right) \quad.
+\end{eqnarray}
+
+As well as being a very flexible prior, this Gaussian process model is also
+appealing because it is very well studied in the machine learning literature.
+In particular, there is a very efficient algorithm---elliptical slice sampling
+(ESS; \citealt{ess})---for generating samples from the posterior probability
+function in \eq{posterior}.
+As mentioned above, we actually go one step further and assume uniform
+(hyper-)priors on $\ln\mean$ and $\ln\smoothpars$ in order to sample from the
+joint posterior
+\begin{eqnarray}
+p(\ratepars,\,\mean,\,\smoothpars\,|\,\{\data_k\})
+\end{eqnarray}
+using Algorithm 1 from \citet{ess-hyper}.
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[p]
 \begin{center}
@@
 
+\bibitem[Gibson \etal(2012)]{gibson-gp}
+Gibson, N.~P., Aigrain, S., Roberts, S., \etal\ 2012, \mnras, 419, 2683
+
 \bibitem[Hogg \etal(2013)]{hoggwhitepaper}
@@
 
+\bibitem[Murray \etal(2010)]{ess}
+Murray, I., Prescott Adams, R., \& MacKay, D.~J.~C.\ 2010, JMLR: W\&CP, 9, 541
+(\arxiv{1001.0175})
+
+\bibitem[Murray \& Prescott Adams(2010)]{ess-hyper}
+Murray, I., \& Prescott Adams, R.\ 2010, Advances in Neural Information
+Processing Systems, 23, 1723
+(\arxiv{1006.0868})
+
 \bibitem[Petigura \etal(2013a)]{petigura-a}
@@
 Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})
+
+\bibitem[Rasmussen \& Williams(2006)]{gp}
+Rasmussen, C.~E. \& Williams, C.~K.~I.\ 2006
+Gaussian Processes for Machine Learning, MIT Press
+(\href{http://www.gaussianprocess.org/gpml/}{online})
 

### 2014-05-14 document/ms.tex
chain stats
some words about transit probability

@@
 
+For all the results below, we run a Markov chain with $10^6$ steps for the bin
+heights and update the hyperparameters every 10 steps.
+We discard the first \dfmtodo{XXXX} samples as burn-in and then use the
+remaining samples as an approximation to the marginalized probability
+distribution for \ratepars.
+We tune the Metropolis--Hastings proposal used to update the hyperparameters
+by hand to get an acceptance fraction of $\sim0.2-0.4$ for those parameters.
+
 \section{Data and completeness function}
@@
 
-\dfmtodo{Talk about how we apply the geometric transit probability.}
+The detection efficiency also depends on the geometric transit probability
+$R_\star/a$.
+Since we're modeling the distribution in the period--radius plane, we need to
+compute the transit probability marginalized over stellar radius and mass.
+This marginalized distribution scales only with the period of the orbit as
+\dfmtodo{$\propto \period^{3/2}$}.
+In theory, this marginalization should be over the true distribution of these
+parameters in the selected stellar catalog but we'll approximate it by
+the empirical distribution.
 

### 2014-05-14 document/code/results.py
hacking

@@ -68,9 +68,14 @@ else:
 
 # Load the extrapolated value.
 factor = (np.log(400) - np.log(200)) * (np.log(2) - np.log(1))
-ext = np.array(open(os.path.join(bp, "extrap.txt"), "r").read().split(),
-               dtype=float) / 42557.0
+if os.path.exists(fn):
+    ext = np.array(open(os.path.join(bp, "extrap.txt"), "r").read().split(),
+                   dtype=float) / 42557.0
+else:
+    ext = np.array([0.057, -0.022, 0.017])
+
 ext /= factor
+print(ext)
 
 # Compute and plot gamma_earth.
 rates = pop.get_lnrate(samples, [np.log(365.), np.log(1.0)])

### 2014-05-14 document/ms.tex
updating figures and a few more comments about transit geom

@@
 \newcommand{\period}{\ensuremath{P}}
-\newcommand{\completeness}{{\ensuremath{P_\mathrm{c}}}}
-\newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
+\newcommand{\completeness}{{\ensuremath{Q_\mathrm{c}}}}
+\newcommand{\transitprob}{{\ensuremath{Q_\mathrm{t}}}}
 \newcommand{\data}{{\ensuremath{\bvec{x}}}}
@@
 This marginalized distribution scales only with the period of the orbit as
-\dfmtodo{$\propto \period^{3/2}$}.
+$\propto \period^{-2/3}$.
 In theory, this marginalization should be over the true distribution of these
 parameters in the selected stellar catalog but we'll approximate it by
-the empirical distribution.
+the empirical distribution; a reasonable approximation given the size of the
+dataset.
+At a period of 10 days, the median transit probability in the selected sample
+of stars is $5\%$ so we model the transit probability\footnote{We're using the
+letter $Q$ to indicate probabilities since we're already using $P$ to mean
+period.} as a function of period as
+$\transitprob (\period) = 0.05\,(\period-10)^{-2/3}$.
 
@@
 The same as \fig{smooth-results} for \modelb.
-\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
 \figlabel{simulation-results}}
@@
 The same as \fig{smooth-rate} for \modelb.
-\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
 \figlabel{simulation-rate}}

### 2014-05-14 document/ms.tex
removing appendix

@@
 the maximum marginalized likelihood solution for \ratepars\ is a histogram of
-the catalog entries (see \app{histogram} for the derivation).
+the catalog entries.
 The case of non-negligible uncertainties is described by \citet{hogge} using a
@@
 
-For each catalog, we directly apply the inverse-detection-efficiency
-procedure as implemented by \citealt{petigura}\footnote{Our implementation
-exactly reproduces their results when applied to the published catalog.} and
-our probabilistic method, marginalizing over the hyperparameters of the
-Gaussian process regularization.
+For each catalog, we directly apply the inverse-detection-efficiency procedure
+as implemented by \citealt{petigura}\footnote{Our implementation reproduces
+their results when applied to the published catalog.} and our probabilistic
+method, marginalizing over the hyperparameters of the Gaussian process
+regularization.
 \Fig{smooth-results} and \fig{simulation-results} show the results of this
@@
 \acknowledgments
+We would like to thank Erik Petigura (Berkeley) for freely sharing his data
+and code.
 It is a pleasure to thank
-    Erik Petigura (Berkeley),
-    \etc
+\ldots
 for helpful contributions to the ideas and code presented here.
@@
 (grant NNX08AJ48G).
+This research builds on collaborations between astronomers and statisticians
+forged during a three week workshop on ``Modern Statistical and Computational
+Methods for Analysis of Kepler Data'' at SAMSI in June 2013.
 This research made use of the NASA \project{Astrophysics Data System}.
-
-\appendix
-
-\section{Maximum likelihood estimators for noiseless catalogs}
-\sectlabel{histogram}
-
-Equation XXXX gives the likelihood of a catalog $\catalog =
-\{\entry_k\}_{k=1}^K$, conditioned on a particular choice of occurence rate
-parameters \ratepars:
-\begin{eqnarray}
-\ln p(\catalog\,|\,\ratepars) &=& \sum_{k=1}^K \ln \rate(\entry_k)
-    - \int \rate(\entry) \dd\entry \quad.
-\end{eqnarray}
-If we model the occurence rate as a piecewise constant function
-\begin{eqnarray}
-\rate (\entry) &=& \left \{\begin{array}{ll}
-\ratepar_1 & \mathrm{if\,} \entry \in \binarea_1, \\
-\cdots & \cdots \\
-\ratepar_N & \mathrm{if\,} \entry \in \binarea_N, \\
-0 & \mathrm{otherwise}
-\end{array}\right.
-\end{eqnarray}
-where $\binarea_n$ is the volume of the bin $n$, the log likelihood becomes
-\begin{eqnarray}
-\ln p(\catalog\,|\,\ratepars) &=&
-    \sum_{k=1}^K \sum_{n=1}^N \mathbf{1}[\entry_k \in \binarea_n]
-        \ln\ratepar_n
-    - \sum_{n=1}^N \ratepar_n\,\binarea_n \quad.
-\end{eqnarray}
-The maximum likelihood solution to this can be computed analytically
-\begin{eqnarray}
-\frac{\dd \ln p(\catalog\,|\,\ratepars)}{\dd \ratepar_m} &=&
-    \frac{N_m}{\ratepar_m} - \binarea_m \\
-\to \quad \ratepar_m &=& \frac{N_m}{\binarea_m}
-\end{eqnarray}
-where $N_m$ is the number of objects in the bin $m$.
-This result is simply a \emph{normalized histogram} of the points.
-
-Including the completeness function $\completeness(\entry)$ is as simple as
-modeling the observable occurence rate as
-\begin{eqnarray}
-\rate (\entry) &=& \completeness(\entry) \, \left \{\begin{array}{ll}
-\ratepar_1 & \mathrm{if\,} \entry \in \binarea_1, \\
-\cdots & \cdots \\
-\ratepar_N & \mathrm{if\,} \entry \in \binarea_N, \\
-0 & \mathrm{otherwise}
-\end{array}\right. \quad.
-\end{eqnarray}
-This leads to a similar analytic maximum likelihood solution
-\begin{eqnarray}
-\frac{\dd \ln p(\catalog\,|\,\ratepars)}{\dd \ratepar_m} &=&
-    \frac{N_m}{\ratepar_m}
-    - \int_{\binarea_m} \completeness(\entry)\dd\entry \\
-\to \quad \ratepar_m &=& N_m \,
-    \left [\int_{\binarea_m} \completeness(\entry)\dd\entry \right ]^{-1}
-\end{eqnarray}
-where the remaining integral is over the bin $m$.
-This result is \emph{almost} the same as the inverse-detection-efficiency
-procedure but each point is weighted by the bin-averaged efficiency instead of
-the per-object efficiency.
 

### 2014-05-15 document/ms.tex
some notes about papers to cite

@@
 \begin{abstract}
-Although no true extra-solar Earth analog is known, hundreds of planets have
-been found around Sun-like stars that are either Earth-sized but on shorter
-periods, or else on year-long orbits but somewhat bigger than Earth.
-These populations ought to permit a (possibly highly uncertain) extrapolated
-measurement of the probability or rate at which Sun-like stars host true Earth
-analogs.
-Here we present a general probabilistic framework for making justified
+
+An impressive number of exoplanets have been discovered and it is of great
+scientific interest to study what we can say about the exoplanet occurence
+rate function given these discoveries.
+Every catalog is censored by non-trivial selection effects and detection
+efficiencies, and the measurements of the system properties (period, radius,
+\etc) are noisy.
+We present a general hierarchical probabilistic framework for making justified
 inferences about the occurrence rate of exoplanets taking into account survey
 completeness \emph{and observational uncertainties}.
-We demonstrate that our method produces lower variance estimates than the
-standard inverse-detection-efficiency weighting procedure.
+By applying our method to synthetic catalogs, we demonstrate that it produces
+more accurate estimates of the underlying occurence rate than the standard
+inverse-detection-efficiency weighting procedure.
 When applied to an existing catalog of small planet candidates around G dwarfs
@@
 flattening radius distribution near Earth's radius.
-One output of our models is fully marginalized estimates---marginalizing out
-observational uncertainties and all distribution parameters---of the rate
-density of true Earth analogs, expressed as a number of planets per star per
-natural logarithmic interval of period and radius, evaluated at the properties
-of Earth.
-The rate density we infer is an extrapolation; it depends strongly on the
-permitted flexibility of the distribution model, but it tends to come out
-at around ten percent, with large uncertainty.
+Although no true extra-solar Earth analog is known, hundreds of planets have
+been found around Sun-like stars that are either Earth-sized but on shorter
+periods, or else on year-long orbits but somewhat larger.
+Under strong assumptions, it has been shown that these populations permit an
+uncertain extrapolated estimate of the probability or rate at which Sun-like
+stars host true Earth analogs.
+We relax these assumptions by modeling the occurence rate as a
+multidimensional Gaussian process and find that the rate density of Earth
+analogs is about two percent with large uncertainty.
+
 \end{abstract}
@@
 \section{Introduction}
+
+CITATIONS:
+\begin{itemize}
+\item
+\citet{fressin-fp} --- simultaneous occurence rate and false positive rate
+\item
+\citet{morton} --- false positive rate
+\item
+\citet{morton-swift} --- KDE estimator for radius distribution of cool KOIs
+\item
+\citet{howard} --- inv-det-eff for short-period candidates around Sun-like
+stars
+\item
+\citet{dressing} --- inv-det-eff for cool KOIs
+\item
+\citet{dong} --- semi-analytic completeness, likelihood function-ish,
+turn-over at small radii, non-separability
+\item
+\citet{tabachnik} ---- Poisson likelihood
+\item
+\citet{tremaine} --- statistics of multis
+\item
+\citet{youdin} --- Poisson likelihood
+\item
+\citet{catanzarite} --- extrapolation to Eta-Earth assuming complete
+\item
+\citet{traub} --- extrapolation to Eta-Earth assuming complete
+\end{itemize}
 
@@
 
+\bibitem[Catanzarite \& Shao(2011)]{catanzarite}
+Catanzarite, J., \& Shao, M.\ 2011, \apj, 738, 151 (\arxiv{1103.1443})
+
 \bibitem[Dong \& Zhu(2013)]{dong}
@@
 Dressing, C.~D., \& Charbonneau, D.\ 2013, \apj, 767, 95 (\arxiv{1302.1647})
+
+\bibitem[Fressin \etal(2013)]{fressin-fp}
+Fressin, F., Torres, G., Charbonneau, D., \etal\ 2013, \apj, 766, 81
+(\arxiv{1301.0842})
 
@@
 26, 403
+
+\bibitem[Morton \& Johnson(2011)]{morton}
+Morton, T.~D., \& Johnson, J.~A.\ 2011, \apj, 738, 170 (\arxiv{1101.5630})
+
+\bibitem[Morton \& Swift(2013)]{morton-swift}
+Morton, T.~D., \& Swift, J.~J.\ 2013, \arxiv{1303.3013}
 
@@
 
+\bibitem[Traub(2012)]{traub}
+Traub, W.~A.\ 2012, \apj, 745, 20 (\arxiv{1109.4682})
+
 \bibitem[Tremaine \& Dong(2012)]{tremaine}

### 2014-05-15 document/ms.tex
fixing stupid type in transit probability... the code is right though

@@
 multidimensional Gaussian process and find that the rate density of Earth
-analogs is about two percent with large uncertainty.
+analogs is about two percent (per natural logarithmic bin in period and
+radius) with large uncertainty.
 
@@
 period.} as a function of period as
-$\transitprob (\period) = 0.05\,(\period-10)^{-2/3}$.
+\begin{eqnarray}
+\transitprob (\period) &=&
+    0.05\,\left[\frac{\period}{10\,\mathrm{days}}\right]^{-2/3} \quad.
+\end{eqnarray}
 

### 2014-05-15 document/ms.tex
some notes about the median transit probability

@@
 \end{eqnarray}
+This expression is clearly only valid for $\period \gtrsim 1.4\,\mathrm{days}$
+but the dataset that we're using (\citealt{petigura}) explicitly only includes
+periods longer than five days so this isn't a problem.
+It's also worth noting that we're using the \emph{median} transit probability
+(instead of the mean) because it is a more robust estimator in the presence of
+outliers.
+In our experiments, the results do not seem to be very sensitive to this
+choice.
+\dfmtodo{Is there any rough scaling that is obvious?}
 

### 2014-05-19 document/ms.tex
comments from @davidwhogg

@@
 
-This study is novel in a number of ways.
-We create probabilistic information about the full exoplanet population
-\emph{not} by weighting the detected objects by the inverse detection
-efficiency (\citealt{petigura}), but rather by forward modeling the
-observed distribution subject to the detection efficiency
-(\dfmtodo{CITE TREMAINE?} \citealt{dong}).
-Re-weighting the data by inverse selection probability---a method called
-V-max in the quasar and galaxy luminosity function contexts---is not wrong,
-but it produces a higher-variance estimate of the population than a justified
-likelihood approach; it is higher variance because
-the effective number of samples, after
-weighting, can be much smaller than the actual number of samples.
-In this work, we build a justified
-likelihood of the observed data, taking into account the completeness function
-or detection efficiency.
-Our likelihood function is a variable-rate Poisson likelihood; it makes the
-fundamental assumption that the data points are independently drawn from the
-model, but is agnostic about all other properties of the data or distribution
-functions.
-Indeed, the assumptions behind this study here are weaker than those of any
-previous study of the period and radius distribution of the Earth-sized
-exoplanet population (\dfmtodo{CITE TREMAINE?} \citealt{dong, petigura}).
-
-Another new aspect of this study is that it makes use of very flexible
-distribution function models.
-Indeed most of the models we consider are ``non-parametric'' in the weak sense
-that they have enormous numbers of free parameters (not the strong sense of
-having an \emph{infinite} number of free parameters).
-The models are protected from degeneracies and over-fitting by priors.
-In some models we assume that the two-dimensional distribution of planets in
-period and radius is separable; in others we don't.
-We treat all of these parameters as ``nuisance parameters'' when we ask
-questions about the occurrence of Earth analogs.
-That is, our rate conclusions are fully marginalized, and those
-marginalizations are over non-trivial numbers of parameters.
-
-When, in what follows, we ask questions about the radius disribution of
-exoplanets, the distribution parameters are no longer nuisances The
-radius-distribution parameters are what we are trying to measure.
-Even in this case, however, our results will have the period-distribution
-parameters marginalized out.
-
-Perhaps the most important novelty of this study is in its treatment of
-observational uncertainties.
-Although planet periods in our data set are very precisely measured, the
-planet radius measurements have large uncertainties.
-The justified probabilistic approach to including these uncertainties in a
-distribution analysis is to introduce the ``true'' radii as latent variables
-and infer and marginalize them out along with all other nuisance parameters.
-Although this sounds expensive---and it can be---there is a simple approach
-we have been advocating (\cite{hogge}) that makes use of importance sampling.
-We use this approach here; our results are fully marginalized over posterior
-uncertainties in all of the planet radii.
-More will be said about this below; the importance sampling looks like a
-product over data points of sums (averages) over individual-datum radius
-samplings.
-
 \section{A brief introduction to hierarchical inference}
@@
 \kepler?}
-Mathematically, this involves evaluating the so-called \emph{marginalized
-likelihood function}
+Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
+occurence rate \rate---parameterized by the parameters \ratepars---as a
+function of the physical parameters \entry\ (orbital period, planetary radius,
+\etc).
+In this framework, the occurence rate function can be ``parametric''---for
+example, a power law---or a ``non-parametric'' function---such as a histogram
+where the bin heights are the parameters \ratepars.
+The light curve for a particular target $k$ is called $\data_k$ and the
+\emph{true} set of physical parameters for that same object are given by
+$\entry_k$.
+A catalog provides a set of constraints on the parameters $\{\entry_k\}$
+conditioned the set light curves $\{\data_k\}$.
+
+Any inference about the rate function parameters can be formally addressed by
+evaluating the \emph{marginalized likelihood function}
 \begin{eqnarray}\eqlabel{crazylike}
@@
 \end{eqnarray}
-where $\{\data_k\}$ is the set of all light curves, one light curve $\data_k$
-per target $k$, \ratepars\ is a vector of parameters describing the occurence
-rate function $\rate_\ratepars(\entry)$ and $\entry_k$ is the vector of
-physical parameters describing the planetary system (orbital periods,
-radius ratios, stellar radius, \etc) around target $k$.
+where, as we mentioned above, $\{\data_k\}$ is the set of all light curves,
+one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
+parameters describing the occurence rate function $\rate_\ratepars(\entry)$
+and $\entry_k$ is the vector of physical parameters describing the planetary
+system (orbital periods, radius ratios, stellar radius, \etc) around target
+$k$.
 In this equation, our only assumption is that the datasets depend on the
@@
 \emph{observable} rate function $\obs{\rate}_\ratepars$.
-This leads to the well known result (see \citealt{tabachnik,youdin} for some
-of the examples from the exoplanet literature)
+This leads to the previously known result (see \citealt{tabachnik,youdin} for
+some of the examples from the exoplanet literature)
 \begin{eqnarray}\eqlabel{poisson-like}
@@
 As well as being a very flexible prior, this Gaussian process model is also
-appealing because it is very well studied in the machine learning literature.
+appealing because it is well studied in the machine learning literature.
 In particular, there is a very efficient algorithm---elliptical slice sampling
@@
 ratios.
-In future catalogs, this problem would be solved by publishing representations
-of all the posterior probabilities that are being measured.
+In the future, this problem would be solved by publishing a representation of
+\emph{the full posterior PDF for each object in the catalog}.
 In this case, the most useful data product would be \emph{posterior samples
@@
 \end{eqnarray}
-This is also the quantity that \citet{petigura} estimated in different units.
+This is also the quantity that \citet{petigura} estimated but in different
+units.
+That is, \gammaearth\ is the average number of exoplanets around a
+``Sun-like'' star---per natural logarithmic bins in period and
+radius---evaluated at the period and radius of Earth.
 
@@
 
-We have made the first measurement of the Earth-like exoplanet population that
-is based on forward modeling, takes proper account of the observational
-uncertainties on planet radii, and makes use of extremely flexible
-non-separable distributions for period and radius.
-We find \hoggtodo{BLAH BLAH BLAH.}
+\dfmtodo{write this section, eh?}
 
-Although our method makes weaker assumptions than any previous work in this
-area, of course it \emph{does} make strong assumptions.
-It assumes that the data set (from \citealt{petigura}) contains planets and
-only planets (no false positives).
-It assumes that the individual-planet radius uncertainties are also correctly
-determined.
-It assumes that the completeness calculation (again from \citealt{petigura})
-is accurate and smoothly varying with period and radius; it makes no attempt
-to adjust the completeness star-by-star for brightness or variability.
-It assumes that the exoplanets are all independently drawn from the same
-distribution function, with no regard to age, metallicity, or multiplicity.
+% We have made the first measurement of the Earth-like exoplanet population that
+% is based on forward modeling, takes proper account of the observational
+% uncertainties on planet radii, and makes use of extremely flexible
+% non-separable distributions for period and radius.
+% We find \hoggtodo{BLAH BLAH BLAH.}
 
-Although we have taken account of the radius uncertainties in a principled
-way, we did make one important approximation: We ignored planets with central
-values far outside the period and radius box in which we built the model.
-That is, we assumed that there are no very large outliers in radius
-determination.
-This assumption is a complex one; it involves the effective, end-to-end error
-distribution in the data set, but also smoothness of the distribution function
-at or near the edges of the analysis box.
+% Although our method makes weaker assumptions than any previous work in this
+% area, of course it \emph{does} make strong assumptions.
+% It assumes that the data set (from \citealt{petigura}) contains planets and
+% only planets (no false positives).
+% It assumes that the individual-planet radius uncertainties are also correctly
+% determined.
+% It assumes that the completeness calculation (again from \citealt{petigura})
+% is accurate and smoothly varying with period and radius; it makes no attempt
+% to adjust the completeness star-by-star for brightness or variability.
+% It assumes that the exoplanets are all independently drawn from the same
+% distribution function, with no regard to age, metallicity, or multiplicity.
 
-If our inferences are right about true Earth analogs, then we expect that
-there are about XXX transiting Earth analogs hidden in the \kepler\ data set.
-We would be able to find some of these Earth analogs if the effective noise
-level in the data could be reduced by a factor of about XXX.
-We have identified a number of directions along which this noise reduction
-might happen, including but not limited to retrospective data-driven or
-physics-driven recalibrations of the \kepler\ focal plane
-(\citealt{hoggwhitepaper}), new methods for performing aperture photometry,
-and better models for stochastic instrument variability or stochastic
-intrinsic stellar variability that take account of frequency structure
-(\citealt{brewer, carter, roberts}).
-If any (or all) of these deliver substantial improvements for \kepler\
-archival data reanalyses, it's Nobel Prize time!  (Or at least \textsl{The
-Colbert Report}?)
+% Although we have taken account of the radius uncertainties in a principled
+% way, we did make one important approximation: We ignored planets with central
+% values far outside the period and radius box in which we built the model.
+% That is, we assumed that there are no very large outliers in radius
+% determination.
+% This assumption is a complex one; it involves the effective, end-to-end error
+% distribution in the data set, but also smoothness of the distribution function
+% at or near the edges of the analysis box.
+
+% If our inferences are right about true Earth analogs, then we expect that
+% there are about XXX transiting Earth analogs hidden in the \kepler\ data set.
+% We would be able to find some of these Earth analogs if the effective noise
+% level in the data could be reduced by a factor of about XXX.
+% We have identified a number of directions along which this noise reduction
+% might happen, including but not limited to retrospective data-driven or
+% physics-driven recalibrations of the \kepler\ focal plane
+% (\citealt{hoggwhitepaper}), new methods for performing aperture photometry,
+% and better models for stochastic instrument variability or stochastic
+% intrinsic stellar variability that take account of frequency structure
+% (\citealt{brewer, carter, roberts}).
+% If any (or all) of these deliver substantial improvements for \kepler\
+% archival data reanalyses, it's Nobel Prize time!  (Or at least \textsl{The
+% Colbert Report}?)
 

### 2014-05-19 document/ms.tex
a first pass at the introduction

@@
 
-\newcommand{\modela}{\emph{Model A}}
-\newcommand{\modelb}{\emph{Model B}}
+\newcommand{\modela}{\emph{Catalog A}}
+\newcommand{\modelb}{\emph{Catalog B}}
 
@@
 
-CITATIONS:
+% CITATIONS:
+% \begin{itemize}
+% \item
+% \citet{fressin-fp} --- simultaneous occurence rate and false positive rate
+% \item
+% \citet{morton} --- false positive rate
+% \item
+% \citet{morton-swift} --- KDE estimator for radius distribution of cool KOIs
+% \item
+% \citet{howard} --- inv-det-eff for short-period candidates around Sun-like
+% stars
+% \item
+% \citet{dressing} --- inv-det-eff for cool KOIs
+% \item
+% \citet{dong} --- semi-analytic completeness, likelihood function-ish,
+% turn-over at small radii, non-separability
+% \item
+% \citet{tabachnik} ---- Poisson likelihood
+% \item
+% \citet{tremaine} --- statistics of multis
+% \item
+% \citet{youdin} --- Poisson likelihood
+% \item
+% \citet{catanzarite} --- extrapolation to Eta-Earth assuming complete
+% \item
+% \citet{traub} --- extrapolation to Eta-Earth assuming complete
+% \end{itemize}
+
+NASA's \kepler\ mission has enabled the discovery of thousands of exoplanet
+candidates (\dfmtodo{CITE KEPLER}).
+Many of these planets orbit Sun-like stars (\citealt{petigura}), where the
+definition of Sun-like is given in terms of the star's temperature and surface
+gravity.
+Given these catalogs, it is interesting to ask what we can say about the
+occurence rate of exoplanets as a function of their physical parameters
+(period, radius, \etc).
+The two main reasons why this is interesting are that \emph{(a)} theories of
+planet formation make predictions about the shape of this function
+(\dfmtodo{CITE}), and \emph{(b)} it allows us to place probabilistic
+constraints on the rate of Earth analogs\footnote{We define, in detail, what
+we mean by Earth analog below but, in short, we mean Earth-sized planets on
+year-long orbits around Sun-like stars.}.
+
+\citet{petigura} recently published constraints on this function based on
+their independent analysis of the \kepler\ light curves for 42,000 Sun-like
+stars but the study of exoplanet occurence rates has a long history.
+\dfmtodo{Summarize history with lotsa citations.}
+
+A very important component of any study of exoplanet occurence rates is the
+treatment of detection efficiency.
+Qualitatively, in a transit survey, small planets with long periods are much
+harder to detect than large planets orbiting close to their star.
+This effect is degenerate with any inferences about the rate function and it
+can be hard to constraint quantitatively.
+In practice, there are three methods for taking this effect into account:
+\emph{(a)} making conservative cuts on the candidates and assuming that the
+resulting catalog is complete (\citealt{catanzarite, traub}; \dfmtodo{CITE}),
+\emph{(b)} asserting an analytic form for the detection efficiency as a
+function of approximate signal-to-noise (\citealt{howard, dressing, dong,
+fressin-fp, morton-swift}; \dfmtodo{CITE}), and \emph{(c)} determining the
+detection efficiency empirically by injecting synthetic signals into the raw
+data and testing recovery rates (\citealt{petigura-a, petigura}).
+
+In this \paper, we develop a flexible framework for probabilistic inference of
+exoplanet occurence rates that can be applied to incomplete catalogs with
+\emph{non-negligible observational uncertainties}.
+Our method is a generalization of a previously studied method
+(\citealt{hogge}) to account for detection efficiencies.
+We run tests on simulated datasets---comparing results with the standard
+techniques that neglect observational uncertainties---and apply our method to
+a real catalog of small planets transiting Sun-like stars
+(\citealt{petigura}).
+For the purposes of this \paper, we make some strong assumptions---although we
+argue that they are weaker than the implicit assumptions in previous
+studies---but none of these are necessary for the applicability of our method.
+Specifically, we make the following assumptions
 \begin{itemize}
-\item
-\citet{fressin-fp} --- simultaneous occurence rate and false positive rate
-\item
-\citet{morton} --- false positive rate
-\item
-\citet{morton-swift} --- KDE estimator for radius distribution of cool KOIs
-\item
-\citet{howard} --- inv-det-eff for short-period candidates around Sun-like
-stars
-\item
-\citet{dressing} --- inv-det-eff for cool KOIs
-\item
-\citet{dong} --- semi-analytic completeness, likelihood function-ish,
-turn-over at small radii, non-separability
-\item
-\citet{tabachnik} ---- Poisson likelihood
-\item
-\citet{tremaine} --- statistics of multis
-\item
-\citet{youdin} --- Poisson likelihood
-\item
-\citet{catanzarite} --- extrapolation to Eta-Earth assuming complete
-\item
-\citet{traub} --- extrapolation to Eta-Earth assuming complete
+
+\item the candidates in the catalog are independent Poisson draws from the
+censored occurence rate,
+
+\item every candidate is a real exoplanet (there are no false positives),
+
+\item the observational uncertainties on the physical parameters are
+non-negligible but the amplitudes are known (the catalog provides
+probabilistic constraints on the parameters),
+
+\item the detection efficiency of the pipeline is known, and
+
+\item the true occurence rate function is \emph{smooth}\footnote{We give our
+definition of ``smooth'' in more detail below but our model is very flexible
+so this is not a strong restriction.}.
+
 \end{itemize}
-
-\hoggtodo{%
-\kepler\ has been busting out exoplanets (CITE).
-Many of them even appear possibly rocky and habitable (CITE).
-Nonetheless, none yet known are true Earth analogs.
-}
-
-\hoggtodo{%
-This hasn't stopped anyone from estimating the rate at which Sun-like stars
-host Earth-like planets on year-ish period orbits (CITE).
-All of these rate estimates are extrapolations of one kind or another.
-Here we perform an extremely conservative extrapolation of this kind.
-}
-
-\hoggtodo{%
-What would make an extrapolation like this conservative?
-Using \emph{huge bins} does not a conservative estimate make.
-You don't want to assume separabilities you can't justify.
-You want to explore an enormous function space for the exoplanet rates or pdf.
-We are going to do be conservative in these senses here, far more conservative
-than anyone who has come before.
-}
-
-\dfmtodo{%
-Cite \citet{tremaine} for likelihood function including multiplicity.
-}
+The first assumption---conditional independence of the candidates---is
+reasonable since the dataset that we consider explicitly includes only single
+transiting systems (\citealt{petigura}).
+When considering other datasets it will be interesting and important to study
+the distributions of taking multiple transiting systems into account
+(\dfmtodo{CITE}) and this would be possible using a conditional distributions
+(as developed by \citealt{tremaine}).
+The second assumption---neglecting false positives---is also strong and not
+really justified (\citealt{fressin-fp, morton}).
+For this \paper, we will neglect this issue and only comment on the effects
+but the prior distributions published by \citet{fressin-fp} would be directly
+applicable in our method and this would be a very interesting generalization.
 
@@
 but fit using the full general model.
-The first model---\modela---has a smooth occurence surface where both
-distributions are smooth broken power laws.
+The first catalog---\modela--- is generated assuming a smooth occurence
+surface where both distributions are broken power laws.
 The second---\modelb---is designed to be exactly the distribution inferred by

### 2014-05-19 document/ms.tex
expanding on the extrapolation section

@@
 \item the observational uncertainties on the physical parameters are
-non-negligible but the amplitudes are known (the catalog provides
-probabilistic constraints on the parameters),
+non-negligible but known (the catalog provides probabilistic constraints on
+the parameters),
 
@@
 \section{Model generalities}
+\sectlabel{model}
 
@@
 can also be used to constrain the rate of Earth analogs.
-Even though no true analog has been observed, we can make some assumptions
-about the distribution (as described below) and extrapolate to the location of
-the Earth.
 Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
@@
 radius---evaluated at the period and radius of Earth.
+
+Since no Earth analogs have been found any constraints on this rate must be
+extrapolated from the existing observations.
+This extrapolation is generally done by assuming a functional form for the
+occurence rate function, constraining it using the observed candidates and
+extrapolating.
+All published extrapolations are based on rigid models like a power law
+(\citealt{catanzarite, traub}) or linear in the cumulative period distribution
+(\citealt{petigura}).
+All of these results are sensitive to this choice of function and to the other
+implicit decisions (for example the size of radius bin that is considered
+Earth-like by \citealt{petigura}).
+We weaken the assumptions necessary for extrapolation by only assuming that
+the distribution is smooth using the Gaussian process prior described in
+\sect{model}.
+This model will, of course, break down if there is a sharp feature in the
+occurence rate function right at the location of Earth but otherwise this is
+the most conservative extrapolation method published up to this point.
+
+The results of our extrapolation method, when applied to the synthetic
+catalogs described in the previous section, are shown in \fig{smooth-rate} and
+\fig{simulation-rate}.
+We implemented the extrapolation method from \citet{petigura} that assumes a
+linear relation in the cumulative period distribution for small planets and
+compared these results---converting to the correct units---to our method and
+the truth.
+In both cases, our method returns a less precise but more accurate result for
+the occurence rate and the error bars given by the functional extrapolation
+are overly optimistic.
 

### 2014-05-19 document/ms.tex
a few words about the real rate

@@
 analogs is about two percent (per natural logarithmic bin in period and
-radius) with large uncertainty.
+radius) at low signal-to-noise.
 
@@
 \section{Data and completeness function}
+\sectlabel{data}
 
@@
 \caption{%
+{\bf Simulated data}.
 Inferences about the occurence rate function based on the simulated catalog
@@
 \caption{%
+{\bf Simulated data}.
 The same as \fig{smooth-results} for \modelb.
@@
 \section{Extrapolation to Earth}
+\sectlabel{extrap}
 
@@
 \caption{%
+{\bf Simulated data}.
 The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
-different techniques.
+different techniques applied to the \modela\ simulation.
 Applying the method used by \citet{petigura} gives a constraint indicated by
@@
 \caption{%
+{\bf Simulated data}.
 The same as \fig{smooth-rate} for \modelb.
@@
 
-\section{Real data}
+\section{Results from real data}
+
+Having developed this probabilistic framework for occurence rate inferences
+and demonstrating that it produces reasonable results when applied to
+simulated datasets, we now turn to real data.
+As described in \sect{data}, we will use the catalog of small exoplanet
+candidates orbiting Sun-like stars published by \citet{petigura}.
+This is a great test case because those authors empirically measured the
+detection efficiency of their pipeline as a function of the parameters of
+interest.
+
+\Fig{real-results} shows posterior samples from the inferred occurence rate
+distribution as a function of period and radius conditioned on the catalog.
+The marginalized distributions are qualitatively consistent with the occurence
+rate functions measured by \citet{petigura}.
+The MCMC chain used to generate this figure is available online at
+\dfmtodo{ADD SOME URL}.
+
+Our constraint on the rate of Earth analogs (as defined in \sect{extrap}) is
+in tension---even though our result has large fractional uncertainty---with
+the result from \citet{petigura}.
+This is shown in \fig{real-rate} where we compare the marginalized posterior
+PDF for \gammaearth\ to the published value and uncertainty.
+Quantitatively, we find that the rate of Earth analogs---the average number of
+Earth-sized planets orbiting a Sun-like on a year-long period---is
+\begin{eqnarray}
+\gammaearth &=& 0.017^{+0.018}_{-0.009} \quad.
+\end{eqnarray}
+Converted to these units, \citet{petigura} measured $0.119_{-0.046}^{+0.035}$
+for the same quantity (indicated as the vertical lines in \fig{real-rate}).
 

### 2014-05-19 document/ms.tex
I can't spell

@@
 An impressive number of exoplanets have been discovered and it is of great
-scientific interest to study what we can say about the exoplanet occurence
+scientific interest to study what we can say about the exoplanet occurrence
 rate function given these discoveries.
@@
 By applying our method to synthetic catalogs, we demonstrate that it produces
-more accurate estimates of the underlying occurence rate than the standard
+more accurate estimates of the underlying occurrence rate than the standard
 inverse-detection-efficiency weighting procedure.
@@
 stars host true Earth analogs.
-We relax these assumptions by modeling the occurence rate as a
+We relax these assumptions by modeling the occurrence rate as a
 multidimensional Gaussian process and find that the rate density of Earth
@@
 % \item
-% \citet{fressin-fp} --- simultaneous occurence rate and false positive rate
+% \citet{fressin-fp} --- simultaneous occurrence rate and false positive rate
 % \item
@@
 Given these catalogs, it is interesting to ask what we can say about the
-occurence rate of exoplanets as a function of their physical parameters
+occurrence rate of exoplanets as a function of their physical parameters
 (period, radius, \etc).
@@
 their independent analysis of the \kepler\ light curves for 42,000 Sun-like
-stars but the study of exoplanet occurence rates has a long history.
+stars but the study of exoplanet occurrence rates has a long history.
 \dfmtodo{Summarize history with lotsa citations.}
 
-A very important component of any study of exoplanet occurence rates is the
+A very important component of any study of exoplanet occurrence rates is the
 treatment of detection efficiency.
@@
 In this \paper, we develop a flexible framework for probabilistic inference of
-exoplanet occurence rates that can be applied to incomplete catalogs with
+exoplanet occurrence rates that can be applied to incomplete catalogs with
 \emph{non-negligible observational uncertainties}.
@@
 \item the candidates in the catalog are independent Poisson draws from the
-censored occurence rate,
+censored occurrence rate,
 
@@
 
-\item the true occurence rate function is \emph{smooth}\footnote{We give our
+\item the true occurrence rate function is \emph{smooth}\footnote{We give our
 definition of ``smooth'' in more detail below but our model is very flexible
@@
 In this \paper, we are asking the question: \emph{what constraints can we put
-on the occurence rate of exoplanets given all the light curves measured by
+on the occurrence rate of exoplanets given all the light curves measured by
 \kepler?}
 Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
-occurence rate \rate---parameterized by the parameters \ratepars---as a
+occurrence rate \rate---parameterized by the parameters \ratepars---as a
 function of the physical parameters \entry\ (orbital period, planetary radius,
 \etc).
-In this framework, the occurence rate function can be ``parametric''---for
+In this framework, the occurrence rate function can be ``parametric''---for
 example, a power law---or a ``non-parametric'' function---such as a histogram
@@
 one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
-parameters describing the occurence rate function $\rate_\ratepars(\entry)$
+parameters describing the occurrence rate function $\rate_\ratepars(\entry)$
 and $\entry_k$ is the vector of physical parameters describing the planetary
@@
 In this equation, our only assumption is that the datasets depend on the
-occurence rate of exoplanets only through the catalog $\{\entry_k\}$.
+occurrence rate of exoplanets only through the catalog $\{\entry_k\}$.
 In our case, this assumption qualitatively means that the signals found in the
@@
 In the inference procedure described above, we only discussed an abstract
-occurence rate function.
+occurrence rate function.
 To make it more concrete, we'll model the catalog as a Poisson draw from the
@@
 $\rate_\ratepars(\entry)$ is the object that we want to infer: the true
-occurence rate.
+occurrence rate.
 Up to this point, we haven't discussed any specific functional form for
 $\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
-whether we model the occurence rate as (for example) a broken power law or
+whether we model the occurrence rate as (for example) a broken power law or
 a histogram.
@@
 It lets us efficiently compute the \emph{marginalized likelihood of the entire
-set of light curves for a particular occurence rate function}.
+set of light curves for a particular occurrence rate function}.
 
@@
 
-If we model the occurence rate as a histogram with $J$ fixed bin volumes
+If we model the occurrence rate as a histogram with $J$ fixed bin volumes
 $\binarea_j$
@@
 Since it is the maximum likelihood solution, this result is the minimum
-variance estimator for the occurence rate in the limit of negligible catalog
+variance estimator for the occurrence rate in the limit of negligible catalog
 uncertainties.
@@
 
-\paragraph{Occurence rate parameterization \& priors}
-For the remainder of this \paper, we also model the occurence rate as a
+\paragraph{occurrence rate parameterization \& priors}
+For the remainder of this \paper, we also model the occurrence rate as a
 two-dimensional histogram with fixed logarithmic bins in period and radius.
@@
 likelihood result is no longer analytic.
-Therefore, if we want to compute the ``best-fit'' occurence rate, we can use a
+Therefore, if we want to compute the ``best-fit'' occurrence rate, we can use a
 standard non-linear optimization algorithm.
@@
 In this case, we're especially interested in probabilistic constraints on the
-occurence rate; not just the best-fit model.
+occurrence rate; not just the best-fit model.
 To do this, we apply a prior\footnote{This is a slight misnomer because
 we also we also marginalize over the parameters of this model.} $p(\ratepars)$
-on the occurence rate parameters and generate samples from the posterior
+on the occurrence rate parameters and generate samples from the posterior
 probability
@@
 This model encodes our prior belief that, on the grid scale that we consider,
-the occurence rate should be smooth but it is otherwise very flexible about
+the occurrence rate should be smooth but it is otherwise very flexible about
 the form of the function.
@@
 start by ``observing'' a few synthetic catalogs from qualitatively different
-known occurence rate functions.
+known occurrence rate functions.
 For each of these simulations, we'll take the completeness function computed
@@
 but fit using the full general model.
-The first catalog---\modela--- is generated assuming a smooth occurence
+The first catalog---\modela--- is generated assuming a smooth occurrence
 surface where both distributions are broken power laws.
@@
 analysis in both cases.
-In particular, the side panels compare the marginalized occurence rate
+In particular, the side panels compare the marginalized occurrence rate
 functions in period and radius to the true functions that were used to
 generate the catalogs.
-\Fig{smooth-results} shows that even if the true occurence rate is a smooth
+\Fig{smooth-results} shows that even if the true occurrence rate is a smooth
 function, the rate inferred by the inverse-detection-efficiency method can
@@
 In this first example---where the true distribution is well described by our
-Gaussian process model---the probabilistic inference of the occurence rate is
+Gaussian process model---the probabilistic inference of the occurrence rate is
 both more precise and accurate.
@@
 In this case, \fig{simulation-results} shows that the probabilistic
-constraints on the occurence rate are less precise \emph{but more accurate}
+constraints on the occurrence rate are less precise \emph{but more accurate}
 than results using the inverse-detection-efficiency method.
@@
 {\bf Simulated data}.
-Inferences about the occurence rate function based on the simulated catalog
+Inferences about the occurrence rate function based on the simulated catalog
 \modela.
@@
 function (\citealt{petigura}), and the grayscale shows the median posterior
-occurence surface.
+occurrence surface.
 \emph{Top and left:} the red dashed line shows the \emph{true} distribution
@@
 results of the inverse-detection-efficiency procedure, and the histograms are
-posterior samples from the marginalized occurence rate as inferred by our
+posterior samples from the marginalized occurrence rate as inferred by our
 method.
@@
 This extrapolation is generally done by assuming a functional form for the
-occurence rate function, constraining it using the observed candidates and
+occurrence rate function, constraining it using the observed candidates and
 extrapolating.
@@
 This model will, of course, break down if there is a sharp feature in the
-occurence rate function right at the location of Earth but otherwise this is
+occurrence rate function right at the location of Earth but otherwise this is
 the most conservative extrapolation method published up to this point.
@@
 In both cases, our method returns a less precise but more accurate result for
-the occurence rate and the error bars given by the functional extrapolation
+the occurrence rate and the error bars given by the functional extrapolation
 are overly optimistic.
@@
 
-Having developed this probabilistic framework for occurence rate inferences
+Having developed this probabilistic framework for occurrence rate inferences
 and demonstrating that it produces reasonable results when applied to
@@
 
-\Fig{real-results} shows posterior samples from the inferred occurence rate
+\Fig{real-results} shows posterior samples from the inferred occurrence rate
 distribution as a function of period and radius conditioned on the catalog.
-The marginalized distributions are qualitatively consistent with the occurence
-rate functions measured by \citet{petigura}.
+The marginalized distributions are qualitatively consistent with the
+occurrence rate functions measured by \citet{petigura}.
 The MCMC chain used to generate this figure is available online at

### 2014-05-20 document/ms.tex
listing assumptions in discussion

@@
 
-\dfmtodo{write this section, eh?}
+In our analysis we make a few simplifying assumptions.
+Each of these could be relaxed and \dfmtodo{BLAH BLAH BLAH}.
+\begin{itemize}
+
+\item {\bf Conditional independence}\quad
+We assumed that every object in the catalog was a conditionally independent
+Poisson draw from the observable occurrence rate function.
+This is an interesting assumption to consider when applying this method to a
+different catalog where multiple transiting systems are included.
+In practice, the best first step towards relaxing this assumption is probably
+to follow \citet{tremaine} and assume that the mutual inclination distribution
+is the only source of conditional dependence between planets.
+For this \paper, the assumption of conditional independence is justified
+because the dataset explicitly includes only systems with a single transiting
+exoplanet.
+
+\item {\bf False positives}\quad
+
+\item {\bf Known observational uncertainties}\quad
+
+\item {\bf Given empirical detection efficiency}\quad
+
+\item {\bf Smooth rate function}\quad
+
+\end{itemize}
 

### 2014-05-20 document/ms.tex
definition of gamma earth in introduction and summary in intro

@@
 
+Conditioned on our assumptions and the choices made in planet detection,
+vetting and characterization pipeline (\citealt{petigura-a, petigura}), we
+constrain the rate of small exoplanets orbiting Sun-like stars.
+As part of this analysis we also place probabilistic constraints on the rate
+of Earth analogs \gammaearth\ which we define as \emph{the number density of
+planets per natural logarithmic bin in period and radius evaluated at the
+location of Earth}
+\begin{eqnarray}
+\gammaearth &=&
+\left.\frac{\dd N}{\dd\ln\period\,\dd\ln\radius}\right|
+_{\radius=\radius_\oplus,\,\period=\period_\oplus}\quad.
+\end{eqnarray}
+Since no Earth analogs have been detected, this constraint requires an
+extrapolation in both period and radius.
+\citet{petigura} performed this extrapolation by assuming that the period
+distribution of planets in a small bin in radius is flat and extrapolated from
+their sample.
+We relax this assumption and extrapolate by only assuming that the
+occurrence rate function is ``smooth'' in period and radius.
+We enforce this constraint by applying a flexible Gaussian process
+regularization to the bin heights.
+
+In the next section, we introduce the concept of hierarchical inference, how
+it applies to the problem at hand, and we derive the importance sampling
+approximation to the marginalized likelihood (this is a generalization of the
+method from \citealt{hogge}).
+Then, in \sect{model}, we describe how to include detection efficiencies in
+this model and derive a probabilistic motivation for the standard
+inverse-detection-efficiency weighting procedure.
+After testing our method on synthetic catalogs in \sect{valid}, we use the
+catalog of planet candidates and the empirically determined detection
+efficiency from \citet{petigura} to measure the occurrence rate of Earth
+analogs.
+
 \section{A brief introduction to hierarchical inference}
@@
 \section{Validation using synthetic catalogs}
+\sectlabel{valid}
 
@@
 
+Our constraints are less confident---in a signal-to-noise sense---than
+previous measurements (made using stronger assumptions; \citealt{petigura})
+but, even so, our result is
+
 % We have made the first measurement of the Earth-like exoplanet population that

### 2014-05-21 document/figures/results/make_plots.py
making plots

@@ -1,0 +1,74 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import h5py
+import numpy as np
+import matplotlib.pyplot as pl
+from scipy.misc import logsumexp
+from matplotlib.ticker import FormatStrFormatter
+
+with h5py.File("samples.h5", "r") as f:
+    samples = f["ln_occurrence_rate_samples"][...]
+    ln_period_bin_edges = f["ln_period_bin_edges"][...]
+    ln_radius_bin_edges = f["ln_radius_bin_edges"][...]
+
+bins = [ln_period_bin_edges, ln_radius_bin_edges]
+labels = ["R / R_\oplus", "P / \mathrm{day}"]
+slices = [
+    [slice(0, 4), slice(4, 7), slice(7, 10), slice(10, 13)],
+    [slice(0, 3), slice(3, 5), slice(5, 7)]
+]
+colors = "brgc"
+linestyles = ["-", "-", "-", "-"]
+
+
+def plot_results(a):
+    x = bins[a % 2]
+    s = [slice(None), slice(None), slice(None)]
+
+    fig = pl.figure()
+    ax = fig.add_subplot(111)
+    ax_right = ax.twinx()
+
+    for i, (blah, ls, c) in enumerate(zip([slice(None)] + slices[a % 2],
+                                          ["-"] + linestyles, "k" + colors)):
+        s[a] = blah
+        samps = logsumexp(samples[s], axis=a)
+        y = np.mean(samps, axis=0) - np.log(42557.0)
+        yerr = np.std(samps, axis=0)
+
+        if i == 0:
+            txt = "all"
+        else:
+            rng = np.exp(min(bins[a-1][blah])), np.exp(max(bins[a-1][blah]))
+            txt = r"({0:.1f}, {1:.1f})".format(*rng)
+
+        ax_right.plot(np.array(zip(x[:-1], x[1:])).flatten(),
+                      np.array(zip(y, y)).flatten(),
+                      color=c, ls=ls, lw=0.5*(len(slices[a % 2])-i)+1)
+        ax_right.errorbar(0.5*(x[:-1]+x[1:])+(i-0.5*len(slices[a % 2]))*0.03,
+                          y, yerr=yerr, fmt="+", color=c, capsize=0)
+
+    ax_right.set_yticklabels([])
[…]

### 2014-05-22 document/ms.tex
making plots
some comments from @davidwhogg

@@
 periods, or else on year-long orbits but somewhat larger.
-Under strong assumptions, it has been shown that these populations permit an
-uncertain extrapolated estimate of the probability or rate at which Sun-like
-stars host true Earth analogs.
-We relax these assumptions by modeling the occurrence rate as a
-multidimensional Gaussian process and find that the rate density of Earth
-analogs is about two percent (per natural logarithmic bin in period and
-radius) at low signal-to-noise.
+Under strong assumptions, these data have been used to make an extrapolated
+estimate of the probability or rate at which Sun-like stars host Earth
+analogs.
+We relax these assumptions, only requiring that the occurrence rate be a
+smooth function of period and radius (using a Gaussian process) and find that
+the rate of Earth analogs per star is about 0.02 (per natural logarithmic bin
+in period and radius) with large uncertainty.
 
@@
 As part of this analysis we also place probabilistic constraints on the rate
-of Earth analogs \gammaearth\ which we define as \emph{the number density of
-planets per natural logarithmic bin in period and radius evaluated at the
-location of Earth}
+of Earth analogs \gammaearth\ which we define as \emph{the number of planets
+per star per natural logarithmic bin in period and radius, evaluated at the
+period and radius of Earth}
 \begin{eqnarray}
@@
 \citet{petigura} performed this extrapolation by assuming that the period
-distribution of planets in a small bin in radius is flat and extrapolated from
-their sample.
-We relax this assumption and extrapolate by only assuming that the
-occurrence rate function is ``smooth'' in period and radius.
-We enforce this constraint by applying a flexible Gaussian process
+distribution of planets in a small bin in radius is flat, obtaining
+$\gammaearth \approx 0.12$.
+We relax this assumption and extrapolate by only assuming that the occurrence
+rate is a smooth function of period and radius; we find a much lower value for
+\gammaearth.
+We enforce the smoothness constraint by applying a flexible Gaussian process
 regularization to the bin heights.
+
+\dfmtodo{Justify our choice of definition. Not habitability.}
 
@@
 function in \eq{posterior}.
+Our implementation is based on Jo Bovy's BSD licensed ESS
+code\footnote{\url{https://github.com/jobovy/bovy_mcmc/blob/master/bovy_mcmc/%
+elliptical_slice.py}}.
 As mentioned above, we actually go one step further and assume uniform
@@
 
+\dfmtodo{Justify our choice of definition. Not habitability.}
+
 As well as inferring the ocurrence distribution of exoplanets, this dataset
@@
 \begin{center}
+\includegraphics[width=0.6\textwidth]{figures/results/period.pdf}
+\end{center}
+\caption{%
+The period distribution.
+\figlabel{period}}
+\end{figure}
+
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{figures/results/radius.pdf}
+\end{center}
+\caption{%
+The radius distribution.
+\figlabel{radius}}
+\end{figure}
+
+\begin{figure}[p]
+\begin{center}
 \includegraphics[width=0.6\textwidth]{figures/results/rate.pdf}
@@
 \section{Discussion}
+
+Probabilistic framework justified by MATHS.
+
+Disagrees in detail with Petigura esp.\ extrapolation.
+
+Agrees qualitatively with Petigura, Dong, Howard, \etc\ but hard to make
+quantitative comparisons to other ones (pipeline, stars, \etc).
+
+Why did we use the definition of \gammaearth?
+
+Our constraints are less confident---in a signal-to-noise sense---than
+previous measurements (made using stronger assumptions; \citealt{petigura})
+but, even so, our result is
+
+\dfmtodo{%
+Predicted number of observable Earth analogs in the \kepler\ dataset.
+Low but not zero.
+}
 
@@
 
-Our constraints are less confident---in a signal-to-noise sense---than
-previous measurements (made using stronger assumptions; \citealt{petigura})
-but, even so, our result is
-
-% We have made the first measurement of the Earth-like exoplanet population that
-% is based on forward modeling, takes proper account of the observational
-% uncertainties on planet radii, and makes use of extremely flexible
-% non-separable distributions for period and radius.
-% We find \hoggtodo{BLAH BLAH BLAH.}
-
-% Although our method makes weaker assumptions than any previous work in this
-% area, of course it \emph{does} make strong assumptions.
-% It assumes that the data set (from \citealt{petigura}) contains planets and
-% only planets (no false positives).
-% It assumes that the individual-planet radius uncertainties are also correctly
-% determined.
-% It assumes that the completeness calculation (again from \citealt{petigura})
-% is accurate and smoothly varying with period and radius; it makes no attempt
-% to adjust the completeness star-by-star for brightness or variability.
-% It assumes that the exoplanets are all independently drawn from the same
-% distribution function, with no regard to age, metallicity, or multiplicity.
-
-% Although we have taken account of the radius uncertainties in a principled
-% way, we did make one important approximation: We ignored planets with central
-% values far outside the period and radius box in which we built the model.
-% That is, we assumed that there are no very large outliers in radius
-% determination.
-% This assumption is a complex one; it involves the effective, end-to-end error
-% distribution in the data set, but also smoothness of the distribution function
-% at or near the edges of the analysis box.
-
-% If our inferences are right about true Earth analogs, then we expect that
-% there are about XXX transiting Earth analogs hidden in the \kepler\ data set.
-% We would be able to find some of these Earth analogs if the effective noise
-% level in the data could be reduced by a factor of about XXX.
-% We have identified a number of directions along which this noise reduction
-% might happen, including but not limited to retrospective data-driven or
-% physics-driven recalibrations of the \kepler\ focal plane
-% (\citealt{hoggwhitepaper}), new methods for performing aperture photometry,
-% and better models for stochastic instrument variability or stochastic
-% intrinsic stellar variability that take account of frequency structure
-% (\citealt{brewer, carter, roberts}).
-% If any (or all) of these deliver substantial improvements for \kepler\
-% archival data reanalyses, it's Nobel Prize time!  (Or at least \textsl{The
-% Colbert Report}?)
+Under these assumptions and the dataset, \etc, we stand by this analysis but
+really this is a model of the planets in Kepler convolved with any latent
+effects of the TERRA pipeline\ldots
+
+A major effect that isn't taken into account is the highest signal-to-noise
+blah blah blah\ldots our number is prolly low.
+
+\dfmtodo{%
+{\bf Advice:} \quad
+Remind us that we want a representation of the full posterior (samples)
+because that would let us do better things with the completeness (modeled as
+r/R, \etc).
+Must explicitly include priors too.
+}
 
@@
 It is a pleasure to thank
-\ldots
+Jo Bovy (IAS),
 for helpful contributions to the ideas and code presented here.

### 2014-05-28 document/figures/results/make_plots.py
making plot with linear radius bins

@@ -3,6 +3,7 @@
 
 import h5py
 import numpy as np
+from triangle import quantile
 import matplotlib.pyplot as pl
 from scipy.misc import logsumexp
 from matplotlib.ticker import FormatStrFormatter
@@ -26,28 +27,48 @@ def plot_results(a):
     x = bins[a % 2]
     s = [slice(None), slice(None), slice(None)]
 
+    x0 = np.linspace(np.exp(x[0]), np.exp(x[-1]), 5000)[:-1]
+    inds = np.digitize(np.log(x0), x) - 1
+
     fig = pl.figure()
     ax = fig.add_subplot(111)
     ax_right = ax.twinx()
+
+    fig2 = pl.figure()
+    ax2 = fig2.add_subplot(111)
+    ax_right2 = ax2.twinx()
 
     for i, (blah, ls, c) in enumerate(zip([slice(None)] + slices[a % 2],
                                           ["-"] + linestyles, "k" + colors)):
         s[a] = blah
         samps = logsumexp(samples[s], axis=a)
-        y = np.mean(samps, axis=0) - np.log(42557.0)
-        yerr = np.std(samps, axis=0)
+        # y = np.mean(samps, axis=0) - np.log(42557.0)
+        # yerr = np.std(samps, axis=0)
+        q = np.array([quantile(_ - np.log(42557.0), [0.16, 0.5, 0.84])
+                      for _ in samps.T])
+        y = q[:, 1]
+        yerr = np.vstack((q[:, 1] - q[:, 0], q[:, 2] - q[:, 1]))
 
         if i == 0:
             txt = "all"
         else:
             rng = np.exp(min(bins[a-1][blah])), np.exp(max(bins[a-1][blah]))
             txt = r"({0:.1f}, {1:.1f})".format(*rng)
+            print(txt)
 
         ax_right.plot(np.array(zip(x[:-1], x[1:])).flatten(),
                       np.array(zip(y, y)).flatten(),
                       color=c, ls=ls, lw=0.5*(len(slices[a % 2])-i)+1)
         ax_right.errorbar(0.5*(x[:-1]+x[1:])+(i-0.5*len(slices[a % 2]))*0.03,
                           y, yerr=yerr, fmt="+", color=c, capsize=0)
+
+        # Plot the linear plot.
+        y0 = y[inds] - np.log(x0)
[…]

### 2014-05-28 document/ms.tex
keywords
incorporating comments from @ruthangus

@@
 \keywords{%
-exoplanets: sickness
+methods: data analysis
 ---
-exoplanets: eta-Earth
+methods: statistical
 ---
-code: open-source
+catalogs
 ---
-keywords: made-up-by-Hogg
+planetary systems
+---
+stars: statistics
 }
@@
 treatment of detection efficiency.
-Qualitatively, in a transit survey, small planets with long periods are much
-harder to detect than large planets orbiting close to their star.
+Speaking qualitatively, in a transit survey, small planets with long periods
+are much harder to detect than large planets orbiting close to their star.
 This effect is degenerate with any inferences about the rate function and it
@@
 transiting systems (\citealt{petigura}).
-When considering other datasets it will be interesting and important to study
-the distributions of taking multiple transiting systems into account
-(\dfmtodo{CITE}) and this would be possible using a conditional distributions
-(as developed by \citealt{tremaine}).
+When considering other datasets it will be interesting and important to take
+multiple transiting systems into account (\dfmtodo{CITE}) and this would be
+possible using a conditional distributions (as developed by
+\citealt{tremaine}).
 The second assumption---neglecting false positives---is also strong and not
@@
 A catalog provides a set of constraints on the parameters $\{\entry_k\}$
-conditioned the set light curves $\{\data_k\}$.
+conditioned on the set light curves $\{\data_k\}$.
 
@@
 This is true because the catalog provides constraints on the posterior
-probability of the parameters $\{\entry_k\}$ under some choice of ``interim
-prior'' \interim
+probability of the parameters $\{\entry_k\}$ under some choice of
+\emph{interim prior} \interim
 \begin{eqnarray}\eqlabel{crazypost}
@@
 \end{eqnarray}
-It turns out that we can use this posterior to simplify \eq{crazylike} to a
-form that can, in many common cases, be evaluated efficiently.
+The interim prior $p(\{\entry_k\}\,|\,\interim)$ is the prior
+function---parameterized by \interim---that the author of the catalog choose
+when making their measurements.
+
+It turns out that we can use these posterior measurements to simplify
+\eq{crazylike} to a form that can, in many common cases, be evaluated
+efficiently.
 To find this result, multiply the integrand in \eq{crazylike} by
@@
 posterior constraints on radius ratio and stellar radius, we can't directly
-compute the importance weights $p(\{\entry_k^\}\,|\,\interim)$ needed for
+compute the importance weights $p(\{\entry_k\}\,|\,\interim)$ needed for
 \eq{importance}.
@@
 A huge benefit of this dataset is that Erik Petigura and collaborators
-published a rigorous analysis the empirical end-to-end completeness of their
-transit search pipeline.
+published a rigorous analysis of the empirical end-to-end completeness of
+their transit search pipeline.
 Instead of choosing a functional form for the detection efficiency of the
@@
 injected synthetic signals of known period and radius into the raw aperture
-photometry and determining the empirical recovery rate after the full analysis.
+photometry and determined the empirical recovery rate after the full analysis.
 
@@
 introduced by the finite number of samples.
-
 The largest injected signal had a radius of $16\,R_\oplus$ but, because of the
@@
 $R_\star/a$.
-Since we're modeling the distribution in the period--radius plane, we need to
+Since we are modeling the distribution in the period--radius plane, we need to
 compute the transit probability marginalized over stellar radius and mass.
@@
 dataset.
-At a period of 10 days, the median transit probability in the selected sample
-of stars is $5\%$ so we model the transit probability\footnote{We're using the
-letter $Q$ to indicate probabilities since we're already using $P$ to mean
-period.} as a function of period as
+At a period of 10 days\footnote{This period is chosen arbitrarily because the
+power law only needs to be normalized at one point.}, the median transit
+probability in the selected sample of stars is $5.061\%$ so we model the
+transit probability\footnote{We are using the letter $Q$ to indicate
+probabilities since we are already using $P$ to mean period.} as a function of
+period as
 \begin{eqnarray}
 \transitprob (\period) &=&
-    0.05\,\left[\frac{\period}{10\,\mathrm{days}}\right]^{-2/3} \quad.
+    0.05061\,\left[\frac{\period}{10\,\mathrm{days}}\right]^{-2/3} \quad.
 \end{eqnarray}
 This expression is clearly only valid for $\period \gtrsim 1.4\,\mathrm{days}$
-but the dataset that we're using (\citealt{petigura}) explicitly only includes
-periods longer than five days so this isn't a problem.
-It's also worth noting that we're using the \emph{median} transit probability
+but the dataset that we are using (\citealt{petigura}) explicitly only
+includes periods longer than five days so this is not a problem.
+It is also worth noting that we're using the \emph{median} transit probability
 (instead of the mean) because it is a more robust estimator in the presence of
@@
 \end{eqnarray}
-This is also the quantity that \citet{petigura} estimated but in different
-units.
 That is, \gammaearth\ is the average number of exoplanets around a
@@
 radius---evaluated at the period and radius of Earth.
-
-Since no Earth analogs have been found any constraints on this rate must be
+This is also the quantity that \citet{petigura} estimated but in different
+units.
+
+Since no Earth analogs have been found, any constraints on this rate must be
 extrapolated from the existing observations.
-This extrapolation is generally done by assuming a functional form for the
-occurrence rate function, constraining it using the observed candidates and
-extrapolating.
+This is generally done by assuming a functional form for the occurrence rate
+function, constraining it using the observed candidates and extrapolating.
 All published extrapolations are based on rigid models like a power law
@@
 It is a pleasure to thank
+Ruth Angus (Oxford),
 Jo Bovy (IAS),

### 2014-05-28 document/ms.tex
incorporating some of @timothydmorton's comments

@@
 %  * eqnarrys ONLY
-%  * ALWAYS cite inside parens () and use \citealt{}; no in-line cites
 %  * that is all.
@@
 \newcommand{\eq}[1]{\Eq{#1}}
+\newcommand{\eqalt}[1]{Equation~\ref{eq:#1}}
 \newcommand{\eqlabel}[1]{\label{eq:#1}}
@@
 We present a general hierarchical probabilistic framework for making justified
-inferences about the occurrence rate of exoplanets taking into account survey
+inferences about the occurrence rate of exoplanets, taking into account survey
 completeness \emph{and observational uncertainties}.
@@
 analogs.
-We relax these assumptions, only requiring that the occurrence rate be a
-smooth function of period and radius (using a Gaussian process) and find that
-the rate of Earth analogs per star is about 0.02 (per natural logarithmic bin
-in period and radius) with large uncertainty.
+We relax some of these assumptions, only requiring that the occurrence rate be
+a smooth function of period and radius (using a Gaussian process) and find
+that the rate of Earth analogs per star is about 0.02 (per natural logarithmic
+bin in period and radius) with large uncertainty.
 
@@
 candidates (\dfmtodo{CITE KEPLER}).
+While many of these candidates have not been confirmed as bona fide planets,
+there is evidence that the false positive rate is low (\citealt{morton,
+fressin-fp}), enabling conclusions about the population of planets based on
+the catalog of candidates.
 Many of these planets orbit Sun-like stars (\citealt{petigura}), where the
@@
 
+\dfmtodo{From Tim: In a new paragraph here, I think we should say something
+like ``In addition, there are two philosophically different ways of applying
+this detection efficiency to derive planet occurrence rates.''
+ Then first describe the inv-det-eff (cite Howard2012, Dong+Zhu2012,
+Swift+2013, Dress+Charb2013), then the forward-modeling concept, advocated by
+Youdin (2011) in a max-like framework.}
+
+In every previous study of the occurrence rate function, the authors have
+assumed that the measurement uncertainties are negligible.
+This assumption is not justified because these uncertainties---especially on
+measurements of the stellar parameters---can be large compared to the scales
+of interest (\dfmtodo{give a number and CITE}).
 In this \paper, we develop a flexible framework for probabilistic inference of
@@
 (\citealt{petigura}).
+
 For the purposes of this \paper, we make some strong assumptions---although we
@@
 studies---but none of these are necessary for the applicability of our method.
-Specifically, we make the following assumptions
+Specifically, we make the following assumptions:
 \begin{itemize}
@@
 \citealt{tremaine}).
-The second assumption---neglecting false positives---is also strong and not
-really justified (\citealt{fressin-fp, morton}).
+The second assumption---neglecting false positives---is also strong and only
+weakly justified by estimates of low false positive rates in the \kepler\ data
+(\citealt{fressin-fp, morton}).
 For this \paper, we will neglect this issue and only comment on the effects
@@
 \dfmtodo{Justify our choice of definition. Not habitability.}
+
+\dfmtodo{We should explain why our number is a lower limit here.}
 
@@
 one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
-parameters describing the occurrence rate function $\rate_\ratepars(\entry)$
-and $\entry_k$ is the vector of physical parameters describing the planetary
-system (orbital periods, radius ratios, stellar radius, \etc) around target
-$k$.
+parameters describing the population occurrence rate function
+$\rate_\ratepars(\entry)$ and $\entry_k$ is the vector of physical parameters
+describing the planetary system (orbital periods, radius ratios, stellar
+radius, \etc) around target $k$.
 In this equation, our only assumption is that the datasets depend on the
@@
 distributions from which they are drawn.
-It is worth noting that---as we will discuss further below---the catalog only
-provides \emph{probabilistic} constraints on $\{\entry_k\}$.
+It is worth emphasizing that---as we will discuss further below---the catalog
+only provides \emph{probabilistic constraints} on $\{\entry_k\}$; not perfect
+delta-function measurements.
 
@@
 
+\dfmtodo{Can we incorporate Tim's comment: There's the additional level that
+the catalog also distills detections from non-detections.  I don't know if
+there's a place to state that here, I just thought of it.  Because the whole
+"survey" is a collection of detections and non-detections.  Might be beyond
+the scope of this paragraph; just thought I'd mention.}
+
 It turns out that we can use these posterior measurements to simplify
@@
 show that not many samples are needed to get an accurate estimate.
+\Eq{importance} is the \emph{importance sampling approximation} to the
+integral in \eq{simplemarglike} where the trial density is the posterior
+probability for the catalog measurements.
 
@@
 transit probability and any other detection efficiencies.
-In practice, it probably makes sense to model the observable rate as
+In practice, we can model the observable rate as
 \begin{eqnarray}
@@
 Now, we can substitute \eq{poisson-like} into \eq{simplemarglike} and apply
-the importance sampling approximation to derive the following expression for
-the marginalized likelihood
+the importance sampling approximation (\eqalt{importance}) to derive the
+following expression for the marginalized likelihood
 \begin{eqnarray}\eqlabel{money}

### 2014-05-30 document/figures/results/make_plots.py
making figures not ugly

@@ -6,7 +6,8 @@ import numpy as np
 from triangle import quantile
 import matplotlib.pyplot as pl
 from scipy.misc import logsumexp
-from matplotlib.ticker import FormatStrFormatter
+from matplotlib.font_manager import FontProperties
+from matplotlib.ticker import FormatStrFormatter, MaxNLocator
 
 with h5py.File("samples.h5", "r") as f:
     samples = f["ln_occurrence_rate_samples"][...]
@@ -16,11 +17,13 @@ with h5py.File("samples.h5", "r") as f:
 bins = [ln_period_bin_edges, ln_radius_bin_edges]
 labels = ["R / R_\oplus", "P / \mathrm{day}"]
 slices = [
-    [slice(0, 4), slice(4, 7), slice(7, 10), slice(10, 13)],
-    [slice(0, 3), slice(3, 5), slice(5, 7)]
+    [slice(0, 4), slice(4, 8), slice(8, 12)],
+    [slice(0, 2), slice(2, 4), slice(4, 6)]
 ]
-colors = "brgc"
-linestyles = ["-", "-", "-", "-"]
+colors = ["#222222", "#222222", "#222222", "#222222"]
+# colors = "brgc"
+linestyles = ["--", ":", "-", "-."]
+linewidths = [0.75, 0.75, 0.75, 0.75]
 
 
 def plot_results(a):
@@ -30,20 +33,20 @@ def plot_results(a):
     x0 = np.linspace(np.exp(x[0]), np.exp(x[-1]), 5000)[:-1]
     inds = np.digitize(np.log(x0), x) - 1
 
-    fig = pl.figure()
+    fig = pl.figure(figsize=(6, 4))
     ax = fig.add_subplot(111)
     ax_right = ax.twinx()
 
-    fig2 = pl.figure()
+    fig2 = pl.figure(figsize=(6, 4))
     ax2 = fig2.add_subplot(111)
     ax_right2 = ax2.twinx()
 
-    for i, (blah, ls, c) in enumerate(zip([slice(None)] + slices[a % 2],
-                                          ["-"] + linestyles, "k" + colors)):
+    for i, (blah, ls, lw, c) in enumerate(zip([slice(None)] + slices[a % 2],
+                                              ["-"] + linestyles,
+                                              [1.25] + linewidths,
+                                              ["#222222"] + colors)):
         s[a] = blah
         samps = logsumexp(samples[s], axis=a)
[…]

### 2014-05-30 document/figures/results/make_plots.py
changing units of plots

@@ -29,6 +29,7 @@ linewidths = [0.75, 0.75, 0.75, 0.75]
 def plot_results(a):
     x = bins[a % 2]
     s = [slice(None), slice(None), slice(None)]
+    s2 = [None, None, None]
 
     x0 = np.linspace(np.exp(x[0]), np.exp(x[-1]), 5000)[:-1]
     inds = np.digitize(np.log(x0), x) - 1
@@ -46,7 +47,9 @@ def plot_results(a):
                                               [1.25] + linewidths,
                                               ["#222222"] + colors)):
         s[a] = blah
-        samps = logsumexp(samples[s], axis=a)
+        s2[a] = blah
+        samps = logsumexp(samples[s] + np.log(np.diff(bins[a - 1]))[s2],
+                          axis=a)
         q = np.array([quantile(_ - np.log(42557.0), [0.16, 0.5, 0.84])
                       for _ in samps.T])
         y = q[:, 1]
@@ -87,7 +90,7 @@ def plot_results(a):
     ax_right.set_yticklabels([])
     ax.set_xlim(min(x), max(x))
     ax.set_xlabel(r"$\ln {0}$".format(labels[a-1]))
-    ax.set_ylabel(r"$N_\mathrm{avg} / N_\star$")
+    ax.set_ylabel(r"$\Gamma (\ln {0})$".format(labels[a-1]))
     ax.xaxis.set_major_locator(MaxNLocator(5))
 
     a2 = ax.twiny()
@@ -98,7 +101,6 @@ def plot_results(a):
 
     ax.set_ylim(np.exp(ax_right.get_ylim()))
     ax.set_yscale("log")
-    ax.set_ylabel(r"$N_\mathrm{avg} / N_\star$")
 
     fig.subplots_adjust(bottom=0.17, top=0.83, right=0.7, left=0.17)
     prop = FontProperties()
@@ -107,12 +109,12 @@ def plot_results(a):
 
     if a == 1:
         ax_right2.set_xlim(0, 8)
-        ax_right2.set_ylim(-7, 1)
+        ax_right2.set_ylim(-7.5, 0.2)
 
     ax_right2.set_yticklabels([])
     ax2.set_ylim(np.exp(ax_right2.get_ylim()))
     ax2.set_yscale("log")
-    ax2.set_ylabel(r"$N_\mathrm{avg} / N_\star$")
[…]

### 2014-05-30 document/ms.tex
changing units of plots
incorporating a few more comments from tim

@@
 where the weight for object $\entry_k$ is $1/\completeness(\entry_k)$.
-This makes intuitive sense but (to our knowledge) does not have a clear
+This makes intuitive sense but the standard formulation does not have a clear
 probabilistic justification or interpretation.
+\hoggtodo{Find the papers that Tremaine suggested that we look at re:vmax and
+discuss the motivation in those cases.}
 
@@
 equal to zero, we find the maximum likelihood result
-\begin{eqnarray}
-\ratepar_j &=& \frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
+\begin{eqnarray}\eqlabel{ml-ide}
+{\ratepar_j}^* &=&
+\frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
 \end{eqnarray}
 where $N_j$ is the number of objects that fall within the bin $j$.
-Since it is the maximum likelihood solution, this result is the minimum
-variance estimator for the occurrence rate in the limit of negligible catalog
-uncertainties.
-This is \emph{almost} the same as the inverse-detection-efficiency result and
-in various limiting cases, they would provide the same result.
-\dfmtodo{Finish explaining why this result is better under the specified
-assumptions.}
-
-\paragraph{occurrence rate parameterization \& priors}
+We estimate the uncertainty $\delta\ratepar_j$ on this value by examining the
+curvature of the log-likelihood function near the maximum and find
+\begin{eqnarray}
+\frac{\delta\ratepar_j}{{\ratepar_j}^*} &=& \frac{1}{\sqrt{N_j}} \quad.
+\end{eqnarray}
+
+Since \eq{ml-ide} gives the maximum likelihood solution, this result is the
+minimum-variance estimator for the occurrence rate when the completeness
+function $\completeness(\entry)$ is known and the catalog uncertainties are
+negligible.
+This is \emph{almost} the same as the inverse-detection-efficiency procedure
+and if the completeness function is constant in each bin \binarea, both
+methods give the same result.
+If the completeness function is not piecewise-constant on the same grid as the
+occurrence rate inference, the inverse-detection-efficiency procedure will
+give a higher variance estimate of the occurrence rate.
+
+\paragraph{Occurrence rate parameterization \& priors}
 For the remainder of this \paper, we also model the occurrence rate as a
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/results/period.pdf}
+\includegraphics{figures/results/period.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/results/radius.pdf}
+\includegraphics{figures/results/radius.pdf}
 \end{center}
@@
 \figlabel{radius}}
+\end{figure}
+
+\begin{figure}[p]
+\begin{center}
+\includegraphics{figures/results/linear-radius.pdf}
+\end{center}
+\caption{%
+The radius distribution.
+\figlabel{linear-radius}}
 \end{figure}

### 2014-05-30 document/ms.tex
more discussion of GP 'prior'

@@
 When we include observational uncertainties---using \eq{money}---the maximum
-likelihood result is no longer analytic.
+likelihood result is no longer analytic; \eq{ml-ide} is no longer applicable
+when the uncertainties are non-negligible.
 Therefore, if we want to compute the ``best-fit'' occurrence rate, we can use a
@@
 occurrence rate; not just the best-fit model.
-To do this, we apply a prior\footnote{This is a slight misnomer because
-we also we also marginalize over the parameters of this model.} $p(\ratepars)$
-on the occurrence rate parameters and generate samples from the posterior
-probability
+To do this, we must apply a prior $p(\ratepars)$ on the occurrence rate
+parameters and generate samples from the posterior probability
 \begin{eqnarray}\eqlabel{posterior}
@@
 
-Given our model of the rate function as a collection of bins, we choose to
-model the prior as a Gaussian process (\citealt{gp,gibson-gp}) on the
-logarithmic bin heights.
+There is a lot of flexibility in the choice of functional form of
+$p(\ratepars)$.
+In the well-sampled parts of parameter space there are a lot of
+detected planets and the choice of prior makes little difference, but in the
+regions that we care about, the detection efficiency is low and applying a
+prior that captures our beliefs about the occurrence rate is necessary.
+This will be especially important when we extrapolate the occurrence rate
+function to the location of Earth---in \sect{extrap}---where no exoplanets
+have been found.
+Therefore, instead of using an uninformative prior, we want to use a prior
+that encourages the occurrence rate to be ``smooth'' but it should be flexible
+enough to capture structure that is supported by the data.
+To achieve this, we model the logarithmic bin heights as being drawn from a
+Gaussian process (\citealt{gp, gibson-gp, dfm-gp}).
 This model encodes our prior belief that, on the grid scale that we consider,
@@
 the form of the function.
-Mathematically, the prior is written as
-\begin{eqnarray}
-p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \\
-&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{I},\,
+
+Mathematically, the Gaussian process density is
+\begin{eqnarray}
+p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \nonumber\\
+&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{1},\,
 \cov(\{\binarea_j\},\,\smoothpars)\right]
-\end{eqnarray}
-where $\mathcal{N}(\cdot;\,\mean\,\bvec{I},\,\cov)$ is a $J$-dimensional
-Gaussian with a constant mean \mean\ and covariance matrix \cov\ that depends
-on the bin locations $\{\binarea_j\}$ and a set of hyperparameters
-$\smoothpars = (\smooth_0,\,\smooth_\period,\,\smooth_\radius)$.
+\eqlabel{gp}
+\end{eqnarray}
+where $\mathcal{N}(\cdot;\,\mean\,\bvec{1},\,\cov)$ is a $J$-dimensional
+Gaussian\footnote{$J$ is the total number of bins.} with a constant mean
+\mean\ and covariance matrix \cov\ that depends on the bin locations
+$\{\binarea_j\}$ and a set of hyperparameters $\smoothpars = (\smooth_0,\,
+\smooth_\period,\,\smooth_\radius)$.
 The covariance function that we use is an anisotropic, axis-aligned
@@
 
-As well as being a very flexible prior, this Gaussian process model is also
-appealing because it is well studied in the machine learning literature.
-In particular, there is a very efficient algorithm---elliptical slice sampling
-(ESS; \citealt{ess})---for generating samples from the posterior probability
-function in \eq{posterior}.
-Our implementation is based on Jo Bovy's BSD licensed ESS
+The Gaussian process model for the bin heights given in \eq{gp} very flexible
+but the results will depend on the values of the hyperparameters \mean\ and
+\smoothpars.
+Therefore, instead of fixing these parameters to specific values, we add
+another level to our hierarchical probabilistic model and marginalize over
+this choice.
+In other words, we apply priors---uniform in the logarithm---on \mean\ and
+\smoothpars, and sample from the joint posterior
+\begin{eqnarray} \eqlabel{joint-posterior}
+p(\ratepars,\,\mean,\,\smoothpars\,|\,\{\data_k\}) &\propto&
+    p(\mean,\,\smoothpars)\,p(\ratepars\,|\,\mean,\,\smoothpars)\,
+    p(\{\data_k\}\,|\,\ratepars) \quad.
+\end{eqnarray}
+
+There is a very efficient algorithm called elliptical slice sampling (ESS;
+\citealt{ess, ess-hyper}) for sampling the bin heights \ratepars\ from the
+density in \eq{joint-posterior}.
+Our implementation is adapted from Jo Bovy's BSD licensed ESS
 code\footnote{\url{https://github.com/jobovy/bovy_mcmc/blob/master/bovy_mcmc/%
 elliptical_slice.py}}.
-As mentioned above, we actually go one step further and assume uniform
-(hyper-)priors on $\ln\mean$ and $\ln\smoothpars$ in order to sample from the
-joint posterior
-\begin{eqnarray}
-p(\ratepars,\,\mean,\,\smoothpars\,|\,\{\data_k\})
-\end{eqnarray}
-using Algorithm 1 from \citet{ess-hyper}.
+To simultaneously marginalize over the hyperparameter choice, we use the
+Metropolis--Hastings update from Algorithm 1 in \citet{ess-hyper}.
+We tune the Metropolis--Hastings proposal by hand until we get an acceptance
+fraction of $\sim0.2-0.4$ for the hyperparameters.
 
@@
 heights and update the hyperparameters every 10 steps.
-We discard the first \dfmtodo{XXXX} samples as burn-in and then use the
-remaining samples as an approximation to the marginalized probability
+We only keep the final $2 \times 10^5$ steps and discard the earlier samples
+as burn-in.
+The remaining samples provide an approximation to the marginalized probability
 distribution for \ratepars.
-We tune the Metropolis--Hastings proposal used to update the hyperparameters
-by hand to get an acceptance fraction of $\sim0.2-0.4$ for those parameters.
 
@@
 
+% Just a little bit of semi-justified self-citation.
+\bibitem[Ambikasaran \etal(2014)]{dfm-gp}
+Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D.~W.,
+\& O'Neil, M.\ 2014, \arxiv{1403.6015}
+
 \bibitem[Brewer \& Stello(2009)]{brewer}

### 2014-05-31 document/ms.tex
working on the introduction

@@
 \title{%
-  Inferring the distribution and rate of Earth-like exoplanets \\
-  from noisy individual-planet inferences
+    Exoplanet occurrence rate inference \& the rate density of Earth analogs
+    from noisy incomplete catalogs
 }
+% \title{%
+%   Inferring the distribution and rate of Earth-like exoplanets \\
+%   from noisy individual-planet inferences
+% }
 
@@
     David~W.~Hogg\altaffilmark{\nyu,\mpia},
-    Timothy~D.~Morton\altaffilmark{\princeton},
-    \etal
+    Timothy~D.~Morton\altaffilmark{\princeton}
 }
@@
 \altaffiltext{\princeton}{Department of Astrophysics, Princeton University,
-                          Princeton, NJ, 08540, USA} % CHECK THIS ZIP CODE
+                          Princeton, NJ, 08544, USA}
 
@@
 
-% CITATIONS:
-% \begin{itemize}
-% \item
-% \citet{fressin-fp} --- simultaneous occurrence rate and false positive rate
-% \item
-% \citet{morton} --- false positive rate
-% \item
-% \citet{morton-swift} --- KDE estimator for radius distribution of cool KOIs
-% \item
-% \citet{howard} --- inv-det-eff for short-period candidates around Sun-like
-% stars
-% \item
-% \citet{dressing} --- inv-det-eff for cool KOIs
-% \item
-% \citet{dong} --- semi-analytic completeness, likelihood function-ish,
-% turn-over at small radii, non-separability
-% \item
-% \citet{tabachnik} ---- Poisson likelihood
-% \item
-% \citet{tremaine} --- statistics of multis
-% \item
-% \citet{youdin} --- Poisson likelihood
-% \item
-% \citet{catanzarite} --- extrapolation to Eta-Earth assuming complete
-% \item
-% \citet{traub} --- extrapolation to Eta-Earth assuming complete
-% \end{itemize}
-
 NASA's \kepler\ mission has enabled the discovery of thousands of exoplanet
-candidates (\dfmtodo{CITE KEPLER}).
+candidates (\citealt{kepler-catalog}).
 While many of these candidates have not been confirmed as bona fide planets,
@@
 (period, radius, \etc).
-The two main reasons why this is interesting are that \emph{(a)} theories of
-planet formation make predictions about the shape of this function
-(\dfmtodo{CITE}), and \emph{(b)} it allows us to place probabilistic
-constraints on the rate of Earth analogs\footnote{We define, in detail, what
-we mean by Earth analog below but, in short, we mean Earth-sized planets on
-year-long orbits around Sun-like stars.}.
+Observational constraints on the occurrence rate function can inform theories
+of planet formation (\dfmtodo{CITE}) and place probabilistic bounds on the
+rate of Earth analogs\footnote{We define, in detail, what we mean by Earth
+analog below but, in short, we mean Earth-sized planets on year-long orbits
+around Sun-like stars.}.
 
@@
 their independent analysis of the \kepler\ light curves for 42,000 Sun-like
-stars but the study of exoplanet occurrence rates has a long history.
-\dfmtodo{Summarize history with lotsa citations.}
+stars.
+This study was especially novel because the authors developed an independent
+planet search pipeline (TERRA; \citealt{petigura-a}) and determined the
+detection efficiency of their analysis empirically by injecting synthetic
+signals into real light curves measured by \kepler.
+The occurrence rate function determined by \citet{petigura} agreed
+qualitatively with previous studies of small planets orbiting Sun-like stars
+(\citealt{dong}).
+In particular, both papers describe a ``flattening'' rate function (in
+logarithmic radius) for planets around Earth's radius.
+Even though no Earth analogs were discovered in their search, \citet{petigura}
+used the small candidates that the did find to place an extrapolated
+constraint on the frequency of Earth-like exoplanets, assuming a flat
+occurrence rate in logarithmic period.
 
@@
 \emph{(a)} making conservative cuts on the candidates and assuming that the
-resulting catalog is complete (\citealt{catanzarite, traub}; \dfmtodo{CITE}),
+resulting catalog is complete (\citealt{catanzarite, traub, tremaine}),
 \emph{(b)} asserting an analytic form for the detection efficiency as a
-function of approximate signal-to-noise (\citealt{howard, dressing, dong,
-fressin-fp, morton-swift}; \dfmtodo{CITE}), and \emph{(c)} determining the
-detection efficiency empirically by injecting synthetic signals into the raw
-data and testing recovery rates (\citealt{petigura-a, petigura}).
-
-\dfmtodo{From Tim: In a new paragraph here, I think we should say something
-like ``In addition, there are two philosophically different ways of applying
-this detection efficiency to derive planet occurrence rates.''
- Then first describe the inv-det-eff (cite Howard2012, Dong+Zhu2012,
-Swift+2013, Dress+Charb2013), then the forward-modeling concept, advocated by
-Youdin (2011) in a max-like framework.}
+function of approximate signal-to-noise (\citealt{youdin, howard, dressing,
+dong, fressin-fp, morton-swift}), and \emph{(c)} determining the detection
+efficiency empirically by injecting synthetic signals into the raw data and
+testing recovery rates (\citealt{petigura-a, petigura}).
+
+There are two pedagogically different methods that are commonly used to infer
+the occurrence rate from a catalog and a detection efficiency specification.
+The first is an intuitive method that we will refer to as
+inverse-detection-efficiency procedure and the second is based on the
+likelihood function of the catalog given a parametric rate function.
+The inverse-detection-efficiency method involves making a histogram of the
+objects in the catalog where each point is weighted by its inverse detection
+probability.
+This method is very popular in the literature (\citealt{howard, dong,
+dressing, swift, petigura}) even though it is not motivated probabilistically.
+The alternative ``likelihood'' method models the catalog as a Poisson
+realization of the \emph{observable} rate function of exoplanets taking the
+survey detection efficiencies and transit probabilities into account.
+This technique has been used to constrain parametric models---a broken power
+law, for example---for the occurrence rate function (\citealt{tabachnik,
+youdin, dong}).
+There are many reasons to prefer a result based on an optimization of a
+likelihood function and as we show in \sect{model}, if we model the occurrence
+rate as a piecewise constant function, the inverse-detection-efficiency method
+can be derived as a special case of the likelihood method with a specific set
+of assumptions.
 
@@
 measurements of the stellar parameters---can be large compared to the scales
-of interest (\dfmtodo{give a number and CITE}).
+of interest.
 In this \paper, we develop a flexible framework for probabilistic inference of
@@
 \emph{non-negligible observational uncertainties}.
-Our method is a generalization of a previously studied method
-(\citealt{hogge}) to account for detection efficiencies.
+We generalize the inference method introduced by \citet{hogge} to account for
+survey detection efficiencies.
 We run tests on simulated datasets---comparing results with the standard
@@
 argue that they are weaker than the implicit assumptions in previous
-studies---but none of these are necessary for the applicability of our method.
-Specifically, we make the following assumptions:
+studies.
+None of these assumptions are necessary for the applicability of our method
+but they do simplify the procedure.
+Specifically, we assume that
 \begin{itemize}
@@
 When considering other datasets it will be interesting and important to take
-multiple transiting systems into account (\dfmtodo{CITE}) and this would be
-possible using a conditional distributions (as developed by
+multiple transiting systems into account (\citealt{lissauer, tremaine, fang})
+and this would be possible using a conditional distributions (as developed by
 \citealt{tremaine}).
@@
 applicable in our method and this would be a very interesting generalization.
+
+We must emphasize one very important consequence of our assumptions.
+We assume that the catalog of exoplanet candidates is only missing planets
+with probabilities given by the empirical detection efficiency but this
+doesn't take into account the fact that the catalog doesn't include multiple
+transiting systems.
+A large fraction of the transiting planets discovered by the \kepler\ transit
+search pipeline are members of multiple transiting systems (see
+\citealt{lissauer}, for example).
+Since \citet{petigura} only detected at most one planet per system, their
+catalog is actually a list of planet candidates \emph{without a more
+detectable companion}.
+The global effects of this selection is not trivial and an in-depth discussion
+is beyond the scope of this \paper\ but all of the results should be
+interpreted with this caveat in mind.
 
@@
 regularization to the bin heights.
-
-\dfmtodo{Justify our choice of definition. Not habitability.}
-
-\dfmtodo{We should explain why our number is a lower limit here.}
 
@@
 Gaussian\footnote{$J$ is the total number of bins.} with a constant mean
-\mean\ and covariance matrix \cov\ that depends on the bin locations
+\mean\ and covariance matrix \cov\ that depends on the bin centers
 $\{\binarea_j\}$ and a set of hyperparameters $\smoothpars = (\smooth_0,\,
@@
 \end{eqnarray}
+Strictly speaking, in this model, $p(\ratepars\,|\,\mean,\,\smoothpars)$ can't
+really be called a ``prior'' anymore and the constraints on the bin heights
+\ratepars\ are no longer independent.
 
@@
 For each candidate, \citet{petigura} used Markov chain Monte Carlo to sample
-the posterior probability density---presumably assuming uniform priors---for
-the radius ratio, transit duration, and impact parameter.
+the posterior probability density for the radius ratio, transit duration, and
+impact parameter assuming uninformative uniform priors.
 They then incorporated the uncertainties in the stellar radius and published
@@
 For the purposes of this \paper, we neglect the counting uncertainties
-introduced by the finite number of samples.
+introduced by the finite number of samples used to estimate the completeness.
 The largest injected signal had a radius of $16\,R_\oplus$ but, because of the
@@
 Given our domain knowledge of how detection efficiency depends on the physical
-parameters, it might seem silly to measure the survey completeness in period
-and radius instead of radius ratio or signal-to-noise.
+parameters, the intuitive choice would be to measure the survey completeness
+in radius ratio or signal-to-noise instead of period and radius.
 It is also likely that a change of coordinates would yield a higher precision
@@
 includes periods longer than five days so this is not a problem.
-It is also worth noting that we're using the \emph{median} transit probability
-(instead of the mean) because it is a more robust estimator in the presence of
-outliers.
-In our experiments, the results do not seem to be very sensitive to this
-choice.
-\dfmtodo{Is there any rough scaling that is obvious?}
+We're using the \emph{median} transit probability (instead of the mean)
+because it is a more robust estimator in the presence of outliers but in our
+experiments, the results do not seem to be very sensitive to this choice.
 
@@
 
-For each catalog, we directly apply the inverse-detection-efficiency procedure
-as implemented by \citealt{petigura}\footnote{Our implementation reproduces
-their results when applied to the published catalog.} and our probabilistic
-method, marginalizing over the hyperparameters of the Gaussian process
-regularization.
+For each catalog, we directly apply both the inverse-detection-efficiency
+procedure as implemented by \citealt{petigura}\footnote{Our implementation
+reproduces their results when applied to the published catalog.} and our
+probabilistic method, marginalizing over the hyperparameters of the Gaussian
+process regularization.
 \Fig{smooth-results} and \fig{simulation-results} show the results of this
@@
 In this case, \fig{simulation-results} shows that the probabilistic
-constraints on the occurrence rate are less precise \emph{but more accurate}
-than results using the inverse-detection-efficiency method.
-At this point, the distributions inferred by these two different techniques
-don't seem drastically inconsistent but the difference are most obvious as the
-survey completeness goes to zero: exactly the region of interest.
+constraints on the occurrence rate are less precise but more accurate than
+results using the inverse-detection-efficiency method.
+This effect is most apparent in the parts of parameter space where the
+detection efficiency is low---long period and small radius.
+
+\dfmtodo{Add more discussion here about how (because it is high variance) the
+inv-det-eff method tends to find a lot of structure in the occurrence rate
+function even when there shouldn't be.}
 
@@
 All of these results are sensitive to this choice of function and to the other
-implicit decisions (for example the size of radius bin that is considered
-Earth-like by \citealt{petigura}).
+implicit decisions (for example the size of bin that is considered Earth-like
+by \citealt{petigura}).
+
 We weaken the assumptions necessary for extrapolation by only assuming that
-the distribution is smooth using the Gaussian process prior described in
+the distribution is smooth using the Gaussian process model described in
 \sect{model}.
+Under this model, the occurrence rate in bins where no objects have been
+detected will be constrained by the height of nearby bins.
+
 This model will, of course, break down if there is a sharp feature in the
@@
 \fig{simulation-rate}.
-We implemented the extrapolation method from \citet{petigura} that assumes a
+We implemented the extrapolation method from \citet{petigura}.
+
+ that assumes a
 linear relation in the cumulative period distribution for small planets and
@@
 
+\bibitem[Batalha \etal(2013)]{kepler-catalog}
+Batalha, N.~M., Rowe, J.~F., Bryson, S.~T., et al.\ 2013, \apjs, 204, 24
+(\arxiv{1202.5852})
+
 \bibitem[Brewer \& Stello(2009)]{brewer}
@@
 Dressing, C.~D., \& Charbonneau, D.\ 2013, \apj, 767, 95 (\arxiv{1302.1647})
+
+\bibitem[Fang \& Margot(2012)]{fang}
+Fang, J., \& Margot, J.-L.\ 2012, \apj, 761, 92 (\arxiv{1207.5250})
 
@@
 26, 403
+
+\bibitem[Lissauer \etal(2011)]{lissauer}
+Lissauer, J.~J., Ragozzine, D., Fabrycky, D.~C., et al.\ 2011, \apjs, 197, 8
+(\arxiv{1102.0543})
 
@@
 
+\bibitem[Swift \etal(2013)]{swift}
+Swift, J.~J., Johnson, J.~A., Morton, T.~D., et al.\ 2013, \apj, 764, 105
+(\arxiv{1301.0023})
+
 \bibitem[Tabachnik \& Tremaine(2002)]{tabachnik}

### 2014-06-01 document/ms.tex
total refactor of abstract, small title tweaks

@@
 \title{%
-    Exoplanet occurrence rate inference \& the rate density of Earth analogs
-    from noisy incomplete catalogs
+    Exoplanet population inference \& the rate of Earth analogs
+    from noisy, incomplete catalogs
 }
@@
 \newcommand{\mpia}{3}
-\newcommand{\princeton}{4}
-\newcommand{\berkeley}{5}
+\newcommand{\cds}{4}
+\newcommand{\princeton}{5}
+\newcommand{\berkeley}{6}
 \author{%
     Daniel~Foreman-Mackey\altaffilmark{1,\nyu},
-    David~W.~Hogg\altaffilmark{\nyu,\mpia},
+    David~W.~Hogg\altaffilmark{\nyu,\mpia,\cds},
     Timothy~D.~Morton\altaffilmark{\princeton}
@@
                           K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{\cds}      {Center for Data Science,
+                          New York University,
+                          4 Washington Place, New York, NY, 10003, USA}
 \altaffiltext{\princeton}{Department of Astrophysics, Princeton University,
@@
 
-An impressive number of exoplanets have been discovered and it is of great
-scientific interest to study what we can say about the exoplanet occurrence
-rate function given these discoveries.
-Every catalog is censored by non-trivial selection effects and detection
-efficiencies, and the measurements of the system properties (period, radius,
-\etc) are noisy.
-We present a general hierarchical probabilistic framework for making justified
-inferences about the occurrence rate of exoplanets, taking into account survey
-completeness \emph{and observational uncertainties}.
-By applying our method to synthetic catalogs, we demonstrate that it produces
-more accurate estimates of the underlying occurrence rate than the standard
-inverse-detection-efficiency weighting procedure.
-When applied to an existing catalog of small planet candidates around G dwarfs
-(Petigura \etal\ 2013), our method confirms previous measurements of a
-flattening radius distribution near Earth's radius.
-Although no true extra-solar Earth analog is known, hundreds of planets have
+Many exoplanets have been discovered and characterized;
+it is now possible to study the population as a whole.
+Although no true extra-Solar Earth analog is known, hundreds of planets have
 been found around Sun-like stars that are either Earth-sized but on shorter
 periods, or else on year-long orbits but somewhat larger.
-Under strong assumptions, these data have been used to make an extrapolated
-estimate of the probability or rate at which Sun-like stars host Earth
-analogs.
-We relax some of these assumptions, only requiring that the occurrence rate be
-a smooth function of period and radius (using a Gaussian process) and find
-that the rate of Earth analogs per star is about 0.02 (per natural logarithmic
+Under strong assumptions, these exoplanet catalogs have been used to make an
+extrapolated estimate of the probability or rate at which Sun-like stars host
+Earth analogs.
+Complicating these studies are that
+every catalog is censored by non-trivial selection effects and detection
+efficiencies, and every system property measurement (period, radius,
+\etc) is noisy.
+Here we present a general hierarchical probabilistic framework for making justified
+inferences about the occurrence rate of exoplanets, taking into account survey
+completeness and, for the first time, \emph{observational uncertainties}.
+We are able to make fewer assumptions about the distribution than previous studies;
+we only require that the occurrence rate be a smooth function of period
+and radius (employing a Gaussian process).
+By applying our method to synthetic catalogs, we demonstrate that it produces
+more accurate estimates of the whole population than standard procedures based
+on weighting by inverse detection efficiency.
+We apply the method to an existing catalog of small planet candidates around
+G dwarf stars (Petigura \etal\ 2013).
+We confirm a previous result that the radius distribution changes slope near Earth radius.
+We find
+that the rate of Earth analogs is about 0.02 (per star per natural logarithmic
 bin in period and radius) with large uncertainty.
+This number is much smaller than previous estimates made with the same data but
+stronger assumptions.
 

### 2014-06-01 document/ms.tex
style note
small intro tweaks

@@
 %  * line breaks at the ends of sentences
+%  * Sun Solar Earth all capitalized when referring to our peeps, even as in "extra-Solar"
 %  * eqnarrys ONLY
@@
 (period, radius, \etc).
-Observational constraints on the occurrence rate function can inform theories
+Observational constraints on the population can inform theories
 of planet formation (\dfmtodo{CITE}) and place probabilistic bounds on the
-rate of Earth analogs\footnote{We define, in detail, what we mean by Earth
-analog below but, in short, we mean Earth-sized planets on year-long orbits
+rate of Earth analogs\footnote{We define, in detail, what we mean by ``Earth
+analog'' below but, roughly, we mean Earth-sized planets on year-long orbits
 around Sun-like stars.}.
@@
 This assumption is not justified because these uncertainties---especially on
-measurements of the stellar parameters---can be large compared to the scales
-of interest.
+measurements (like exoplanet radius) that depend on the stellar
+parameters---can be large compared to the scales of interest.
 In this \paper, we develop a flexible framework for probabilistic inference of
@@
 
-For the purposes of this \paper, we make some strong assumptions---although we
+For the purposes of this \paper, we make some strong assumptions, although we
 argue that they are weaker than the implicit assumptions in previous
 studies.
-None of these assumptions are necessary for the applicability of our method
-but they do simplify the procedure.
+None of these assumptions is necessary for the validity of our general method
+but they do simplify the specific procedures we employ.
 Specifically, we assume that
@@
 transiting systems (\citealt{petigura}).
+This is a limitation of the input catalog.
 When considering other datasets it will be interesting and important to take
@@
 We assume that the catalog of exoplanet candidates is only missing planets
-with probabilities given by the empirical detection efficiency but this
+with probabilities given by the empirical detection efficiency.
+In detail this must be false; the detection efficiency we use
 doesn't take into account the fact that the catalog doesn't include multiple
@@
 per star per natural logarithmic bin in period and radius, evaluated at the
-period and radius of Earth}
+period and radius of Earth} % Prof Tremaine, you owe us a beer.
 \begin{eqnarray}
@@
 $\gammaearth \approx 0.12$.
-We relax this assumption and extrapolate by only assuming that the occurrence
+We relax this assumption and extrapolate only by assuming that the occurrence
 rate is a smooth function of period and radius; we find a much lower value for
@@
 
-In the next section, we introduce the concept of hierarchical inference, how
+In the next \sectionname, we introduce the concept of hierarchical inference, how
 it applies to the problem at hand, and we derive the importance sampling

### 2014-06-01 document/ms.tex
adding arxivid for gibson paper
fixing @davidwhogg's problems with 'The Rules'

@@
 %  * line breaks at the ends of sentences
-%  * Sun Solar Earth all capitalized when referring to our peeps, even as in "extra-Solar"
+%  * Sun Solar Earth all capitalized when referring to our peeps, even as in
+%    "extra-Solar"
 %  * eqnarrys ONLY
@@
 
-Many exoplanets have been discovered and characterized;
-it is now possible to study the population as a whole.
+Many exoplanets have been discovered and characterized; it is now possible to
+study the population as a whole.
 Although no true extra-Solar Earth analog is known, hundreds of planets have
@@
 Earth analogs.
-Complicating these studies are that
-every catalog is censored by non-trivial selection effects and detection
-efficiencies, and every system property measurement (period, radius,
-\etc) is noisy.
-Here we present a general hierarchical probabilistic framework for making justified
-inferences about the occurrence rate of exoplanets, taking into account survey
-completeness and, for the first time, \emph{observational uncertainties}.
-We are able to make fewer assumptions about the distribution than previous studies;
-we only require that the occurrence rate be a smooth function of period
-and radius (employing a Gaussian process).
+Complicating these studies are that every catalog is censored by non-trivial
+selection effects and detection efficiencies, and every system property
+measurement (period, radius, \etc) is noisy.
+Here we present a general hierarchical probabilistic framework for making
+justified inferences about the occurrence rate of exoplanets, taking into
+account survey completeness and, for the first time, \emph{observational
+uncertainties}.
+We are able to make fewer assumptions about the distribution than previous
+studies; we only require that the occurrence rate be a smooth function of
+period and radius (employing a Gaussian process).
 By applying our method to synthetic catalogs, we demonstrate that it produces
@@
 on weighting by inverse detection efficiency.
-We apply the method to an existing catalog of small planet candidates around
-G dwarf stars (Petigura \etal\ 2013).
-We confirm a previous result that the radius distribution changes slope near Earth radius.
-We find
-that the rate of Earth analogs is about 0.02 (per star per natural logarithmic
-bin in period and radius) with large uncertainty.
-This number is much smaller than previous estimates made with the same data but
-stronger assumptions.
+We apply the method to an existing catalog of small planet candidates around G
+dwarf stars (Petigura \etal\ 2013).
+We confirm a previous result that the radius distribution changes slope near
+Earth radius.
+We find that the rate of Earth analogs is about 0.02 (per star per natural
+logarithmic bin in period and radius) with large uncertainty.
+This number is much smaller than previous estimates made with the same data
+but stronger assumptions.
 
@@
 
-In the next \sectionname, we introduce the concept of hierarchical inference, how
+In the next Section, we introduce the concept of hierarchical inference, how
 it applies to the problem at hand, and we derive the importance sampling
@@
 Gibson, N.~P., Aigrain, S., Roberts, S., \etal\ 2012, \mnras, 419, 2683
+(\arxiv{1109.3251})
 

### 2014-06-03 document/ms.tex
fixing #9
pair-writing with hogg

@@
 \newcommand{\project}[1]{{\sffamily #1}}
-\newcommand{\emcee}{\project{emcee}}
 \newcommand{\kepler}{\project{Kepler}}
+\newcommand{\terra}{\project{TERRA}}
 \newcommand{\license}{MIT License}
@@
 \newcommand{\etc}{\foreign{etc.}}
+\newcommand{\True}{\foreign{True}}
+\newcommand{\Truth}{\foreign{Truth}}
 
@@
 \title{%
-    Exoplanet population inference \& the rate of Earth analogs
+    Exoplanet population inference and the rate of Earth analogs \\
     from noisy, incomplete catalogs
 }
-% \title{%
-%   Inferring the distribution and rate of Earth-like exoplanets \\
-%   from noisy individual-planet inferences
-% }
 
@@
 Earth analogs.
-Complicating these studies are that every catalog is censored by non-trivial
-selection effects and detection efficiencies, and every system property
-measurement (period, radius, \etc) is noisy.
+These studies are complicated by the fact that every catalog is censored by
+non-trivial selection effects and detection efficiencies, and every property
+(period, radius, \etc) is measured noisily.
 Here we present a general hierarchical probabilistic framework for making
@@
 This study was especially novel because the authors developed an independent
-planet search pipeline (TERRA; \citealt{petigura-a}) and determined the
+planet search pipeline (\terra; \citealt{petigura-a}) and determined the
 detection efficiency of their analysis empirically by injecting synthetic
@@
 
-\item the true occurrence rate function is \emph{smooth}\footnote{We give our
+\item the \True\footnote{In this \paper, we use ``\True'' to describe an
+observable (for example, the exoplanet occurrence rate function) that would be
+trivially measured in the limit of very high signal-to-noise data.
+We use ``true'' to describe a simulation quantity with a value exactly known
+to us.} occurrence rate function is \emph{smooth}\footnote{We give our
 definition of ``smooth'' in more detail below but our model is very flexible
@@
 The light curve for a particular target $k$ is called $\data_k$ and the
-\emph{true} set of physical parameters for that same object are given by
+\True\ set of physical parameters for that same object are given by
 $\entry_k$.
@@
 an unimportant constant---without ever looking at the raw data again.
-This is true because the catalog provides constraints on the posterior
+This is permitted because the catalog provides constraints on the posterior
 probability of the parameters $\{\entry_k\}$ under some choice of
@@
 The interim prior $p(\{\entry_k\}\,|\,\interim)$ is the prior
-function---parameterized by \interim---that the author of the catalog choose
+function---parameterized by \interim---that the author of the catalog chose
 when making their measurements.
-
-\dfmtodo{Can we incorporate Tim's comment: There's the additional level that
-the catalog also distills detections from non-detections.  I don't know if
-there's a place to state that here, I just thought of it.  Because the whole
-"survey" is a collection of detections and non-detections.  Might be beyond
-the scope of this paragraph; just thought I'd mention.}
 
@@
 In practice, we can model the observable rate as
-\begin{eqnarray}
+\begin{eqnarray}\eqlabel{obs-rate}
 \obs{\rate}_\ratepars(\entry) &=&
@@
 \end{eqnarray}
-where $\completeness(\entry)$ is the detection efficiency at \entry\ and
-$\rate_\ratepars(\entry)$ is the object that we want to infer: the true
-occurrence rate.
+where $\completeness(\entry)$ is the detection efficiency (including transit
+probability) at \entry\ and $\rate_\ratepars(\entry)$ is the object that we
+want to infer: the \True\ occurrence rate.
 Up to this point, we haven't discussed any specific functional form for
@@
 a histogram.
+
+The observed rate \obs{\rate} is a quantitative description of the rate at
+which planets appear in the \citet{petigura} catalog; it is not a description
+of the \True\ rate of exoplanets.
+Inasmuch as the detection efficiency $\completeness(\entry)$ is calculated
+correctly, the function $\rate_\ratepars(\entry)$ will represent the \True\
+rate of exoplanets, at least where there is support in the data.
+In practice, an estimate of the detection efficiency will not include every
+decision or effect in the pipeline.
+One example in this case is that the completeness function does not account
+for the explicit removal of second (and third, and so on) planets.
+In principle a more complete description of the detection efficiency can be
+included and as this function becomes more accurate, our inferences about the
+\True\ rate $\rate_\ratepars(\entry)$ will be less biased.
 
@@
 density in \eq{joint-posterior}.
+In practice, for problems with this specific structure, ESS outperforms more
+traditional MCMC methods commonly employed in astrophysics
+(e.g.,~\citealt{emcee}).
 Our implementation is adapted from Jo Bovy's BSD licensed ESS
@@
 $\propto \period^{-2/3}$.
-In theory, this marginalization should be over the true distribution of these
-parameters in the selected stellar catalog but we'll approximate it by
+In theory, this marginalization should be over the \True\ distribution of
+these parameters in the selected stellar catalog but we'll approximate it by
 the empirical distribution; a reasonable approximation given the size of the
@@
 generate the catalogs.
-\Fig{smooth-results} shows that even if the true occurrence rate is a smooth
+\Fig{smooth-results} shows that even if the \True\ occurrence rate is a smooth
 function, the rate inferred by the inverse-detection-efficiency method can
@@
 Ruth Angus (Oxford),
-Jo Bovy (IAS),
+Jo Bovy (IAS), and
+Eric Ford (PSU)
 for helpful contributions to the ideas and code presented here.
@@
 
+\bibitem[Foreman-Mackey \etal(2012)]{emcee}
+\dfmtodo{THE FUCK}
+
 \bibitem[Fressin \etal(2013)]{fressin-fp}

### 2014-06-04 document/ms.tex
adding correct emcee ref

@@
 occurrence surface.
-\emph{Top and left:} the red dashed line shows the \emph{true} distribution
-that was used to generate the catalog, the points with error bars show the
-results of the inverse-detection-efficiency procedure, and the histograms are
-posterior samples from the marginalized occurrence rate as inferred by our
-method.
+\emph{Top and left:} the red dashed line shows the true distribution that was
+used to generate the catalog, the points with error bars show the results of
+the inverse-detection-efficiency procedure, and the histograms are posterior
+samples from the marginalized occurrence rate as inferred by our method.
 \figlabel{smooth-results}}
@@
 \caption{%
+{\bf Real data}.
 The same as \fig{smooth-results} when applied to the observed data from
 \citet{petigura}.
+\emph{Center:} the points with error bars show the catalog measurements, the
+contours show the survey completeness function, and the grayscale shows the
+median posterior occurrence surface.
+\emph{Top and left:} the points with error bars show the results of the
+inverse-detection-efficiency procedure, and the histograms are posterior
+samples from the marginalized occurrence rate as inferred by our method.
 \figlabel{real-results}}
@@
 \caption{%
-The period distribution.
+{\bf Real data}.
+The occurrence rate as a function of logarithmic period integrated over bins
+in logarithmic radius.
+The lines with error bars show the posterior sample median and 68th
+percentile and the line style specifies the radius bin.
+The period distribution for the largest planets in the sample
+($8 \le R/R_\oplus < 32$) continues to increase (as a function of
+$\ln\period$) for all periods while the distribution seems to flatten and
+turn over at periods around 50 days.
 \figlabel{period}}
@@
 \caption{%
-The radius distribution.
+{\bf Real data}.
+The occurrence rate as a function of logarithmic radius integrated over bins
+in logarithmic period.
+The lines with error bars show the posterior sample median and 68th
+percentile and the line style specifies the period bin.
+The distributions in all the period bins are qualitatively consistent and
+there are plausibly features near $R\sim3\,R_\oplus$ and $R\sim10\,R_\oplus$.
 \figlabel{radius}}
@@
 \caption{%
-The radius distribution.
+{\bf Real data}.
+The same as \fig{radius} but presented as a density in radius instead of
+logarithmic radius.
 \figlabel{linear-radius}}
@@
 \bibitem[Foreman-Mackey \etal(2012)]{emcee}
-\dfmtodo{THE FUCK}
+Foreman-Mackey, D., Hogg, D.~W., Lang, D., \& Goodman, J.\ 2013, \pasp, 125,
+306 (\arxiv{1202.3665})
 

### 2014-06-04 document/ms.tex
more discussion of extrapolation method

@@
 
-\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
+\newcommand{\figref}[1]{\ref{fig:#1}}
+\newcommand{\Fig}[1]{Figure~\figref{#1}}
 \newcommand{\fig}[1]{\Fig{#1}}
@@
 per star per natural logarithmic bin in period and radius, evaluated at the
-period and radius of Earth} % Prof Tremaine, you owe us a beer.
+period and radius of Earth}
+% A certain distinguished prof at the IAS owes us beer.
 \begin{eqnarray}
@@
 
-\dfmtodo{Add more discussion here about how (because it is high variance) the
-inv-det-eff method tends to find a lot of structure in the occurrence rate
-function even when there shouldn't be.}
+When applied to either simulated catalog, the inverse-detection-efficiency
+method gives a high-variance estimate of the true occurrence rate.
+One effect of this variance is that the inferred distribution will appear to
+have more small-scale structure than the true underlying distribution.
 
@@
 
-\dfmtodo{Justify our choice of definition. Not habitability.}
-
-As well as inferring the ocurrence distribution of exoplanets, this dataset
+As well as inferring the occurrence distribution of exoplanets, this dataset
 can also be used to constrain the rate of Earth analogs.
-Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
-location of Earth}
-\begin{eqnarray}
+Explicitly, we constrain the occurrence rate of exoplanets orbiting
+``Sun-like'' stars, as defined by \citet{petigura}, evaluated at the location
+of Earth:
+\begin{eqnarray}\eqlabel{gammaearth}
 \gammaearth &=& \rate (\ln\period_\oplus,\,\ln\radius_\oplus) \\
@@
 That is, \gammaearth\ is the average number of exoplanets around a
-``Sun-like'' star---per natural logarithmic bins in period and
-radius---evaluated at the period and radius of Earth.
+Sun-like star---per natural logarithmic bins in period and radius---evaluated
+at the period and radius of Earth.
 This is also the quantity that \citet{petigura} estimated but in different
 units.
+
+In \eq{gammaearth}, we define ``Earth analog'' in terms of measurable
+quantities with no mention of habitability or composition.
+To some readers, this might seem an unsatisfying choice.
+That being said, the composition of an exoplanet is notoriously difficult to
+measure even with large uncertainty and any definition of habitability is
+still extremely subjective.
+With this in mind, we stick to the observable definition for this \paper.
 
@@
 function, constraining it using the observed candidates and extrapolating.
-All published extrapolations are based on rigid models like a power law
-(\citealt{catanzarite, traub}) or linear in the cumulative period distribution
-(\citealt{petigura}).
-All of these results are sensitive to this choice of function and to the other
-implicit decisions (for example the size of bin that is considered Earth-like
-by \citealt{petigura}).
+All published extrapolations are based on rigid models of the occurrence rate
+function (for example, a power law) fit to the catalog and evaluated at the
+location of Earth (\citealt{catanzarite, traub}).
+\citet{petigura} used their catalog of planet candidates to constrain the rate
+of Earth analogs---using a definition similar to our definition in
+\eq{gammaearth}---assuming an extremely rigid model of the occurrence rate
+function: \emph{flat in logarithmic period}.
+These results are sensitive to the choice of extrapolation function and the
+specific definition of ``Earth analog''.
 
 We weaken the assumptions necessary for extrapolation by only assuming that
-the distribution is smooth using the Gaussian process model described in
-\sect{model}.
+the distribution is smooth using the Gaussian process regularization described
+in \sect{model}.
 Under this model, the occurrence rate in bins where no objects have been
-detected will be constrained by the height of nearby bins.
-
-This model will, of course, break down if there is a sharp feature in the
-occurrence rate function right at the location of Earth but otherwise this is
-the most conservative extrapolation method published up to this point.
-
-The results of our extrapolation method, when applied to the synthetic
-catalogs described in the previous section, are shown in \fig{smooth-rate} and
-\fig{simulation-rate}.
-We implemented the extrapolation method from \citet{petigura}.
-
- that assumes a
-linear relation in the cumulative period distribution for small planets and
-compared these results---converting to the correct units---to our method and
-the truth.
+detected will be constrained---with large uncertainty---by the heights of
+nearby bins.
+Therefore, even though there are no candidates that qualify as Earth analogs,
+we simply fit our model of the occurrence rate function in a large enough
+region of parameter space (including Earth) and compute the posterior
+constraints on \gammaearth.
+This works because the Gaussian process regularization actually captures our
+prior beliefs about the shape of the occurrence rate function.
+This model---and any other extrapolation---will, of course, break down if
+there is an unmeasured sharp feature in the occurrence rate function near the
+location of Earth and our method is the most conservative extrapolation
+technique published to date.
+
+For comparison, we also implemented and applied the extrapolation technique
+applied by \citet{petigura}.
+Their method assumes that, for small planets ($1 \le \radius/\radius_\oplus <
+2$) on long periods ($\period > 50\,\mathrm{days}$), the occurrence rate flat
+in logarithmic period or, equivalently, the cumulative rate is linear.
+\citet{petigura} used the candidates in their catalog to estimate the slope of
+the empirical cumulative period distribution and used that function to
+extrapolate.
+Instead of defining \gammaearth\ differentially, as we did in \eq{gammaearth},
+\citet{petigura} constrained the integral of the occurrence rate over a box in
+period and radius ($1 \le \radius/\radius_\oplus < 2$ and $200 \le
+\period/\mathrm{day} < 400$).
+Since their model implicitly assumes a constant occurrence rate across the
+bin, the differential rate is just their number divided by the bin area.
+
+Figures~\figref{smooth-rate} and~\figref{simulation-rate} compare our results
+and the results of the \citet{petigura} extrapolation procedure when applied
+to the synthetic catalogs.
+Since these catalogs were simulated from a known population model, we know the
+true value of \gammaearth\ and it is indicated in the Figures with a vertical
+gray line.
 In both cases, our method returns a less precise but more accurate result for
@@
 are overly optimistic.
+One major effect that leads to this bias is that the period distribution is
+not flat.
+Restricting the result to only include uniform models is equivalent to
+applying an extremely informative prior that doesn't have enough freedom to
+capture the complexity of the problem.
+As a result, the posterior constraints on \gammaearth\ are dominated by this
+prior choice and the resulting uncertainties are much smaller than they should
+be.
 
@@
 
+We directly applied our method to the \citet{petigura} sample and generated
+MCMC samples from the posterior probability for the occurrence rate bin
+heights, marginalizing over the hyperparameters of the Gaussian process model.
+The resulting MCMC chain is available online at \dfmtodo{ADD SOME URL}.
+
 \Fig{real-results} shows posterior samples from the inferred occurrence rate
@@
 The marginalized distributions are qualitatively consistent with the
-occurrence rate functions measured by \citet{petigura}.
-The MCMC chain used to generate this figure is available online at
-\dfmtodo{ADD SOME URL}.
+occurrence rate functions measured using the inverse-detection-efficiency
+method with larger uncertainties.
+
+The period distribution integrated over various radius ranges is shown in
+\fig{period}.
+In agreement with \citet{dong}, we find that the period distribution of large
+planets ($R > 8\,R_\oplus$) is inconsistent with the distribution of smaller
+planets.
+The occurrence rate of large planets appears to monotonically increase as a
+function of log period while the distribution for small planets seems to turn
+over at a relatively short period (around 50 days) and decrease for larger
+periods.
+
+The equivalent Figures for the radius distribution (integrated over period
+bins) are shown in Figures~\figref{radius} and~\figref{linear-radius}.
 

### 2014-06-05 document/ms.tex
massive refactor
more discussion

@@
 In the future, this problem would be solved by publishing a representation of
-\emph{the full posterior PDF for each object in the catalog}.
+\emph{the full posterior density functino for each object in the catalog}.
 In this case, the most useful data product would be \emph{posterior samples
@@
 
-The equivalent Figures for the radius distribution (integrated over period
-bins) are shown in Figures~\figref{radius} and~\figref{linear-radius}.
+The equivalent Figures for the radius distribution are shown in
+Figures~\figref{radius} and~\figref{linear-radius}.
+\Fig{radius} shows the log-radius occurrence rate integrated over various
+logarithmic bins in period.
+The distributions in each period bin are qualitatively consistent; the
+rate is dominated by small planets (around two Earth radii) with potential
+``features'' near $\radius\sim3\radius_\oplus$ and $\radius\sim
+10\radius_\oplus$.
+These features appear in every period bin and they were also detected---using
+a completely different dataset and technique---by \citet{dong} and a similar
+result is visible in the occurrence rate determined by \citet[][their Figure
+7]{fressin-fp} at low signal-to-noise.
+\Fig{linear-radius} shows the same result but presented as a function of
+linear radius.
+In these coordinates, the occurrence rate in a single bin is no longer
+uniform; instead, scales as inverse radius.
 
@@
 This is shown in \fig{real-rate} where we compare the marginalized posterior
-PDF for \gammaearth\ to the published value and uncertainty.
-Quantitatively, we find that the rate of Earth analogs---the average number of
-Earth-sized planets orbiting a Sun-like on a year-long period---is
+probability function for \gammaearth\ to the published value and uncertainty.
+Quantitatively, we find that the rate of Earth analogs is
 \begin{eqnarray}
@@
 
-% Just a little bit of semi-justified self-citation.
 \bibitem[Ambikasaran \etal(2014)]{dfm-gp}

### 2014-06-06 document/ms.tex
rate density

@@
 Under strong assumptions, these exoplanet catalogs have been used to make an
-extrapolated estimate of the probability or rate at which Sun-like stars host
-Earth analogs.
+extrapolated estimate of the rate at which Sun-like stars host Earth analogs.
 These studies are complicated by the fact that every catalog is censored by
@@
 Here we present a general hierarchical probabilistic framework for making
-justified inferences about the occurrence rate of exoplanets, taking into
-account survey completeness and, for the first time, \emph{observational
+justified inferences about the population of exoplanets, taking into account
+survey completeness and, for the first time, \emph{observational
 uncertainties}.
 We are able to make fewer assumptions about the distribution than previous
-studies; we only require that the occurrence rate be a smooth function of
-period and radius (employing a Gaussian process).
+studies; we only require that the occurrence rate density be a smooth function
+of period and radius (employing a Gaussian process).
 By applying our method to synthetic catalogs, we demonstrate that it produces
@@
 Earth radius.
-We find that the rate of Earth analogs is about 0.02 (per star per natural
-logarithmic bin in period and radius) with large uncertainty.
+We find that the rate density of Earth analogs is about 0.02 (per star per
+natural logarithmic bin in period and radius) with large uncertainty.
 This number is much smaller than previous estimates made with the same data
@@
 Given these catalogs, it is interesting to ask what we can say about the
-occurrence rate of exoplanets as a function of their physical parameters
+population of exoplanets as a function of their physical parameters
 (period, radius, \etc).
-Observational constraints on the population can inform theories
-of planet formation (\dfmtodo{CITE}) and place probabilistic bounds on the
-rate of Earth analogs\footnote{We define, in detail, what we mean by ``Earth
-analog'' below but, roughly, we mean Earth-sized planets on year-long orbits
-around Sun-like stars.}.
+Observational constraints on the population can inform theories of planet
+formation (\dfmtodo{CITE}) and place probabilistic bounds on the abundance of
+Earth analogs\footnote{For our purposes, an ``Earth analog'' is an Earth-sized
+exoplanet orbiting a Sun-like star with a year-long period.}.
 
@@
 constraint on the frequency of Earth-like exoplanets, assuming a flat
-occurrence rate in logarithmic period.
-
-A very important component of any study of exoplanet occurrence rates is the
+occurrence rate density in logarithmic period.
+
+A very important component of any study of exoplanet populations is the
 treatment of detection efficiency.
@@
 are much harder to detect than large planets orbiting close to their star.
-This effect is degenerate with any inferences about the rate function and it
+This effect is degenerate with any inferences about the rate density and it
 can be hard to constraint quantitatively.
@@
 efficiency empirically by injecting synthetic signals into the raw data and
-testing recovery rates (\citealt{petigura-a, petigura}).
+testing recovery (\citealt{petigura-a, petigura}).
 
 There are two pedagogically different methods that are commonly used to infer
-the occurrence rate from a catalog and a detection efficiency specification.
+the occurrence rate density from a catalog and a detection efficiency
+specification.
 The first is an intuitive method that we will refer to as
 inverse-detection-efficiency procedure and the second is based on the
-likelihood function of the catalog given a parametric rate function.
+likelihood function of the catalog given a parametric rate density.
 The inverse-detection-efficiency method involves making a histogram of the
@@
 The alternative ``likelihood'' method models the catalog as a Poisson
-realization of the \emph{observable} rate function of exoplanets taking the
+realization of the \emph{observable} rate density of exoplanets taking the
 survey detection efficiencies and transit probabilities into account.
 This technique has been used to constrain parametric models---a broken power
-law, for example---for the occurrence rate function (\citealt{tabachnik,
+law, for example---for the occurrence rate density (\citealt{tabachnik,
 youdin, dong}).
@@
 likelihood function and as we show in \sect{model}, if we model the occurrence
-rate as a piecewise constant function, the inverse-detection-efficiency method
-can be derived as a special case of the likelihood method with a specific set
-of assumptions.
-
-In every previous study of the occurrence rate function, the authors have
+rate density as a piecewise constant function, the
+inverse-detection-efficiency method can be derived as a special case of the
+likelihood method with a specific set of assumptions.
+
+In every previous study of exoplanet occurrence rates, the authors have
 assumed that the measurement uncertainties are negligible.
@@
 In this \paper, we develop a flexible framework for probabilistic inference of
-exoplanet occurrence rates that can be applied to incomplete catalogs with
-\emph{non-negligible observational uncertainties}.
+exoplanet occurrence rate density that can be applied to incomplete catalogs
+with \emph{non-negligible observational uncertainties}.
 We generalize the inference method introduced by \citet{hogge} to account for
@@
 
-\item the candidates in the catalog are independent Poisson draws from the
-censored occurrence rate,
+\item the candidates in the catalog are independent draws from an
+inhomogeneous Poisson process set by the censored occurrence rate density,
 
@@
 \item the \True\footnote{In this \paper, we use ``\True'' to describe an
-observable (for example, the exoplanet occurrence rate function) that would be
+observable (for example, the exoplanet occurrence rate density) that would be
 trivially measured in the limit of very high signal-to-noise data.
 We use ``true'' to describe a simulation quantity with a value exactly known
-to us.} occurrence rate function is \emph{smooth}\footnote{We give our
+to us.} occurrence rate density is \emph{smooth}\footnote{We give our
 definition of ``smooth'' in more detail below but our model is very flexible
@@
 vetting and characterization pipeline (\citealt{petigura-a, petigura}), we
-constrain the rate of small exoplanets orbiting Sun-like stars.
+constrain the rate density of small exoplanets orbiting Sun-like stars.
 As part of this analysis we also place probabilistic constraints on the rate
-of Earth analogs \gammaearth\ which we define as \emph{the number of planets
-per star per natural logarithmic bin in period and radius, evaluated at the
-period and radius of Earth}
+density\footnote{In this \paper, we use the word ``rate'' to indicate the
+dimensionless expectation value of a Poisson process and the words ``rate
+density'' to indicate a quantity that must be integrated over a finite bin in
+period and radius to deliver a rate.} of Earth analogs \gammaearth\ which we
+define as \emph{the expected number of planets per star per natural
+logarithmic bin in period and radius, evaluated at the period and radius of
+Earth}
 % A certain distinguished prof at the IAS owes us beer.
@@
 We relax this assumption and extrapolate only by assuming that the occurrence
-rate is a smooth function of period and radius; we find a much lower value for
-\gammaearth.
+rate density is a smooth function of period and radius; we find a much lower
+value for \gammaearth.
 We enforce the smoothness constraint by applying a flexible Gaussian process
@@
 catalog of planet candidates and the empirically determined detection
-efficiency from \citet{petigura} to measure the occurrence rate of Earth
-analogs.
+efficiency from \citet{petigura} to measure the occurrence rate density of
+Earth analogs.
 
@@
 In this \paper, we are asking the question: \emph{what constraints can we put
-on the occurrence rate of exoplanets given all the light curves measured by
-\kepler?}
+on the occurrence rate density of exoplanets given all the light curves
+measured by \kepler?}
 Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
-occurrence rate \rate---parameterized by the parameters \ratepars---as a
-function of the physical parameters \entry\ (orbital period, planetary radius,
-\etc).
-In this framework, the occurrence rate function can be ``parametric''---for
+occurrence rate density \rate---parameterized by the parameters \ratepars---as
+a function of the physical parameters \entry\ (orbital period, planetary
+radius, \etc).
+In this framework, the occurrence rate density can be ``parametric''---for
 example, a power law---or a ``non-parametric'' function---such as a histogram
@@
 
-Any inference about the rate function parameters can be formally addressed by
+Any inference about the rate density parameters can be formally addressed by
 evaluating the \emph{marginalized likelihood function}
@@
 one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
-parameters describing the population occurrence rate function
+parameters describing the population occurrence rate density
 $\rate_\ratepars(\entry)$ and $\entry_k$ is the vector of physical parameters
@@
 In this equation, our only assumption is that the datasets depend on the
-occurrence rate of exoplanets only through the catalog $\{\entry_k\}$.
+rate density of exoplanets only through the catalog $\{\entry_k\}$.
 In our case, this assumption qualitatively means that the signals found in the
@@
 If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
-constant rate function---where the bin heights are the parameters---and if the
+constant rate density---where the bin heights are the parameters---and if the
 uncertainties on the catalog are negligible compared to the bin widths then
@@
 In the inference procedure described above, we only discussed an abstract
-occurrence rate function.
+occurrence rate density.
 To make it more concrete, we'll model the catalog as a Poisson draw from the
-\emph{observable} rate function $\obs{\rate}_\ratepars$.
+\emph{observable} rate density $\obs{\rate}_\ratepars$.
 This leads to the previously known result (see \citealt{tabachnik,youdin} for
@@
 A further assumption that we have made in \eq{poisson-like} is that each
-object is an independent sample from the rate function.
+object is an independent sample from the inhomogeneous Poisson process set by
+the rate density.
 This assumption---though it is often made---cannot be correct when we consider
@@
 
-The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate of
-exoplanets that you would expect to observe taking into account the geometric
-transit probability and any other detection efficiencies.
-In practice, we can model the observable rate as
+The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate
+density of exoplanets that you would expect to observe taking into account the
+geometric transit probability and any other detection efficiencies.
+In practice, we can model the observable rate density as
 \begin{eqnarray}\eqlabel{obs-rate}
@@
 probability) at \entry\ and $\rate_\ratepars(\entry)$ is the object that we
-want to infer: the \True\ occurrence rate.
+want to infer: the \True\ occurrence rate density.
 Up to this point, we haven't discussed any specific functional form for
 $\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
-whether we model the occurrence rate as (for example) a broken power law or
+whether we model the rate density as (for example) a broken power law or
 a histogram.
 
-The observed rate \obs{\rate} is a quantitative description of the rate at
-which planets appear in the \citet{petigura} catalog; it is not a description
-of the \True\ rate of exoplanets.
+The observed rate density \obs{\rate} is a quantitative description of the
+rate density at which planets appear in the \citet{petigura} catalog; it is
+not a description of the \True\ rate density of exoplanets.
 Inasmuch as the detection efficiency $\completeness(\entry)$ is calculated
 correctly, the function $\rate_\ratepars(\entry)$ will represent the \True\
-rate of exoplanets, at least where there is support in the data.
+rate density of exoplanets, at least where there is support in the data.
 In practice, an estimate of the detection efficiency will not include every
@@
 included and as this function becomes more accurate, our inferences about the
-\True\ rate $\rate_\ratepars(\entry)$ will be less biased.
+\True\ rate density $\rate_\ratepars(\entry)$ will be less biased.
 
@@
 It lets us efficiently compute the \emph{marginalized likelihood of the entire
-set of light curves for a particular occurrence rate function}.
+set of light curves for a particular occurrence rate density}.
 
@@
 
-If we model the occurrence rate as a histogram with $J$ fixed bin volumes
-$\binarea_j$
+If we model the occurrence rate density as a histogram with $J$ fixed bin
+volumes $\binarea_j$
 \begin{eqnarray}
@@
 Since \eq{ml-ide} gives the maximum likelihood solution, this result is the
-minimum-variance estimator for the occurrence rate when the completeness
-function $\completeness(\entry)$ is known and the catalog uncertainties are
-negligible.
+minimum-variance estimator for the rate density when the completeness function
+$\completeness(\entry)$ is known and the catalog uncertainties are negligible.
 This is \emph{almost} the same as the inverse-detection-efficiency procedure
@@
 If the completeness function is not piecewise-constant on the same grid as the
-occurrence rate inference, the inverse-detection-efficiency procedure will
-give a higher variance estimate of the occurrence rate.
-
-\paragraph{Occurrence rate parameterization \& priors}
-For the remainder of this \paper, we also model the occurrence rate as a
+rate density inference, the inverse-detection-efficiency procedure will
+give a higher variance estimate of the occurrence rate density.
+
+\paragraph{Occurrence rate density parameterization \& priors}
+For the remainder of this \paper, we also model the rate density as a
 two-dimensional histogram with fixed logarithmic bins in period and radius.
@@
 when the uncertainties are non-negligible.
-Therefore, if we want to compute the ``best-fit'' occurrence rate, we can use a
+Therefore, if we want to compute the ``best-fit'' rate density, we can use a
 standard non-linear optimization algorithm.
@@
 In this case, we're especially interested in probabilistic constraints on the
-occurrence rate; not just the best-fit model.
-To do this, we must apply a prior $p(\ratepars)$ on the occurrence rate
+occurrence rate density; not just the best-fit model.
+To do this, we must apply a prior $p(\ratepars)$ on the rate density
 parameters and generate samples from the posterior probability
@@
 regions that we care about, the detection efficiency is low and applying a
-prior that captures our beliefs about the occurrence rate is necessary.
-This will be especially important when we extrapolate the occurrence rate
+prior that captures our beliefs about the rate density is necessary.
+This will be especially important when we extrapolate the rate density
 function to the location of Earth---in \sect{extrap}---where no exoplanets
@@
 Therefore, instead of using an uninformative prior, we want to use a prior
-that encourages the occurrence rate to be ``smooth'' but it should be flexible
-enough to capture structure that is supported by the data.
+that encourages the occurrence rate density to be ``smooth'' but it should be
+flexible enough to capture structure that is supported by the data.
 To achieve this, we model the logarithmic bin heights as being drawn from a
@@
 This model encodes our prior belief that, on the grid scale that we consider,
-the occurrence rate should be smooth but it is otherwise very flexible about
-the form of the function.
+the rate density should be smooth but it is otherwise very flexible about the
+form of the function.
 
@@
 injected synthetic signals of known period and radius into the raw aperture
-photometry and determined the empirical recovery rate after the full analysis.
+photometry and determined the empirical recovery after the full analysis.
 
@@
 start by ``observing'' a few synthetic catalogs from qualitatively different
-known occurrence rate functions.
+known occurrence rate density functions.
 For each of these simulations, we'll take the completeness function computed
@@
 (\citealt{poisson}) but for our purposes, we'll simply consider a piecewise
-constant rate function evaluated on a fine grid in log-period and log-radius.
+constant rate density evaluated on a fine grid in log-period and log-radius.
 For this discrete function, the generative procedure is simple;
@@
 {\item draw Poisson random integer $K_i\sim\mathrm{Poissson}(\obs{\rate}_i)$
-with the observable rate evaluated in the cell, and}
+with the observable rate density in the cell, and}
 {\item distribute $K_i$ catalog entries in the cell randomly.}
@@
 
-We generate synthetic catalogs from two qualitatively different rate
+We generate synthetic catalogs from two qualitatively different rate density
 functions.
@@
 but fit using the full general model.
-The first catalog---\modela--- is generated assuming a smooth occurrence
+The first catalog---\modela---is generated assuming a smooth occurrence
 surface where both distributions are broken power laws.
@@
 In particular, the side panels compare the marginalized occurrence rate
-functions in period and radius to the true functions that were used to
+density in period and radius to the true functions that were used to
 generate the catalogs.
-\Fig{smooth-results} shows that even if the \True\ occurrence rate is a smooth
-function, the rate inferred by the inverse-detection-efficiency method can
+\Fig{smooth-results} shows that even if the \True\ rate density is a smooth
+function, the density inferred by the inverse-detection-efficiency method can
 appear to have sharp features.
 In this first example---where the true distribution is well described by our
-Gaussian process model---the probabilistic inference of the occurrence rate is
-both more precise and accurate.
-
-In the second example, the true rate function includes a sharp feature chosen
+Gaussian process model---the probabilistic inference of the occurrence rate
+density is both more precise and accurate.
+
+In the second example, the true rate density includes a sharp feature chosen
 to reproduce the result published by \citet{petigura}.
 In this case, \fig{simulation-results} shows that the probabilistic
-constraints on the occurrence rate are less precise but more accurate than
+constraints on the rate density are less precise but more accurate than
 results using the inverse-detection-efficiency method.
@@
 When applied to either simulated catalog, the inverse-detection-efficiency
-method gives a high-variance estimate of the true occurrence rate.
+method gives a high-variance estimate of the true occurrence rate density.
 One effect of this variance is that the inferred distribution will appear to
@@
 {\bf Simulated data}.
-Inferences about the occurrence rate function based on the simulated catalog
-\modela.
+Inferences about the rate density based on the simulated catalog \modela.
 \emph{Center:} the points with error bars show the exoplanet candidates in the
@@
 the inverse-detection-efficiency procedure, and the histograms are posterior
-samples from the marginalized occurrence rate as inferred by our method.
+samples from the marginalized rate density as inferred by our method.
 \figlabel{smooth-results}}
@@
 As well as inferring the occurrence distribution of exoplanets, this dataset
-can also be used to constrain the rate of Earth analogs.
-Explicitly, we constrain the occurrence rate of exoplanets orbiting
-``Sun-like'' stars, as defined by \citet{petigura}, evaluated at the location
-of Earth:
+can also be used to constrain the rate density of Earth analogs.
+Explicitly, we constrain the occurrence rate density of exoplanets orbiting
+``Sun-like'' stars\footnote{In this \paper, we adopt the \citet{petigura}
+sample of G-stars as our definition of ``Sun-like''.}, evaluated at the
+location of Earth:
 \begin{eqnarray}\eqlabel{gammaearth}
@@
 \end{eqnarray}
-That is, \gammaearth\ is the average number of exoplanets around a
-Sun-like star---per natural logarithmic bins in period and radius---evaluated
-at the period and radius of Earth.
-This is also the quantity that \citet{petigura} estimated but in different
-units.
+That is, \gammaearth\ is the rate density of exoplanets around a Sun-like
+star (expected number of planets per star per natural logarithm of period per
+natural logarithm of radius), evaluated at the period and radius of Earth.
 
@@
 
-Since no Earth analogs have been found, any constraints on this rate must be
-extrapolated from the existing observations.
+Since no Earth analogs have been found, any constraints on this density must
+be extrapolated from the existing observations.
 This is generally done by assuming a functional form for the occurrence rate
-function, constraining it using the observed candidates and extrapolating.
+density, constraining it using the observed candidates and extrapolating.
 All published extrapolations are based on rigid models of the occurrence rate
-function (for example, a power law) fit to the catalog and evaluated at the
+density (for example, a power law) fit to the catalog and evaluated at the
 location of Earth (\citealt{catanzarite, traub}).
 \citet{petigura} used their catalog of planet candidates to constrain the rate
-of Earth analogs---using a definition similar to our definition in
-\eq{gammaearth}---assuming an extremely rigid model of the occurrence rate
-function: \emph{flat in logarithmic period}.
+of Earth analogs in a specific period--radius bin assuming an extremely rigid
+model: \emph{flat in logarithmic period}.
 These results are sensitive to the choice of extrapolation function and the
@@
 in \sect{model}.
-Under this model, the occurrence rate in bins where no objects have been
-detected will be constrained---with large uncertainty---by the heights of
-nearby bins.
+Under this model, the occurrence rate density at periods and radii where no
+objects have been detected will be constrained---with large uncertainty---by
+the heights of nearby bins.
 Therefore, even though there are no candidates that qualify as Earth analogs,
-we simply fit our model of the occurrence rate function in a large enough
+we simply fit our model of the occurrence rate density in a large enough
 region of parameter space (including Earth) and compute the posterior
@@
 This works because the Gaussian process regularization actually captures our
-prior beliefs about the shape of the occurrence rate function.
+prior beliefs about the shape of the rate density function.
 This model---and any other extrapolation---will, of course, break down if
-there is an unmeasured sharp feature in the occurrence rate function near the
+there is an unmeasured sharp feature in the occurrence rate density near the
 location of Earth and our method is the most conservative extrapolation
@@
 Their method assumes that, for small planets ($1 \le \radius/\radius_\oplus <
-2$) on long periods ($\period > 50\,\mathrm{days}$), the occurrence rate flat
-in logarithmic period or, equivalently, the cumulative rate is linear.
+2$) on long periods ($\period > 50\,\mathrm{days}$), the occurrence rate
+density is a flat function of logarithmic period or, equivalently, the
+cumulative rate is linear.
 \citet{petigura} used the candidates in their catalog to estimate the slope of
@@
 Instead of defining \gammaearth\ differentially, as we did in \eq{gammaearth},
-\citet{petigura} constrained the integral of the occurrence rate over a box in
+\citet{petigura} constrained the integral of the rate density over a box in
 period and radius ($1 \le \radius/\radius_\oplus < 2$ and $200 \le
 \period/\mathrm{day} < 400$).
-Since their model implicitly assumes a constant occurrence rate across the
-bin, the differential rate is just their number divided by the bin area.
+Since their model implicitly assumes a constant rate density across the bin,
+the differential rate is just their number divided by the bin area.
+This rate density (rate divided by bin volume) is what is shown as a
+comparison to our results in the figures.
 
@@
 In both cases, our method returns a less precise but more accurate result for
-the occurrence rate and the error bars given by the functional extrapolation
+the rate density and the error bars given by the functional extrapolation
 are overly optimistic.
@@
 {\bf Simulated data}.
-The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
+The extrapolated rate density of Earth analogs \gammaearth\ as inferred by the
 different techniques applied to the \modela\ simulation.
@@
 The histogram is the MCMC estimate of our posterior constraint on this rate
-and the true value is indicated as the thick gray vertical line.
+density and the true value is indicated as the thick gray vertical line.
 \figlabel{smooth-rate}}
@@
 
-Having developed this probabilistic framework for occurrence rate inferences
-and demonstrating that it produces reasonable results when applied to
-simulated datasets, we now turn to real data.
+Having developed this probabilistic framework for exoplanet population
+inferences and demonstrating that it produces reasonable results when applied
+to simulated datasets, we now turn to real data.
 As described in \sect{data}, we will use the catalog of small exoplanet
@@
 We directly applied our method to the \citet{petigura} sample and generated
-MCMC samples from the posterior probability for the occurrence rate bin
-heights, marginalizing over the hyperparameters of the Gaussian process model.
+MCMC samples from the posterior probability for the occurrence rate density
+step heights, marginalizing over the hyperparameters of the Gaussian process
+model.
 The resulting MCMC chain is available online at \dfmtodo{ADD SOME URL}.
@@
 \Fig{real-results} shows posterior samples from the inferred occurrence rate
-distribution as a function of period and radius conditioned on the catalog.
+density as a function of period and radius conditioned on the catalog.
 The marginalized distributions are qualitatively consistent with the
-occurrence rate functions measured using the inverse-detection-efficiency
+occurrence rate density measured using the inverse-detection-efficiency
 method with larger uncertainties.
@@
 planets.
-The occurrence rate of large planets appears to monotonically increase as a
+The rate density of large planets appears to monotonically increase as a
 function of log period while the distribution for small planets seems to turn
@@
 Figures~\figref{radius} and~\figref{linear-radius}.
-\Fig{radius} shows the log-radius occurrence rate integrated over various
-logarithmic bins in period.
+\Fig{radius} shows the log-radius occurrence rate density integrated over
+various logarithmic bins in period.
 The distributions in each period bin are qualitatively consistent; the
-rate is dominated by small planets (around two Earth radii) with potential
-``features'' near $\radius\sim3\radius_\oplus$ and $\radius\sim
+rate density is dominated by small planets (around two Earth radii) with
+potential ``features'' near $\radius\sim3\radius_\oplus$ and $\radius\sim
 10\radius_\oplus$.
@@
 linear radius.
-In these coordinates, the occurrence rate in a single bin is no longer
+In these coordinates, the rate density in a single bin is no longer
 uniform; instead, scales as inverse radius.
 
-Our constraint on the rate of Earth analogs (as defined in \sect{extrap}) is
-in tension---even though our result has large fractional uncertainty---with
-the result from \citet{petigura}.
+Our constraint on the rate density of Earth analogs (as defined in
+\sect{extrap}) is in tension---even though our result has large fractional
+uncertainty---with the result from \citet{petigura}.
 This is shown in \fig{real-rate} where we compare the marginalized posterior
 probability function for \gammaearth\ to the published value and uncertainty.
-Quantitatively, we find that the rate of Earth analogs is
+Quantitatively, we find that the rate density of Earth analogs is
 \begin{eqnarray}
@@
 inverse-detection-efficiency procedure, and the histograms are posterior
-samples from the marginalized occurrence rate as inferred by our method.
+samples from the marginalized rate density as inferred by our method.
 \figlabel{real-results}}
@@
 {\bf Real data}.
-The occurrence rate as a function of logarithmic period integrated over bins
-in logarithmic radius.
+The occurrence rate density as a function of logarithmic period integrated
+over bins in logarithmic radius.
 The lines with error bars show the posterior sample median and 68th
@@
 {\bf Real data}.
-The occurrence rate as a function of logarithmic radius integrated over bins
-in logarithmic period.
+The occurrence rate density as a function of logarithmic radius integrated
+over bins in logarithmic period.
 The lines with error bars show the posterior sample median and 68th
@@
 \caption{%
-The same as \fig{smooth-rate} but applied to the catalog from \citet{petigura}.
+The extrapolated rate density of Earth analogs \gammaearth\ (the same as
+\fig{smooth-rate} but applied to the catalog from \citealt{petigura}).
+The histogram is the MCMC estimate of our posterior constraint on this rate
+density.
+The vertical black line with error bars shown as dashed lines is the result
+from \citet{petigura} converted to a rate density by dividing by their bin
+volume.
 \figlabel{real-rate}}

### 2014-06-09 document/ms.tex
integrated rates

@@
     \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
+\newcommand{\densityunit}{{\ensuremath{\mathrm{nat}^{-2}}}}
 
@@
 \begin{eqnarray}
-\gammaearth &=& 0.017^{+0.018}_{-0.009} \quad.
-\end{eqnarray}
-Converted to these units, \citet{petigura} measured $0.119_{-0.046}^{+0.035}$
-for the same quantity (indicated as the vertical lines in \fig{real-rate}).
+\gammaearth &=& 0.017^{+0.018}_{-0.009}~\densityunit
+\end{eqnarray}
+where the ``\densityunit'' indicates that this quantity is a rate density, per
+natural logarithmic period per natural logarithmic radius.
+Converted to these units, \citet{petigura} measured
+$0.119_{-0.046}^{+0.035}~\densityunit$ for the same quantity (indicated as the
+vertical lines in \fig{real-rate}).
+This rate density is \emph{exactly} what Petigura's extrapolation model
+predicts but, for comparison, we can also integrate our inferred rate density
+over their choice of ``Earth-like'' bin ($200 \le \period/\mathrm{day} < 400$
+and $1 \le \radius/\radius_\oplus < 2$) to find a \emph{rate} of
+Earth-analogs.
+The published rate is $0.057_{-0.017}^{+0.022}$ (\citealt{petigura}) and our
+posterior constraint is
+\begin{eqnarray}
+\int_{\period=200\,\mathrm{day}}^{400\,\mathrm{day}}
+\int_{\radius=1\,\radius_\oplus}^{2\,\radius_\oplus}
+\rate_\ratepars (\ln\period,\,\ln\radius)
+\dd[\ln\radius]
+\dd[\ln\period]
+&=&
+0.016_{-0.007}^{+0.011}
+\quad.
+\end{eqnarray}
 

### 2014-06-09 document/figures/results/make_plots.py

@@ -14,6 +14,15 @@ with h5py.File("samples.h5", "r") as f:
     ln_period_bin_edges = f["ln_period_bin_edges"][...]
     ln_radius_bin_edges = f["ln_radius_bin_edges"][...]
 
+# Integrate the rate over EP's bin.
+ep_inds = (slice(None), 5, slice(2, 4))
+ep_area = (np.diff(ln_period_bin_edges)[ep_inds[1]]
+           * np.diff(ln_radius_bin_edges)[ep_inds[2]])
+rate = np.sum(ep_area * np.exp(samples[ep_inds] - np.log(42557.0)), axis=1)
+q = quantile(rate, [0.16, 0.5, 0.84])
+print("Integrated rate: {0:.3f}_{{-{1:.3f}}}^{{+{2:.3f}}}"
+      .format(q[1], *(np.diff(q))))
+
 bins = [ln_period_bin_edges, ln_radius_bin_edges]
 labels = ["R / R_\oplus", "P / \mathrm{day}"]
 slices = [
@@ -108,8 +117,8 @@ def plot_results(a):
     ax_right.legend(loc="center left", bbox_to_anchor=(1, 0.5), prop=prop)
 
     if a == 1:
-        ax_right2.set_xlim(0, 8)
-        ax_right2.set_ylim(-7.5, 0.2)
+        ax_right2.set_xlim(0, 16)
+        ax_right2.set_ylim(-8.9, 0.2)
 
     ax_right2.set_yticklabels([])
     ax2.set_ylim(np.exp(ax_right2.get_ylim()))

### 2014-06-09 document/ms.tex
mentioning that our results are different

@@
 Quantitatively, we find that the rate density of Earth analogs is
-\begin{eqnarray}
+\begin{eqnarray}\eqlabel{ge-result}
 \gammaearth &=& 0.017^{+0.018}_{-0.009}~\densityunit
@@
 and $1 \le \radius/\radius_\oplus < 2$) to find a \emph{rate} of
-Earth-analogs.
+Earth analogs.
 The published rate is $0.057_{-0.017}^{+0.022}$ (\citealt{petigura}) and our
@@
 \end{eqnarray}
+
+Our inferred rate density of Earth analogs (\eqalt{ge-result}) is not
+consistent with the previously published result even though we use
+\emph{exactly the same dataset}.
+This inconsistency is due to the different assumptions made.
+In particular, for their analysis, \citet{petigura} asserted that the period
+distribution for small planets on long orbits is flat in logarithmic period.
+If we enforce the same constraint, we reproduce their result but we relax this
+assumption, allowing the data to constrain the period distribution, we find
+evidence for a \emph{decreasing} period distribution as a function of
+logarithmic period (see \fig{period}).
 

### 2014-06-09 document/code/results.py
saving samples

@@ -18,6 +18,8 @@ import numpy as np
 import cPickle as pickle
 import matplotlib.pyplot as pl
 
+from emcee.autocorr import integrated_time
+
 import load_data
 
 bp = sys.argv[1]
@@ -57,12 +59,29 @@ pl.clf()
 pl.plot(lnprob)
 pl.savefig(os.path.join(bp, "time-lnprob.png"))
 
-samples = samples[-200000:, :]  # [::50, :]
+nstar = 42557.0
+ntot = 200000
+samples = samples[-ntot:, :]  # [::50, :]
+
+# Reformat the samples and save the samples.
+thin_by = int(np.min(integrated_time(samples, axis=0)))
+thinned = samples[::thin_by, :]
+grids = thinned.reshape((len(thinned), pop.shape[0], pop.shape[1]))
+print(grids.shape)
+print([b.shape for b in pop.bins])
+
+with h5py.File(os.path.join(bp, "samples.h5"), "w") as f:
+    f.create_dataset("hyperparameter_samples", data=hyper[-ntot:][::thin_by])
+    f.create_dataset("ln_occurrence_rate_samples", data=grids)
+    f.create_dataset("ln_period_bin_edges", data=pop.bins[0])
+    f.create_dataset("ln_radius_bin_edges", data=pop.bins[1])
+
+assert 0
 
 # Load the true gamma earth if it exists.
 fn = os.path.join(bp, "gamma.txt")
 if os.path.exists(fn):
-    gamma = [np.exp(float(open(fn).read())) / 42557.0]
+    gamma = [np.exp(float(open(fn).read())) / nstar]
 else:
     gamma = None  # [5.7 / 100, 1.7 / 100, 2.2 / 100]
 

### 2014-06-09 document/ms.tex

@@
 
+We have developed a hierarchical probabilistic framework for inferring the
+population of exoplanets based on noisy incomplete catalogs.
+This method incorporates systematic treatment of observational uncertainties
+and detection efficiency.
+One major benefit of this framework is that it provides the best possible
+constraints on the population under the stated assumptions.
+After demonstrating the validity of our method on two qualitatively different
+synthetic exoplanet catalogs, we run our inference on a published catalog of
+small exoplanet candidates orbiting Sun-like stars (\citealt{petigura}) to
+determine the occurrence rate density of exoplanets as a function of period
+and radius.
+We extrapolate this measurement to the location of Earth and constrain the
+rate density of Earth analogs with large error bars.
+
 Probabilistic framework justified by MATHS.

### 2014-06-10 document/ms.tex
pair writing with Hogg

@@
 \newcommand{\sect}[1]{\Sect{#1}}
+\newcommand{\sectalt}[1]{\ref{sect:#1}}
 \newcommand{\App}[1]{Appendix~\ref{sect:#1}}
@@
 \title{%
-    Exoplanet population inference and the rate of Earth analogs \\
-    from noisy, incomplete catalogs
+    Exoplanet population inference and the abundance \\
+    of Earth analogs from noisy, incomplete catalogs
 }
@@
 
-Many exoplanets have been discovered and characterized; it is now possible to
-study the population as a whole.
-Although no true extra-Solar Earth analog is known, hundreds of planets have
-been found around Sun-like stars that are either Earth-sized but on shorter
-periods, or else on year-long orbits but somewhat larger.
-Under strong assumptions, these exoplanet catalogs have been used to make an
+No true extra-Solar Earth analog is known.
+Hundreds of planets have been found around Sun-like stars that are either
+Earth-sized but on shorter periods, or else on year-long orbits but somewhat
+larger.
+Under strong assumptions, exoplanet catalogs have been used to make an
 extrapolated estimate of the rate at which Sun-like stars host Earth analogs.
@@
 non-trivial selection effects and detection efficiencies, and every property
-(period, radius, \etc) is measured noisily.
+(period, radius, \etc)\ is measured noisily.
 Here we present a general hierarchical probabilistic framework for making
@@
 Observational constraints on the population can inform theories of planet
-formation (\dfmtodo{CITE}) and place probabilistic bounds on the abundance of
-Earth analogs\footnote{For our purposes, an ``Earth analog'' is an Earth-sized
+formation and place probabilistic bounds on the abundance of Earth
+analogs\footnote{For our purposes, an ``Earth analog'' is an Earth-sized
 exoplanet orbiting a Sun-like star with a year-long period.}.
 
-\citet{petigura} recently published constraints on this function based on
-their independent analysis of the \kepler\ light curves for 42,000 Sun-like
+\citet{petigura} recently published an exoplanet population analysis based on
+an independent analysis of the \kepler\ light curves for 42,000 Sun-like
 stars.
@@
 Even though no Earth analogs were discovered in their search, \citet{petigura}
-used the small candidates that the did find to place an extrapolated
+used the small candidates that they did find to place an extrapolated
 constraint on the frequency of Earth-like exoplanets, assuming a flat
@@
 This effect is degenerate with any inferences about the rate density and it
-can be hard to constraint quantitatively.
+can be hard to constrain quantitatively.
 In practice, there are three methods for taking this effect into account:
@@
 
-There are two pedagogically different methods that are commonly used to infer
+There are two qualitatively different methods that are commonly used to infer
 the occurrence rate density from a catalog and a detection efficiency
@@
 The first is an intuitive method that we will refer to as
-inverse-detection-efficiency procedure and the second is based on the
-likelihood function of the catalog given a parametric rate density.
+``inverse-detection-efficiency'' and the second is based on the likelihood
+function of the catalog given a parametric rate density.
 The inverse-detection-efficiency method involves making a histogram of the
@@
 dressing, swift, petigura}) even though it is not motivated probabilistically.
-The alternative ``likelihood'' method models the catalog as a Poisson
-realization of the \emph{observable} rate density of exoplanets taking the
-survey detection efficiencies and transit probabilities into account.
+The alternative likelihood method models the catalog as a Poisson realization
+of the \emph{observable} rate density of exoplanets taking the survey
+detection efficiencies and transit probabilities into account.
 This technique has been used to constrain parametric models---a broken power
@@
 with \emph{non-negligible observational uncertainties}.
-We generalize the inference method introduced by \citet{hogge} to account for
-survey detection efficiencies.
+Our method takes the form of a hierarchical probabilistic (Bayesian)
+inference.
+We generalize the method introduced by \citet{hogge} to account for survey
+detection efficiencies.
 We run tests on simulated datasets---comparing results with the standard
@@
 multiple transiting systems into account (\citealt{lissauer, tremaine, fang})
-and this would be possible using a conditional distributions (as developed by
+and this would be possible using conditional distributions (as developed by
 \citealt{tremaine}).
@@
 For this \paper, we will neglect this issue and only comment on the effects
-but the prior distributions published by \citet{fressin-fp} would be directly
-applicable in our method and this would be a very interesting generalization.
+but the prior distributions published by \citet{fressin-fp} could be directly
+applied in a generalization of our method.
 
@@
 
-Conditioned on our assumptions and the choices made in planet detection,
+Conditioned on our assumptions and the choices made in the planet detection,
 vetting and characterization pipeline (\citealt{petigura-a, petigura}), we
@@
 density'' to indicate a quantity that must be integrated over a finite bin in
-period and radius to deliver a rate.} of Earth analogs \gammaearth\ which we
+period and radius to deliver a rate.} of Earth analogs \gammaearth, which we
 define as \emph{the expected number of planets per star per natural
@@
 We relax this assumption and extrapolate only by assuming that the occurrence
-rate density is a smooth function of period and radius; we find a much lower
-value for \gammaearth.
+rate density is a smooth function of period and radius; we find lower values
+for \gammaearth.
 We enforce the smoothness constraint by applying a flexible Gaussian process
@@
 inverse-detection-efficiency weighting procedure.
-After testing our method on synthetic catalogs in \sect{valid}, we use the
-catalog of planet candidates and the empirically determined detection
-efficiency from \citet{petigura} to measure the occurrence rate density of
-Earth analogs.
+We test our method on synthetic catalogs in Sections~\sectalt{valid}
+and~\sectalt{extrap}.
+In \sect{real}, we use the catalog of planet candidates and the empirically
+determined detection efficiency from \citet{petigura} to measure the
+occurrence rate density of Earth analogs.
 
@@
 where the bin heights are the parameters \ratepars.
-The light curve for a particular target $k$ is called $\data_k$ and the
-\True\ set of physical parameters for that same object are given by
-$\entry_k$.
+The light curve for a particular target $k$ is $\data_k$ and the \True\ set of
+physical parameters for that same object is given by $\entry_k$.
 A catalog provides a set of constraints on the parameters $\{\entry_k\}$
@@
 
-In other words, a catalog is a dimensionality reduction of the raw data with
-all the relevant information retained.
+In other words, we treat a catalog as being a dimensionality reduction of the
+raw data with all the relevant information retained.
 In the context of \kepler, the catalog reduces the set of downloaded pixel
@@
 The constraints in \eq{crazypost} can always be---and often are---propagated
-as a list of $N$ samples $\{\entry_k^{(n)}\}$ from the posterior
+as a list of $N$ samples $\{\entry_k\}^{(n)}$ from the posterior
 \begin{eqnarray}
@@
     \frac{1}{N} \sum_{n=1}^N
-    \frac{p(\{\entry_k^{(n)}\}\,|\,\ratepars)}
-         {p(\{\entry_k^{(n)}\}\,|\,\interim)} \quad.
+    \frac{p(\{\entry_k\}^{(n)}\,|\,\ratepars)}
+         {p(\{\entry_k\}^{(n)}\,|\,\interim)} \quad.
 \end{eqnarray}
 This is very efficient to compute as long as an evaluation of
-$p(\{\entry_k^{(n)}\}\,|\,\ratepars)$ is not very expensive.
-That being said, this could be a high variance estimator depending on the
-initial choice of $p(\{\entry_k^{(n)}\}\,|\,\interim)$ but in our case, we'll
-show that not many samples are needed to get an accurate estimate.
+$p(\{\entry_k\}^{(n)}\,|\,\ratepars)$ is not very expensive.
+That being said, this could be a high variance estimator, depending on the
+initial choice of $p(\{\entry_k\}\,|\,\interim)$.
+Additionally, the support of $p(\{\entry_k\}\,|\,\ratepars)$ in $\{\entry_k\}$
+space is restricted to be narrower than that of
+$p(\{\entry_k\}\,|\,\interim)$.
 \Eq{importance} is the \emph{importance sampling approximation} to the
@@
 the validity of this method.
-Instead, we could use a functional form for it and infer the parameters
-(\dfmtodo{cite all the papers that do the ramp, step function, whatever}).
+Instead, we could use a functional form for it and even infer its parameters
+along with the parameters of the rate density.
 
@@
 probabilistic justification or interpretation.
-\hoggtodo{Find the papers that Tremaine suggested that we look at re:vmax and
-discuss the motivation in those cases.}
 
@@
 rate density inference, the inverse-detection-efficiency procedure will
-give a higher variance estimate of the occurrence rate density.
+give a higher variance estimate of the occurrence rate density than
+\eq{ml-ide}.
+In this \paper, we don't use either of these inverse-detection-efficiency
+procedures because we are righteous.
 
@@
 \section{Results from real data}
+\sectlabel{real}
 
@@
 One major benefit of this framework is that it provides the best possible
-constraints on the population under the stated assumptions.
+probabilistic measurements of the population under the stated assumptions.
 After demonstrating the validity of our method on two qualitatively different
@@
 rate density of Earth analogs with large error bars.
-
-Probabilistic framework justified by MATHS.
-
-Disagrees in detail with Petigura esp.\ extrapolation.
-
-Agrees qualitatively with Petigura, Dong, Howard, \etc\ but hard to make
-quantitative comparisons to other ones (pipeline, stars, \etc).
-
-Why did we use the definition of \gammaearth?
-
-Our constraints are less confident---in a signal-to-noise sense---than
-previous measurements (made using stronger assumptions; \citealt{petigura})
-but, even so, our result is
-
-\dfmtodo{%
-Predicted number of observable Earth analogs in the \kepler\ dataset.
-Low but not zero.
-}
+In order to perform this extrapolation, we don't assume a specific functional
+form for the rate density instead, we only assume that it is a smooth function
+of logarithmic period and radius.
+
+The occurrence rate density function that we infer is qualitatively consistent
+with previously published results using different inference techniques
+(\citealt{dong, fressin-fp, petigura}).
+In particular, we find (see \fig{radius}) previously recorded features in the
+radius distribution around $\radius\sim 3\,\radius_\oplus$ and $\radius\sim
+10\,\radius_\oplus$, although not at high signal-to-noise.
+We find that the period distributions for planets in different radius bins are
+different, in qualitative agreement with previous results (\citealt{dong}).
+\Fig{period} shows that larger planets tend to be on longer periods than
+smaller planets.
+
+In detail, any inference of the exoplanet population depends on the choices
+made in the pipeline that creates the catalog.
+Our results are based on a catalog that only includes single transiting
+systems (\citealt{petigura}) so our conclusions---and theirs---must be
+considered with this in mind.
+In particular, the exclusion of second planets must suppress the inferred rate
+density, especially at long periods and small radii.
+
+Our extrapolation of the rate density to the location of Earth is more general
+and conservative than any previously published method.
+Using this method, we find a rate density of Earth analogs that is
+inconsistent with the result published by \citet{petigura}.
+This discrepancy can be largely attributed to the rigidity of the assumptions
+about the period distribution.
+Our extrapolation is less confident than previous measurements.
+Again, this difference is due to the fact that we allow a much more flexible
+extrapolation function.
+\dfmtodo{This is another illustration that the use of flexible models makes
+probabilistic data analysis more conservative.}
+
+In contrast to previous work (\dfmtodo{CITE?}), we don't define ``Earth
+analog'' in terms of habitability or composition.
+Instead, we advocate for a definition in terms of more directly observable
+quantities (in this case, period and radius).
+Furthermore, we define \gammaearth\ as a rate density (per star per
+logarithmic period per logarithmic radius) so that its value doesn't depend on
+choices about the ``Earth-like'' bin.
 
 In our analysis we make a few simplifying assumptions.
-Each of these could be relaxed and \dfmtodo{BLAH BLAH BLAH}.
 \begin{itemize}
@@
 \item {\bf False positives}\quad
+Makes our estimates overestimates.
+Go to a mixture model.
 
 \item {\bf Known observational uncertainties}\quad
+Assumed Gaussian. Would be better with samples.
 
 \item {\bf Given empirical detection efficiency}\quad
+Single vs multi.
+In detail there are limitations: radius vs ratio. Stellar params.
 
 \item {\bf Smooth rate function}\quad
+No reason for the function to be smooth.
+Phase transitions.
+Composition etc.
 
 \end{itemize}
-
-Under these assumptions and the dataset, \etc, we stand by this analysis but
-really this is a model of the planets in Kepler convolved with any latent
-effects of the TERRA pipeline\ldots
-
-A major effect that isn't taken into account is the highest signal-to-noise
-blah blah blah\ldots our number is prolly low.
+Our assumptions are severe but we believe that this is the most conservative
+population inference method currently on the market.
+
+\dfmtodo{%
+Predicted number of observable Earth analogs in the \kepler\ dataset.
+Low but not zero.
+}
 
@@
 for helpful contributions to the ideas and code presented here.
-This project was partially supported by the NSF (grant AST-0908357), and NASA
-(grant NNX08AJ48G).
-This research builds on collaborations between astronomers and statisticians
-forged during a three week workshop on ``Modern Statistical and Computational
-Methods for Analysis of Kepler Data'' at SAMSI in June 2013.
+This project was partially supported by the NSF (grant AST-0908357), NASA
+(grant NNX08AJ48G), and the Moore--Sloan Data Science Environment at NYU.
+This research builds on ideas generated at a three-week workshop on ``Modern
+Statistical and Computational Methods for Analysis of Kepler Data'' at SAMSI
+in 2013 June.
 This research made use of the NASA \project{Astrophysics Data System}.

### 2014-06-10 document/ms.tex
discussing assumptions

@@
 as a list of $N$ samples $\{\entry_k\}^{(n)}$ from the posterior
-\begin{eqnarray}
+\begin{eqnarray}\eqlabel{samples}
 \{\entry_k\}^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
@@
 
+\citet{petigura} did not publish or share posterior samples of their
+measurements of the physical parameter (\eqalt{samples}).
+They did publish a list of periods, radii and radius uncertainties based on
+their analysis.
+Assuming that there is no measurement uncertainty on the period measurement
+and that the radius posterior is Gaussian in linear radius (with a standard
+deviation given by the published uncertainty), we draw 256 samples for
+$\entry_k$ and use these as an approximation to the posterior probability
+function.
+
 A huge benefit of this dataset is that Erik Petigura and collaborators
@@
 In the future, this problem would be solved by publishing a representation of
-\emph{the full posterior density functino for each object in the catalog}.
+\emph{the full posterior density function for each object in the catalog}.
 In this case, the most useful data product would be \emph{posterior samples
@@
 
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
-\end{center}
-\caption{%
-{\bf Simulated data}.
-Inferences about the rate density based on the simulated catalog \modela.
-\emph{Center:} the points with error bars show the exoplanet candidates in the
-simulated incomplete catalog, the contours show the survey completeness
-function (\citealt{petigura}), and the grayscale shows the median posterior
-occurrence surface.
-\emph{Top and left:} the red dashed line shows the true distribution that was
-used to generate the catalog, the points with error bars show the results of
-the inverse-detection-efficiency procedure, and the histograms are posterior
-samples from the marginalized rate density as inferred by our method.
-\figlabel{smooth-results}}
-\end{figure}
-
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
-\end{center}
-\caption{%
-{\bf Simulated data}.
-The same as \fig{smooth-results} for \modelb.
-\figlabel{simulation-results}}
-\end{figure}
-
 \section{Extrapolation to Earth}
@@
 be.
-
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
-\end{center}
-\caption{%
-{\bf Simulated data}.
-The extrapolated rate density of Earth analogs \gammaearth\ as inferred by the
-different techniques applied to the \modela\ simulation.
-Applying the method used by \citet{petigura} gives a constraint indicated by
-the vertical black line with error bars shown as dashed lines.
-The histogram is the MCMC estimate of our posterior constraint on this rate
-density and the true value is indicated as the thick gray vertical line.
-\figlabel{smooth-rate}}
-\end{figure}
-
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
-\end{center}
-\caption{%
-{\bf Simulated data}.
-The same as \fig{smooth-rate} for \modelb.
-\figlabel{simulation-rate}}
-\end{figure}
 
@@
 
+\section{Discussion}
+
+We have developed a hierarchical probabilistic framework for inferring the
+population of exoplanets based on noisy incomplete catalogs.
+This method incorporates systematic treatment of observational uncertainties
+and detection efficiency.
+One major benefit of this framework is that it provides the best possible
+probabilistic measurements of the population under the stated assumptions.
+After demonstrating the validity of our method on two qualitatively different
+synthetic exoplanet catalogs, we run our inference on a published catalog of
+small exoplanet candidates orbiting Sun-like stars (\citealt{petigura}) to
+determine the occurrence rate density of exoplanets as a function of period
+and radius.
+We extrapolate this measurement to the location of Earth and constrain the
+rate density of Earth analogs with large error bars.
+In order to perform this extrapolation, we don't assume a specific functional
+form for the rate density instead, we only assume that it is a smooth function
+of logarithmic period and radius.
+
+The occurrence rate density function that we infer is qualitatively consistent
+with previously published results using different inference techniques
+(\citealt{dong, fressin-fp, petigura}).
+In particular, we find (see \fig{radius}) previously recorded features in the
+radius distribution around $\radius\sim 3\,\radius_\oplus$ and $\radius\sim
+10\,\radius_\oplus$, although not at high signal-to-noise.
+We find that the period distributions for planets in different radius bins are
+different, in qualitative agreement with previous results (\citealt{dong}).
+\Fig{period} shows that larger planets tend to be on longer periods than
+smaller planets.
+
+In detail, any inference of the exoplanet population depends on the choices
+made in the pipeline that creates the catalog.
+Our results are based on a catalog that only includes single transiting
+systems (\citealt{petigura}) so our conclusions---and theirs---must be
+considered with this in mind.
+In particular, the exclusion of second planets must suppress the inferred rate
+density, especially at long periods and small radii.
+
+Our extrapolation of the rate density to the location of Earth is more general
+and conservative than any previously published method.
+Using this method, we find a rate density of Earth analogs that is
+inconsistent with the result published by \citet{petigura}.
+This discrepancy can be largely attributed to the rigidity of the assumptions
+about the period distribution.
+Our extrapolation is less confident than previous measurements.
+Again, this difference is due to the fact that we allow a much more flexible
+extrapolation function.
+\dfmtodo{This is another illustration that the use of flexible models makes
+probabilistic data analysis more conservative.}
+
+In contrast to previous work, we don't define ``Earth analog'' in terms of
+habitability or composition.
+Instead, we advocate for a definition in terms of more directly observable
+quantities (in this case, period and radius).
+Furthermore, we define \gammaearth\ as a rate density (per star per
+logarithmic period per logarithmic radius) so that its value doesn't depend on
+choices about the ``Earth-like'' bin.
+
+In our analysis we make a few simplifying assumptions.
+Every assumption has an effect on the results and could be relaxed as an
+extension of this project.
+For completeness, we list and discuss the effects of our assumptions below.
+\begin{itemize}
+
+\item {\bf Conditional independence}\quad
+We assume that every object in the catalog was a conditionally independent
+draw from the observable occurrence rate density.
+This is probably a bad assumption when applying this method to a different
+catalog where multiple transiting systems are included.
+In practice, the best first step towards relaxing this assumption is probably
+to follow \citet{tremaine} and assume that the mutual inclination distribution
+is the only source of conditional dependence between planets.
+For this \paper, the assumption of conditional independence is justified
+because the dataset explicitly includes only systems with a single transiting
+exoplanet.
+
+\item {\bf False positives}\quad
+In our inferences, we assume that all of the candidates in the catalog are
+\True\ exoplanets.
+The rate of false positives in the \kepler\ catalog has been shown to be low
+but not negligible (\citealt{morton, fressin-fp}).
+Since some of the objects in the catalog are probably false positives, our
+inferences about the occurrence rate density are biased high but without
+explicitly including a model of false positives, it's hard to say in detail
+what effect this would have on the distributions.
+In an extension of this work, we could incorporate the effects of false
+positives by switching to a mixture model where each object is modeled as a
+mixture of \True\ exoplanet and false positive (\citealt{hogg-line}).
+
+\item {\bf Known observational uncertainties}\quad
+To apply the importance sampling approximation to the published catalog, we
+assume that the measurement uncertainties are known and, in this case,
+Gaussian.
+The assumption of normally distributed uncertainties could be relaxed given
+a sampling representation of the posterior probability function for the
+physical parameters (period, radius, \etc).
+It is also possible to relax the noise model---to include effects like the
+underestimation of uncertainties---but inference would be substantially more
+computationally expensive.
+
+\item {\bf Given empirical detection efficiency}\quad
+\citet{petigura} determined the end-to-end detection efficiency of their
+planet detection pipeline as a function of \True\ period and radius by
+injecting synthetic signals into real light curves and testing recovery.
+We used these simulations as an exact representation of the detection
+efficiency of the catalog but there are several missing components.
+The biggest effect is probably the fact that this formulation doesn't include
+the fact that only the \emph{most detectable signal} in each light curve was
+included in the catalog.
+This bias will have the largest effect in the parts of parameter space where
+the baseline detection efficiency is lowest: at long periods and small radius.
+As a result, our inferences (and the results from \citealt{petigura}) about
+the occurrence rate of small planets on long periods should be seen as
+\emph{lower limits}.
+In detail there is another limitation due to the fact that the stellar
+parameters are only known noisily and the transit light curve only constrains
+the radius ratio.
+This means that the marginalized detection efficiency should be measured as a
+function of radius ratio and the interpretation in terms of \True\ radius is
+only approximately correct.
+Given the size of the dataset and the number of injection simulations, this
+effect should be small.
+
+\item {\bf Smooth rate function}\quad
+Throughout our analysis, we make the prior assumption that the occurrence rate
+density is a smooth function of logarithmic period and radius.
+This model is useful because it allows us to make probabilistically justified
+inferences about the exoplanet population in regions of parameter space with
+low detection efficiency.
+The assumption that the rate density should be smooth is intuitive but there
+is no theoretical indication that it must be true at all scales.
+That being said, the Gaussian process regularization that we use to enforce
+smoothness is flexible enough to capture substantial departures from smooth if
+they were supported by the data.
+
+\end{itemize}
+Our assumptions are severe but we believe that this is the most conservative
+population inference method currently on the market.
+
+\dfmtodo{%
+Predicted number of observable Earth analogs in the \kepler\ dataset.
+Low but not zero.
+}
+
+\dfmtodo{%
+{\bf Advice:} \quad
+Remind us that we want a representation of the full posterior (samples)
+because that would let us do better things with the completeness (modeled as
+r/R, \etc).
+Must explicitly include priors too.
+}
+
+All of the code used in this project is available from
+\url{http://github.com/dfm/exopop} under the MIT open-source software license.
+This code (plus some dependencies) can be run to re-generate all of the
+figures and results in this \paper; this version of the paper was generated
+with git commit \texttt{\githash} (\gitdate).
+
+\acknowledgments
+We would like to thank Erik Petigura (Berkeley) for freely sharing his data
+and code.
+It is a pleasure to thank
+Ruth Angus (Oxford),
+Jo Bovy (IAS), and
+Eric Ford (PSU)
+for helpful contributions to the ideas and code presented here.
+This project was partially supported by the NSF (grant AST-0908357), NASA
+(grant NNX08AJ48G), and the Moore--Sloan Data Science Environment at NYU.
+This research builds on ideas generated at a three-week workshop on ``Modern
+Statistical and Computational Methods for Analysis of Kepler Data'' at SAMSI
+in 2013 June.
+This research made use of the NASA \project{Astrophysics Data System}.
+
+\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}
+\begin{thebibliography}{}\raggedright
+
+\bibitem[Adams, Murray \& MacKay(2009)]{poiss-gp}
+Adams, R.~P., Murray, I., \& MacKay, D.~J.~C.\ 2009, ICML, 2009, 9
+(\href{http://homepages.inf.ed.ac.uk/imurray2/pub/09poisson/}{online})
+
+\bibitem[Ambikasaran \etal(2014)]{dfm-gp}
+Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D.~W.,
+\& O'Neil, M.\ 2014, \arxiv{1403.6015}
+
+\bibitem[Batalha \etal(2013)]{kepler-catalog}
+Batalha, N.~M., Rowe, J.~F., Bryson, S.~T., et al.\ 2013, \apjs, 204, 24
+(\arxiv{1202.5852})
+
+\bibitem[Brewer \& Stello(2009)]{brewer}
+Brewer, B.~J., \& Stello, D.\ 2009, \mnras, 395, 2226 (\arxiv{0902.3907})
+
+\bibitem[Carter \& Winn(2009)]{carter}
+Carter, J.~A., \& Winn, J.~N.\ 2009, \apj, 704, 51 (\arxiv{0909.0747})
+
+\bibitem[Catanzarite \& Shao(2011)]{catanzarite}
+Catanzarite, J., \& Shao, M.\ 2011, \apj, 738, 151 (\arxiv{1103.1443})
+
+\bibitem[Dong \& Zhu(2013)]{dong}
+Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 (\arxiv{1212.4853})
+
+\bibitem[Dressing  \& Charbonneau(2013)]{dressing}
+Dressing, C.~D., \& Charbonneau, D.\ 2013, \apj, 767, 95 (\arxiv{1302.1647})
+
+\bibitem[Fang \& Margot(2012)]{fang}
+Fang, J., \& Margot, J.-L.\ 2012, \apj, 761, 92 (\arxiv{1207.5250})
+
+\bibitem[Foreman-Mackey \etal(2012)]{emcee}
+Foreman-Mackey, D., Hogg, D.~W., Lang, D., \& Goodman, J.\ 2013, \pasp, 125,
+306 (\arxiv{1202.3665})
+
+\bibitem[Fressin \etal(2013)]{fressin-fp}
+Fressin, F., Torres, G., Charbonneau, D., \etal\ 2013, \apj, 766, 81
+(\arxiv{1301.0842})
+
+\bibitem[Gibson \etal(2012)]{gibson-gp}
+Gibson, N.~P., Aigrain, S., Roberts, S., \etal\ 2012, \mnras, 419, 2683
+(\arxiv{1109.3251})
+
+\bibitem[Hogg \etal(2013)]{hoggwhitepaper}
+Hogg, D.~W., Angus, R., Barclay, T., et al.\ 2013, \arxiv{1309.0653}
+
+\bibitem[Hogg \etal(2010a)]{hogg-line}
+Hogg, D.~W., Bovy, J., \& Lang, D.\ 2010a, \arxiv{1008.4686}
+
+\bibitem[Hogg \etal(2010b)]{hogge}
+Hogg, D.~W., Myers, A.~D., \& Bovy, J.\ 2010b, \apj, 725, 2166
+(\arxiv{1008.4146})
+
+\bibitem[Howard \etal(2012)]{howard}
+Howard, A.~W., Marcy, G.~W., Bryson, S.~T., et al.\ 2012, \apjs, 201, 15
+(\arxiv{1103.2541})
+
+\bibitem[Lewis \& Shedler(1979)]{poisson}
+Lewis, P.~A.~W., \& Shedler, G.~S.\ 1979, Naval Research Logistics Quarterly,
+26, 403
+
+\bibitem[Lissauer \etal(2011)]{lissauer}
+Lissauer, J.~J., Ragozzine, D., Fabrycky, D.~C., et al.\ 2011, \apjs, 197, 8
+(\arxiv{1102.0543})
+
+\bibitem[Morton \& Johnson(2011)]{morton}
+Morton, T.~D., \& Johnson, J.~A.\ 2011, \apj, 738, 170 (\arxiv{1101.5630})
+
+\bibitem[Morton \& Swift(2013)]{morton-swift}
+Morton, T.~D., \& Swift, J.~J.\ 2013, \arxiv{1303.3013}
+
+\bibitem[Murray \etal(2010)]{ess}
+Murray, I., Prescott Adams, R., \& MacKay, D.~J.~C.\ 2010, JMLR: W\&CP, 9, 541
+(\arxiv{1001.0175})
+
+\bibitem[Murray \& Prescott Adams(2010)]{ess-hyper}
+Murray, I., \& Prescott Adams, R.\ 2010, Advances in Neural Information
+Processing Systems, 23, 1723
+(\arxiv{1006.0868})
+
+\bibitem[Petigura \etal(2013a)]{petigura-a}
+Petigura, E.~A., Marcy, G.~W., \& Howard, A.~W.\ 2013a, \apj, 770, 69
+(\arxiv{1304.0460})
+
+\bibitem[Petigura \etal(2013b)]{petigura}
+Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013b,
+Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})
+
+\bibitem[Rasmussen \& Williams(2006)]{gp}
+Rasmussen, C.~E. \& Williams, C.~K.~I.\ 2006
+Gaussian Processes for Machine Learning, MIT Press
+(\href{http://www.gaussianprocess.org/gpml/}{online})
+
+\bibitem[Roberts \etal(2013)]{roberts}
+Roberts, S., McQuillan, A., Reece, S., \& Aigrain, S.\ 2013, \mnras, 435, 3639
+(\arxiv{1308.3644})
+
+\bibitem[Swift \etal(2013)]{swift}
+Swift, J.~J., Johnson, J.~A., Morton, T.~D., et al.\ 2013, \apj, 764, 105
+(\arxiv{1301.0023})
+
+\bibitem[Tabachnik \& Tremaine(2002)]{tabachnik}
+Tabachnik, S., \& Tremaine, S.\ 2002, \mnras, 335, 151
+(\arxiv{astro-ph/0107482})
+
+\bibitem[Traub(2012)]{traub}
+Traub, W.~A.\ 2012, \apj, 745, 20 (\arxiv{1109.4682})
+
+\bibitem[Tremaine \& Dong(2012)]{tremaine}
+Tremaine, S., \& Dong, S.\ 2012, \aj, 143, 94 (\arxiv{1106.5403})
+
+\bibitem[Youdin(2011)]{youdin}
+Youdin, A.~N.\ 2011, \apj, 742, 38 (\arxiv{1105.1782})
+
+\end{thebibliography}
+
+\clearpage
+
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
+\end{center}
+\caption{%
+{\bf Simulated data}.
+Inferences about the rate density based on the simulated catalog \modela.
+\emph{Center:} the points with error bars show the exoplanet candidates in the
+simulated incomplete catalog, the contours show the survey completeness
+function (\citealt{petigura}), and the grayscale shows the median posterior
+occurrence surface.
+\emph{Top and left:} the red dashed line shows the true distribution that was
+used to generate the catalog, the points with error bars show the results of
+the inverse-detection-efficiency procedure, and the histograms are posterior
+samples from the marginalized rate density as inferred by our method.
+\figlabel{smooth-results}}
+\end{figure}
+
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
+\end{center}
+\caption{%
+{\bf Simulated data}.
+The same as \fig{smooth-results} for \modelb.
+\figlabel{simulation-results}}
+\end{figure}
+
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
+\end{center}
+\caption{%
+{\bf Simulated data}.
+The extrapolated rate density of Earth analogs \gammaearth\ as inferred by the
+different techniques applied to the \modela\ simulation.
+Applying the method used by \citet{petigura} gives a constraint indicated by
+the vertical black line with error bars shown as dashed lines.
+The histogram is the MCMC estimate of our posterior constraint on this rate
+density and the true value is indicated as the thick gray vertical line.
+\figlabel{smooth-rate}}
+\end{figure}
+
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
+\end{center}
+\caption{%
+{\bf Simulated data}.
+The same as \fig{smooth-rate} for \modelb.
+\figlabel{simulation-rate}}
+\end{figure}
+
 \begin{figure}[p]
@@
 
-\section{Discussion}
-
-We have developed a hierarchical probabilistic framework for inferring the
-population of exoplanets based on noisy incomplete catalogs.
-This method incorporates systematic treatment of observational uncertainties
-and detection efficiency.
-One major benefit of this framework is that it provides the best possible
-probabilistic measurements of the population under the stated assumptions.
-After demonstrating the validity of our method on two qualitatively different
-synthetic exoplanet catalogs, we run our inference on a published catalog of
-small exoplanet candidates orbiting Sun-like stars (\citealt{petigura}) to
-determine the occurrence rate density of exoplanets as a function of period
-and radius.
-We extrapolate this measurement to the location of Earth and constrain the
-rate density of Earth analogs with large error bars.
-In order to perform this extrapolation, we don't assume a specific functional
-form for the rate density instead, we only assume that it is a smooth function
-of logarithmic period and radius.
-
-The occurrence rate density function that we infer is qualitatively consistent
-with previously published results using different inference techniques
-(\citealt{dong, fressin-fp, petigura}).
-In particular, we find (see \fig{radius}) previously recorded features in the
-radius distribution around $\radius\sim 3\,\radius_\oplus$ and $\radius\sim
-10\,\radius_\oplus$, although not at high signal-to-noise.
-We find that the period distributions for planets in different radius bins are
-different, in qualitative agreement with previous results (\citealt{dong}).
-\Fig{period} shows that larger planets tend to be on longer periods than
-smaller planets.
-
-In detail, any inference of the exoplanet population depends on the choices
-made in the pipeline that creates the catalog.
-Our results are based on a catalog that only includes single transiting
-systems (\citealt{petigura}) so our conclusions---and theirs---must be
-considered with this in mind.
-In particular, the exclusion of second planets must suppress the inferred rate
-density, especially at long periods and small radii.
-
-Our extrapolation of the rate density to the location of Earth is more general
-and conservative than any previously published method.
-Using this method, we find a rate density of Earth analogs that is
-inconsistent with the result published by \citet{petigura}.
-This discrepancy can be largely attributed to the rigidity of the assumptions
-about the period distribution.
-Our extrapolation is less confident than previous measurements.
-Again, this difference is due to the fact that we allow a much more flexible
-extrapolation function.
-\dfmtodo{This is another illustration that the use of flexible models makes
-probabilistic data analysis more conservative.}
-
-In contrast to previous work (\dfmtodo{CITE?}), we don't define ``Earth
-analog'' in terms of habitability or composition.
-Instead, we advocate for a definition in terms of more directly observable
-quantities (in this case, period and radius).
-Furthermore, we define \gammaearth\ as a rate density (per star per
-logarithmic period per logarithmic radius) so that its value doesn't depend on
-choices about the ``Earth-like'' bin.
-
-In our analysis we make a few simplifying assumptions.
-\begin{itemize}
-
-\item {\bf Conditional independence}\quad
-We assumed that every object in the catalog was a conditionally independent
-Poisson draw from the observable occurrence rate function.
-This is an interesting assumption to consider when applying this method to a
-different catalog where multiple transiting systems are included.
-In practice, the best first step towards relaxing this assumption is probably
-to follow \citet{tremaine} and assume that the mutual inclination distribution
-is the only source of conditional dependence between planets.
-For this \paper, the assumption of conditional independence is justified
-because the dataset explicitly includes only systems with a single transiting
-exoplanet.
-
-\item {\bf False positives}\quad
-Makes our estimates overestimates.
-Go to a mixture model.
-
-\item {\bf Known observational uncertainties}\quad
-Assumed Gaussian. Would be better with samples.
-
-\item {\bf Given empirical detection efficiency}\quad
-Single vs multi.
-In detail there are limitations: radius vs ratio. Stellar params.
-
-\item {\bf Smooth rate function}\quad
-No reason for the function to be smooth.
-Phase transitions.
-Composition etc.
-
-\end{itemize}
-Our assumptions are severe but we believe that this is the most conservative
-population inference method currently on the market.
-
-\dfmtodo{%
-Predicted number of observable Earth analogs in the \kepler\ dataset.
-Low but not zero.
-}
-
-\dfmtodo{%
-{\bf Advice:} \quad
-Remind us that we want a representation of the full posterior (samples)
-because that would let us do better things with the completeness (modeled as
-r/R, \etc).
-Must explicitly include priors too.
-}
-
-All of the code used in this project is available from
-\url{http://github.com/dfm/exopop} under the MIT open-source software license.
-This code (plus some dependencies) can be run to re-generate all of the
-figures and results in this \paper; this version of the paper was generated
-with git commit \texttt{\githash} (\gitdate).
-
-\acknowledgments
-We would like to thank Erik Petigura (Berkeley) for freely sharing his data
-and code.
-It is a pleasure to thank
-Ruth Angus (Oxford),
-Jo Bovy (IAS), and
-Eric Ford (PSU)
-for helpful contributions to the ideas and code presented here.
-This project was partially supported by the NSF (grant AST-0908357), NASA
-(grant NNX08AJ48G), and the Moore--Sloan Data Science Environment at NYU.
-This research builds on ideas generated at a three-week workshop on ``Modern
-Statistical and Computational Methods for Analysis of Kepler Data'' at SAMSI
-in 2013 June.
-This research made use of the NASA \project{Astrophysics Data System}.
-
-\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}
-\begin{thebibliography}{}\raggedright
-
-\bibitem[Adams, Murray \& MacKay(2009)]{poiss-gp}
-Adams, R.~P., Murray, I., \& MacKay, D.~J.~C.\ 2009, ICML, 2009, 9
-(\href{http://homepages.inf.ed.ac.uk/imurray2/pub/09poisson/}{online})
-
-\bibitem[Ambikasaran \etal(2014)]{dfm-gp}
-Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D.~W.,
-\& O'Neil, M.\ 2014, \arxiv{1403.6015}
-
-\bibitem[Batalha \etal(2013)]{kepler-catalog}
-Batalha, N.~M., Rowe, J.~F., Bryson, S.~T., et al.\ 2013, \apjs, 204, 24
-(\arxiv{1202.5852})
-
-\bibitem[Brewer \& Stello(2009)]{brewer}
-Brewer, B.~J., \& Stello, D.\ 2009, \mnras, 395, 2226 (\arxiv{0902.3907})
-
-\bibitem[Carter \& Winn(2009)]{carter}
-Carter, J.~A., \& Winn, J.~N.\ 2009, \apj, 704, 51 (\arxiv{0909.0747})
-
-\bibitem[Catanzarite \& Shao(2011)]{catanzarite}
-Catanzarite, J., \& Shao, M.\ 2011, \apj, 738, 151 (\arxiv{1103.1443})
-
-\bibitem[Dong \& Zhu(2013)]{dong}
-Dong, S., \& Zhu, Z.\ 2013, \apj, 778, 53 (\arxiv{1212.4853})
-
-\bibitem[Dressing  \& Charbonneau(2013)]{dressing}
-Dressing, C.~D., \& Charbonneau, D.\ 2013, \apj, 767, 95 (\arxiv{1302.1647})
-
-\bibitem[Fang \& Margot(2012)]{fang}
-Fang, J., \& Margot, J.-L.\ 2012, \apj, 761, 92 (\arxiv{1207.5250})
-
-\bibitem[Foreman-Mackey \etal(2012)]{emcee}
-Foreman-Mackey, D., Hogg, D.~W., Lang, D., \& Goodman, J.\ 2013, \pasp, 125,
-306 (\arxiv{1202.3665})
-
-\bibitem[Fressin \etal(2013)]{fressin-fp}
-Fressin, F., Torres, G., Charbonneau, D., \etal\ 2013, \apj, 766, 81
-(\arxiv{1301.0842})
-
-\bibitem[Gibson \etal(2012)]{gibson-gp}
-Gibson, N.~P., Aigrain, S., Roberts, S., \etal\ 2012, \mnras, 419, 2683
-(\arxiv{1109.3251})
-
-\bibitem[Hogg \etal(2013)]{hoggwhitepaper}
-Hogg, D.~W., Angus, R., Barclay, T., et al.\ 2013, \arxiv{1309.0653}
-
-\bibitem[Hogg \etal(2010)]{hogge}
-Hogg, D.~W., Myers, A.~D., \& Bovy, J.\ 2010, \apj, 725, 2166
-(\arxiv{1008.4146})
-
-\bibitem[Howard \etal(2012)]{howard}
-Howard, A.~W., Marcy, G.~W., Bryson, S.~T., et al.\ 2012, \apjs, 201, 15
-(\arxiv{1103.2541})
-
-\bibitem[Lewis \& Shedler(1979)]{poisson}
-Lewis, P.~A.~W., \& Shedler, G.~S.\ 1979, Naval Research Logistics Quarterly,
-26, 403
-
-\bibitem[Lissauer \etal(2011)]{lissauer}
-Lissauer, J.~J., Ragozzine, D., Fabrycky, D.~C., et al.\ 2011, \apjs, 197, 8
-(\arxiv{1102.0543})
-
-\bibitem[Morton \& Johnson(2011)]{morton}
-Morton, T.~D., \& Johnson, J.~A.\ 2011, \apj, 738, 170 (\arxiv{1101.5630})
-
-\bibitem[Morton \& Swift(2013)]{morton-swift}
-Morton, T.~D., \& Swift, J.~J.\ 2013, \arxiv{1303.3013}
-
-\bibitem[Murray \etal(2010)]{ess}
-Murray, I., Prescott Adams, R., \& MacKay, D.~J.~C.\ 2010, JMLR: W\&CP, 9, 541
-(\arxiv{1001.0175})
-
-\bibitem[Murray \& Prescott Adams(2010)]{ess-hyper}
-Murray, I., \& Prescott Adams, R.\ 2010, Advances in Neural Information
-Processing Systems, 23, 1723
-(\arxiv{1006.0868})
-
-\bibitem[Petigura \etal(2013a)]{petigura-a}
-Petigura, E.~A., Marcy, G.~W., \& Howard, A.~W.\ 2013a, \apj, 770, 69
-(\arxiv{1304.0460})
-
-\bibitem[Petigura \etal(2013b)]{petigura}
-Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013b,
-Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})
-
-\bibitem[Rasmussen \& Williams(2006)]{gp}
-Rasmussen, C.~E. \& Williams, C.~K.~I.\ 2006
-Gaussian Processes for Machine Learning, MIT Press
-(\href{http://www.gaussianprocess.org/gpml/}{online})
-
-\bibitem[Roberts \etal(2013)]{roberts}
-Roberts, S., McQuillan, A., Reece, S., \& Aigrain, S.\ 2013, \mnras, 435, 3639
-(\arxiv{1308.3644})
-
-\bibitem[Swift \etal(2013)]{swift}
-Swift, J.~J., Johnson, J.~A., Morton, T.~D., et al.\ 2013, \apj, 764, 105
-(\arxiv{1301.0023})
-
-\bibitem[Tabachnik \& Tremaine(2002)]{tabachnik}
-Tabachnik, S., \& Tremaine, S.\ 2002, \mnras, 335, 151
-(\arxiv{astro-ph/0107482})
-
-\bibitem[Traub(2012)]{traub}
-Traub, W.~A.\ 2012, \apj, 745, 20 (\arxiv{1109.4682})
-
-\bibitem[Tremaine \& Dong(2012)]{tremaine}
-Tremaine, S., \& Dong, S.\ 2012, \aj, 143, 94 (\arxiv{1106.5403})
-
-\bibitem[Youdin(2011)]{youdin}
-Youdin, A.~N.\ 2011, \apj, 742, 38 (\arxiv{1105.1782})
-
-\end{thebibliography}
-
 \end{document}

### 2014-06-10 document/ms.tex
final version

@@
 period as
-\begin{eqnarray}
+\begin{eqnarray}\eqlabel{transitprob}
 \transitprob (\period) &=&
@@
 extrapolation function.
-\dfmtodo{This is another illustration that the use of flexible models makes
-probabilistic data analysis more conservative.}
+This is another illustration that, against the standard data analysis
+folklore, the correct use of flexible models is \emph{conservative}.
 
@@
 
-\dfmtodo{%
-Predicted number of observable Earth analogs in the \kepler\ dataset.
-Low but not zero.
-}
-
-\dfmtodo{%
-{\bf Advice:} \quad
-Remind us that we want a representation of the full posterior (samples)
-because that would let us do better things with the completeness (modeled as
-r/R, \etc).
-Must explicitly include priors too.
-}
+Under the assumptions that we have made here, our inference of the occurrence
+rate density of exoplanets places a probabilistic constraint on the number of
+transiting Earth analogs in the existing \kepler\ dataset.
+If we adopt the definition of ``Earth-like'' from \citet[][$200 \le
+\period/\mathrm{day} < 400$ and $1 \le \radius/\radius_\oplus < 2$]{petigura},
+and integrate the product inferred rate density function and the geometric
+transit probability (\eqalt{transitprob}) over this bin, we find that there
+should be
+\begin{eqnarray}\eqlabel{ntransit}
+N_{\oplus,\,\mathrm{transiting}} &=& 9.2_{-4.0}^{+5.9}
+\end{eqnarray}
+Earth-like exoplanets transiting the stars in the sample of Sun-like stars
+chosen by \citet{petigura}.
+This is an exciting result because it means that, if we can improve the
+sensitivity of exoplanet search pipelines to small planets orbiting on long
+periods then we should find some Earth analogs in the existing data.
+Furthermore, because of the treatment of multiple transiting systems in the
+catalog, the number in \eq{ntransit} is probably a lower limit on the expected
+number of Earth-like exoplanets transiting Sun-like stars!
+
+Some of the caveats on the results in this paper are due to assumptions made
+for computational simplicity but a much more robust study would be possible
+given a complete representation of the posterior probability function for the
+physical parameters in the catalog.
+The use of MCMC to fit models to observations is becoming standard practice in
+astronomy and the results in many catalogs (including \citealt{petigura}) are
+given as statistics computed on posterior samplings.
+For the sake of hierarchical inferences like the method presented here, it
+would be very useful if the authors of upcoming catalogs also published
+samples from these distributions \emph{along with the value of their prior
+function evaluated at each sample}.
+In this spirit, we have released the results of this paper as posterior
+samplings for the occurrence rate density function (\dfmtodo{ADD SOME URL}).
 

### 2014-06-10 document/figures/results/make_plots.py

@@ -18,10 +18,20 @@ with h5py.File("samples.h5", "r") as f:
 ep_inds = (slice(None), 5, slice(2, 4))
 ep_area = (np.diff(ln_period_bin_edges)[ep_inds[1]]
            * np.diff(ln_radius_bin_edges)[ep_inds[2]])
+factor = np.diff(ln_radius_bin_edges)[ep_inds[2]] * 0.05061 * 10 ** (2./3)
+factor *= -3.0 * np.diff(ln_period_bin_edges ** (-2./3))[ep_inds[1]] / 2
+
 rate = np.sum(ep_area * np.exp(samples[ep_inds] - np.log(42557.0)), axis=1)
 q = quantile(rate, [0.16, 0.5, 0.84])
 print("Integrated rate: {0:.3f}_{{-{1:.3f}}}^{{+{2:.3f}}}"
       .format(q[1], *(np.diff(q))))
+
+num = np.sum(factor * np.exp(samples[ep_inds]), axis=1)
+q = quantile(num, [0.16, 0.5, 0.84])
+print("Integrated observable number: {0:.3f}_{{-{1:.3f}}}^{{+{2:.3f}}}"
+      .format(q[1], *(np.diff(q))))
+
+assert 0
 
 bins = [ln_period_bin_edges, ln_radius_bin_edges]
 labels = ["R / R_\oplus", "P / \mathrm{day}"]

### 2014-06-10 document/ms.tex
a few more comments from @ruthangus

@@
 \documentclass[12pt,preprint]{aastex}
+
+\pdfoutput=1
 
@@
 probability of the parameters $\{\entry_k\}$ under some choice of
-\emph{interim prior} \interim
+\emph{interim prior}
 \begin{eqnarray}\eqlabel{crazypost}
@@
 
-The Gaussian process model for the bin heights given in \eq{gp} very flexible
-but the results will depend on the values of the hyperparameters \mean\ and
-\smoothpars.
+The Gaussian process model for the bin heights given in \eq{gp} is very
+flexible but the results will depend on the values of the hyperparameters
+\mean\ and \smoothpars.
 Therefore, instead of fixing these parameters to specific values, we add
@@
 In order to perform this extrapolation, we don't assume a specific functional
-form for the rate density instead, we only assume that it is a smooth function
-of logarithmic period and radius.
+form for the rate density.
+Instead, we only assume that it is a smooth function of logarithmic period and
+radius.
 

### 2014-06-10 document/publish.py
publish script

@@ -1,0 +1,39 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+import os
+import glob
+import shutil
+from datetime import date
+from itertools import imap
+from subprocess import check_call
+
+datestamp = date.today().strftime("%m%d")
+TMPDIR = "fore" + datestamp
+outfn = TMPDIR+".tar"
+tex = open("ms.tex", "r").read()
+
+try:
+    os.makedirs(TMPDIR)
+except os.error:
+    pass
+
+
+def rename(fn):
+    a, b = os.path.split(fn)
+    return os.path.split(a)[1] + "-" + b, fn
+
+
+for a, b in imap(rename, glob.glob("figures/*/*.pdf")):
+    shutil.copyfile(b, os.path.join(TMPDIR, a))
+    tex = tex.replace(b, a)
+
+shutil.copyfile("vc.tex", os.path.join(TMPDIR, "vc.tex"))
+open(os.path.join(TMPDIR, "ms.tex"), "w").write(tex)
+check_call(" ".join(["cd", TMPDIR+";",
+                     "tar", "-cf", os.path.join("..", outfn), "*"]),
+           shell=True)
+shutil.rmtree(TMPDIR)
+
+print("Wrote file: '{0}'".format(outfn))

### 2014-06-10 document/ms.tex
comments from @benmontet

@@
 
-\newcommand{\project}[1]{{\sffamily #1}}
+\newcommand{\project}[1]{\emph{#1}}
 \newcommand{\kepler}{\project{Kepler}}
@@
 
-No true extra-Solar Earth analog is known.
+No true extrasolar Earth analog is known.
 Hundreds of planets have been found around Sun-like stars that are either
@@
 youdin, dong}).
-There are many reasons to prefer a result based on an optimization of a
-likelihood function and as we show in \sect{model}, if we model the occurrence
-rate density as a piecewise constant function, the
-inverse-detection-efficiency method can be derived as a special case of the
-likelihood method with a specific set of assumptions.
+A result based on an optimization of a likelihood function will---under any
+stated assumptions---saturate the Cram\'er--Rao bound and give an estimator of
+the population with the lowest possible variance.
+As we show in \sect{model}, if we model the occurrence rate density as a
+piecewise constant function, the inverse-detection-efficiency method can be
+derived as a special case of the likelihood method with a specific set of
+assumptions.
 
@@
 \end{eqnarray}
-We can use these samples to approximately compute the integral in
-\eq{simplemarglike} as
+We can use these samples and the Monte Carlo integral approximation to
+estimate the marginalized likelihood from \eq{simplemarglike}---up to an
+irrelevant constant---as
 \begin{eqnarray}\eqlabel{importance}
-\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
-    \frac{1}{N} \sum_{n=1}^N
+p(\{\data_k\}\,|\,\ratepars) &\approx&
+    \frac{\mathcal{Z}}{N} \sum_{n=1}^N
     \frac{p(\{\entry_k\}^{(n)}\,|\,\ratepars)}
-         {p(\{\entry_k\}^{(n)}\,|\,\interim)} \quad.
-\end{eqnarray}
+         {p(\{\entry_k\}^{(n)}\,|\,\interim)}
+\end{eqnarray}
+where the constant $\mathcal{Z} = p(\{\data_k\}\,|\,\interim)$ is not a
+function of the parameters.
 This is very efficient to compute as long as an evaluation of
-$p(\{\entry_k\}^{(n)}\,|\,\ratepars)$ is not very expensive.
+$p(\{\entry_k\}\,|\,\ratepars)$ is not very expensive.
 That being said, this could be a high variance estimator, depending on the
-initial choice of $p(\{\entry_k\}\,|\,\interim)$.
+number of independent samples $N$ and the initial choice of
+$p(\{\entry_k\}\,|\,\interim)$.
 Additionally, the support of $p(\{\entry_k\}\,|\,\ratepars)$ in $\{\entry_k\}$
@@
 $p(\{\entry_k\}\,|\,\interim)$.
+Besides this caveat, in the limit of infinite samples, the approximation in
+\eq{importance} becomes exact.
 \Eq{importance} is the \emph{importance sampling approximation} to the
@@
 the objects independently.
+This is a somewhat subtle point if we were to consider targets with more than
+one transiting planet---a point that we will return to below---but for the
+considerations of the dataset considered in this \paper, it is a justified
+simplification.
 In other words, we have per-object posterior samples
@@
 In this \paper, we don't use either of these inverse-detection-efficiency
-procedures because we are righteous.
+procedures because they both fundamentally rely on the assumption of
+\emph{negligible observational uncertainties}.
+By applying \eq{money}, we derive the probabilistic generalization of this
+model in the next section but the result is no longer analytic.
 
@@
 as burn-in.
-The remaining samples provide an approximation to the marginalized probability
+This results in approximately 4000 independent posterior samples estimated
+using the empirical integrated autocorrelation time of the chain
+(\citealt{goodman}).
+These samples provide an approximation to the marginalized probability
 distribution for \ratepars.
@@
 quantities with no mention of habitability or composition.
-To some readers, this might seem an unsatisfying choice.
-That being said, the composition of an exoplanet is notoriously difficult to
-measure even with large uncertainty and any definition of habitability is
-still extremely subjective.
+This might seem unsatisfying but the composition of an exoplanet is
+notoriously difficult to measure even with large uncertainty and any
+definition of habitability is still extremely subjective.
 With this in mind, we stick to the observable definition for this \paper.
@@
 Ruth Angus (Oxford),
-Jo Bovy (IAS), and
-Eric Ford (PSU)
+Jo Bovy (IAS),
+Eric Ford (PSU), and
+Ben Montet (Caltech/CfA)
 for helpful contributions to the ideas and code presented here.
@@
 (\arxiv{1109.3251})
+
+\bibitem[Goodman~\&\ Weare(2010)]{goodman}
+Goodman,~J. \& Weare,\ J.,
+2010, Comm.\ App.\ Math.\ Comp.\ Sci., 5, 65
 

### 2014-06-11 document/ms.tex
more comments from tim

@@
 considered with this in mind.
-In particular, the exclusion of second planets must suppress the inferred rate
-density, especially at long periods and small radii.
+In particular, the exclusion of second (and third, and so on) planets must
+suppress the inferred rate density, especially at long periods and small
+radii.
 
@@
 In an extension of this work, we could incorporate the effects of false
-positives by switching to a mixture model where each object is modeled as a
-mixture of \True\ exoplanet and false positive (\citealt{hogg-line}).
+positives by switching to a mixture model (see \citealt{hogg-line}, for
+example) where each object is modeled as a mixture of \True\ exoplanet and
+false positive.
+In this mixture model, the false positives would be represented using prior
+distributions similar to those used by \citet{morton-fp} or
+\citet{fressin-fp}.
 
@@
 Eric Ford (PSU), and
-Ben Montet (Caltech/CfA)
+Ben Montet (Caltech/Harvard)
 for helpful contributions to the ideas and code presented here.
@@
 (\arxiv{1102.0543})
+
+\bibitem[Morton(2012)]{morton-fp}
+Morton, T.~D.\ 2012, \apj, 761, 6 (\arxiv{1206.1568})
 

### 2014-06-11 document/ms.tex
adding figshare urls

@@
 The catalogs generated from these two models are shown in \fig{smooth-results}
-and \fig{simulation-results}, respectively and the data are available online
-at \dfmtodo{ADD SOME URL}.
+and \fig{simulation-results}, respectively and the data are available
+online\footnote{\url{%
+http://figshare.com/articles/Exoplanet_population_inference/1051864}}.
 
@@
 model.
-The resulting MCMC chain is available online at \dfmtodo{ADD SOME URL}.
+The resulting MCMC chain is available online\footnote{\url{%
+http://figshare.com/articles/Exoplanet_population_inference/1051864}}.
 
@@
 In this spirit, we have released the results of this paper as posterior
-samplings for the occurrence rate density function (\dfmtodo{ADD SOME URL}).
+samplings\footnote{\url{%
+http://figshare.com/articles/Exoplanet_population_inference/1051864}} for the
+occurrence rate density function.
 

### 2014-06-11 document/ms.tex
bin->step

@@
 \citet{petigura} recently published an exoplanet population analysis based on
-an independent analysis of the \kepler\ light curves for 42,000 Sun-like
+an independent analysis of the \kepler\ light curves for 42,557 Sun-like
 stars.
@@
 If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
-constant rate density---where the bin heights are the parameters---and if the
+constant rate density---where the step heights are the parameters---and if the
 uncertainties on the catalog are negligible compared to the bin widths then
@@
 \rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
-\ratepar_1 & \entry \in \binarea_1,\\
-\ratepar_2 & \entry \in \binarea_2,\\
+\exp(\ratepar_1) & \entry \in \binarea_1,\\
+\exp(\ratepar_2) & \entry \in \binarea_2,\\
 \cdots \\
-\ratepar_J & \entry \in \binarea_J,\\
+\exp(\ratepar_J) & \entry \in \binarea_J,\\
 0 & \mathrm{otherwise}
@@
     \sum_{k=1}^K \sum_{j=1}^J \mathbf{1}[\entry_k \in
-        \binarea_j]\,\ln\completeness(\entry_k)\,\ratepar_j
-    -\sum_{j=1}^J\ratepar_j\,\int_{\binarea_j} \completeness(\entry)\dd\entry
+        \binarea_j]\,[\ln\completeness(\entry_k)+\ratepar_j]
+    -\sum_{j=1}^J\exp(\ratepar_j)\,
+        \int_{\binarea_j} \completeness(\entry)\dd\entry
 \end{eqnarray}
@@
 \begin{eqnarray}\eqlabel{ml-ide}
-{\ratepar_j}^* &=&
+\exp ({\ratepar_j}^*) &=&
 \frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
@@
 \begin{eqnarray}
-\frac{\delta\ratepar_j}{{\ratepar_j}^*} &=& \frac{1}{\sqrt{N_j}} \quad.
+\frac{\delta\exp({\ratepar_j}^*)}{\exp({\ratepar_j}^*)} &=&
+    \frac{1}{\sqrt{N_j}}
+\quad.
 \end{eqnarray}
@@
 flexible enough to capture structure that is supported by the data.
-To achieve this, we model the logarithmic bin heights as being drawn from a
+To achieve this, we model the logarithmic step heights as being drawn from a
 Gaussian process (\citealt{gp, gibson-gp, dfm-gp}).
@@
 p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \nonumber\\
-&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{1},\,
+&=& \mathcal{N} \left[\ratepars;\,\mean\,\bvec{1},\,
 \cov(\{\binarea_j\},\,\smoothpars)\right]
@@
 
-The Gaussian process model for the bin heights given in \eq{gp} is very
+The Gaussian process model for the step heights given in \eq{gp} is very
 flexible but the results will depend on the values of the hyperparameters
@@
 Strictly speaking, in this model, $p(\ratepars\,|\,\mean,\,\smoothpars)$ can't
-really be called a ``prior'' anymore and the constraints on the bin heights
-\ratepars\ are no longer independent.
+really be called a ``prior'' anymore and the constraints on the step heights
+are no longer independent.
 
 There is a very efficient algorithm called elliptical slice sampling (ESS;
-\citealt{ess, ess-hyper}) for sampling the bin heights \ratepars\ from the
+\citealt{ess, ess-hyper}) for sampling the step heights \ratepars\ from the
 density in \eq{joint-posterior}.
@@
 
-For all the results below, we run a Markov chain with $10^6$ steps for the bin
+For all the results below, we run a Markov chain with $10^6$ steps for the
 heights and update the hyperparameters every 10 steps.

### 2014-06-11 document/ms.tex
a few typos etc

@@
 \section{Introduction}
+\sectlabel{intro}
 
@@
 \citet{petigura} recently published an exoplanet population analysis based on
-an independent analysis of the \kepler\ light curves for 42,557 Sun-like
-stars.
-This study was especially novel because the authors developed an independent
+an independent study of the \kepler\ light curves for 42,557 Sun-like stars.
+This study was especially novel because the authors developed their own
 planet search pipeline (\terra; \citealt{petigura-a}) and determined the
@@
 signals into real light curves measured by \kepler.
-The occurrence rate function determined by \citet{petigura} agreed
+The occurrence rate function determined by \citet{petigura} agrees
 qualitatively with previous studies of small planets orbiting Sun-like stars
@@
 youdin, dong}).
-A result based on an optimization of a likelihood function will---under any
-stated assumptions---saturate the Cram\'er--Rao bound and give an estimator of
-the population with the lowest possible variance.
 As we show in \sect{model}, if we model the occurrence rate density as a
@@
 but they do simplify the specific procedures we employ.
-Specifically, we assume that
+We assume that
 \begin{itemize}
@@
 detectable companion}.
-The global effects of this selection is not trivial and an in-depth discussion
-is beyond the scope of this \paper\ but all of the results should be
-interpreted with this caveat in mind.
+The global effects of this selection are not trivial and an in-depth
+discussion is beyond the scope of this \paper\ but all of the results should
+be interpreted with this caveat in mind.
 
@@
 In our case, this assumption qualitatively means that the signals found in the
-light curves depend only on the actual planet properties and not on the
-distributions from which they are drawn.
+light curves depend only on the actual properties of the planet and star, and
+not on the distributions from which they are drawn.
 It is worth emphasizing that---as we will discuss further below---the catalog
@@
 p(\{\data_k\}\,|\,\ratepars) &\approx&
-    \frac{\mathcal{Z}}{N} \sum_{n=1}^N
+    \frac{Z_\interim}{N} \sum_{n=1}^N
     \frac{p(\{\entry_k\}^{(n)}\,|\,\ratepars)}
@@
 \end{eqnarray}
-where the constant $\mathcal{Z} = p(\{\data_k\}\,|\,\interim)$ is not a
-function of the parameters.
+where the constant $Z_\interim = p(\{\data_k\}\,|\,\interim)$ is not a
+function of the parameters \ratepars.
 This is very efficient to compute as long as an evaluation of
-$p(\{\entry_k\}\,|\,\ratepars)$ is not very expensive.
-That being said, this could be a high variance estimator, depending on the
-number of independent samples $N$ and the initial choice of
-$p(\{\entry_k\}\,|\,\interim)$.
+$p(\{\entry_k\}\,|\,\ratepars)$ is not expensive.
+That being said, \eq{importance} could be a high variance estimator of
+\eq{simplemarglike}, depending on the number of independent samples $N$ and
+the initial choice of $p(\{\entry_k\}\,|\,\interim)$.
 Additionally, the support of $p(\{\entry_k\}\,|\,\ratepars)$ in $\{\entry_k\}$
@@
 occurrence rate density.
-To make it more concrete, we'll model the catalog as a Poisson draw from the
-\emph{observable} rate density $\obs{\rate}_\ratepars$.
+To make it more concrete, we'll model the catalog as a draw from the
+inhomogeneous Poisson process set by the \emph{observable} rate density
+$\obs{\rate}_\ratepars$.
 This leads to the previously known result (see \citealt{tabachnik,youdin} for
@@
 number of observable exoplanets in the sample.
-A further assumption that we have made in \eq{poisson-like} is that each
-object is an independent sample from the inhomogeneous Poisson process set by
-the rate density.
-This assumption---though it is often made---cannot be correct when we consider
-systems of multiple planets.
-It is possible to relax this assumption (see \citealt{tremaine} for a related
-technique) but this generalization is beyond the scope of the current \paper.
 
@@
 $\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
-whether we model the rate density as (for example) a broken power law or
+whether we model the rate density as, for example, a broken power law or
 a histogram.
@@
 For the results in this \paper, we will assume that the completeness function
-$\completeness(\entry)$ is known empirically but that is not a requirement for
-the validity of this method.
-Instead, we could use a functional form for it and even infer its parameters
-along with the parameters of the rate density.
+$\completeness(\entry)$ is known empirically on a grid in period and radius
+but that is not a requirement for the validity of this method.
+Instead, we could use a functional form for the completeness and even infer
+its parameters along with the parameters of the rate density.
 
@@
 
-There is a very efficient algorithm called elliptical slice sampling (ESS;
+There is an efficient algorithm called elliptical slice sampling (ESS;
 \citealt{ess, ess-hyper}) for sampling the step heights \ratepars\ from the
@@
 as burn-in.
-This results in approximately 4000 independent posterior samples estimated
-using the empirical integrated autocorrelation time of the chain
-(\citealt{goodman}).
+By estimating the empirical integrated autocorrelation time of the chain
+(\citealt{goodman}), we find that the resulting chain has $\gtrsim4000$
+independent posterior samples.
 These samples provide an approximation to the marginalized probability
@@
 these parameters in the selected stellar catalog but we'll approximate it by
-the empirical distribution; a reasonable approximation given the size of the
+the empirical distribution; a reasonable simplification given the size of the
 dataset.
@@
 explore any biases introduced by ignoring the observational uncertainties, we
-start by ``observing'' a few synthetic catalogs from qualitatively different
+start by ``observing'' two synthetic catalogs from qualitatively different
 known occurrence rate density functions.
-For each of these simulations, we'll take the completeness function computed
-by \citet{petigura} as given.
+For each of these simulations, we take the completeness function computed by
+\citet{petigura} as given.
 In general, \eq{poisson-like} can be sampled using a procedure called thinning
@@
 model: \emph{flat in logarithmic period}.
-These results are sensitive to the choice of extrapolation function and the
-specific definition of ``Earth analog''.
+These results are all sensitive to the choice of extrapolation function and
+the specific definition of ``Earth analog''.
 
@@
 Since these catalogs were simulated from a known population model, we know the
-true value of \gammaearth\ and it is indicated in the Figures with a vertical
+true value of \gammaearth\ and it is indicated in the figures with a vertical
 gray line.
@@
 
-The equivalent Figures for the radius distribution are shown in
+The equivalent results for the radius distribution are shown in
 Figures~\figref{radius} and~\figref{linear-radius}.
@@
 10\radius_\oplus$.
-These features appear in every period bin and they were also detected---using
-a completely different dataset and technique---by \citet{dong} and a similar
-result is visible in the occurrence rate determined by \citet[][their Figure
-7]{fressin-fp} at low signal-to-noise.
+These features appear in every period bin.
+They were also detected---using a completely different dataset and
+technique---by \citet{dong} and a similar result is visible in the occurrence
+rate determined by \citet[][their Figure 7]{fressin-fp} at low
+signal-to-noise.
 \Fig{linear-radius} shows the same result but presented as a function of
@@
 distribution for small planets on long orbits is flat in logarithmic period.
-If we enforce the same constraint, we reproduce their result but we relax this
-assumption, allowing the data to constrain the period distribution, we find
-evidence for a \emph{decreasing} period distribution as a function of
+If we enforce the same constraint, we reproduce their result but when we relax
+this assumption, allowing the data to constrain the period distribution, we
+find evidence for a \emph{decreasing} period distribution as a function of
 logarithmic period (see \fig{period}).
@@
 One major benefit of this framework is that it provides the best possible
-probabilistic measurements of the population under the stated assumptions.
+probabilistic measurements of the population under the assumptions listed in
+\sect{intro} and repeated below.
 After demonstrating the validity of our method on two qualitatively different
@@
 small exoplanet candidates orbiting Sun-like stars (\citealt{petigura}) to
-determine the occurrence rate density of exoplanets as a function of period
+determine the occurrence rate density these planets as a function of period
 and radius.
@@
 and conservative than any previously published method.
-Using this method, we find a rate density of Earth analogs that is
-inconsistent with the result published by \citet{petigura}.
+We find a rate density of Earth analogs that is inconsistent with the result
+published by \citet{petigura}.
 This discrepancy can be largely attributed to the rigidity of the assumptions
 about the period distribution.
-Our extrapolation is less confident than previous measurements.
+Our extrapolation is also less confident than previous measurements.
 Again, this difference is due to the fact that we allow a much more flexible
@@
 \item {\bf Conditional independence}\quad
-We assume that every object in the catalog was a conditionally independent
+We assume that every object in the catalog is a conditionally independent
 draw from the observable occurrence rate density.
-This is probably a bad assumption when applying this method to a different
-catalog where multiple transiting systems are included.
+This is a bad assumption when applying this method to a different catalog
+where multiple transiting systems are included.
 In practice, the best first step towards relaxing this assumption is probably
@@
 The biggest effect is probably the fact that this formulation doesn't include
-the fact that only the \emph{most detectable signal} in each light curve was
-included in the catalog.
-This bias will have the largest effect in the parts of parameter space where
-the baseline detection efficiency is lowest: at long periods and small radius.
+the selection of only the \emph{most detectable signal} in each light curve.
+This bias will be largest in the parts of parameter space where the baseline
+detection efficiency is lowest: at long periods and small radius.
 As a result, our inferences (and the results from \citealt{petigura}) about
-the occurrence rate of small planets on long periods should be seen as
-\emph{lower limits}.
+the occurrence rate of small planets on long periods is probably
+\emph{underestimated} relative to \Truth.
 In detail there is another limitation due to the fact that the stellar
@@
 and integrate the product inferred rate density function and the geometric
-transit probability (\eqalt{transitprob}) over this bin, we find that there
-should be
+transit probability (\eqalt{transitprob}) over this bin, we find that the
+expected number of Earth-like exoplanets transiting the stars in the sample of
+Sun-like stars chosen by \citet{petigura} is
 \begin{eqnarray}\eqlabel{ntransit}
@@
 \end{eqnarray}
-Earth-like exoplanets transiting the stars in the sample of Sun-like stars
-chosen by \citet{petigura}.
+where the uncertainties are only on the expectation value and don't include
+the Poisson sampling variance.
 This is an exciting result because it means that, if we can improve the
@@
 Furthermore, because of the treatment of multiple transiting systems in the
-catalog, the number in \eq{ntransit} is probably a lower limit on the expected
-number of Earth-like exoplanets transiting Sun-like stars!
+catalog, the \True\ expected number of transiting Earth-like exoplanets
+orbiting Sun-like stars is probably larger than the values in \eq{ntransit}!
 

### 2014-06-11 document/ms.tex
Addressing #13

@@
 physical parameters (period, radius, \etc).
-It is also possible to relax the noise model---to include effects like the
-underestimation of uncertainties---but inference would be substantially more
-computationally expensive.
+There is recent evidence that the stellar radii of \kepler\ targets might, on
+average, be underestimated (\citealt{bastien}), introducing another source of
+noise.
+It is possible to relax the noise model and include effects like this but
+inference would be substantially more computationally expensive.
 
@@
 
+\bibitem[Bastien \etal(2014)]{bastien}
+Bastien, F.~A., Stassun, K.~G., \& Pepper, J.\ 2014, \apjl, 788, L9
+(\arxiv{1405.0940})
+
 \bibitem[Batalha \etal(2013)]{kepler-catalog}

### 2014-06-11 document/ms.tex
updating figshare urls

@@
 and \fig{simulation-results}, respectively and the data are available
-online\footnote{\url{%
-http://figshare.com/articles/Exoplanet_population_inference/1051864}}.
+online\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.1051864}}.
 
@@
 The resulting MCMC chain is available online\footnote{\url{%
-http://figshare.com/articles/Exoplanet_population_inference/1051864}}.
+http://dx.doi.org/10.6084/m9.figshare.1051864}}.
 
@@
 In this spirit, we have released the results of this paper as posterior
-samplings\footnote{\url{%
-http://figshare.com/articles/Exoplanet_population_inference/1051864}} for the
-occurrence rate density function.
+samplings\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.1051864}} for
+the occurrence rate density function.
 

### 2014-06-11 document/ms.tex
a few comments from @mrtommyb

@@
 We confirm a previous result that the radius distribution changes slope near
-Earth radius.
+Earth's radius.
 We find that the rate density of Earth analogs is about 0.02 (per star per
@@
 NASA's \kepler\ mission has enabled the discovery of thousands of exoplanet
-candidates (\citealt{kepler-catalog}).
+candidates (\citealt{kepler-catalog, burke}).
 While many of these candidates have not been confirmed as bona fide planets,
@@
 efficiency empirically by injecting synthetic signals into the raw data and
-testing recovery (\citealt{petigura-a, petigura}).
+testing recovery (\citealt{pixel, petigura-a, petigura}).
 
@@
 sensitivity of exoplanet search pipelines to small planets orbiting on long
-periods then we should find some Earth analogs in the existing data.
+periods, then we should find some Earth analogs in the existing data.
 Furthermore, because of the treatment of multiple transiting systems in the
@@
 \bibitem[Batalha \etal(2013)]{kepler-catalog}
-Batalha, N.~M., Rowe, J.~F., Bryson, S.~T., et al.\ 2013, \apjs, 204, 24
+Batalha, N.~M., Rowe, J.~F., Bryson, S.~T., \etal\ 2013, \apjs, 204, 24
 (\arxiv{1202.5852})
@@
 
+\bibitem[Burke \etal(2014)]{burke}
+Burke, C.~J., Bryson, S.~T., Mullally, F., \etal\ 2014, \apjs, 210, 19
+(\arxiv{1312.5358})
+
 \bibitem[Carter \& Winn(2009)]{carter}
@@
 Catanzarite, J., \& Shao, M.\ 2011, \apj, 738, 151 (\arxiv{1103.1443})
+
+\bibitem[Christiansen \etal(2013)]{pixel}
+Christiansen, J.~L., Clarke, B.~D., Burke, C.~J., \etal\ 2013, \apjs, 207, 35
+(\arxiv{1303.0255})
 
@@
 \bibitem[Hogg \etal(2013)]{hoggwhitepaper}
-Hogg, D.~W., Angus, R., Barclay, T., et al.\ 2013, \arxiv{1309.0653}
+Hogg, D.~W., Angus, R., Barclay, T., \etal\ 2013, \arxiv{1309.0653}
 
@@
 \bibitem[Howard \etal(2012)]{howard}
-Howard, A.~W., Marcy, G.~W., Bryson, S.~T., et al.\ 2012, \apjs, 201, 15
+Howard, A.~W., Marcy, G.~W., Bryson, S.~T., \etal\ 2012, \apjs, 201, 15
 (\arxiv{1103.2541})
@@
 \bibitem[Lissauer \etal(2011)]{lissauer}
-Lissauer, J.~J., Ragozzine, D., Fabrycky, D.~C., et al.\ 2011, \apjs, 197, 8
+Lissauer, J.~J., Ragozzine, D., Fabrycky, D.~C., \etal\ 2011, \apjs, 197, 8
 (\arxiv{1102.0543})
@@
 \bibitem[Swift \etal(2013)]{swift}
-Swift, J.~J., Johnson, J.~A., Morton, T.~D., et al.\ 2013, \apj, 764, 105
+Swift, J.~J., Johnson, J.~A., Morton, T.~D., \etal\ 2013, \apj, 764, 105
 (\arxiv{1301.0023})

### 2014-06-11 document/ms.tex
adding @mrtommyb to the acks

@@
 Ruth Angus (Oxford),
+Tom Barclay (NASA Ames),
 Jo Bovy (IAS),

### 2014-06-12 code/load_data.py
documentation

@@ -1,0 +1,93 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+"""
+This module eases interaction with the datasets and monkey patches the import
+path. It's "research code"... give me a break!
+
+"""
+
+from __future__ import division, print_function
+
+import os
+import sys
+import h5py
+import numpy as np
+import pandas as pd
+
+d = os.path.dirname
+base = d(os.path.abspath(__file__))
+sys.path.insert(0, base)
+from population import SavedCensoringFunction
+
+bp = os.path.join(d(base), "data")
+
+# Hard-coded transit probability scaling.
+P = 10.0
+transit_lnprob0 = -2.98353340397
+ln_period0 = np.log(P)
+
+# Un-comment the following lines to recompute the transit probability if you
+# get a list of Petigura's injections.
+# G = 2945.4625385377644
+# stars = pd.read_hdf(os.path.join(bp, "stlr.h5"), "stlr")
+# transit_lnprobs = np.log(stars.Rstar) + (np.log(4*np.pi*np.pi)
+#                                          - np.log(G*P*P*stars.Mstar)) / 3
+# transit_lnprob0 = np.median(transit_lnprobs)
+
+
+def load_detection_efficiency():
+    """
+    Load a pre-computed detection efficiency grid.
+
+    """
+    with h5py.File(os.path.join(bp, "completeness.h5"), "r") as f:
+        bins = [f["ln_period_bin_edges"][...],
+                f["ln_radius_bin_edges"][...]]
+        lnprob = f["ln_detect_eff"][...]
+        lncompleteness = f["ln_completeness"][...]
+    return SavedCensoringFunction(bins, lnprob, lncompleteness)
+
+
+def load_completenes_sim(rp_func=np.log, per_rng=np.log([5, 400]),
+                         rp_rng=None, K=50000):
+    """
+    This function will only work if you request Petigura's completeness
+    simulations from him and save them in the ``data`` directory.
+
+    """
+    if rp_rng is None:
+        rp_rng = rp_func([0.5, 16])
+    sim = pd.read_hdf(os.path.join(bp, "mcDV.h5"), "mcDV")
+    m = sim.found * sim.bDV
[…]

### 2014-06-12 code/main.py

@@ -1,0 +1,205 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+"""
+This script runs the analysis on either the real catalog or a simulated one.
+To run on the real data, call:
+
+    python main.py
+
+and to run on a simulation, call:
+
+    python main.py path/to/the/simulated/catalog/
+
+All of the MCMC tuning parameters are hard coded. Sorry!
+
+"""
+
+from __future__ import division, print_function
+
+import os
+import h5py
+import triangle
+import numpy as np
+import cPickle as pickle
+import matplotlib.pyplot as pl
+
+from load_data import load_candidates, load_detection_efficiency
+from population import ProbabilisticModel, Dataset, Population
+
+
+def inverse_detection_efficiency(pop, censor, catalog, err, truth=None):
+    c = np.log(catalog)
+    ind = [np.digitize(x, b) for x, b in zip(c.T, censor.bins)]
+    weights = np.exp(-censor.lnprob[ind])
+    val, x, y = np.histogram2d(c[:, 0], c[:, 1], pop.bins, weights=weights)
+    var, x, y = np.histogram2d(c[:, 0], c[:, 1], pop.bins, weights=weights**2)
+
+    val[~np.isfinite(val)] = 0.0
+    var[~np.isfinite(var)] = 0.0
+
+    # Build the model for plotting.
+    v = pop.initial()
+    lg = np.log(val).flatten()
+    m = np.isfinite(lg)
+    lg[~m] = 0.0
+    v[-len(lg):] = lg
+
+    # Compute the marginalized errorbars.
+    marg = [np.sum(val, axis=(i+1) % 2) for i in range(2)]
+    norm = [np.sum(a * np.diff(pop.bins[i])) for i, a in enumerate(marg)]
+    literature = [(pop.bins[i], a/norm[i],
+                   np.sqrt(np.sum(var, axis=(i+1) % 2))/norm[i])
+                  for i, a in enumerate(marg)]
+
+    # Plot the results.
+    labels = ["$\ln T/\mathrm{day}$", "$\ln R/R_\oplus$"]
+    top_axes = ["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"]
+    fig = pop.plot_2d(v, censor=censor, catalog=np.log(catalog),
+                      err=err, true=truth, labels=labels,
[…]

### 2014-06-12 code/population.py

@@ -4,7 +4,7 @@
 from __future__ import division, print_function
 
 __all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",
-           "Dataset", "CensoringFunction"]
+           "Dataset", "CensoringFunction", "SavedCensoringFunction"]
 
 from itertools import izip
 
@@ -600,6 +600,14 @@ class CensoringFunction(object):
     def get_lncompleteness(self, samples):
         i = self.index(samples)
         return self.lncompleteness[i]
+
+
+class SavedCensoringFunction(CensoringFunction):
+
+    def __init__(self, bins, lnprob, lncompleteness):
+        self.bins = bins
+        self.lnprob = lnprob
+        self.lncompleteness = lncompleteness
 
 
 class ProbabilisticModel(object):

### 2014-06-12 code/results.py

@@ -1,0 +1,145 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+"""
+This script loads the results of a `main.py` run and performs a consistent
+analysis generating the main figures from the paper and thinning the MCMC
+chain.
+
+Usage:
+
+    python results.py path/to/mcmc/output/
+
+"""
+
+from __future__ import division, print_function
+
+import os
+import sys
+import h5py
+import triangle
+import numpy as np
+import cPickle as pickle
+import matplotlib.pyplot as pl
+
+from emcee.autocorr import integrated_time
+
+# HACK: make sure that the import path is set up properly.
+import load_data
+load_data = load_data
+
+bp = sys.argv[1]
+model, catalog, err, truth, labels, top_axes, literature = \
+    pickle.load(open(os.path.join(bp, "model.pkl")))
+pop = model.population
+
+# Load the original catalog if it's there.
+fn = os.path.join(bp, "catalog.pkl")
+if os.path.exists(fn):
+    catalog, err, truth = \
+        pickle.load(open(os.path.join(bp, "catalog.pkl")))
+
+    # Hack to deal with negative radii.
+    m = catalog[:, 1] > 0
+    catalog = catalog[m]
+    rerr = [np.log(catalog[:, 1]) - np.log(np.abs(catalog[:, 1]-err[m, 1])),
+            np.log(catalog[:, 1]+err[m, 1]) - np.log(catalog[:, 1])]
+    err = [0, rerr]
+
+labels = ["$\ln P/\mathrm{day}$", "$\ln R/R_\oplus$"]
+top_axes = ["$P\,[\mathrm{days}]$", "$R\,[R_\oplus]$"]
+
+
+with h5py.File(os.path.join(bp, "results.h5")) as f:
+    i = int(f.attrs["iteration"])
+    samples = f["samples"][:i, :]
+    hyper = f["hyper"][:i, :]
+    lnprob = f["lnprob"][:i]
+
+for i in range(hyper.shape[1]):
+    pl.clf()
+    pl.plot(hyper[:, i])
+    pl.savefig(os.path.join(bp, "time-hyper-{0:03d}.png".format(i)))
+
+pl.clf()
+pl.plot(lnprob)
+pl.savefig(os.path.join(bp, "time-lnprob.png"))
+
+nstar = 42557.0
[…]

### 2014-06-12 code/simulate.py

@@ -1,0 +1,126 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+"""
+This script generates synthetic catalogs from known occurrence rate densities.
+
+Usage:
+    python simulate.py path/to/output/directory
+
+Run it with the "--smooth" flag to sample for the smooth model (Catalog A in
+the paper) and without for Catalog B.
+
+"""
+
+from __future__ import division, print_function
+
+try:
+    from savefig import monkey_patch
+except ImportError:
+    pass
+else:
+    monkey_patch()
+
+import os
+import numpy as np
+import cPickle as pickle
+from itertools import product
+from scipy.misc import logsumexp
+
+from load_data import load_detection_efficiency, load_candidates
+from population import SeparablePopulation, Histogram, BrokenPowerLaw
+
+
+def main(args, state=None, smooth=False):
+    bp, seed = args
+    if seed is not None:
+        np.random.seed(seed)
+    try:
+        os.makedirs(bp)
+    except os.error:
+        pass
+
+    if state is not None:
+        np.random.set_state(state)
+
+    # Save the random state.
+    pickle.dump(np.random.get_state(),
+                open(os.path.join(bp, "state.pkl"), "wb"), -1)
+
+    # Load the censoring function.
+    censor = load_detection_efficiency()
+
+    # Load the candidates to get fractional uncertainties.
+    inds, mu, sig = load_candidates()
+    ferr = sig[:, 1] / mu[:, 1]
+
+    # The values from EP's paper (+some made up numbers).
+    lpb, lrb = censor.bins
+    x, y = lpb[::4], lrb[::4]
+    p_vals = np.log(np.array([8.9, 13.7, 15.8, 15.2, 13.3, 12.2]))
+    r_vals = np.log(np.array([11, 11.5, 12, 14.2, 18.6, 5.9, 1.9, 1, 0.9, 0.7,
+                              0.5, 0.5]))
+
+    # Normalize the underlying distribution.
+    p_vals -= logsumexp(p_vals + np.log(np.diff(x)))
+    r_vals -= logsumexp(r_vals + np.log(np.diff(y)))
+
+    # Build a synthetic population.
+    norm = 10.5
+    if smooth:
+        truth = [norm, 0.5, -0.2, 4.0, 0.8, -1.5, 1.0]
+        pdist = BrokenPowerLaw(lpb)
[…]

### 2014-06-12 code/population.py
one more docstring

@@ -1,6 +1,15 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
 
+"""
+This module implements the probabilistic hierarchical model from
+Foreman-Mackey, Hogg & Morton (2014). It should (mostly) work in arbitrary
+latent dimensions but it has only been tested in 2D.
+
+Please cite our paper if you use this code.
+
+"""
+
 from __future__ import division, print_function
 
 __all__ = ["BrokenPowerLaw", "Histogram", "Population", "SeparablePopulation",

### 2014-06-12 document/code/fake_noise.py
cleaning up

@@ -1,129 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-try:
-    from savefig import monkey_patch
-except ImportError:
-    pass
-else:
-    monkey_patch()
-
-import os
-import numpy as np
-import cPickle as pickle
-from itertools import product
-from scipy.misc import logsumexp
-
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates)
-from population import (CensoringFunction, SeparablePopulation, Histogram,
-                        BrokenPowerLaw)
-
-
-def main(args, state=None, smooth=False):
-    bp, seed = args
-    if seed is not None:
-        np.random.seed(seed)
-    try:
-        os.makedirs(bp)
-    except os.error:
-        pass
-
-    if state is not None:
-        np.random.set_state(state)
-
-    # Save the random state.
-    pickle.dump(np.random.get_state(),
-                open(os.path.join(bp, "state.pkl"), "wb"), -1)
-
-    # Bins chosen to include EP's bins.
-    per_rng, np_bins = np.log([6.25, 400.0]), 4*6
-    rp_rng, nr_bins = np.log([0.5, 32]), 4*12
-
-    # Load the data.
-    ln_P_inj, ln_R_inj, recovered = load_completenes_sim(per_rng=per_rng,
-                                                         rp_rng=rp_rng)
-    tlp = lambda lnp, lnr: transit_lnprob0 - 2.*(lnp - ln_period0)/3
-
-    # Set up the censoring function.
-    censor = CensoringFunction(np.vstack((ln_P_inj, ln_R_inj)).T, recovered,
-                               bins=(np_bins, nr_bins),
-                               range=[per_rng, rp_rng],
-                               transit_lnprob_function=tlp)
-
-    # Load the candidates to get fractional uncertainties.
-    inds, mu, sig = load_candidates()
-    ferr = sig[:, 1] / mu[:, 1]
-
-    # The values from EP's paper (+some made up numbers).
-    lpb, lrb = censor.bins
-    x, y = lpb[::4], lrb[::4]
-    p_vals = np.log(np.array([8.9, 13.7, 15.8, 15.2, 13.3, 12.2]))
[…]

### 2014-06-12 document/code/load_data.py

@@ -1,65 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-import os
-import sys
-import numpy as np
-import pandas as pd
-
-d = os.path.dirname
-base = d(d(d(os.path.abspath(__file__))))
-sys.path.insert(0, base)
-import population
-population = population
-
-bp = os.path.join(base, "data")
-
-G = 2945.4625385377644
-P = 10.0
-# stars = pd.read_hdf(os.path.join(bp, "stlr.h5"), "stlr")
-# transit_lnprobs = np.log(stars.Rstar) + (np.log(4*np.pi*np.pi)
-#                                          - np.log(G*P*P*stars.Mstar)) / 3
-# transit_lnprob0 = np.median(transit_lnprobs)
-transit_lnprob0 = -2.98353340397
-ln_period0 = np.log(P)
-
-
-def load_completenes_sim(rp_func=np.log, per_rng=np.log([5, 400]),
-                         rp_rng=None, K=50000):
-    if rp_rng is None:
-        rp_rng = rp_func([0.5, 16])
-    sim = pd.read_hdf(os.path.join(bp, "mcDV.h5"), "mcDV")
-    m = sim.found * sim.bDV
-    x, y, z = np.log(sim.inj_P), rp_func(sim.inj_Rp), m
-    if rp_rng[1] > rp_func(16):
-        x = np.append(x, np.random.uniform(per_rng[0], per_rng[1], K))
-        y = np.append(y, np.random.uniform(rp_func(16.), rp_rng[1], K))
-        z = np.append(z, np.ones(K, dtype=bool))
-    return x, y, z
-
-
-def load_candidates():
-    lines = open(os.path.join(bp, "table_ekoi836.tex")).readlines()
-    data = np.array([[l.split("&")[i] for i in (0, 2, 12, 13)] for l in lines
-                     if l.split("&")[4].strip() == "P"],
-                    dtype=float)
-    return (np.array(data[:, 0], dtype=int), data[:, 1:3],
-            np.vstack([np.zeros(len(data)), data[:, 3]]).T)
-
-
-def load_petigura_bins(mylog=np.log):
[…]

### 2014-06-12 document/code/main.py

@@ -1,218 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-try:
-    from savefig import monkey_patch
-except ImportError:
-    pass
-else:
-    monkey_patch()
-
-import os
-import h5py
-import triangle
-import numpy as np
-import cPickle as pickle
-import matplotlib.pyplot as pl
-
-from load_data import (transit_lnprob0, ln_period0, load_completenes_sim,
-                       load_candidates)
-from population import (CensoringFunction,
-                        ProbabilisticModel, Dataset, Population)
-
-
-def inverse_detection_efficiency(pop, censor, catalog, err, truth=None):
-    c = np.log(catalog)
-    ind = [np.digitize(x, b) for x, b in zip(c.T, censor.bins)]
-    weights = np.exp(-censor.lnprob[ind])
-    val, x, y = np.histogram2d(c[:, 0], c[:, 1], pop.bins, weights=weights)
-    var, x, y = np.histogram2d(c[:, 0], c[:, 1], pop.bins, weights=weights**2)
-
-    val[~np.isfinite(val)] = 0.0
-    var[~np.isfinite(var)] = 0.0
-
-    # Build the model for plotting.
-    v = pop.initial()
-    lg = np.log(val).flatten()
-    m = np.isfinite(lg)
-    lg[~m] = 0.0
-    v[-len(lg):] = lg
-
-    # Compute the marginalized errorbars.
-    marg = [np.sum(val, axis=(i+1) % 2) for i in range(2)]
-    norm = [np.sum(a * np.diff(pop.bins[i])) for i, a in enumerate(marg)]
-    literature = [(pop.bins[i], a/norm[i],
-                   np.sqrt(np.sum(var, axis=(i+1) % 2))/norm[i])
-                  for i, a in enumerate(marg)]
-
-    # Plot the results.
-    labels = ["$\ln T/\mathrm{day}$", "$\ln R/R_\oplus$"]
-    top_axes = ["$T\,[\mathrm{days}]$", "$R\,[R_\oplus]$"]
-    fig = pop.plot_2d(v, censor=censor, catalog=np.log(catalog),
-                      err=err, true=truth, labels=labels,
-                      top_axes=top_axes, literature=literature)
-
-    # Extrapolate to the Earth.
-    o = np.argsort(c[:, 0])
[…]

### 2014-06-12 document/code/results.py

@@ -1,141 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-try:
-    from savefig import monkey_patch
-except ImportError:
-    pass
-else:
-    monkey_patch()
-
-import os
-import sys
-import h5py
-import triangle
-import numpy as np
-import cPickle as pickle
-import matplotlib.pyplot as pl
-
-from emcee.autocorr import integrated_time
-
-import load_data
-
-bp = sys.argv[1]
-model, catalog, err, truth, labels, top_axes, literature = \
-    pickle.load(open(os.path.join(bp, "model.pkl")))
-pop = model.population
-
-# Load the original catalog if it's there.
-fn = os.path.join(bp, "catalog.pkl")
-if os.path.exists(fn):
-    catalog, err, truth = \
-        pickle.load(open(os.path.join(bp, "catalog.pkl")))
-
-    # Hack to deal with negative radii.
-    m = catalog[:, 1] > 0
-    catalog = catalog[m]
-    rerr = [np.log(catalog[:, 1]) - np.log(np.abs(catalog[:, 1]-err[m, 1])),
-            np.log(catalog[:, 1]+err[m, 1]) - np.log(catalog[:, 1])]
-    err = [0, rerr]
-
-labels = ["$\ln P/\mathrm{day}$", "$\ln R/R_\oplus$"]
-top_axes = ["$P\,[\mathrm{days}]$", "$R\,[R_\oplus]$"]
-
-
-with h5py.File(os.path.join(bp, "results.h5")) as f:
-    i = int(f.attrs["iteration"])
-    samples = f["samples"][:i, :]
-    hyper = f["hyper"][:i, :]
-    lnprob = f["lnprob"][:i]
-
-for i in range(hyper.shape[1]):
-    pl.clf()
-    pl.plot(hyper[:, i])
-    pl.savefig(os.path.join(bp, "time-hyper-{0:03d}.png".format(i)))
-
-pl.clf()
-pl.plot(lnprob)
-pl.savefig(os.path.join(bp, "time-lnprob.png"))
-
-nstar = 42557.0
-ntot = 200000
-samples = samples[-ntot:, :]  # [::50, :]
-
-# Reformat the samples and save the samples.
-thin_by = int(np.min(integrated_time(samples, axis=0)))
[…]

### 2014-08-11 code/main.py
adding option to ignore uncertainties

@@ -92,7 +92,9 @@ def inverse_detection_efficiency(pop, censor, catalog, err, truth=None):
     return val, var, literature, fig, fig2, (q[1], np.diff(q))
 
 
-def main(bp, real_data, ep_bins=False):
+def main(bp, real_data, ep_bins=False, ignore_uncert=False):
+    if ignore_uncert:
+        bp = "{0}-no-uncert".format(bp)
     try:
         os.makedirs(bp)
     except os.error:
@@ -111,10 +113,13 @@ def main(bp, real_data, ep_bins=False):
         truth = None
         m = np.log(catalog[:, 0]) > np.min(x)
         catalog, err = catalog[m], err[m]
+        if ignore_uncert:
+            err = np.zeros_like(err)
     else:
         catalog, err, truth = \
             pickle.load(open(os.path.join(bp, "catalog.pkl")))
-    dataset = Dataset.sample(catalog, err, samples=256, censor=censor,
+    K = 1 if ignore_uncert else 256
+    dataset = Dataset.sample(catalog, err, samples=K, censor=censor,
                              functions=[np.log, np.log])
     print("{0} entries in catalog".format(dataset.catalogs.shape[1]))
 
@@ -199,7 +204,11 @@ def main(bp, real_data, ep_bins=False):
 
 if __name__ == "__main__":
     import sys
+    ignore = False
+    if "--ignore" in sys.argv:
+        sys.argv.remove("--ignore")
+        ignore = True
     if len(sys.argv) > 1:
-        main(sys.argv[1], False)
+        main(sys.argv[1], False, ignore_uncert=ignore)
     else:
-        main("results", True)
+        main("results", True, ignore_uncert=ignore)

### 2014-08-11 document/ms.tex
acking the referee
samsi ack

@@
 for helpful contributions to the ideas and code presented here.
+We would also like to acknowlege the anonymous referee for suggestions that
+substantially improved the paper.
 This project was partially supported by the NSF (grant AST-0908357), NASA
 (grant NNX08AJ48G), and the Moore--Sloan Data Science Environment at NYU.
-This research builds on ideas generated at a three-week workshop on ``Modern
-Statistical and Computational Methods for Analysis of Kepler Data'' at SAMSI
-in 2013 June.
+This research builds on ideas generated at a three-week workshop supported by
+NSF Grant DMS-1127914 to the Statistical and Applied Mathematical Sciences
+Institute.
 This research made use of the NASA \project{Astrophysics Data System}.

### 2014-08-21 document/ms.tex
moving derivation of inverse-detection-efficiency to appendix

@@
 
-\paragraph{Inverse-detection-efficiency}
-It's now interesting to take a brief aside and discuss the connection between
-our model and the commonly used inverse-detection-efficiency procedure
-(\citealt{howard,dressing,petigura}).
-This procedure involves making a weighted histogram of the catalog entries
-where the weight for object $\entry_k$ is $1/\completeness(\entry_k)$.
-This makes intuitive sense but the standard formulation does not have a clear
-probabilistic justification or interpretation.
-
-If we model the occurrence rate density as a histogram with $J$ fixed bin
-volumes $\binarea_j$
-\begin{eqnarray}
-\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
-\exp(\ratepar_1) & \entry \in \binarea_1,\\
-\exp(\ratepar_2) & \entry \in \binarea_2,\\
-\cdots \\
-\exp(\ratepar_J) & \entry \in \binarea_J,\\
-0 & \mathrm{otherwise}
-\end{array}\right.
-\end{eqnarray}
-then \eq{poisson-like} becomes
-\begin{eqnarray}
-\ln p(\{\entry_k\}\,|\,\ratepars) &=&
-    \sum_{k=1}^K \sum_{j=1}^J \mathbf{1}[\entry_k \in
-        \binarea_j]\,[\ln\completeness(\entry_k)+\ratepar_j]
-    -\sum_{j=1}^J\exp(\ratepar_j)\,
-        \int_{\binarea_j} \completeness(\entry)\dd\entry
-\end{eqnarray}
-where the indicator function $\mathbf{1}[\cdot]$ is one if $\cdot$ is true and
-zero otherwise.
-If we take the gradient of this function with respect to \ratepars\ and set it
-equal to zero, we find the maximum likelihood result
-\begin{eqnarray}\eqlabel{ml-ide}
-\exp ({\ratepar_j}^*) &=&
-\frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
-\end{eqnarray}
-where $N_j$ is the number of objects that fall within the bin $j$.
-We estimate the uncertainty $\delta\ratepar_j$ on this value by examining the
-curvature of the log-likelihood function near the maximum and find
-\begin{eqnarray}
-\frac{\delta\exp({\ratepar_j}^*)}{\exp({\ratepar_j}^*)} &=&
-    \frac{1}{\sqrt{N_j}}
-\quad.
-\end{eqnarray}
-
-Since \eq{ml-ide} gives the maximum likelihood solution, this result is the
-minimum-variance estimator for the rate density when the completeness function
-$\completeness(\entry)$ is known and the catalog uncertainties are negligible.
-This is \emph{almost} the same as the inverse-detection-efficiency procedure
-and if the completeness function is constant in each bin \binarea, both
-methods give the same result.
-If the completeness function is not piecewise-constant on the same grid as the
-rate density inference, the inverse-detection-efficiency procedure will
-give a higher variance estimate of the occurrence rate density than
-\eq{ml-ide}.
-In this \paper, we don't use either of these inverse-detection-efficiency
-procedures because they both fundamentally rely on the assumption of
-\emph{negligible observational uncertainties}.
-By applying \eq{money}, we derive the probabilistic generalization of this
-model in the next section but the result is no longer analytic.
-
 \paragraph{Occurrence rate density parameterization \& priors}
@@
 
+\appendix
+
+\section*{APPENDIX}
+
+\section{Inverse-detection-efficiency}
+
+It's now interesting to take a brief aside and discuss the connection between
+our model and the commonly used inverse-detection-efficiency procedure
+(\citealt{howard,dressing,petigura}).
+This procedure involves making a weighted histogram of the catalog entries
+where the weight for object $\entry_k$ is $1/\completeness(\entry_k)$.
+This makes intuitive sense but the standard formulation does not have a clear
+probabilistic justification or interpretation.
+
+If we model the occurrence rate density as a histogram with $J$ fixed bin
+volumes $\binarea_j$
+\begin{eqnarray}
+\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
+\exp(\ratepar_1) & \entry \in \binarea_1,\\
+\exp(\ratepar_2) & \entry \in \binarea_2,\\
+\cdots \\
+\exp(\ratepar_J) & \entry \in \binarea_J,\\
+0 & \mathrm{otherwise}
+\end{array}\right.
+\end{eqnarray}
+then \eq{poisson-like} becomes
+\begin{eqnarray}
+\ln p(\{\entry_k\}\,|\,\ratepars) &=&
+    \sum_{k=1}^K \sum_{j=1}^J \mathbf{1}[\entry_k \in
+        \binarea_j]\,[\ln\completeness(\entry_k)+\ratepar_j]
+    -\sum_{j=1}^J\exp(\ratepar_j)\,
+        \int_{\binarea_j} \completeness(\entry)\dd\entry
+\end{eqnarray}
+where the indicator function $\mathbf{1}[\cdot]$ is one if $\cdot$ is true and
+zero otherwise.
+If we take the gradient of this function with respect to \ratepars\ and set it
+equal to zero, we find the maximum likelihood result
+\begin{eqnarray}\eqlabel{ml-ide}
+\exp ({\ratepar_j}^*) &=&
+\frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
+\end{eqnarray}
+where $N_j$ is the number of objects that fall within the bin $j$.
+We estimate the uncertainty $\delta\ratepar_j$ on this value by examining the
+curvature of the log-likelihood function near the maximum and find
+\begin{eqnarray}
+\frac{\delta\exp({\ratepar_j}^*)}{\exp({\ratepar_j}^*)} &=&
+    \frac{1}{\sqrt{N_j}}
+\quad.
+\end{eqnarray}
+
+Since \eq{ml-ide} gives the maximum likelihood solution, this result is the
+minimum-variance estimator for the rate density when the completeness function
+$\completeness(\entry)$ is known and the catalog uncertainties are negligible.
+This is \emph{almost} the same as the inverse-detection-efficiency procedure
+and if the completeness function is constant in each bin \binarea, both
+methods give the same result.
+If the completeness function is not piecewise-constant on the same grid as the
+rate density inference, the inverse-detection-efficiency procedure will
+give a higher variance estimate of the occurrence rate density than
+\eq{ml-ide}.
+In this \paper, we don't use either of these inverse-detection-efficiency
+procedures because they both fundamentally rely on the assumption of
+\emph{negligible observational uncertainties}.
+By applying \eq{money}, we derive the probabilistic generalization of this
+model in the next section but the result is no longer analytic.
+
 \newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}

### 2014-08-21 document/ms.tex
adjusting wording of intro discussion of inv-det-eff

@@
 youdin, dong}).
-As we show in \sect{model}, if we model the occurrence rate density as a
-piecewise constant function, the inverse-detection-efficiency method can be
-derived as a special case of the likelihood method with a specific set of
-assumptions.
+In this \paper, we start from the likelihood method but model the rate density
+non-parametrically as a piecewise-constant step function.
+Using this formulation of the problem, we derive a generalization that takes
+observational uncertainties into account.
+In \app{inv-det-eff}, we show that the inverse-detection-efficiency method can
+be derived as a special case of the likelihood method in the limit of a
+smoothly varying completeness function.
 
@@
 Jo Bovy (IAS),
-Eric Ford (PSU), and
-Ben Montet (Caltech/Harvard)
+Eric Ford (PSU),
+David Kipping (CfA),
+Ben Montet (Caltech/Harvard), and
+Scott Tremaine (IAS)
 for helpful contributions to the ideas and code presented here.
@@
 \section{Inverse-detection-efficiency}
+\sectlabel{inv-det-eff}
 

### 2014-08-21 document/ms.tex
re-wording of summary paragraph

@@
 \newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
-\newcommand{\appropto}{\mathrel{\vcenter{
-  \offinterlineskip\halign{\hfil$##$\cr
-    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
 \newcommand{\densityunit}{{\ensuremath{\mathrm{nat}^{-2}}}}
@@
 
-In the next Section, we introduce the concept of hierarchical inference, how
-it applies to the problem at hand, and we derive the importance sampling
-approximation to the marginalized likelihood (this is a generalization of the
-method from \citealt{hogge}).
-Then, in \sect{model}, we describe how to include detection efficiencies in
-this model and derive a probabilistic motivation for the standard
-inverse-detection-efficiency weighting procedure.
+In the next Section, we introduce our method for incorporating the effects of
+observational uncertainties into the derivation of occurrence rates.
+The technical term for this procedure is \emph{hierarchical inference} and
+while a general discussion of this field is beyond the scope of this \paper,
+in the next section, we present the basic probabilistic question and derive a
+computationally tractable method for inferring the population of exoplanets.
+In \sect{model}, we explain how to incorporate the effects of survey
+completeness into our inferences and derive the key equation for our method:
+\eq{money}.
 We test our method on synthetic catalogs in Sections~\sectalt{valid}
@@
 determined detection efficiency from \citet{petigura} to measure the
-occurrence rate density of Earth analogs.
+occurrence rate density of small planets with long orbital periods.
 
@@
 for helpful contributions to the ideas and code presented here.
-We would also like to acknowlege the anonymous referee for suggestions that
+We would also like to acknowledge the anonymous referee for suggestions that
 substantially improved the paper.

### 2014-08-21 document/ms.tex
clearer (hopefully) description of interim priors

@@
 statistics} of the data then we can, in theory, compute \eq{crazylike}---up to
-an unimportant constant---without ever looking at the raw data again.
-This is permitted because the catalog provides constraints on the posterior
-probability of the parameters $\{\entry_k\}$ under some choice of
-\emph{interim prior}
+an unimportant constant---without ever looking at the raw data again!
+This is important because the high-dimensional integral in \eq{crazylike}
+won't generally have an analytic solution and each evaluation of the
+per-object likelihood $p(\data_k\,|\,\entry_k)$ is expensive, making numerical
+methods intractable.
+
+Instead, we will reuse the hard work that went into building the catalog.
+We must first notice that each entry in a catalog is a representation of the
+posterior probability
 \begin{eqnarray}\eqlabel{crazypost}
-p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
-\frac{p(\{\data_k\}\,|\,\{\entry_k\})\,p(\{\entry_k\}\,|\,\interim)}
-     {p(\{\data_k\}\,|\,\interim)} \quad.
-\end{eqnarray}
-The interim prior $p(\{\entry_k\}\,|\,\interim)$ is the prior
-function---parameterized by \interim---that the author of the catalog chose
-when making their measurements.
-
-It turns out that we can use these posterior measurements to simplify
-\eq{crazylike} to a form that can, in many common cases, be evaluated
-efficiently.
+p(\entry_k\,|\,\data_k,\,\interim) &=&
+\frac{p(\data_k\,|\,\entry_k)\,p(\entry_k\,|\,\interim)}
+     {p(\data_k\,|\,\interim)}
+\end{eqnarray}
+of the parameters $\entry_k$ conditioned on the observations of that object
+$\data_k$.
+The notation \interim\ is a reminder that the catalog was produced under a
+specific choice of some---probably ``uninformative''---\emph{interim prior}
+$p(\entry_k\,|\,\interim)$.
+This prior was chosen by the author of the catalog and it is different from
+the distribution that we want to infer: $p(\entry_k\,|\,\ratepars)$.
+
+Now, we can use these posterior measurements to simplify \eq{crazylike} to a
+form that can, in many common cases, be evaluated efficiently.
 To find this result, multiply the integrand in \eq{crazylike} by
@@
 \frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
-     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
+     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)} &=&
+\prod_{k=1}^K
+\frac{p(\entry_k\,|\,\data_k,\,\interim)}
+     {p(\entry_k\,|\,\data_k,\,\interim)}
 \end{eqnarray}

### 2014-08-22 document/ms.tex
moving likelihood method discussion earlier

@@
 
+\section{The likelihood method}
+\sectlabel{likelihood}
+
+The first ingredient for any probabilistic inference is a likelihood function;
+a description of the probability of observing a specific dataset given a set
+of model parameters.
+In this particular project, the dataset is a catalog of exoplanet measurements
+and the model parameters are the values that set the shape and normalization
+of the occurrence rate density.
+Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
+occurrence rate density \rate---parameterized by the parameters \ratepars---as
+a function of the physical parameters \entry\ (orbital period, planetary
+radius, \etc).
+In this framework, the occurrence rate density can be ``parametric''---for
+example, a power law---or a ``non-parametric'' function---such as a histogram
+where the bin heights are the parameters \ratepars.
+
+We'll model the catalog as a draw from the inhomogeneous Poisson process set
+by the \emph{observable} rate density $\obs{\rate}_\ratepars$.
+This leads to the previously known result (see \citealt{tabachnik,youdin} for
+some of the examples from the exoplanet literature)
+\begin{eqnarray}\eqlabel{poisson-like}
+p(\{\entry_k\}\,|\,\ratepars) &=&
+    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
+    \prod_{k=1}^K \obs{\rate}_\ratepars (\entry_k)\quad.
+\end{eqnarray}
+In this equation, the integral in the normalization term is the expected
+number of observable exoplanets in the sample.
+
+The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate
+density of exoplanets that you would expect to observe taking into account the
+geometric transit probability and any other detection efficiencies.
+In practice, we can model the observable rate density as
+\begin{eqnarray}\eqlabel{obs-rate}
+\obs{\rate}_\ratepars(\entry) &=&
+    \completeness(\entry)\,\rate_\ratepars(\entry)
+\end{eqnarray}
+where $\completeness(\entry)$ is the detection efficiency (including transit
+probability) at \entry\ and $\rate_\ratepars(\entry)$ is the object that we
+want to infer: the \True\ occurrence rate density.
+We haven't yet discussed any specific functional form for
+$\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
+whether we model the rate density as, for example, a broken power law or a
+histogram.
+
+The observed rate density \obs{\rate} is a quantitative description of the
+rate density at which planets appear in the \citet{petigura} catalog; it is
+not a description of the \True\ rate density of exoplanets.
+Inasmuch as the detection efficiency $\completeness(\entry)$ is calculated
+correctly, the function $\rate_\ratepars(\entry)$ will represent the \True\
+rate density of exoplanets, at least where there is support in the data.
+In practice, an estimate of the detection efficiency will not include every
+decision or effect in the pipeline.
+One example in this case is that the completeness function does not account
+for the explicit removal of second (and third, and so on) planets.
+In principle a more complete description of the detection efficiency can be
+included and as this function becomes more accurate, our inferences about the
+\True\ rate density $\rate_\ratepars(\entry)$ will be less biased.
+
+For the results in this \paper, we will assume that the completeness function
+$\completeness(\entry)$ is known empirically on a grid in period and radius
+but that is not a requirement for the validity of this method.
+Instead, we could use a functional form for the completeness and even infer
+its parameters along with the parameters of the rate density.
+
+Finally, for all the results in this \paper, we'll model the rate density as
+a piecewise constant step function
+\begin{eqnarray}\eqlabel{rate-model}
+\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
+\exp(\ratepar_1) & \entry \in \binarea_1,\\
+\exp(\ratepar_2) & \entry \in \binarea_2,\\
+\cdots \\
+\exp(\ratepar_J) & \entry \in \binarea_J,\\
+0 & \mathrm{otherwise}
+\end{array}\right.
+\end{eqnarray}
+where the parameters $\ratepar_j$ are the log step heights and the bin edges
+are fixed \foreign{a priori}.
+
+In \app{inv-det-eff}, we use the model in \eq{rate-model} and neglect
+observational uncertainties, we can derive the analytic maximum likelihood
+solution for the step heights.
+This result is similar to and just as simple as the
+inverse-detection-efficiency method and it is guaranteed to provide a lower
+variance estimate of the rate density than the standard procedure.
+
+One major benefit of expressing the problem of occurrence rate inference
+probabilistically is that it can now be formally extended to include the
+effects of observational uncertainties.
+
 \section{A brief introduction to hierarchical inference}
@@
 The notation \interim\ is a reminder that the catalog was produced under a
-specific choice of some---probably ``uninformative''---\emph{interim prior}
+specific choice of a---probably ``uninformative''---\emph{interim prior}
 $p(\entry_k\,|\,\interim)$.
 This prior was chosen by the author of the catalog and it is different from
-the distribution that we want to infer: $p(\entry_k\,|\,\ratepars)$.
+the likelihood $p(\entry_k\,|\,\ratepars)$ from \eq{poisson-like}.
 
@@
 \sectlabel{model}
-
-In the inference procedure described above, we only discussed an abstract
-occurrence rate density.
-To make it more concrete, we'll model the catalog as a draw from the
-inhomogeneous Poisson process set by the \emph{observable} rate density
-$\obs{\rate}_\ratepars$.
-This leads to the previously known result (see \citealt{tabachnik,youdin} for
-some of the examples from the exoplanet literature)
-\begin{eqnarray}\eqlabel{poisson-like}
-p(\{\entry_k\}\,|\,\ratepars) &=&
-    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
-    \prod_{k=1}^K \obs{\rate}_\ratepars (\entry_k)\quad.
-\end{eqnarray}
-In this equation, the integral in the normalization term is the expected
-number of observable exoplanets in the sample.
-
-The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate
-density of exoplanets that you would expect to observe taking into account the
-geometric transit probability and any other detection efficiencies.
-In practice, we can model the observable rate density as
-\begin{eqnarray}\eqlabel{obs-rate}
-\obs{\rate}_\ratepars(\entry) &=&
-    \completeness(\entry)\,\rate_\ratepars(\entry)
-\end{eqnarray}
-where $\completeness(\entry)$ is the detection efficiency (including transit
-probability) at \entry\ and $\rate_\ratepars(\entry)$ is the object that we
-want to infer: the \True\ occurrence rate density.
-Up to this point, we haven't discussed any specific functional form for
-$\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
-whether we model the rate density as, for example, a broken power law or
-a histogram.
-
-The observed rate density \obs{\rate} is a quantitative description of the
-rate density at which planets appear in the \citet{petigura} catalog; it is
-not a description of the \True\ rate density of exoplanets.
-Inasmuch as the detection efficiency $\completeness(\entry)$ is calculated
-correctly, the function $\rate_\ratepars(\entry)$ will represent the \True\
-rate density of exoplanets, at least where there is support in the data.
-In practice, an estimate of the detection efficiency will not include every
-decision or effect in the pipeline.
-One example in this case is that the completeness function does not account
-for the explicit removal of second (and third, and so on) planets.
-In principle a more complete description of the detection efficiency can be
-included and as this function becomes more accurate, our inferences about the
-\True\ rate density $\rate_\ratepars(\entry)$ will be less biased.
-
-For the results in this \paper, we will assume that the completeness function
-$\completeness(\entry)$ is known empirically on a grid in period and radius
-but that is not a requirement for the validity of this method.
-Instead, we could use a functional form for the completeness and even infer
-its parameters along with the parameters of the rate density.
 

### 2014-08-23 document/ms.tex
summarizing

@@
 
-In the next Section, we introduce our method for incorporating the effects of
-observational uncertainties into the derivation of occurrence rates.
+In the next Section, we summarize the likelihood method for exoplanet
+population inference and in \sect{hier}, we describe how to include the
+effects of observational uncertainties.
 The technical term for this procedure is \emph{hierarchical inference} and
 while a general discussion of this field is beyond the scope of this \paper,
-in the next section, we present the basic probabilistic question and derive a
-computationally tractable method for inferring the population of exoplanets.
-In \sect{model}, we explain how to incorporate the effects of survey
-completeness into our inferences and derive the key equation for our method:
-\eq{money}.
+in the \sect{hier}, we present the basic probabilistic question and derive a
+computationally tractable inference procedure.
+In \sect{model}, we summarize the technique and derive the key equation for
+our method: \eq{money}.
 We test our method on synthetic catalogs in Sections~\sectalt{valid}
@@
 \section{A brief introduction to hierarchical inference}
-
-In this \paper, we are asking the question: \emph{what constraints can we put
-on the occurrence rate density of exoplanets given all the light curves
-measured by \kepler?}
-Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
-occurrence rate density \rate---parameterized by the parameters \ratepars---as
-a function of the physical parameters \entry\ (orbital period, planetary
-radius, \etc).
-In this framework, the occurrence rate density can be ``parametric''---for
-example, a power law---or a ``non-parametric'' function---such as a histogram
-where the bin heights are the parameters \ratepars.
-The light curve for a particular target $k$ is $\data_k$ and the \True\ set of
-physical parameters for that same object is given by $\entry_k$.
-A catalog provides a set of constraints on the parameters $\{\entry_k\}$
-conditioned on the set light curves $\{\data_k\}$.
-
-Any inference about the rate density parameters can be formally addressed by
-evaluating the \emph{marginalized likelihood function}
+\sectlabel{hier}
+
+The general question that we are trying to answer in this \paper\ is:
+\emph{what constraints can we put on the occurrence rate density of exoplanets
+given all the light curves measured by \kepler?}
+In the case of negligible measurement uncertainties, this is equivalent to
+optimizing \eq{poisson-like} but when this approximation is no longer valid,
+we must instead compute the \emph{marginalized likelihood}
 \begin{eqnarray}\eqlabel{crazylike}
@@
 \end{eqnarray}
-where, as we mentioned above, $\{\data_k\}$ is the set of all light curves,
-one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
-parameters describing the population occurrence rate density
-$\rate_\ratepars(\entry)$ and $\entry_k$ is the vector of physical parameters
-describing the planetary system (orbital periods, radius ratios, stellar
-radius, \etc) around target $k$.
+where $\{\data_k\}$ is the set of all light curves, one light curve $\data_k$
+per target $k$, \ratepars\ is the vector of parameters describing the
+population occurrence rate density $\rate_\ratepars(\entry)$ and $\entry_k$ is
+the vector of physical parameters describing the planetary system (orbital
+periods, radius ratios, stellar radius, \etc) around target $k$.
 In this equation, our only assumption is that the datasets depend on the
@@
 
-In other words, we treat a catalog as being a dimensionality reduction of the
-raw data with all the relevant information retained.
+In other words, we treat the catalog as being a dimensionality reduction of
+the raw data with all the relevant information retained.
 In the context of \kepler, the catalog reduces the set of downloaded pixel
@@
     \frac{\obs{\rate}_\ratepars (\entry_k^{(n)})}
-         {p(\entry_k^{(n)}\,|\,\interim)} \quad.
-\end{eqnarray}
+         {p(\entry_k^{(n)}\,|\,\interim)}
+\end{eqnarray}
+where the values $\{{\entry_k}^{(n)}\}$ are samples drawn from the posterior
+probability
+\begin{eqnarray}
+{\entry_k}^{(n)} &\sim& p(\entry_k\,|\,\data_k,\,\interim)
+\end{eqnarray}
+as described in the previous section.
+\Eq{money} is the \emph{money equation} for our method.
+It lets us efficiently compute the \emph{marginalized likelihood of the entire
+set of light curves for a particular occurrence rate density}.
+
 In this equation, we're making the further assumption that the catalog treated
@@
 one transiting planet---a point that we will return to below---but for the
-considerations of the dataset considered in this \paper, it is a justified
+considerations of the dataset considered here, it is a justified
 simplification.
-In other words, we have per-object posterior samples
-\begin{eqnarray}
-\entry_k^{(n)} &\sim& p(\entry_k\,|\,\data_k,\,\interim) \quad.
-\end{eqnarray}
-\Eq{money} is the \emph{money equation} for our method.
-It lets us efficiently compute the \emph{marginalized likelihood of the entire
-set of light curves for a particular occurrence rate density}.
-
-\paragraph{Occurrence rate density parameterization \& priors}
-For the remainder of this \paper, we also model the rate density as a
+
+For the remainder of this \paper, we model the rate density as a
 two-dimensional histogram with fixed logarithmic bins in period and radius.
 When we include observational uncertainties---using \eq{money}---the maximum
-likelihood result is no longer analytic; \eq{ml-ide} is no longer applicable
-when the uncertainties are non-negligible.
+likelihood result is no longer analytic.
 Therefore, if we want to compute the ``best-fit'' rate density, we can use a
 standard non-linear optimization algorithm.
+
 In the regions of parameter space that we tend to care about, the completeness

### 2014-08-23 document/ms.tex
note to self
re-writing inverse-detection-efficiency derivation

@@
 transiting systems (\citealt{petigura}).
-This is a limitation of the input catalog.
-When considering other datasets it will be interesting and important to take
-multiple transiting systems into account (\citealt{lissauer, tremaine, fang})
-and this would be possible using conditional distributions (as developed by
-\citealt{tremaine}).
 The second assumption---neglecting false positives---is also strong and only
@@
 In practice, an estimate of the detection efficiency will not include every
-decision or effect in the pipeline.
-One example in this case is that the completeness function does not account
-for the explicit removal of second (and third, and so on) planets.
-In principle a more complete description of the detection efficiency can be
-included and as this function becomes more accurate, our inferences about the
-\True\ rate density $\rate_\ratepars(\entry)$ will be less biased.
+decision or effect in the pipeline and as this function becomes more accurate,
+our inferences about the \True\ rate density $\rate_\ratepars(\entry)$ will be
+less biased.
 
@@
 \end{eqnarray}
-where the parameters $\ratepar_j$ are the log step heights and the bin edges
-are fixed \foreign{a priori}.
+where the parameters $\ratepar_j$ are the log step heights and the bins
+$\binarea_j$ are fixed \foreign{a priori}.
 
@@
 experiments, the results do not seem to be very sensitive to this choice.
+
+\dfmtodo{Include Kipping's comments about eccentricity.}
 
@@
 
-In detail, any inference of the exoplanet population depends on the choices
-made in the pipeline that creates the catalog.
-Our results are based on a catalog that only includes single transiting
-systems (\citealt{petigura}) so our conclusions---and theirs---must be
-considered with this in mind.
-In particular, the exclusion of second (and third, and so on) planets must
-suppress the inferred rate density, especially at long periods and small
-radii.
+% In detail, any inference of the exoplanet population depends on the choices
+% made in the pipeline that creates the catalog.
+% Our results are based on a catalog that only includes single transiting
+% systems (\citealt{petigura}) so our conclusions---and theirs---must be
+% considered with this in mind.
+% In particular, the exclusion of second (and third, and so on) planets must
+% suppress the inferred rate density, especially at long periods and small
+% radii.
 
@@
 for helpful contributions to the ideas and code presented here.
-We would also like to acknowledge the anonymous referee for suggestions that
+We would also like to acknowledge the anonymous referee and the Scientific
+Editor, Eric Feigelson, for suggestions that
 substantially improved the paper.
@@
 
-It's now interesting to take a brief aside and discuss the connection between
-our model and the commonly used inverse-detection-efficiency procedure
-(\citealt{howard,dressing,petigura}).
-This procedure involves making a weighted histogram of the catalog entries
-where the weight for object $\entry_k$ is $1/\completeness(\entry_k)$.
-This makes intuitive sense but the standard formulation does not have a clear
-probabilistic justification or interpretation.
+One huge benefit of the inverse-detection-efficiency procedure is its
+simplicity.
+Therefore, it's worth noting that there is a probabilistically justified
+procedure that will always provide less biased results while being only
+marginally more complicated.
+
+The standard procedure involves making a weighted histogram of the catalog
+entries where the weight for object $\entry_k$ is $1/\completeness(\entry_k)$.
+This makes intuitive sense but it does not have a clear probabilistic
+justification or interpretation.
+As we will show below, the maximum likelihood result involves weighting the
+points by the inverse of the \emph{integral} of the completeness function over
+the bin area.
+
+To motivate this derivation, let's start by considering the following
+pathological example: a single bin where the completeness sharply drops from
+one to zero halfway across the bin.
+If we observe $K$ objects in this bin, we would have observed about $2K$
+objects in a complete sample.
+If we apply the inverse-detection-efficiency procedure to this dataset, each
+sample will get unit weight because they are all found in the part of the bin
+where the completeness is one.
+Therefore, we would \emph{underestimate} the true rate in the bin by half.
+It's clear in this specific case that giving the points a weight of two would
+give a better solution and we'll derive the general result below.
 
 If we model the occurrence rate density as a histogram with $J$ fixed bin
-volumes $\binarea_j$
-\begin{eqnarray}
-\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
-\exp(\ratepar_1) & \entry \in \binarea_1,\\
-\exp(\ratepar_2) & \entry \in \binarea_2,\\
-\cdots \\
-\exp(\ratepar_J) & \entry \in \binarea_J,\\
-0 & \mathrm{otherwise}
-\end{array}\right.
-\end{eqnarray}
-then \eq{poisson-like} becomes
+volumes $\binarea_j$ (\eqalt{rate-model}) then \eq{poisson-like} becomes
 \begin{eqnarray}
@@
 zero otherwise.
-If we take the gradient of this function with respect to \ratepars\ and set it
+Taking the gradient of this function with respect to \ratepars\ and setting it
 equal to zero, we find the maximum likelihood result
@@
 \exp ({\ratepar_j}^*) &=&
-\frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
-\end{eqnarray}
-where $N_j$ is the number of objects that fall within the bin $j$.
+\frac{K_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
+\end{eqnarray}
+where $K_j$ is the number of objects that fall within the bin $j$.
 We estimate the uncertainty $\delta\ratepar_j$ on this value by examining the
@@
 \begin{eqnarray}
-\frac{\delta\exp({\ratepar_j}^*)}{\exp({\ratepar_j}^*)} &=&
-    \frac{1}{\sqrt{N_j}}
+\frac{\delta\exp({\ratepar_j}^*)}{\exp({\ratepar_j}^*)}
+&=& \frac{1}{\sqrt{K_j}}
 \quad.
@@
 
-Since \eq{ml-ide} gives the maximum likelihood solution, this result is the
-minimum-variance estimator for the rate density when the completeness function
-$\completeness(\entry)$ is known and the catalog uncertainties are negligible.
-This is \emph{almost} the same as the inverse-detection-efficiency procedure
-and if the completeness function is constant in each bin \binarea, both
-methods give the same result.
-If the completeness function is not piecewise-constant on the same grid as the
-rate density inference, the inverse-detection-efficiency procedure will
-give a higher variance estimate of the occurrence rate density than
-\eq{ml-ide}.
-In this \paper, we don't use either of these inverse-detection-efficiency
-procedures because they both fundamentally rely on the assumption of
-\emph{negligible observational uncertainties}.
-By applying \eq{money}, we derive the probabilistic generalization of this
-model in the next section but the result is no longer analytic.
+In our pathological example from above, the integral of the completeness
+function over the bin is $1/2$, giving each sample the expected weight of $2$.
+In more realistic cases, where the completeness function varies smoothly, the
+inverse-detection-efficiency result will begin to agree with \eq{ml-ide} but
+the severity of this bias will be very problem dependent.
+Therefore, if you have a dataset with negligible observational uncertainties,
+we recommend that you always apply \eq{ml-ide} instead of the standard
+inverse-detection-efficiency procedure.
+As the uncertainties become more significant, there is no longer an analytic
+result and the method derived in this \paper\ is necessary.
 

### 2014-08-25 code/main.py
updating figures to new run

@@ -118,7 +118,7 @@ def main(bp, real_data, ep_bins=False, ignore_uncert=False):
     else:
         catalog, err, truth = \
             pickle.load(open(os.path.join(bp, "catalog.pkl")))
-    K = 1 if ignore_uncert else 256
+    K = 1 if ignore_uncert else 512
     dataset = Dataset.sample(catalog, err, samples=K, censor=censor,
                              functions=[np.log, np.log])
     print("{0} entries in catalog".format(dataset.catalogs.shape[1]))

### 2014-08-25 code/results.py

@@ -76,6 +76,16 @@ grids = thinned.reshape((len(thinned), pop.shape[0], pop.shape[1]))
 print(grids.shape)
 print([b.shape for b in pop.bins])
 
+
+def xmap(f, i):
+    return (f(*x) for x in i)
+
+h_mu = np.mean(np.exp(0.5 * hyper[-ntot:][::thin_by, 2:]), axis=0)
+h_std = np.std(np.exp(0.5 * hyper[-ntot:][::thin_by, 2:]), axis=0)
+print("Hyper:")
+print("\n".join(xmap("{0} ± {1}".format, zip(h_mu, h_std))))
+print()
+
 with h5py.File(os.path.join(bp, "samples.h5"), "w") as f:
     f.create_dataset("hyperparameter_samples", data=hyper[-ntot:][::thin_by])
     f.create_dataset("ln_occurrence_rate_samples", data=grids)
@@ -123,13 +133,14 @@ ax.axvline(np.log(ext[0]), color="k")
 # ax.axvline(c, color="k", ls="dashed")
 # ax.axvline(a, color="k", ls="dashed")
 
+ax.set_xlim(-7, -1)
 ax.set_xlabel(r"$\ln \Gamma_\oplus$")
 ax.set_ylabel(r"$p(\ln \Gamma_\oplus)$")
 
 a2 = ax.twiny()
-a2.set_xlim(100 * np.exp(ax.get_xlim()))
+a2.set_xlim(np.exp(ax.get_xlim()))
 a2.set_xscale("log")
-a2.set_xlabel(r"$\Gamma_\oplus\,[\%]$")
+a2.set_xlabel(r"$\Gamma_\oplus$")
 
 fig.subplots_adjust(top=0.85)
 

### 2014-08-25 document/figures/results/make_plots.py
new figures

@@ -31,8 +31,6 @@ q = quantile(num, [0.16, 0.5, 0.84])
 print("Integrated observable number: {0:.3f}_{{-{1:.3f}}}^{{+{2:.3f}}}"
       .format(q[1], *(np.diff(q))))
 
-assert 0
-
 bins = [ln_period_bin_edges, ln_radius_bin_edges]
 labels = ["R / R_\oplus", "P / \mathrm{day}"]
 slices = [

### 2014-08-25 document/ms.tex
Adding comparison plot

@@
 \Sigma^{-1} &=& \left(\begin{array}{cc}
-1/\smooth_\period & 0 \\
-0 & 1/\smooth_\radius
+1/\smooth_\period^2 & 0 \\
+0 & 1/\smooth_\radius^2
 \end{array}\right) \quad.
@@
 
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/comparison.pdf}
+\end{center}
+\caption{%
+Comparison of various estimates of \gammaearth.
+From the top, the first value is the number published by \citet{petigura} and
+converted to consistent units.
+The second point shows the value implied by the power law model for the
+occurrence rate of $1-2\,R_\oplus$ planets from \citet{dong}.
+The point labeled ``linear extrapolation'' is the result of modeling the
+distribution of small planets ($1-2\,R_\oplus$) on long periods
+($50-400\,\mathrm{days}$) but allowing the period distribution to be
+\emph{linear} instead of \emph{uniform}.
+The ``negligible uncertainties'' value is given by applying the model
+developed in this \paper\ but with the error bars on radius artificially
+set to zero.
+Finally, the bottom point is the result of our full analysis.
+\figlabel{comparison}}
+\end{figure}
+
 \end{document}

### 2014-08-25 document/ms.tex
actually using more importance samples

@@
 and that the radius posterior is Gaussian in linear radius (with a standard
-deviation given by the published uncertainty), we draw 256 samples for
+deviation given by the published uncertainty), we draw 512 samples for
 $\entry_k$ and use these as an approximation to the posterior probability
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
+\includegraphics[width=0.8\textwidth]{figures/smooth/rate.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
+\includegraphics[width=0.8\textwidth]{figures/simulation/rate.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/results/rate.pdf}
+\includegraphics[width=0.8\textwidth]{figures/results/rate.pdf}
 \end{center}

### 2014-08-27 document/ms.tex
updated numbers
expanding discussion of comparison to previous work

@@
 
-\dfmtodo{Include Kipping's comments about eccentricity.}
+Implicit in the expression for the transit probability in \eq{transitprob} is
+the assumption that all of the planets are on circular orbits.
+Recently, \citet{kipping} demonstrated that when eccentric orbits are
+included, our given value is an underestimate by about 10\%.
+This effect will propagate directly to our inferred rate densities.
+Even though the degeneracy is not exact---due to our choice of priors on the
+rate density parameters---it is not a bad approximation to assume that it is
+and scale the results down by your preferred factor.
+The right thing to do would be to marginalize over this effect directly during
+inference but that exercise is beyond the scope of the current \paper.
 
@@
 
-In \eq{gammaearth}, we define ``Earth analog'' in terms of measurable
-quantities with no mention of habitability or composition.
+In \eq{gammaearth}, we use the symbol $\Gamma$ instead of the more commonly
+used $\eta$ since we define ``Earth analog'' in terms of measurable quantities
+with no mention of habitability or composition.
 This might seem unsatisfying but the composition of an exoplanet is
@@
 \begin{eqnarray}\eqlabel{ge-result}
-\gammaearth &=& 0.017^{+0.018}_{-0.009}~\densityunit
+\gammaearth &=& 0.019^{+0.019}_{-0.010}~\densityunit
 \end{eqnarray}
@@
 &=&
-0.016_{-0.007}^{+0.011}
+0.019_{-0.008}^{+0.010}
 \quad.
@@
 
+\dfmtodo{Do we want to add a comment about giant planets? Or maybe some sort
+of cumulative distribution plot/code snippet?}
+
+\section{Comparison with previous work}
+
 Our inferred rate density of Earth analogs (\eqalt{ge-result}) is not
-consistent with the previously published result even though we use
-\emph{exactly the same dataset}.
-This inconsistency is due to the different assumptions made.
-In particular, for their analysis, \citet{petigura} asserted that the period
-distribution for small planets on long orbits is flat in logarithmic period.
-If we enforce the same constraint, we reproduce their result but when we relax
-this assumption, allowing the data to constrain the period distribution, we
-find evidence for a \emph{decreasing} period distribution as a function of
-logarithmic period (see \fig{period}).
+consistent with previously published results.
+In particular, our result is completely inconsistent with the earlier result
+based on \emph{exactly the same dataset} (\citealt{petigura}).
+This inconsistency is due to the different assumptions made the detailed cause
+merits some investigation.
+The two key differences between our analysis and previous work are \emph{(a)}
+the form of the extrapolation function and \emph{(b)} the presence of
+measurement uncertainties on the planet radii.
+
+To make their estimate of \gammaearth, \citet{petigura} asserted a flat
+distribution in logarithmic period for small planets.
+Our results suggest that the data \emph{do not support} this assumption (see
+\fig{period}).
+We find that the data require a \emph{decreasing} period distribution in the
+relevant range.
+A similar result was also found by \citet{dong} and it is apparent in Figure 2
+of \citet{petigura}.
+
+To test the significance of the choice of extrapolation function, we relax the
+assumption of a uniform period distribution and allow the distribution to be
+linear in the same range ($R=1-2\,R_\oplus$ and $P=50-400\mathrm{d}$).
+Under this model, the likelihood of the catalog of planets in this range can
+be calculated using \eq{poisson-like}.
+We apply uniform priors in the physically allowed range of slopes and
+intercepts for this distribution and estimate the posterior probability for
+the extrapolated rate using MCMC (\citealt{emcee}).
+This results give a much more uncertain and substantially lower estimate for
+the rate of Earth analogs
+\begin{eqnarray}
+\Gamma_\oplus &=& 0.072^{+0.088}_{-0.047} \quad.
+\end{eqnarray}
+With the large error bars, this result is consistent with both results (see
+\fig{comparison} where this value is labeled ``linear extrapolation'') but it
+does not fully account for the discrepancy.
+
+To examine the effects of measurement uncertainties, we repeat our analysis
+with the error bars on the radii artificially set to zero, keeping everything
+else the same.
+This analysis (labeled ``negligible uncertainties'' in \fig{comparison}) gives
+the result
+\begin{eqnarray}
+\Gamma_\oplus &=& 0.040^{+0.031}_{-0.019} \quad.
+\end{eqnarray}
+This result is relatively more precise and higher than our final result and
+consistent with the value obtained with linear extrapolation.
+This confirms the hypothesis that the discrepancy between our result and the
+previously published values is the combined result of both of our key
+generalizations.
+
+For comparison, we have also included the value of \gammaearth\ implied by
+\citet[][their Table 2]{dong}.
+This result is based on a power law fit to the period distribution of small
+planets ($R=1-2\,R_\oplus$) on long periods ($P=10-250\,\mathrm{d}$) in a
+different catalog (\citealt{kepler-catalog}) with a parametric completeness
+model.
+There are a few factors to consider when comparing to this to our analysis.
+Firstly, while \citet{dong} did use a somewhat flexible extrapolation
+function than (a power law in log period), it is still very restrictive
+considering the large range of periods; a broken power law might have been
+more applicable.
+Furthermore, their analysis did not incorporate the effects of measurement
+uncertainties.
+Finally, unlike the \citet{petigura}, the \citet{kepler-catalog} catalog used
+by \citet{dong} includes multiple transiting systems.
+As mentioned previously, the effect of this selection is hard to determine
+without further investigation but it should, intuitively, cause any inference
+based on the \citet{petigura} sample an underestimate of the \True\ rate.
 
@@
 \begin{eqnarray}\eqlabel{ntransit}
-N_{\oplus,\,\mathrm{transiting}} &=& 9.2_{-4.0}^{+5.9}
+N_{\oplus,\,\mathrm{transiting}} &=& 10.6_{-4.5}^{+5.9}
 \end{eqnarray}
@@
 
+\bibitem[Kipping(2014)]{kipping}
+Kipping, D.~M.\ 2014, \arxiv{1408.1393}
+
 \bibitem[Lewis \& Shedler(1979)]{poisson}

### 2014-08-28 document/ms.tex
fixing Petigura number

@@
 Converted to these units, \citet{petigura} measured
-$0.119_{-0.046}^{+0.035}~\densityunit$ for the same quantity (indicated as the
+$0.119_{-0.035}^{+0.046}~\densityunit$ for the same quantity (indicated as the
 vertical lines in \fig{real-rate}).
@@
 of cumulative distribution plot/code snippet?}
+
+\dfmtodo{Add sentence listing posterior constraints on hyperparameters.}
 

### 2014-08-28 document/figures/make_comparison.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+import numpy as np
+import matplotlib.pyplot as pl
+from matplotlib.ticker import MaxNLocator
+
+values = [
+    ("Petigura \emph{et al.}\ (2013)", 0.119, 0.046, 0.035),
+    ("Dong \& Zhu (2013)", 0.086, 0.048, 0.031),
+    ("linear extrapolation", 0.072, 0.087, 0.046),
+    ("negligible uncertainties",
+     0.0397591956934, 0.0309881584307, 0.0190620520124),
+    ("Foreman-Mackey \emph{et al.}\ (2014)",
+     0.0188817594233, 0.0193141328059, 0.0104440671941),
+]
+
+fig = pl.figure(figsize=(8, 4))
+for i, v in enumerate(values):
+    pl.plot(np.log(v[1]), i, "ok")
+    pl.plot(np.log([v[1]+v[2], v[1]-v[3]]), [i, i], "k")
+
+pl.gca().set_yticklabels([""] + [v[0] for v in values])
+pl.xlabel(r"$\ln\Gamma_\oplus$")
+fig.subplots_adjust(left=0.48, bottom=0.17, right=0.97, top=0.98)
+pl.ylim(4.5, -0.5)
+pl.gca().xaxis.set_major_locator(MaxNLocator(5))
+pl.savefig("comparison.pdf")

### 2014-08-28 document/ms.tex
typos

@@
 while a general discussion of this field is beyond the scope of this \paper,
-in the \sect{hier}, we present the basic probabilistic question and derive a
+in \sect{hier}, we present the basic probabilistic question and derive a
 computationally tractable inference procedure.
@@
 occurrence rate density of small planets with long orbital periods.
+
+Sections~\sectalt{likelihood} and~\sectalt{hier} provide a general pedagogical
+introduction to the methods used in this \paper.
+Readers looking to \emph{implement} a population inference are directed to
+\app{inv-det-eff} if measurement uncertainties are negligible or \sect{model}
+(especially \eqalt{money}) for problems with non-negligible uncertainties.
+Readers interested in our results---the inferred population of exoplanets and
+Earth-analogs---can safely skip to \sect{real} and continue to the discussion
+in \sect{comparison}.
 
@@
 
-Finally, for all the results in this \paper, we'll model the rate density as
-a piecewise constant step function
+Finally, we model the rate density as a piecewise constant step function
 \begin{eqnarray}\eqlabel{rate-model}
@@
 $\binarea_j$ are fixed \foreign{a priori}.
-
-In \app{inv-det-eff}, we use the model in \eq{rate-model} and neglect
-observational uncertainties, we can derive the analytic maximum likelihood
-solution for the step heights.
+In \app{inv-det-eff}, we use this parameterization and derive the analytic
+maximum likelihood solution for the step heights.
 This result is similar to and just as simple as the
@@
 the raw data with all the relevant information retained.
-In the context of \kepler, the catalog reduces the set of downloaded pixel
-time series (approximately 70,000 data points for the typical \kepler\ target)
-to probabilistic constraints on a handful of physical parameters---\entry\
-from above---like the orbital period and planetary radius.
+In the context of \kepler, the catalog reduces the set of downloaded time
+series (approximately 70,000 data points for the typical \kepler\ target) to
+probabilistic constraints on a handful of physical parameters---\entry\ from
+above---like the orbital period and planetary radius.
 If we take this set of parameters $\{\entry_k\}$ as \emph{sufficient
@@
 set of light curves for a particular occurrence rate density}.
+
+\dfmtodo{Add a few more pedagogical words here.}
 
@@
 there is an unmeasured sharp feature in the occurrence rate density near the
-location of Earth and our method is the most conservative extrapolation
+location of Earth but our method is the most conservative extrapolation
 technique published to date.
@@
 Since their model implicitly assumes a constant rate density across the bin,
-the differential rate is just their number divided by the bin area.
+the differential rate is just their number divided by the bin volume.
 This rate density (rate divided by bin volume) is what is shown as a
@@
 function of log period while the distribution for small planets seems to turn
-over at a relatively short period (around 50 days) and decrease for larger
+over at a relatively short period (around 50 days) and decrease for longer
 periods.
@@
 \section{Comparison with previous work}
+\sectlabel{comparison}
 
@@
 based on \emph{exactly the same dataset} (\citealt{petigura}).
-This inconsistency is due to the different assumptions made the detailed cause
-merits some investigation.
+This inconsistency is due to the different assumptions made and the detailed
+cause merits some investigation.
 The two key differences between our analysis and previous work are \emph{(a)}
-the form of the extrapolation function and \emph{(b)} the presence of
+the form of the extrapolation function, and \emph{(b)} the presence of
 measurement uncertainties on the planet radii.
@@
 There are a few factors to consider when comparing to this to our analysis.
-Firstly, while \citet{dong} did use a somewhat flexible extrapolation
-function than (a power law in log period), it is still very restrictive
-considering the large range of periods; a broken power law might have been
-more applicable.
+Firstly, while \citet{dong} fit a power law in log period, this is still a
+very restrictive model when considering this large range of periods.
+A broken power law might be more applicable.
 Furthermore, their analysis did not incorporate the effects of measurement
@@
 without further investigation but it should, intuitively, cause any inference
-based on the \citet{petigura} sample an underestimate of the \True\ rate.
+based on the \citet{petigura} sample to be an underestimate of the \True\
+rate.
 
@@
 published by \citet{petigura}.
-This discrepancy can be largely attributed to the rigidity of the assumptions
-about the period distribution.
+This discrepancy can be attributed to both the rigidity of the assumptions
+about the period distribution and the effects of non-negligible measurement
+uncertainties.
 Our extrapolation is also less confident than previous measurements.

### 2014-08-28 code/population.py
adding colorbar to figures

@@ -171,7 +171,8 @@ class Population(object):
 
     def plot_2d(self, thetas, ranges=None, censor=None, catalog=None,
                 err=None, labels=None, top_axes=None, literature=None,
-                lit_style={}, true=None, true_style={}, alpha=0.2):
+                lit_style={}, true=None, true_style={}, alpha=0.2,
+                norm=1.0, vmin=None, vmax=None):
         assert len(self.base) == 2
 
         # Pre-compute the ranges and allowed ranges.
@@ -201,9 +202,17 @@ class Population(object):
 
         # Plot the occurence image.
         img = np.median(grids, axis=0).T
-        m = ax.pcolor(self.base[0], self.base[1], np.exp(img),
-                      cmap="gray", alpha=0.6)
+        m = ax.pcolor(self.base[0], self.base[1], np.exp(img) / norm,
+                      cmap="gray", alpha=0.6, vmin=vmin, vmax=vmax)
         m.set_rasterized(True)
+
+        # Make the colorbar axis.
+        ax_cb = pl.axes([0.74, 0.72, 0.02, 0.18])
+        cb = pl.colorbar(m, ax_cb)
+        cb.locator = MaxNLocator(5)
+        cb.update_ticks()
+        ax_cb.set_ylabel(r"$\Gamma(w)$", rotation=270)
+        ax_cb.yaxis.set_label_coords(6.0, 0.5)
 
         # Plot the occurence histograms.
         ys = [logsumexp(grids

### 2014-08-28 code/results.py

@@ -151,6 +151,6 @@ fig.savefig(os.path.join(bp, "rate.pdf"))
 somesamples = samples[np.random.randint(len(samples), size=50), :]
 fig = pop.plot_2d(somesamples, censor=model.censor, catalog=np.log(catalog),
                   err=err, true=truth, labels=labels, top_axes=top_axes,
-                  literature=literature)
+                  literature=literature, norm=nstar, vmin=0, vmax=None)
 fig.savefig(os.path.join(bp, "results.png"), dpi=300)
 fig.savefig(os.path.join(bp, "results.pdf"))

### 2014-08-28 code/population.py
adding contour labels to plots

@@ -19,6 +19,7 @@ from itertools import izip
 
 import numpy as np
 
+import scipy.ndimage
 from scipy.misc import logsumexp
 from scipy.linalg import cho_factor, cho_solve
 
@@ -238,8 +239,12 @@ class Population(object):
         if censor is not None:
             x, y = censor.bins
             z = np.exp(censor.lncompleteness[1:-1, 1:-1])
-            ax.contour(x[:-1]+0.5*np.diff(x), y[:-1]+0.5*np.diff(y), z.T, 3,
-                       colors="k", linewidths=1, alpha=0.6)
+            z = scipy.ndimage.filters.gaussian_filter(z, 1.5)
+            c = ax.contour(x[:-1]+0.5*np.diff(x), y[:-1]+0.5*np.diff(y), z.T,
+                           3, colors="k", linewidths=1, alpha=0.6, vmin=0,
+                           vmax=1)
+            locs = [(5.8, 0.7), (5.8, 1.0), (5.8, 2.7)]
+            pl.clabel(c, fontsize=12, inline=1, manual=locs, fmt="%.2f")
 
         # Plot the data.
         if catalog is not None:

### 2014-08-28 document/ms.tex
comments from morton

@@
 inference but that exercise is beyond the scope of the current \paper.
+To complicate matters, the detection probability of a transit is also a
+non-trivial function of the duration.
+To account for this effect, so non-circular orbits should also be injected
+when measuring the survey completeness.
 
@@
 catalog, the \True\ expected number of transiting Earth-like exoplanets
-orbiting Sun-like stars is probably larger than the values in \eq{ntransit}!
+orbiting Sun-like stars is almost certainly larger than the values in
+\eq{ntransit}!
 

### 2014-08-28 code/results.py
reversed Petigura errorbars again

@@ -80,10 +80,16 @@ print([b.shape for b in pop.bins])
 def xmap(f, i):
     return (f(*x) for x in i)
 
+print("Hyper:")
+
+h_mu = np.mean(hyper[-ntot:][::thin_by, :2], axis=0)
+h_std = np.std(hyper[-ntot:][::thin_by, :2], axis=0)
+print("\n".join(xmap("{0} ± {1}".format, zip(h_mu, h_std))))
+
 h_mu = np.mean(np.exp(0.5 * hyper[-ntot:][::thin_by, 2:]), axis=0)
 h_std = np.std(np.exp(0.5 * hyper[-ntot:][::thin_by, 2:]), axis=0)
-print("Hyper:")
 print("\n".join(xmap("{0} ± {1}".format, zip(h_mu, h_std))))
+
 print()
 
 with h5py.File(os.path.join(bp, "samples.h5"), "w") as f:
@@ -105,7 +111,7 @@ if os.path.exists(fn):
     ext = np.array(open(os.path.join(bp, "extrap.txt"), "r").read().split(),
                    dtype=float) / 42557.0
 else:
-    ext = np.array([0.057, -0.022, 0.017])
+    ext = np.array([0.057, -0.017, 0.022])
 
 ext /= factor
 print(ext)

### 2014-08-28 document/ms.tex

@@
 \newcommand{\gammaearth}{{\ensuremath{\rate_\oplus}}}
+
+\newcommand{\resultsurl}{\url{http://dx.doi.org/10.6084/m9.figshare.1051864}}
 
@@
 and \fig{simulation-results}, respectively and the data are available
-online\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.1051864}}.
+online\footnote{\resultsurl}.
 
@@
 model.
-The resulting MCMC chain is available online\footnote{\url{%
-http://dx.doi.org/10.6084/m9.figshare.1051864}}.
+The resulting MCMC chain is available online\footnote{\resultsurl}.
 
@@
 
-\dfmtodo{Do we want to add a comment about giant planets? Or maybe some sort
-of cumulative distribution plot/code snippet?}
-
-\dfmtodo{Add sentence listing posterior constraints on hyperparameters.}
+Although they are mainly nuisance parameters, we also obtain posterior
+constraints on the hyperparameters \mean\ and \smoothpars.
+In particular, the constraints on the length scales in $\ln \period$ and $\ln
+\radius$ are $\smooth_\period = 3.65 \pm 1.03$ and $\smooth_\radius = 0.65 \pm
+0.12$ respectively.
+Both of these scales are larger than a bin in their respective dimension.
+For completeness we also find the following constraints on the other
+hyperparameters
+\begin{eqnarray}
+\mean = 5.44 \pm 1.56 &\quad& \ln\smooth_0 = 1.68 \pm 0.72 \quad.
+\end{eqnarray}
+The MCMC chains used to compute these values is available
+online\footnote{\resultsurl}.
 
@@
 In this spirit, we have released the results of this paper as posterior
-samplings\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.1051864}} for
+samplings\footnote{\resultsurl} for
 the occurrence rate density function.

### 2014-08-31 document/ms.tex
updating results DOI
adding comparison figure and changing labels (#19)

@@
 
-\newcommand{\resultsurl}{\url{http://dx.doi.org/10.6084/m9.figshare.1051864}}
+\newcommand{\resultsurl}{\url{http://dx.doi.org/10.5281/zenodo.11507}}
 
@@
 
-\dfmtodo{Add a few more pedagogical words here.}
-
 In this equation, we're making the further assumption that the catalog treated
@@
 \begin{eqnarray}
-\mean = 5.44 \pm 1.56 &\quad& \ln\smooth_0 = 1.68 \pm 0.72 \quad.
+\mean = 5.44 \pm 1.56 &\quad\mathrm{and}\quad&
+\ln\smooth_0 = 1.68 \pm 0.72 \quad.
 \end{eqnarray}
@@
 else the same.
-This analysis (labeled ``negligible uncertainties'' in \fig{comparison}) gives
+This analysis (labeled ``uncertainties ignored'' in \fig{comparison}) gives
 the result
@@
 In this spirit, we have released the results of this paper as posterior
-samplings\footnote{\resultsurl} for
-the occurrence rate density function.
+samplings\footnote{\resultsurl} for the occurrence rate density function.
 
@@
 \emph{linear} instead of \emph{uniform}.
-The ``negligible uncertainties'' value is given by applying the model
+The ``uncertainties ignored'' value is given by applying the model
 developed in this \paper\ but with the error bars on radius artificially

### 2014-08-31 document/figures/make_comparison.py

@@ -10,7 +10,7 @@ values = [
     ("Petigura \emph{et al.}\ (2013)", 0.119, 0.046, 0.035),
     ("Dong \& Zhu (2013)", 0.086, 0.048, 0.031),
     ("linear extrapolation", 0.072, 0.087, 0.046),
-    ("negligible uncertainties",
+    ("uncertainties ignored",
      0.0397591956934, 0.0309881584307, 0.0190620520124),
     ("Foreman-Mackey \emph{et al.}\ (2014)",
      0.0188817594233, 0.0193141328059, 0.0104440671941),

### 2014-09-01 document/publish.py
getting all the figures for publication

@@ -25,7 +25,8 @@ def rename(fn):
     return os.path.split(a)[1] + "-" + b, fn
 
 
-for a, b in imap(rename, glob.glob("figures/*/*.pdf")):
+for a, b in imap(rename, glob.glob("figures/*/*.pdf")
+                 + glob.glob("figures/*.pdf")):
     shutil.copyfile(b, os.path.join(TMPDIR, a))
     tex = tex.replace(b, a)
 
